### 2021-04-02 tex/targetvol.tex
latex

\documentclass[a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{comment}
\usepackage{dcolumn,multirow,abstract,fancyhdr}
\usepackage{geometry}
\usepackage{mathtools} % serve per fare le formule matematiche 
\usepackage{textcomp}
\usepackage{amsfonts}
\usepackage{dsfont}
\usepackage{upgreek}
\geometry{a4paper, top=5cm, bottom=5cm, left=4.5cm, right=4.5cm} %
%\pagestyle{fancy} 
\renewcommand{\headrulewidth}{0pt}
\fancyhead[R]{}
\lhead{\footnotesize\slshape Daluiso, Pallavicini, Nastasi, Polo, Optimal Strategies for Target Volatility Contracts}
\chead{}
\rhead{\thepage}
\lfoot{}
\cfoot{}
\rfoot{}
\usepackage[font=scriptsize,labelfont=bf]{caption}
\usepackage[pdfa]{hyperref}
\hypersetup{colorlinks=true,citecolor=black,filecolor=black,linkcolor=black, urlcolor=black}
\usepackage{cleveref}
\usepackage{natbib}
\setcitestyle{authoryear,open={(},close={)}}
\newtheorem{theorem}{Theorem}[section] 
\newtheorem{lem}{Lemma}[section]
\newtheorem{defi}{Definition}[section]
\newtheorem{prop}{Proposition}[section]
\newtheorem{rem}{Remark}[section]
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\newenvironment{eqsys}{\begin{equation}\begin{dcases}}{\end{dcases}\end{equation}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{page}{1} % start with first page

\title{Optimal Strategies for Target Volatility Contracts}

\author{
Roberto Daluiso\thanks{Banca IMI Milan, roberto.daluiso@bancaimi.com}
\and 
  Andrea Pallavicini\thanks{Imperial College London and Banca IMI Milan, a.pallavicini@imperial.ac.uk} 
\and
  Emanuele Nastasi\thanks{Exprivia, emanuele.nastasi@exprivia.com}
  \and Stefano Polo \thanks{Intesa Sanpaolo Milan, stefano.polo@intesasanpaolo.com}% indicates same as 2nd thanks
}

\date{} % leave empty
\begin{document} % goes here

\maketitle

\begin{abstract}
...
\end{abstract}

\smallskip
\noindent
\textbf{JEL classification codes:} C63, G13.\\
\textbf{AMS classification codes:} 65C05, 91G20, 91G60.\\
\textbf{Keywords:} Asset Allocation, Target Volatility, Funding Costs, Hedging Cost, Reinforcement Learning, Proximal policy optimization. 
\clearpage
\tableofcontents
 % \vspace{\fill\footnotesize \noindent The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.}
\clearpage
\pagestyle{fancy}  
\section{Introduction}
In the recent years portfolio managers are exposed to very low interest rates and quickly changing market volatilities. An effective solution to control risks under such environment is given by target volatility strategies (TVSs) (also known as constant volatility targeting) which are able to preserve the portfolio at a predetermined level of volatility. A TVS is  a  portfolio  of  risky  assets  (typically  equities)  and  a  risk-free  asset dynamically re-balanced with the aim of maintaining the overall portfolio volatility level closed to some target value. This products were initially offered in the Asian markets, see for instance the reports of \cite{Chew} and \cite{XUE} which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \cite{Morrison}.

In literature TVSs are tested to investigate their performances in term of realized returns, see for instance \cite{Hocquard28} and \cite{PERCHET}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim}. Moreover, in pricing literature derivative contracts on TVS are studied, see in particular \cite{Di_Graziano}, \cite{Grasselli}, and \cite{Albeverio}.

In this contribution we study a problem related to TVS. We deal with the funding costs coming from hedging the risky assets underlying the TVS. We consider the point of view of a bank selling a protection to a portfolio manager on the capital invested in a TVS. the portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS. Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario. Hence, the pricing problem becomes a dynamical control problem over the risky portfolio composition. In our contribution we describe the dynamical control problem, and we derive an analytical solution in the simple case of the risky assets driven by a Black-Scholes (BS) model. Then, we tackle the problem in the general case by using reinforcement learning (RL) algorithms.
\\

The paper is organized ad follows. In \Cref{sec:TVS} we describe the dynamics of a TVS in presence of valuation adjustments. Then, in \Cref{sec:Derivative} we derive the analytical results, and in \Cref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem. We conclude the paper with \Cref{sec:Numerical_Results} where we present the numerical results obtained in this work.

\section{Target Volatility Strategy}\label{sec:TVS}
In a TVS the fund manager selects an allocation strategy aiming at stabilizing the portfolio volatility to a target level. Clients investing in the fund pay a running fee for the service and their capital is protected. The fund manager usually buys an option on the TVS to ensure capital protection. For instance, the capital can be protected by buying a put option. In this case, we can write the net asset value (NAV) $A_t$ of the strategy as given by\footnote{Here we neglect discounting factors.}
\begin{equation}
    A_t \coloneqq \max\{V_t,K\} = V_t + (K-V_t)^+,
\end{equation}
where $V_t$ is the price process of the strategy, and $K$ is the guaranteed capital. On the other hand, the fund manager ca n replicate the payoff by means of the put-call parity by investing the capital in a low-risk asset and buying a call on the strategy
\begin{equation}
    A_t = K + (V_t-K)^+.
\end{equation}
In this way the TVS is only defined in the two contracts client-fund and fund-bank. The fund manager is not implementing the strategy by trading in the market, and he is not subject to additional costs to access the market. On the other way, the bank is paying such costs since she is actively hedging the call option sold to the manager.

The bank trading activity implemented to actively hedge the option requires funding the collateral procedures of the hedging instruments along with any lending/borrowing fee. The price of a financial product sold by the bank is modified to include any valuation adjustment due to the trading activity. We proceed by defining the price process for the TVS so that we can highlights the impact of valuation adjustments.
\subsection{The strategy Price Process}
We work on a filtered probability space $\left(\Omega, \mathcal{F}, \{\mathcal{F}_t\}_{t\geq0}, \mathbb{P}  \right)$ satisfying the usual assumptions for a market model, where $\mathbb{P}$ is the physical probability measure representing the actual distribution of supply and demand shocks on equities prices.

We consider a fund trading a basket of $n$ risky securities with price process $S_t^i$ with $i=1,\dots,n$ funded with a cash account $B_t$ accruing at $r_t$. Any dividend paid by the securities is re-invested in the fund, so that we limit our analysis to total return securities, namely we assume that holding the security is self-financing. Here, we assume that the TVS is implemented in continuous time, even if in the practice we can implement the strategy only on a discrete set of dates. We introduce the deflated gain process $\Bar{G}_t^i$ associated to the risky securities as given by
\begin{equation}
    \Bar{G}_t^i \coloneqq \Bar{S}_t^i + \Bar{D}_t^i,
\end{equation}
where we define the deflated price\footnote{We use bar notation for deflated quantities: processes expressed in terms of $B_t$.} and cumulative dividend processes as 
\begin{equation}
    \bar{S}_t^i \coloneqq \frac{S_t^i}{B_t}, \quad \bar{D}^i_t \coloneqq \int_0^t\frac{d\pi_u^i}{B_u} + \int_0^t\frac{d\psi_u^i}{B_u},
\end{equation}
where $\pi^i_t$ represents the cumulative contractual-coupon process paid by the security, and $\psi_t^i$ represents the cumulative valuation adjustments.

Valuation adjustments (XVAs) is a topic widely discussed in the literature. We refer to \cite{Brigo} for a discussion. Since fund managers allocating TVS usually rely on Equity assets, here we use the results of \cite{Gabrielli} which  analyze the valuation adjustments for equity products. We can write
\begin{equation}
    \psi_t^i \coloneqq \int_0^t S_u^i\mu_u^i du,
\label{eq:XVA_equity}\end{equation}
where we call $\mu_t^i$ cost of carry, which basically represents the hedging costs for the $i$-th security.

Then, we introduce the strategy price process $V_t$, and we define the deflated gain process $\Bar{G}_t^V$ as given by
\begin{equation}
    \Bar{G}_t^V \coloneqq \frac{V_t}{B_t} + \int_0^t\frac{V_u\phi_u}{B_u}du,
\end{equation}
where $\phi_t$ are the running fees earned by the fund manager for his activity. We assume that the strategy is self-financing, so that we can write
\begin{equation}
    d\Bar{G}_t^V = q_t \cdot d\bar{G}_t,
\label{eq:self_financing}\end{equation}
where $q_t^i$ is the quantity invested in the $i$-th security\footnote{In all formulae we use dot notation for scalar product between vectors, i.e. $a \cdot b = \sum_i a_i b_i$, or between matrix and vector, i.e. $A \cdot b = \sum_j a_{ij}b_j$ or $b \cdot A = \sum_i b_i a_{ij}$.}.

Now, in order to prevent arbitrages, we assume the existence of a risk-neutral measure $\mathbb{Q}$ under which the deflated gain processes of all traded security are martingales. Under this assumption we are able to derive the drift conditions on the security price processes, and in turn on the strategy price process.
\begin{equation}
    \forall T >t \quad \bar{G}_t^i = \mathbb{E}_t \left[\bar{G}_T^i\right] \quad\Longrightarrow \quad dS_t^i = r_tS_t^idt-d\pi_t^i-d\psi_t^i + dM_t^i,
\label{eq:risk_neutral}\end{equation}
where $M_t^i$ are martingale under $\mathbb{Q}$. If we substitute this expression for the security dynamics into the definition of the strategy we can check that the price process of the strategy is accruing at a cash account rate rate $r_t$ compensated for the fund manager fees
\begin{equation}
    dV_t = V_t(r_t-\phi_t)dt + dM_t^V,
\end{equation}
with $M_t^V$ martingale under $\mathbb{Q}$. Notice that,as expected from non-arbitrage considerations, the coupons paid by each security appear only in the drift of the security price process, but they do not impact the drift of the strategy.

Yet, the strategy priced by $V_t$ cannot be described in the contract between the parties, since \Cref{eq:self_financing} depends via the security gain processes on the valuation adjustment $\psi_t^i$, which is specific of the investor pricing the strategy. Thus, the TVS defined in the contract will be
\begin{equation}
    d\bar{I}_t \coloneqq q_t \cdot \left(d\bar{S}_t + \frac{d\pi_t}{B_t} \right) - \bar{I}_t\phi_t dt \quad \text{with }I_0 = V_0, 
\label{eq:TVS_first}\end{equation}
leading to the following price process dynamics
\begin{equation}
    dI_t = I_t(r_t-\phi_t)dt -q_t\cdot \psi_t +dM_t^I,
\end{equation}
with $M_t^I$ martingale under $\mathbb{Q}$. In this case we observe that $I_t$ depends explicitly both on the valuation adjustments and on the allocation strategy. Indeed, if we substitute the valuation adjustments with their explicit expression (\Cref{eq:XVA_equity}), we get
\begin{equation}
    dI_t = I_t(r_t-\phi_t)dt - q_t \cdot S_t \mu_t dt + dM_t^I,
\end{equation}
where we can see the dependency on cost of carry $\mu_t^i$.
\subsection{The Volatility Targeting Constraint}
In a typical TVS the fund manager selects a risky-asset portfolio with a specific time-dependent allocation strategy expressed by means of the vector of relative weights $\alpha_t$, along with a risk-free asset, which we can identify with the bank account $B_t$. Usually TVSs are total-return products; thus we are justified in assuming $\pi_t=0$. Thus we can write \Cref{eq:TVS_first} as given by
\begin{equation}
    \frac{dI_t}{I_t}= \omega_t \alpha_t \cdot \frac{dS_t}{S_t} + \left(1-\omega_t  \alpha_t \cdot \mathds{1} \right)\frac{dB_t}{B_t}- \phi_t dt,
\label{TVS_elegant}\end{equation}
where $\mathds{1}$ is a $n$-dimensional vector of ones and $\omega_t\in[0,1]$ is determined so that the strategy log-normal volatility is kept constant, namely
\begin{equation}
        \omega_t: \quad \mathrm{Var}_t[dI_t] = \bar{\sigma}^2I_t^2dt,
\end{equation}
where $\Bar{\sigma}$ is the target volatility value.

We assume a generic continuous semi-martingale dynamics under the risk-neutral measure for the underlying securities, so that we can write
In practice, this means that the fund manager will select a risky-portfolio choosing $\alpha_t$ equities from the universe where he can trade and after that his choices will be scaled by the automatic target volatility algorithm\footnote{We recall that the universe of assets where the manager can trade and the value of $\bar{\sigma}$ are written in the contract.} $\omega_t$.

To derive the expression for $\omega_t$ we need to assume a generic continuous semi-martingales dynamics under the risk-neutral measure for the underlying securities, so that we can write \Cref{eq:risk_neutral} as 
\begin{equation}
    \frac{dS_t^i}{S_t^i} = \left(r_t - \mu_t^i \right)dt + \nu_t^i \cdot dW_t,
\label{Equity_process}\end{equation}
where $\nu_t$ is an adapted matrix process ensuring the existence of a solution for the SDE and $W_t$ is a $n$-dimensional vector of Brownian motions under $\mathbb{Q}$. Under these assumptions we can derive an expression for $\omega_t$, and we get\footnote{In all formulae the norm for a vector $a$ is defined as $\|a\|\coloneqq\sqrt{a\cdot a}$.}
\begin{equation}
    \omega_t = \frac{\Bar{\sigma}}{\|\alpha_t\cdot \nu_t \|}.
\end{equation}
Hence, putting this last result in the dynamics of $I_t$ we obtain

\begin{equation}
     \frac{dI_t}{I_t} = \left(r_t -  \phi_t     - \frac{\bar{\sigma} \alpha_t}{\|\alpha_t \cdot \nu_t \|}  \cdot \mu_t \right)dt+ \frac{\bar{\sigma}\alpha_t }{\|\alpha_t \cdot \nu_t \|} \cdot \nu_t \cdot dW_t,
\label{eq:TVS_last}\end{equation}
where we can see, as expected, that the strategy grows at the risk-free rate but for adjustments due to valuation adjustments and fees.
\section{Derivative Pricing}\label{sec:Derivative}
A derivative contract on the TVS with maturity $T$ can be defined as
\begin{equation}
    V_0 \coloneqq \sup_\alpha \mathbb{E}_0\left[\int_0^T D(0,u;\zeta)d\pi_u(\alpha)\right],
\end{equation}
where $D(0,T;\zeta)$ is the discount factor with rate $\zeta_t$, inclusive of the derivative valuation adjustments, and $\pi_t$ is the cumulative coupon process paid by the derivative, and it depends on the allocation strategy since in turn the TVS depends on it via the valuation adjustments. We take the supremum over the strategies since we do not have any information on the future activity of the fund manager.
\subsection{European Options}
If the derivative contract depends only on the marginal distribution of $I_t$ at maturity (a European payoff), we are able to prove that exists an optimal strategy, and we are able to calculate it. We consider the following pricing problem
\begin{equation}
    V_0 \coloneqq \sup_\alpha \mathbb{E}_0\left[D(0,T;\zeta)\Phi(I_t(\alpha))\right],
\end{equation}
where $\Phi$ is the payoff function of the derivative. We start by introducing the Markovian projection of the dynamics followed by $I_t$. We name it $I_t^{\text{MP}}$, and we get by applying the Gy\"ongy Lemma
\begin{equation}
    \frac{dI_t^{\text{MP}}}{I_t^{\text{MP}}} \coloneqq \left(r_t - \ell \left(t,I_t^{\text{MP}}\right)\right)dt + \bar{\sigma}dW_t^{\text{MP}},
\label{eq:markovian_projection}\end{equation}
where the local drift is defined as
\begin{equation}
    \ell \left(t,K\right) \coloneqq \bar{\sigma} \mathbb{E}_0\left[\frac{\mu_t \cdot \alpha_t}{\|\alpha_t \cdot \nu_t \|}\bigg|I_t = K\right],
\end{equation}
and $W_t^{\text{MP}}$ is a Brownian motion under the risk-neutral measure. Notice that the diffusion coefficient collapses to the target volatility value $\bar{\sigma}$. Since European payoffs depends only on the marginal distribution at maturity, they can be calculated by means of the Markovian projection $I_t^{\text{MP}}$, namely
\begin{equation}
V_0 \coloneqq \sup_{\alpha} \mathbb{E}_0\left[D_0\left(T\right)\Phi\left(I_T^{\text{MP}}\left(\alpha\right)\right)\right].
\end{equation}
Hence, we have our first result valid only if valuation adjustments can be neglected:
\begin{prop}
A European payoff on the TVS can be calculated by assuming any allocation in the underlying risky basket if all the underlying
securities grow under the risk-neutral measure at the risk-free rate without any valuation adjustment, namely if we can write $\mu_t=0$.
\end{prop}
\begin{rem}[Existence of the Solution in the General Case]
In a more general settings we are not able to find an explicit solution. A proof of the existence of the solution in a general setting is missing. This is a stochastic optimal control problem where by homogeneity we can suppose that $\alpha_t$ lives in a compact domain, namely (a subset of) the unit simplex. A least if $r_t$, $\mu^i_t$ and $\nu^i_t$ are (uniformly) bounded, and if the eigenvalues of $\nu_t$ are (uniformly) bounded away from zero, then drift and diffusion should be uniformly Lipschitz in $\alpha_t$, and classical theorems should exist. 
\end{rem}
\subsection{Stochastic Optimal Control Problem}
In presence of valuation adjustments we need to solve the full optimization problem. We discretize the optimal strategy $\alpha_t$ as
\begin{equation}
    \alpha_t \coloneqq \sum_k \mathbf{1}_{ \{t \in [T_{k-1}, T_k)\}}\alpha_{T_{k-1}},
\label{eq:piecewise_strategy}\end{equation}
according to a time grid $\mathcal{T}:=\{T_0,...,T_k,...,T_m\}$ with $T_0:=t$ the pricing date and $T_m :=T$ the maturity of the option. 
Therefore we can apply the dynamic programming principle to express the optimal $\alpha_t$ at time $T_{k-1}$ as
\begin{equation}
    \alpha_{T_{k-1}}:=\argmax_{\alpha} \left\{\mathbb{E}_{T_{k-1}}\left[D_{T_{k-1}}\left(T_{k} \right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}(\alpha)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right]\right\},
\end{equation}
where $V_{T_k}$ is the option value at time $T_k$ and $X$ is any Markovian state such that the drift and the diffusion coefficient of $I_t$ are a function of $\left(X_t,I_t,\alpha_t\right)$. We calculate $I_{T_k}\left(\alpha_{T_{k-1}}\right)$ for any given strategy $\alpha_{T_{k-1}}$ by a suitable discretization of \eqref{eq:TVS_last} starting from $X_{T_{k-1}}$ and $I_{T_{k-1}}$.

Thus the derivative price is given by:
\begin{equation}
    V_{T_{k-1}}\left(X_{T_{k-1}}, I_{T_{k-1}}\right)=\mathbb{E}_{T_{k-1}}\left[D_{T_{k-1}}\left(T_{k}\right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right],
\label{recursion}\end{equation}
while the iteration starts from maturity date where the boundary condition is set equal to the payoff function:
\begin{equation}
    V_{T_m} = \Phi\left(I_{T_m}\right).
\end{equation}
\subsection{Black and Scholes Model}
In the Black and Scholes model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$. As a consequence, the local drift can be written as
\begin{equation}
     \ell \left(t,K\right) = \bar{\sigma}   \frac{\mu\left(t\right) \cdot \alpha\left(t,K\right)}{\|\alpha\left(t,K\right) \cdot \nu\left(t\right)\|},  
\end{equation}
so that the optimization problem can be solved looking only at the Markovian projection without simulating all the Brownian motions $W_t$. Notice that we are indicating the dependency on time in parenthesis to highlight that in this formula all the quantities are deterministic function of time.

A direct consequence is the following proposition, which is relevant for plain vanilla options on TVS.

\begin{prop}
When the underlying securities follow a Black and Scholes model with deterministic rates, the optimal strategy for a monotone payoff consists in minimizing the local drift function, indipendently of the current state $I_t$
\begin{equation}
    \alpha^*(t) \coloneqq \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|}.
\label{eq:BS_optimal_strategy}\end{equation}
\end{prop}
The absence of stochastic elments in \Cref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori}; in fact one can solve the optimization problema once for all for each $t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities. Once $\alpha^*$ is known, then one can price the payoff by Monte Carlo simulation. 
\subsubsection{Free strategy: closed form solution}
In absence of constraints on $\alpha \coloneqq \alpha(t)$, the problem even admits a closed form solution. We set $\mu \coloneqq \mu(t)$ and $\nu \coloneqq \nu(t)$. By homogeneity we can rewrite \eqref{eq:BS_optimal_strategy} as
\begin{equation}
	\boxed{
		\begin{aligned}
			&\text { minimize } \alpha \cdot \mu\\
			&\text { subject to } \|\alpha \cdot \nu\|^2=1
	\end{aligned}}
\end{equation}
By setting the Lagrangian function associated with the problem
\begin{equation}
	\mathcal{L}\left(\alpha, \lambda\right)=\alpha \cdot \mu-\lambda\left(\|\alpha \cdot \nu\|^2-1\right) \, ,
\end{equation}
we obtain the first order conditions\footnote{We used the property of the Cholesky factor of the covariance matrix $\Sigma = \nu  \nu^T$, where $\nu^T$ is the transpose matrix of $\nu$.}
\begin{eqsys}
	\frac{\partial \mathcal{L}}{\partial \alpha}=\mu -2\lambda \Sigma \cdot \alpha=0 \\
	\frac{\partial \mathcal{L}}{\partial \lambda}= \|\alpha \cdot \nu\|^2-1=0
	\, .
\end{eqsys}
Then, by applying simple algebra, we obtain the analytical form of the free optimal strategy
\begin{equation}
	\alpha^* = \pm \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
\end{equation}
We take the minus sign to get the minimum value of the TVS local drift.

In our work the definition of $\alpha$ is general, thus it is possible to consider the case of constrained allocation strategy; in this case the problem \eqref{eq:BS_optimal_strategy} must be solved numerically.

\section{Reinforcement Learning}\label{sec:RL}
 Reinforcement Learning (RL) describes how an agent behaves in an environment so to maximize some notion of cumulative reward. The actions of the agent as a function of his observations of the environment are termed the agent policy. In our case the policy is the asset allocation weights $\alpha$, while the rewards are the costs generated by hedging a derivative contract on the TVS. Once the agent is trained, and the optimal policy is selected, we can run a Monte Carlo simulation to calculate the option price on the TVS.

 We adopt as learning strategy the PPO algoritm developed in \cite{ppo} and \cite{gae}. 
\cleardoublepage
\bibliographystyle{abbrvnat}
\bibliography{Bib}
\end{document}

### 2021-04-02 reinforcement/baseline_simulation/pricing/pricing.py
Code Speedup: no append functions in LV dynamics

@@ -286,40 +286,22 @@ class LV_model(PricingModel):
             else:
                 return exp(logmartingale)*self.forward
         else:
-           # logmartingale = np.zeros((Nsim.N_fixings,self.Ndim))
-            logX = np.zeros((Nsim,self.Ndim))
-           # correlated_wiener = np.array([])
-            logmartingale = np.array([])
-            vola_t = np.array([])
+            logmartingale = np.zeros((Nsim,N_times,self.Ndim))
+            vola_t = np.zeros((Nsim,N_times,self.Ndim))
             for i in range (N_times):
                 Z = random_gen.randn(Nsim,self.Ndim)
                 ep = corr_chole@Z.T   #matrix of correlated random variables
-               # correlated_wiener = np.append(correlated_wiener,ep)
                 for j in range(self.Ndim):
                     if i ==0:
                         vol = self.vol[j].intelligent_call(0,0.)
-                        #print("Asset "+str(j+1)+"vola ",vol)
-                        vola_t = np.append(vola_t,vol*np.ones(Nsim))
-                        logX[:,j]=-0.5*dt*(vol**2)+vol*sqrt(dt)*ep[j]
-                        logmartingale = np.append(logmartingale,logX[:,j])
+                        vola_t[:,i,j] = vol
+                        logmartingale[:,i,j] = -0.5*dt*(vol**2)+vol*sqrt(dt)*ep[j]
                     elif i!=0:
-                        vol = self.vol[j].intelligent_call(self.time_indexes_matrix[j,i-1],logX[:,j])
-                        vola_t = np.append(vola_t,vol)
-                        #print("Asset "+str(j+1)+"vola ",vol)
-                        logX[:,j]=logX[:,j]-0.5*dt*(vol**2)+vol*sqrt(dt)*ep[j]
-                        logmartingale = np.append(logmartingale,logX[:,j])
-                counter = i+1
-                if counter%self.N_grid == 0:
-                    #logmartingale[:,time_index,:] = logX
-                    time_index +=1
-                    if counter < N_times:
-                        dt = self.dt[time_index]
-            if normalization:
-                return (correlated_wiener.reshape(N_times,self.Ndim,Nsim)).transpose(2,0,1), (vola_t.reshape(N_times,self.Ndim,Nsim)).transpose(2,0,1)
-            else:    
[…]

### 2021-04-06 reinforcement/baseline_simulation/pricing/loadfromtxt.py
code upgrade with last changes

@@ -1,0 +1,23 @@
+from pricing.pricing import DiscountingCurve, EquityForwardCurve, ForwardVariance, LocalVolatilityCurve
+import numpy as np
+
+def LoadFromTxt(asset_name, folder = None):
+    dates_discounts, discounts = np.loadtxt(folder+"/discount_data.txt")
+    D = DiscountingCurve(reference=0.,discounts=discounts,dates=dates_discounts)
+    F, V, LV = [], [], []
+    for name in asset_name:
+        spot = np.loadtxt(folder+"/spot_data"+name+".txt")[0]
+        repo_dates, repo_rates = np.loadtxt(folder+"/repo_data_"+str(name)+".txt")
+        F.append(EquityForwardCurve(reference=0., discounting_curve = D, spot=spot, repo_rates=repo_rates,repo_dates=repo_dates,name=name,act="No"))
+        
+        spot_vola = np.loadtxt(folder+"/vola_data_"+str(name)+".txt")
+        vola_strikes = np.loadtxt(folder+"/strikes_vola_data_"+str(name)+".txt")
+        vola_dates = np.loadtxt(folder+"/maturities_vola_data_"+name+".txt")
+        V.append(ForwardVariance(reference=0.,maturities=vola_dates,strikes=vola_strikes,spot_volatility=spot_vola,strike_interp=spot,act="No", name=name))
+        
+        lv_pars = np.loadtxt(folder+"/LV_param_data_"+str(name)+".txt")
+        lv_money = np.loadtxt(folder+"/LV_money_vola_data_"+str(name)+".txt")
+        lv_dates = np.loadtxt(folder+"/LV_maturities_vola_data_"+name+".txt")
+        LV.append(LocalVolatilityCurve(lv_pars,lv_money,lv_dates,name))
+
+    return D, F, V, LV

### 2021-04-06 reinforcement/baseline_simulation/pricing/pricing.py

@@ -2,8 +2,6 @@ import numpy as np
 from scipy.interpolate import interp1d, PchipInterpolator
 from numpy import exp, sqrt, log
 from scipy.integrate import quad
-import matplotlib.pyplot as plt
-
 """Classes for my simulation"""
 
 class Curve:
@@ -19,14 +17,15 @@ class Curve:
 class EquityForwardCurve(Curve):
 
     def __init__(self, spot=None, reference=None, discounting_curve=None,
-                repo_rates=None, repo_dates=None, act = "No"):
+                repo_rates=None, repo_dates=None, name = None, act = "No"):
         self.spot = spot
+        self.asset_name = name
         self.reference = reference
         self.discounting_curve = discounting_curve
-        if act =="360":
+        if act =="ACT360":
             self.T = ACT_360(repo_dates,self.reference)
             self.T = self.T*360/365
-        elif act == "365":
+        elif act == "ACT365":
             self.T = ACT_365(repo_dates,self.reference)
         else:
             self.T = abs(repo_dates - self.reference)
@@ -38,7 +37,8 @@ class EquityForwardCurve(Curve):
             alpha = ((self.T[i])*(repo_rates[i])-(self.T[i-1])*
                      (repo_rates[i-1]))/(self.T[i]-self.T[i-1])
             self.q_values = np.append(self.q_values,alpha)
-        self.T = np.append(0.,self.T[:-1])
+        if self.T[0] != 0:
+            self.T = np.append(0.,self.T[:-1])
         self.q = interp1d(self.T, self.q_values, kind='previous',fill_value="extrapolate", assume_sorted=False) 
      #   print("Forward repo time grid",self.T)
       #  print("Forward repo rate: ", self.q_values)
@@ -55,12 +55,13 @@ class DiscountingCurve(Curve):
 
     def __init__(self, reference=None, discounts=None, dates=None, act = "No"):
         self.reference = reference
-        if act=="360":
+        if act=="ACT360":
             self.T = ACT_360(dates,self.reference)
-        elif act == "365":
+        elif act == "ACT365":
             self.T = ACT_365(dates,self.reference)
         else:
             self.T = abs(dates - self.reference)
+        r_instant = np.array([])
         if self.T[0] ==0:
             r_zero = np.array([0])   #at reference date the discount is 1
             r_zero = np.append(r_zero,(-1./((self.T[1:])))*log(discounts[1:]))
[…]

### 2021-04-06 reinforcement/baseline_simulation/pricing/targetvol.py

@@ -2,7 +2,7 @@ import numpy as np
 from numpy import exp, sqrt, log
 from pricing.pricing import piecewise_function, Curve, PricingModel, quad_piecewise
 from numpy.linalg import cholesky
-from scipy.optimize import minimize
+from scipy.optimize import minimize, LinearConstraint, Bounds
 from scipy.interpolate import interp1d
 
 
@@ -236,20 +236,35 @@ def optimization_only_long(mu=None, nu=None,seed = 1, N_trial=None, guess = None
     if guess is None and N_trial is None:
         N_trial = 1
     f = loss_function
-    cons = ({'type': 'eq','fun' : lambda x: np.sum(x)-1},{'type': 'ineq','fun' : lambda x: x})
+    # cons = ({'type': 'eq','fun' : lambda x: np.sum(x)-1},{'type': 'ineq','fun' : lambda x: x})
+    A = np.ones(len(mu))
+    x_low = np.array([1.])
+    x_up = np.array([1.])
+    bounds = Bounds(np.zeros(len(mu)),np.ones(len(mu)))
+    cons = LinearConstraint(A,x_low,x_up)
     if guess is None and N_trial is not None:
         np.random.seed(seed)
         r = np.zeros((N_trial,len(mu)))
         valutation = np.zeros(N_trial)
         for i in range (N_trial):
             x0 =np.random.uniform(0.,1.,len(mu))  #initial position for the optimization algorithm
-            res = minimize(f, x0, args=(mu,nu),constraints=cons)#,options={'ftol': 1e-30})
+            res = minimize(f, x0, args=(mu,nu),constraints=cons,bounds=bounds,method="SLSQP")#,options={'ftol': 1e-30})
             r[i] = res.x
             valutation[i] = f(res.x,mu,nu)
         #print("Minumum: ", np.min(valutation))
         return r[np.argmin(valutation)]
-    else:
-        return minimize(f, guess, args=(mu,nu),constraints=cons).x#,options={'ftol': 1e-30})
+    elif guess.ndim==1:
+        return minimize(f, guess, args=(mu,nu),constraints=cons,bounds = bounds, method = "SLSQP").x#,options={'ftol': 1e-30})
+    elif guess.ndim>1:
+        N_trial = len(guess)
+        r = np.zeros((N_trial,len(mu)))
+        valutation = np.zeros(N_trial)
+        for i in range(N_trial):
+            x0 = guess[i]
[…]

### 2021-04-06 reinforcement/baseline_simulation/run.py

@@ -1,85 +1,67 @@
-from mpi4py import MPI
+from pricing.pricing import LV_model
+from pricing.loadfromtxt import LoadFromTxt
+from pricing.targetvol import optimization_only_long, CholeskyTDependent, Strategy,Drift,  Markowitz_solution
 import numpy as np
-from scipy.interpolate import interp1d
-from time import time
-from scipy.optimize import minimize
-import xml.etree.ElementTree as ET
-from pricing.read_market import MarketDataReader, Market_Local_volatility
-from pricing.montecarlo import MC_Analisys, MC
-from pricing.targetvol import Drift
-from pricing.pricing import LocalVolatilityCurve, EquityForwardCurve, ForwardVariance, DiscountingCurve,piecewise_function,Vanilla_PayOff,PricingModel,LV_model
-from pricing.targetvol import optimization_only_long,optimization_limit_position, CholeskyTDependent, Strategy
+try:
+    from mpi4py import MPI
+except ImportError:
+    MPI = None
 
+if MPI is None or MPI.COMM_WORLD.Get_rank()==0:
+    rank = 0
+else:
+    rank = MPI.COMM_WORLD.Get_rank()
 
-comm = MPI.COMM_WORLD
-size = comm.Get_size()
-rank = comm.Get_rank()
-t_start = MPI.Wtime()
-Nsim = int(10)
+Nsim = int(8e3)
 run = 0
-Seed = rank+run
-print("Seme ",Seed)
-restart = 1
+Seed = run+rank#3000
+strategy = "only_long"   #free
+#print("Seme ",Seed)
+I_0 = 1.
+restart =1 
 Black = 0
 daily = 0
 N_equity = 2                                #number of equities
 target_vol = 5./100.
 T = 3.
-title = 'prove/m1/rest_final_price_lvinvestigation_2_asset_equalrepo_long_Nsim_'+str(Nsim)+'_vola'+str(target_vol)+'_maturity'+str(T)+'_Black'+str(Black)+'_daily'+str(daily)
-restart_file = 'prove/m1/rest_final_price_lvinvestigation_2_asset_equalrepo_long_Nsim_10_vola0.05_maturity2.0_Black0_daily0'
+title = 'day_new/d3/final_price_lvinvestigation_2_asset_equalrepo_long_Nsim_'+str(Nsim)+'_vola'+str(target_vol)+'_maturity'+str(T)+'_Black'+str(Black)+'_daily'+str(daily)+strategy
[…]

### 2021-04-06 reinforcement/envs/pricing/pricing.py

@@ -22,10 +22,10 @@ class EquityForwardCurve(Curve):
         self.asset_name = name
         self.reference = reference
         self.discounting_curve = discounting_curve
-        if act =="360":
+        if act =="ACT360":
             self.T = ACT_360(repo_dates,self.reference)
             self.T = self.T*360/365
-        elif act == "365":
+        elif act == "ACT365":
             self.T = ACT_365(repo_dates,self.reference)
         else:
             self.T = abs(repo_dates - self.reference)
@@ -55,9 +55,9 @@ class DiscountingCurve(Curve):
 
     def __init__(self, reference=None, discounts=None, dates=None, act = "No"):
         self.reference = reference
-        if act=="360":
+        if act=="ACT360":
             self.T = ACT_360(dates,self.reference)
-        elif act == "365":
+        elif act == "ACT365":
             self.T = ACT_365(dates,self.reference)
         else:
             self.T = abs(dates - self.reference)
@@ -85,9 +85,9 @@ class ForwardVariance(Curve):  #I calculate the variance and not the volatility 
     def __init__(self, reference=None, spot_volatility=None, strikes=None, maturities=None, strike_interp=None,name = None, act="No"):
         self.reference = reference #pricing date of the implied volatilities
         self.asset_name = name
-        if act=="360":
+        if act=="ACT360":
             self.T = ACT_360(maturities,self.reference)
-        elif act=="365":
+        elif act=="ACT365":
             self.T = ACT_365(maturities,self.reference)
         else:
             self.T = abs(maturities-self.reference)
@@ -289,40 +289,31 @@ class LV_model(PricingModel):
             else:
                 return exp(logmartingale)*self.forward
         else:
-           # logmartingale = np.zeros((Nsim.N_fixings,self.Ndim))
+            logmartingale = np.zeros((Nsim,N_times,self.Ndim))
             logX = np.zeros((Nsim,self.Ndim))
-           # correlated_wiener = np.array([])
-            logmartingale = np.array([])
-            vola_t = np.array([])
+            vola_t = np.zeros((Nsim,N_times,self.Ndim))#np.array([])
             for i in range (N_times):
                 Z = random_gen.randn(Nsim,self.Ndim)
                 ep = corr_chole@Z.T   #matrix of correlated random variables
[…]

### 2021-04-06 reinforcement/envs/pricing/targetvol.py

@@ -2,7 +2,7 @@ import numpy as np
 from numpy import exp, sqrt, log
 from envs.pricing.pricing import piecewise_function, Curve, PricingModel, quad_piecewise
 from numpy.linalg import cholesky
-from scipy.optimize import minimize
+from scipy.optimize import minimize, LinearConstraint, Bounds
 from scipy.interpolate import interp1d
 
 
@@ -87,8 +87,8 @@ class Strategy(Curve):
                 result = optimization_long_short_position(mu(self.T[i]), nu(self.T[i]), long_limit, short_limit,N_trial,seed)
             self.alpha_t[i] = result
         self.a_t = interp1d(self.T, self.alpha_t, axis=0, kind='previous',fill_value="extrapolate", assume_sorted=False)
-        print("Optimal strategy time grid :",self.T)
-        print("Optimal strategy through minimization: ",self.alpha_t)
+       # print("Optimal strategy time grid :",self.T)
+       # print("Optimal strategy through minimization: ",self.alpha_t)
 
 
     def Intuitive_strategy1(self, forward_curves=None, maturity_date=None):
@@ -231,20 +231,40 @@ def loss_function(x,mu,nu):
     """Target function to minimize"""
     return (x@mu)/np.linalg.norm(x@nu)
 
-def optimization_only_long(mu=None, nu=None,seed = None, N_trial=1):
+def optimization_only_long(mu=None, nu=None,seed = 1, N_trial=None, guess = None):
     """Constrained optimization with only long position and sum of weights equal to 1"""
-    np.random.seed(seed)
+    if guess is None and N_trial is None:
+        N_trial = 1
     f = loss_function
-    cons = ({'type': 'eq','fun' : lambda x: np.sum(x)-1},{'type': 'ineq','fun' : lambda x: x})
-    r = np.zeros((N_trial,len(mu)))
-    valutation = np.zeros(N_trial)
-    for i in range (N_trial):
-        x0 =np.random.uniform(0.,1.,len(mu))  #initial position for the optimization algorithm
-        res = minimize(f, x0, args=(mu,nu),constraints=cons)
-        r[i] = res.x
-        valutation[i] = f(res.x,mu,nu)
-    #print("Minumum: ", np.min(valutation))
-    return r[np.argmin(valutation)]
+    # cons = ({'type': 'eq','fun' : lambda x: np.sum(x)-1},{'type': 'ineq','fun' : lambda x: x})
[…]

### 2021-04-06 reinforcement/envs/tvs_env.py

@@ -19,7 +19,7 @@ class TVS_environment(gym.Env):
         self.strike_option = strike_opt
         self.target_vol = target_volatility
         self.T = maturity
-        self.Nsim = 1e4
+        self.Nsim = 1e3
         self.current_time = 0.
         n_observations = 365
         self.time_index = 0

### 2021-04-06 reinforcement/envs/tvs_lv_env.py

@@ -8,6 +8,7 @@ from envs.pricing.closedforms import European_option_closed_form
 from envs.pricing.targetvol import Drift, Strategy, TVSForwardCurve, TargetVolatilityStrategy, optimization_only_long
 from envs.pricing.fake_market_lv import load_fake_market_lv
 from envs.pricing.loadfromtxt import LoadFromTxt
+from envs.pricing.targetvol import optimization_only_long
 from envs.pricing.n_sphere import sign_renormalization
 
 class TVS_LV(gym.Env):
@@ -47,7 +48,7 @@ class TVS_LV(gym.Env):
         self.time_index = 0
         self.current_time = 0.
         self.simulation_index = 0
-        self.Nsim = 2
+        self.Nsim = 1e3
         """Loading market curves"""
         D, F, V, LV = LoadFromTxt(names, folder_name)
         self.correlation_chole = np.linalg.cholesky(correlation)
@@ -57,6 +58,8 @@ class TVS_LV(gym.Env):
         """Preparing the LV model"""
         self.model = LV_model(fixings=self.observation_grid, local_vol_curve=LV, forward_curve=F, N_grid = self.N_euler_grid)
         euler_grid = self.model.time_grid
+        mu_function = Drift(forward_curves=F)
+        self.mu_values  = mu_function(np.append(0.,euler_grid[:-1]))
         self.r_t = D.r_t(np.append(0.,euler_grid[:-1]))
         self.discount = D(self.T)
         self.dt_vector = self.model.dt
@@ -97,7 +100,6 @@ class TVS_LV(gym.Env):
         elif self.constraint == "long_short_limit":
             action = sign_renormalization(action,self.how_long,self.how_short)
             s = np.sum(action)
-                
         self.time_index = self.time_index + 1
         self.current_time = self.observation_grid[self.time_index]
         self.current_logX = self.logX_t[self.time_index]
@@ -108,6 +110,8 @@ class TVS_LV(gym.Env):
             idx = index_plus + i 
             Vola =  self.sigma_t[idx]*self.Identity
             nu = Vola@self.correlation_chole
+            if i==0:
+                action = optimization_only_long(self.mu_values[idx], nu,seed=1, guess = np.array(([0.4,0.6],[0.6,0.4])))        
             omega = self.target_vol/np.linalg.norm(action@nu)
             self.I_t = self.I_t * (1. + omega * action@self.dS_S[idx] + dt * self.r_t[idx]*(1.-omega*s))
         if self.current_time < self.T:

### 2021-04-06 tex/targetvol.tex
Daluiso Bang Bang Proof

@@
-\documentclass[a4paper]{article}
-\usepackage[utf8]{inputenc}
+\documentclass[11pt]{article}
 \usepackage[english]{babel}
-\usepackage{comment}
-\usepackage{dcolumn,multirow,abstract,fancyhdr}
-\usepackage{geometry}
-\usepackage{mathtools} % serve per fare le formule matematiche 
-\usepackage{textcomp}
-\usepackage{amsfonts}
-\usepackage{dsfont}
-\usepackage{upgreek}
-\geometry{a4paper, top=5cm, bottom=5cm, left=4.5cm, right=4.5cm} %
-%\pagestyle{fancy} 
+\usepackage[latin1]{inputenc}
+\usepackage{lmodern}
+\usepackage[T1]{fontenc}
+%\usepackage[outer=5cm, marginparwidth=4.5cm]{geometry}
+%\usepackage{marginnote}
+\usepackage{amssymb,amsmath,amsthm,latexsym,amsfonts,amscd,dsfont,enumerate,bbm,mathabx}
+\usepackage{graphicx,tikz}
+\usepackage{draftwatermark}
+% \usepackage{listings}
+% \usepackage{booktabs}
+\usepackage{mathtools}
+% \usepackage[gen]{eurosym}
+% \usepackage{algorithm}
+% \usepackage{algpseudocode}
+\usepackage{soul}
+
+% \setbeamertemplate{bibliography item}[text]
+% \renewcommand*{\bibfont}{\scriptsize}
+
+% \usepackage{pgfplots}
+% \pgfplotsset{compat=newest}
+
+\definecolor{mylightyellow}{rgb}{1,1,.8}
+\definecolor{mylightgreen}{rgb}{.8,1,.8}
+\definecolor{mydarkred}{RGB}{178,34,34}
+\definecolor{mydarkgreen}{RGB}{34,139,34}
+\definecolor{mydarkblue}{RGB}{72,61,139}
+\definecolor{mydarkyellow}{RGB}{218,165,32}
+
+\definecolor{myblueA}{RGB}{52,41,39}
+\definecolor{myblueB}{RGB}{92,81,109}
+\definecolor{myblueC}{RGB}{132,121,179}
+\definecolor{myblueD}{RGB}{172,161,249}
+\definecolor{myblueE}{RGB}{212,201,255}
+
+\usetikzlibrary{arrows,patterns,shapes,snakes,shadows,fit,backgrounds,positioning,decorations.pathmorphing,external}
+\tikzstyle{nameusd} = [circle, draw, top color=white, bottom color=mydarkblue!50, draw=mydarkblue!75!black!100, drop shadow, minimum height=4em]
+\tikzstyle{nameeur} = [circle, draw, top color=white, bottom color=mydarkred!50, draw=mydarkred!75!black!100, drop shadow, minimum height=4em]
+\tikzstyle{collusd} = [rectangle,fill=mydarkblue!10, inner sep=0.2cm, rounded corners=5mm]
+\tikzstyle{colleur} = [rectangle,fill=mydarkred!10, inner sep=0.2cm, rounded corners=5mm]
+\tikzstyle{fixto} = [draw, -latex']
+\tikzstyle{fixfrom} = [draw, latex'-]
+\tikzstyle{floatto} = [draw, snake=coil, segment aspect=0, line before snake=2ex, line after snake=1ex, -latex']
+\tikzstyle{floatfrom} = [draw, snake=coil, segment aspect=0, line before snake=2ex, line after snake=1ex, latex'-]
+\tikzstyle{investor} = [rectangle, draw, top color=white, bottom color=mydarkyellow!50, draw=mydarkyellow!75!black!100, drop shadow, rounded corners, minimum height=3em, text width=4em, text centered]
+\tikzstyle{trader} = [circle, draw, top color=white, bottom color=blue!30, draw=blue!50!black!100, drop shadow, minimum height=4em]
+\tikzstyle{bank} = [rectangle, draw, top color=white, bottom color=red!20, draw=red!50!black!100, drop shadow, rounded corners, minimum height=3em, text width=4em, text centered]
+\tikzstyle{market} = [rectangle, draw, top color=white, bottom color=green!20, draw=green!50!black!100, drop shadow, rounded corners, minimum height=3em, text width=4em, text centered]
+\tikzstyle{yellowbox} = [rectangle, draw, top color=white, bottom color=mydarkyellow!50, draw=mydarkyellow!75!black!100, drop shadow, rounded corners, text centered]
+\tikzstyle{fadedyellowbox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkyellow!25, draw=mydarkyellow!50, drop shadow, rounded corners, text centered]
+\tikzstyle{redbox} = [rectangle, draw, top color=white, bottom color=mydarkred!50, draw=mydarkred!75!black!100, drop shadow, rounded corners, text centered]
+\tikzstyle{fadedredbox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkred!25, draw=mydarkred!50, drop shadow, rounded corners, text centered]
+\tikzstyle{bluebox} = [rectangle, draw, top color=white, bottom color=mydarkblue!50, draw=mydarkblue!75!black!100, drop shadow, rounded corners, text centered]
+\tikzstyle{fadedbluebox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkblue!25, draw=mydarkblue!50, drop shadow, rounded corners, text centered]
+\tikzstyle{greenbox} = [rectangle, draw, top color=white, bottom color=mydarkgreen!50, draw=mydarkgreen!75!black!100, drop shadow, rounded corners, text centered]
+\tikzstyle{fadedgreenbox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkgreen!25, draw=mydarkgreen!50, drop shadow, rounded corners, text centered]
+
+\newcommand{\tikzmark}[1]{\tikz[remember picture] \node[coordinate] (#1) {#1};}
+
+% \pgfplotstableread{data/total_consumption_example.dat}\swingtotal
+
+\usepackage[]{hyperref}
+\hypersetup{
+	colorlinks=true,breaklinks=true,
+	urlcolor= mydarkblue,linkcolor= mydarkblue,citecolor= mydarkgreen,
+	pdfauthor={Daluiso, R., Pallavicini, A.},
+}
+\usepackage{natbib}
+\setcitestyle{authoryear,open={(},close={)}}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{corollary}{Corollary}[theorem]
+\newtheorem{remark}{Remark}[section]
+\newtheorem{example}[remark]{Example}
+
+\newcommand{\Eq}[1]{{\begin{equation}{#1}\end{equation}}}
+
+\newcommand{\Exo}[1]{\mathbb{E}\!\left[\,#1\,\right]}
+\newcommand{\Ex}[2]{\mathbb{E}_{#1}\!\left[\,#2\,\right]}
+\newcommand{\ExG}[2]{\mathbb{E}\!\left[\,#2\,\left|{\cal G}_{#1}\right.\right]}
+\newcommand{\ExF}[2]{\mathbb{E}\!\left[\,#2\,\left|{\cal F}_{#1}\right.\right]}
+\newcommand{\ExFT}[3]{\mathbb{E}^{#2}\!\left[\,#3\,\left.\right|{\cal F}_{#1}\right]}
+\newcommand{\ExGT}[3]{\mathbb{E}^{#2}\!\left[\,#3\,\left.\right|{\cal G}_{#1}\right]}
+\newcommand{\ExCo}[2]{\mathbb{E}\!\left[\left.\,#1\,\right|\,#2\,\right]}
+\newcommand{\ExC}[3]{\mathbb{E}_{#1}\!\left[\left.\,#2\,\right|\,#3\,\right]}
+\newcommand{\ExCT}[4]{\mathbb{E}_{#1}^{#2}\!\left[\left.\,#3\,\right|\,#4\,\right]}
+\newcommand{\ExT}[3]{\mathbb{E}_{#1}^{#2}\!\left[\,#3\,\right]}
+\newcommand{\Exf}[2]{\hat{\mathbb{E}}_{#1}\!\left[\,#2\,\right]}
+\newcommand{\ExfT}[3]{\hat{\mathbb{E}}_{#1}^{#2}\!\left[\,#3\,\right]}
+\newcommand{\Px}[1]{\mathbb{P}\left\{\,#1\,\right\}}
+\newcommand{\PxC}[2]{\mathbb{P}\!\left\{\,#1\,\left.\right|\,#2\,\right\}}
+\newcommand{\Qx}[1]{\mathbb{Q}\left\{\,#1\,\right\}}
+\newcommand{\QxC}[2]{\mathbb{Q}\!\left\{\,#1\,\left.\right|\,#2\,\right\}}
+\newcommand{\QxCT}[3]{\mathbb{Q}^{#1}\!\left\{\,#2\,\left.\right|\,#3\,\right\}}
+\newcommand{\QxT}[2]{\mathbb{Q}^{#1}\left\{\,#2\,\right\}}
+\newcommand{\Qxf}[1]{\hat{\mathbb{Q}}\left\{\,#1\,\right\}}
+\newcommand{\Var}[2]{\operatorname{Var}_{#1}\!\left[\,#2\,\right]}
+\newcommand{\Covar}[3]{\operatorname{Covar}_{#1}\!\left[\,#2,#3\,\right]}
+\newcommand{\ind}[1]{1_{\{#1\}}}
+\newcommand{\clip}[3]{\operatorname{clip}\!\left({#1},{#2},{#3}\right)}
+\newcommand{\one}{\ensuremath{\mathbbm{1}}}
+\newcommand{\onehalf}{\tfrac{1}{2}}
+
+\newcommand{\tr}[1]{{#1}^{\intercal}} %transpose
+\let\oldnorm\norm   % <-- Store original \norm as \oldnorm
+\let\norm\undefined % <-- "Undefine" \norm
+\DeclarePairedDelimiter\norm{\lVert}{\rVert}
+
+\newcommand{\rec}{R}
+\newcommand{\lgd}{\mbox{L{\tiny GD}}}
+
+\newcommand{\pvcall}{{\mathcal{C}}}
+\newcommand{\pvput}{{\mathcal{P}}}
+
+\usepackage{eurosym}
+\newcommand{\euroms}[0]{\mbox{\small \euro}}
+\usepackage{cleveref}
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
+
+\newenvironment{eqsys}{\begin{equation}\begin{dcases}}{\end{dcases}\end{equation}}
+\usepackage{fancyhdr}
 \renewcommand{\headrulewidth}{0pt}
@@
 \rfoot{}
-\usepackage[font=scriptsize,labelfont=bf]{caption}
-\usepackage[pdfa]{hyperref}
-\hypersetup{colorlinks=true,citecolor=black,filecolor=black,linkcolor=black, urlcolor=black}
-\usepackage{cleveref}
-\usepackage{natbib}
-\setcitestyle{authoryear,open={(},close={)}}
-\newtheorem{theorem}{Theorem}[section] 
-\newtheorem{lem}{Lemma}[section]
-\newtheorem{defi}{Definition}[section]
-\newtheorem{prop}{Proposition}[section]
-\newtheorem{rem}{Remark}[section]
-\DeclareMathOperator*{\argmax}{arg\,max}
-\DeclareMathOperator*{\argmin}{arg\,min}
-\newenvironment{eqsys}{\begin{equation}\begin{dcases}}{\end{dcases}\end{equation}}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\setcounter{page}{1} % start with first page
 
@@
 Hence, we have our first result valid only if valuation adjustments can be neglected:
-\begin{prop}
+\begin{proposition}
 A European payoff on the TVS can be calculated by assuming any allocation in the underlying risky basket if all the underlying
 securities grow under the risk-neutral measure at the risk-free rate without any valuation adjustment, namely if we can write $\mu_t=0$.
-\end{prop}
-\begin{rem}[Existence of the Solution in the General Case]
+\end{proposition}
+\begin{remark}[Existence of the Solution in the General Case]
 In a more general settings we are not able to find an explicit solution.
@@
 A least if $r_t$, $\mu^i_t$ and $\nu^i_t$ are (uniformly) bounded, and if the eigenvalues of $\nu_t$ are (uniformly) bounded away from zero, then drift and diffusion should be uniformly Lipschitz in $\alpha_t$, and classical theorems should exist.
-\end{rem}
+\end{remark}
 \subsection{Stochastic Optimal Control Problem}
@@
 
-\begin{prop}
+\begin{proposition}
 When the underlying securities follow a Black and Scholes model with deterministic rates, the optimal strategy for a monotone payoff consists in minimizing the local drift function, indipendently of the current state $I_t$
@@
 \label{eq:BS_optimal_strategy}\end{equation}
-\end{prop}
+\end{proposition}
 The absence of stochastic elments in \Cref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori}; in fact one can solve the optimization problema once for all for each $t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities.
 Once $\alpha^*$ is known, then one can price the payoff by Monte Carlo simulation.
-\subsubsection{Free strategy: closed form solution}
+\subsubsection{Closed form solution for free strategy}
 In absence of constraints on $\alpha \coloneqq \alpha(t)$, the problem even admits a closed form solution.
@@
 \end{equation}
-we obtain the first order conditions\footnote{We used the property of the Cholesky factor of the covariance matrix $\Sigma = \nu  \nu^T$, where $\nu^T$ is the transpose matrix of $\nu$.}
+we obtain the first order conditions
 \begin{eqsys}
@@
 	\frac{\partial \mathcal{L}}{\partial \lambda}= \|\alpha \cdot \nu\|^2-1=0
-	\, .
+	\, ,
 \end{eqsys}
+where we have set $\Sigma = \nu  \nu^T$, with $\nu^T$ the transpose matrix of $\nu$.
 Then, by applying simple algebra, we obtain the analytical form of the free optimal strategy
@@
 In our work the definition of $\alpha$ is general, thus it is possible to consider the case of constrained allocation strategy; in this case the problem \eqref{eq:BS_optimal_strategy} must be solved numerically.
+\subsubsection{Active Asset (or Bang Bang) Solution}
+A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry are nonnegative and the only constraint on portfolio weights is nonnegativity, which would mean a long only strategy by the fund manager.
+
+\begin{lemma}
+	Let $\mu \in \mathbb{R}^n$ be a vector with nonnegative components, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix, and $\Sigma = \nu \nu^{T}$. Then
+	\begin{equation}\label{eq:inf}
+		\inf_{\alpha \in \mathbb{R}_{+}^n \setminus \{0\}} \frac{\alpha \cdot \mu}{\norm{\alpha \cdot \nu}} = \min_{i\leq n} \frac{\mu_i}{\sqrt{\Sigma_{ii}}};
+	\end{equation}
+	if $\bar{\imath}$ is the index which realizes the min, then the infimimum is realized by a vector concentrated on the $\bar{\imath}$ component: $\alpha_i = \delta_{i\bar{\imath}}$. 
+\end{lemma}
+\begin{proof}
+	Let us first consider the case in which $\mu = \mathbbm{1}$.
+Since the argument of the infimum is zero-homogeneous, normalizing by $\alpha \cdot \mathbbm{1} > 0$ we can restrict to the affine hyperspace $\{\alpha \cdot \mathbbm{1} = 1\}$, where the minimization \eqref{eq:inf} reduces to the maximization of its denominator: the required infimum will be square root of the reciprocal of
+	\[
+	\sup\left\{\norm{\alpha \cdot \nu}^2 \mid \alpha \in \mathbb{R}_{+}^n,\, \alpha \cdot \mathbbm{1} = 1\right\}.
+	\]
+	Now we can note that $\Sigma$ is positive definite, hence $\Sigma_{ij} < \sqrt{\Sigma_{ii}\Sigma_{jj}} \leq \Sigma_{\bar{\imath}\bar{\imath}}$, which implies
+	\[
+	\norm{\alpha \cdot \nu}^2 = \sum_{i,j=1}^{n} \alpha_i \alpha_j \Sigma_{ij} \leq \sum_{i,j=1}^{n} \alpha_i \alpha_j \Sigma_{\bar{\imath}\bar{\imath}} = \Sigma_{\bar{\imath}\bar{\imath}}
+	\]
+	because $\sum_i \alpha_i = 1$. Since we trivially have equality for $\alpha_i = \delta_{i\bar{\imath}}$, this concludes the proof of the case $\mu = \mathbbm{1}$.
+
+	Next, let us consider the case in which all components of $\mu$ are strictly positive, and define $M$ as the diagonal matrix with diagonal $\mu$.
+Then we can rewrite the infimum as a function of $\beta = M\alpha$:
+	\[
+	\inf_{\beta \in \mathbb{R}_{+}^n \setminus \{0\}} \frac{\beta \cdot \mathbbm{1}}{\norm{\beta \cdot M^{-1}\nu}},
+	\]
+	which by the first part of the proof equals 
+	\[
+	\min_{i \leq n} \frac{1}{\sqrt{\tilde{\Sigma}_{ii}}} = \min_{i \leq n} \frac{\mu_i}{\sqrt{\Sigma_{ii}}}, \quad \tilde{\Sigma} := M^{-1}\nu\nu^{T} M^{-1} = M^{-1}\Sigma M^{-1}.
+	\]
+
+	Finally, let us consider the general case in which $\mu$ may have some components equal to zero. For an arbitrary $\epsilon \geq 0$ let us define
+	\[
+	f_{\epsilon}(\alpha) = \frac{\alpha \cdot (\mu+\epsilon)}{\norm{\alpha \cdot \nu}}.
+	\]
+	One can easily note that as $\epsilon \to 0$, $f_{\epsilon}$ tends to $f_{0}$ uniformly on the compact set $\{\alpha \in \mathbb{R}^n_+ \mid \alpha \cdot \mathbbm{1} = 1\}$, so that the minimum converges to the minimum on that set.
+Since we know by homogeneity that the minimum on $\{\alpha \in \mathbb{R}^n_+ \mid \alpha \cdot \mathbbm{1} = 1\}$ equals the minimum on $\mathbb{R}^n_+ \setminus \{0\}$, we conclude
+	\[
+	\inf_{\alpha \in \mathbb{R}_{+}^n \setminus \{0\}} f_0(\alpha) = 
+	\lim_{\epsilon \to 0+} \inf_{\alpha \in \mathbb{R}_{+}^n \setminus \{0\}} f_{\epsilon}(\alpha) = 
+	\lim_{\epsilon \to 0+} \min_{i\leq n} \frac{\mu_i+\epsilon}{\sqrt{\Sigma_{ii}}} =
+	\min_{i\leq n} \frac{\mu_i}{\sqrt{\Sigma_{ii}}}.
+	\]
+\end{proof}
 

### 2021-04-06 reinforcement/baseline_simulation/run.py
Little canghes for Euler discretization

@@ -58,16 +58,16 @@ Identity = np.identity(N_equity)
 names = ["DJ 50 EURO E","S&P 500 NET EUR"]
 D, F, V, LV = LoadFromTxt(names, "FakeSmiles")
 correlation = np.identity(len(names))
-spot_prices = np.array([])
+spot_prices = np.ones(len(names)
 for i in range(len(names)):
-    spot_prices = np.append(spot_prices,F[i].spot)
+    spot_prices[i] = F[i].spot
    # print(F[i].q_values)
 """Preparing the LV model"""
 #print(N_euler_grid)
 model = LV_model(fixings=observation_grid[1:], local_vol_curve=LV, forward_curve=F, N_grid = N_euler_grid)
 euler_grid = model.time_grid
 discount = D(T)
-dt = model.dt[0]
+dt_vector = model.dt
 mu_function = Drift(forward_curves=F)
 mean = np.array([])
 correlation_chole = np.linalg.cholesky(correlation)
@@ -106,6 +106,7 @@ for i in range(Nsim):
     sigma = simulations_Vola[0]
     norm_price = dS_S[0]
     for j  in range(len(observation_grid[1:])):
+        dt = dt_vector[j]
         index_plus = j*N_euler_grid
         for k in range(N_euler_grid):
             idx = index_plus + k 

### 2021-04-06 reinforcement/envs/tvs_lv_env.py

@@ -52,11 +52,11 @@ class TVS_LV(gym.Env):
         """Loading market curves"""
         D, F, V, LV = LoadFromTxt(names, folder_name)
         self.correlation_chole = np.linalg.cholesky(correlation)
-        self.spot_prices = np.array([])
+        self.spot_prices = np.zeros(self.N_equity)
         for i in range(self.N_equity):
-            self.spot_prices = np.append(self.spot_prices,F[i].spot)
+            self.spot_prices[i] = F[i].spot
         """Preparing the LV model"""
-        self.model = LV_model(fixings=self.observation_grid, local_vol_curve=LV, forward_curve=F, N_grid = self.N_euler_grid)
+        self.model = LV_model(fixings=self.observation_grid[1:], local_vol_curve=LV, forward_curve=F, N_grid = self.N_euler_grid)
         euler_grid = self.model.time_grid
         mu_function = Drift(forward_curves=F)
         self.mu_values  = mu_function(np.append(0.,euler_grid[:-1]))
@@ -72,7 +72,7 @@ class TVS_LV(gym.Env):
         self.integral_variance = np.insert(self.integral_variance,0,np.zeros(self.N_equity),axis=0)
         self.integral_variance_sqrt =sqrt(self.integral_variance)
         self.integral_variance_sqrt[0,:] = 1
-
+        self.forwards = np.insert(self.model.forward.T,0,self.spot_prices,axis=0)[self.state_index,:]
         if self.constraint == 'long_short_limit' and (sum_long is None or sum_short is None):
             raise Exception("You should provide the sum limit for short and long position")
         if sum_long is not None and sum_short is not None:
@@ -134,7 +134,7 @@ class TVS_LV(gym.Env):
             S = np.insert(S,0,self.spot_prices,axis=1)
             self.dS_S_simulations = (S[:,1:,:] - S[:,:-1,:])/S[:,:-1,:]
             S_sliced = S[:,self.state_index,:]
-            self.simulations_logX = (log(S_sliced/np.insert(self.model.forward.T,0,self.spot_prices,axis=0)[self.state_index,:])+0.5*self.integral_variance)/self.integral_variance_sqrt
+            self.simulations_logX = (log(S_sliced/self.forwards)+0.5*self.integral_variance)/self.integral_variance_sqrt
             S_sliced = None
             S = None
             self.simulation_index=0

### 2021-04-07 tex/targetvol.tex
new tex

@@
 }
-\usepackage{natbib}
-\setcitestyle{authoryear,open={(},close={)}}
+\usepackage[backend=biber,style=authoryear-comp,maxcitenames=2,sorting=nyt,natbib=true,sortcites=false]{biblatex}
+\addbibresource{Bib.bib}
+%\usepackage{natbib}
+%\setcitestyle{authoryear,open={(},close={)}}
 \newtheorem{theorem}{Theorem}[section]
@@
 Once $\alpha^*$ is known, then one can price the payoff by Monte Carlo simulation.
-\subsubsection{Closed form solution for free strategy}
-In absence of constraints on $\alpha \coloneqq \alpha(t)$, the problem even admits a closed form solution.
-We set $\mu \coloneqq \mu(t)$ and $\nu \coloneqq \nu(t)$.
-By homogeneity we can rewrite \eqref{eq:BS_optimal_strategy} as
+\subsubsection{Black's Closed form solution}
+In absence of constraints on the allocation strategy, we are able to derive a closed form solution to the BS problem \eqref{eq:BS_optimal_strategy}
+\begin{lemma}
+	Let be $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \nu^T$.
+Then the closed solution of the optimization problem \eqref{eq:BS_optimal_strategy} is
 	\begin{equation}
-	\boxed{
+			\alpha^* = - \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
+	\end{equation}
+\end{lemma}
+\begin{proof}
+Since the argument of the minimum \eqref{eq:BS_optimal_strategy} is zero-homogeneous, then we can rewrite the problem as
+\begin{equation}
 		\begin{aligned}
@@
 			&\text { subject to } \|\alpha \cdot \nu\|^2=1
-	\end{aligned}}
+	\end{aligned}
 \end{equation}
@@
 \end{eqsys}
-where we have set $\Sigma = \nu  \nu^T$, with $\nu^T$ the transpose matrix of $\nu$.
 Then, by applying simple algebra, we obtain the analytical form of the free optimal strategy
@@
 We take the minus sign to get the minimum value of the TVS local drift. 
-
-In our work the definition of $\alpha$ is general, thus it is possible to consider the case of constrained allocation strategy; in this case the problem \eqref{eq:BS_optimal_strategy} must be solved numerically.
+\end{proof}
 \subsubsection{Active Asset (or Bang Bang) Solution}
@@
 
- We adopt as learning strategy the PPO algoritm developed in \cite{ppo} and \cite{gae}. 
+ We adopt as learning strategy the PPO algoritm developed in \cite{ppo} and \cite{gae}. We have
+
 \cleardoublepage
-\bibliographystyle{abbrvnat}
-\bibliography{Bib}
+\printbibliography
 \end{document}

### 2021-04-07 reinforcement/baseline_simulation/run.py

@@ -24,8 +24,8 @@ daily = 0
 N_equity = 2                                #number of equities
 target_vol = 5./100.
 T = 3.
-title = 'day_new/d3/final_price_lvinvestigation_2_asset_equalrepo_long_Nsim_'+str(Nsim)+'_vola'+str(target_vol)+'_maturity'+str(T)+'_Black'+str(Black)+'_daily'+str(daily)+strategy
-restart_file = 'day_new/d2/final_price_lvinvestigation_2_asset_equalrepo_long_Nsim_8000_vola0.05_maturity2.0_Black0_daily0only_long'
+title = 'day_new/d'+str(int(T))+'/final_price_lvinvestigation_2_asset_equalrepo_long_Nsim_'+str(Nsim)+'_vola'+str(target_vol)+'_maturity'+str(T)+'_Black'+str(Black)+'_daily'+str(daily)+strategy
+restart_file = 'day_new/d'+str(int(T-1))+'/final_price_lvinvestigation_2_asset_equalrepo_long_Nsim_8000_vola0.05_maturity'+str(T-1)+'_Black'+str(Black)+'_daily'+str(daily)+strategy
 
 if restart:
     I_0 = np.loadtxt(restart_file+"_rank"+str(rank+run)+'.txt')
@@ -58,7 +58,7 @@ Identity = np.identity(N_equity)
 names = ["DJ 50 EURO E","S&P 500 NET EUR"]
 D, F, V, LV = LoadFromTxt(names, "FakeSmiles")
 correlation = np.identity(len(names))
-spot_prices = np.ones(len(names)
+spot_prices = np.ones(len(names))
 for i in range(len(names)):
     spot_prices[i] = F[i].spot
    # print(F[i].q_values)

### 2021-04-08 reinforcement/envs/tvs_lv_env.py
Bang Bang action: discrete action space for 2 equities

@@ -17,6 +17,7 @@ class TVS_LV(gym.Env):
     def __init__(self, N_equity= 2, frequency = "month", target_volatility=5/100, I_0 = 1., strike_opt=1., 
     maturity=2., constraint = "only_long", action_bound=20./100., sum_long = None, sum_short=None):
         """Pricing parameters for the option"""
+        self.bang_bang_action = 1
         self.constraint = constraint
         self.target_vol = target_volatility
         self.I_0 = I_0
@@ -73,18 +74,21 @@ class TVS_LV(gym.Env):
         self.integral_variance_sqrt =sqrt(self.integral_variance)
         self.integral_variance_sqrt[0,:] = 1
         self.forwards = np.insert(self.model.forward.T,0,self.spot_prices,axis=0)[self.state_index,:]
-        if self.constraint == 'long_short_limit' and (sum_long is None or sum_short is None):
-            raise Exception("You should provide the sum limit for short and long position")
-        if sum_long is not None and sum_short is not None:
-            self.sum_long = sum_long
-            self.sum_short = sum_short
-        if self.constraint != "only_long":
-            low_action = np.ones(self.N_equity)*(-abs(action_bound))-1e-6
-            high_action = np.ones(self.N_equity)*abs(action_bound)+1e-6
+        if not self.bang_bang_action:
+            if self.constraint == 'long_short_limit' and (sum_long is None or sum_short is None):
+                raise Exception("You should provide the sum limit for short and long position")
+            if sum_long is not None and sum_short is not None:
+                self.sum_long = sum_long
+                self.sum_short = sum_short
+            if self.constraint != "only_long":
+                low_action = np.ones(self.N_equity)*(-abs(action_bound))-1e-6
+                high_action = np.ones(self.N_equity)*abs(action_bound)+1e-6
+            else:
+                low_action = np.ones(self.N_equity)*1e-7
+                high_action = np.ones(self.N_equity)
+            self.action_space = spaces.Box(low = np.float32(low_action), high = np.float32(high_action))
         else:
-            low_action = np.ones(self.N_equity)*1e-7
-            high_action = np.ones(self.N_equity)
-        self.action_space = spaces.Box(low = np.float32(low_action), high = np.float32(high_action))
+            self.action_space = spaces.Discrete(2)
         high = np.ones(N_equity)*2.5
         low_bound = np.append(-high,0.)
         high_bound = np.append(high,self.T+1./365)
[…]

### 2021-04-10 reinforcement/baseline_simulation/run.py
Code Speedup: no more np.linalg.norm function for the norm of a vector and no more np.delete function (now slicing arrays)

@@ -127,9 +127,9 @@ for i in range(Nsim):
        #         if f_bl>=f_base:
         #            counter = counter+1
          #       counter_tot = counter_tot+1
-            norm = np.linalg.norm(action@nu)
+            prod = action@nu
+            norm = np.sqrt(prod@prod)
             omega = target_vol/norm
-        #    lnI_t = lnI_t +omega*action@norm_price[idx] + (1-omega*np.sum(action))*r_t[idx]*dt
             I_t = I_t * (1. + omega*action@norm_price[idx]  + (1 - omega*np.sum(action))*r_t[idx]*dt  )
     f = open(title+"_rank"+str(rank+run)+'.txt',"a")
     #final_price[i] = I_t

### 2021-04-10 reinforcement/go.py

@@ -127,18 +127,18 @@ if __name__ == '__main__':
        restart_training = False,
        env='TVS_LV-v0',#'TVS_simple-v0',
        alg='ppo2',
-       num_layers='3',
-       num_hidden='6',
+       num_layers='4',
+       num_hidden='8',
        num_env='1',
        lr='3e-4',
-       train_timesteps='2.5e8',
+       train_timesteps='5e4',
        test_episodes='1e5',
        print_episodes='1',
        print_period='64',
-       save_interval='1e8',
+       save_interval='60',    
        #activation='sigmoid',
        value_network='copy',
-       custom_suffix='free25_month_observation_seed176970406'   #test on one_month 6,10,1e6
+       custom_suffix='bangbang_month2_observation_seed176970406'   #test on one_month 6,10,1e6
     )
 
     main(cur_args)

### 2021-04-10 tex/targetvol.tex
Latex

@@
 	urlcolor= mydarkblue,linkcolor= mydarkblue,citecolor= mydarkgreen,
-	pdfauthor={Daluiso, R., Pallavicini, A.},
+	pdfauthor={Daluiso, R., Pallavicini, A., Nastasi, E., Polo, S.},
 }
@@
 \section{Reinforcement Learning}\label{sec:RL}
+As we have seen previously the determination of the derivative price on the TVS requires to  solve a stochastic control problem with continuous actions.
+Among the available standard techniques, Reinforcement Learning (RL) represents a new efficient way to solve stochastic optimal control problems in high dimensions.
+
  Reinforcement Learning (RL) describes how an agent behaves in an environment so to maximize some notion of cumulative reward.

### 2021-04-17 reinforcement/baseline_simulation/run.py
New Market data condition
creation of a new market with two equities: one with a smile and the
other with a displaced diffusion (equal repo). Then little change in the
 baseline run.py: bang bang solution

@@ -118,7 +118,9 @@ for i in range(Nsim):
                 else:
                     mu = mu_values[idx]
                     if strategy=="only_long":
-                        action = optimization_only_long(mu, nu,seed=rank, guess = np.array(([0.4,0.6],[0.6,0.4])))
+                        #action = optimization_only_long(mu, nu,seed=rank, guess = np.array(([0.4,0.6],[0.6,0.4])))
+                        action = np.zeros(N_equity) 
+                        action[np.argmax(sigma_t[idx])] = 1. 
                     else:
                         action =  Markowitz_solution(mu,nu,-1)
      #           action2 = alpha(observation_grid[j])

### 2021-05-07 tex/targetvol.tex
New Market data condition
creation of a new market with two equities: one with a smile and the
other with a displaced diffusion (equal repo). Then little change in the
 baseline run.py: bang bang solution
Correct Values for BS price and Baseline Price+Fix import error of pyGem

@@
 \usepackage[T1]{fontenc}
-%\usepackage[outer=5cm, marginparwidth=4.5cm]{geometry}
+\usepackage[outer=3.5cm, marginparwidth=4.5cm]{geometry}
 %\usepackage{marginnote}
@@
 \usepackage{draftwatermark}
+\usepackage{enumerate} 
 % \usepackage{listings}
@@
 \usepackage{soul}
+\usepackage{comment}
 
@@
 \newcommand{\tr}[1]{{#1}^{\intercal}} %transpose
+\DeclareMathOperator{\Tr}{Tr}
 \let\oldnorm\norm   % <-- Store original \norm as \oldnorm
@@
 
-\title{Optimal Strategies for Target Volatility Contracts}
+\title{Reinforcement Learning for Options on Target Volatility Funds}
 
 \author{
-Roberto Daluiso\thanks{Banca IMI Milan, roberto.daluiso@bancaimi.com}
+	Roberto Daluiso\thanks{Intesa Sanpaolo Milan, roberto.daluiso@intesasanpaolo.com}
 	\and 
-  Andrea Pallavicini\thanks{Imperial College London and Banca IMI Milan, a.pallavicini@imperial.ac.uk} 
+	Emanuele Nastasi\thanks{Marketz S.p.A., emanuele.nastasi@marketz.eu}
+	\and Andrea Pallavicini\thanks{Imperial College London and Intesa Sanpaolo Milan, a.pallavicini@imperial.ac.uk} 
 	\and
-  Emanuele Nastasi\thanks{Exprivia, emanuele.nastasi@exprivia.com}
-  \and Stefano Polo \thanks{Intesa Sanpaolo Milan, stefano.polo@intesasanpaolo.com}% indicates same as 2nd thanks
+	Stefano Polo \thanks{Intesa Sanpaolo Milan, stefano.polo@intesasanpaolo.com}% indicates same as 2nd thanks
 }
 
-\date{} % leave empty
+\date{
+	First Version: March 2021. This Version: \today
+}
 \begin{document} % goes here
@@
 \begin{abstract}
-...
+\noindent In recent years Reinforcement Learning (RL) techniques are gaining popularity in the field of quantitative finance since they are proving to be an efficient way to solve  high-dimensional optimal control problems.
+Our research project is devoted to apply RL to price derivative contracts on target volatility strategies (TVSs), portfolios of risky assets and a risk-free one dynamically rebalanced in order to keep the realized volatility of the portfolio on a certain level.
+The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs for each component requires to solve a stochastic control problem to evaluate the option prices.
+The topic of hedging costs is a novelty never dealt in the TVS literature and we provide a formal description of the entire control problem.
+We tackle the problem by implementing a Reinforcement Learning algorithm to determine the optimal risky portfolio composition leading to the most conservative option price.
+We investigate the problem for two models of the risky asset dynamics: time-dependent Black \& Scholes and Local Volatility.
+In the first case we prove the existence of an analytical solution of the problem; a result that we use as benchmark to perform a series of fine-tuning of the hyper-parameters of the RL algorithm.
+At the end we provide numerical results for the Local Volatility model, for which an \textit{a priori} solution is not available.
 \end{abstract}
@@
 \noindent
-\textbf{JEL classification codes:} C63, G13.\\
+\textbf{JEL classification codes:} C63, C45, C61, G11, G13.\\
 \textbf{AMS classification codes:} 65C05, 91G20, 91G60.\\
-\textbf{Keywords:} Asset Allocation, Target Volatility, Funding Costs, Hedging Cost, Reinforcement Learning, Proximal policy optimization. 
-\clearpage
+\textbf{Keywords:} Reinforcement Learning, Hedging Costs, Funding Costs, Proximal Policy Optimization, Target Volatility, Asset Allocation. 
+\newpage
 \tableofcontents
- % \vspace{\fill\footnotesize \noindent The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.}
-\clearpage
-\pagestyle{fancy}  
+\newpage
+\pagestyle{myheadings} \markboth{}{{\footnotesize Daluiso, Pallavicini, Nastasi, Polo, Reinforcement Learning for Options on Target Volatility Funds}}
 \section{Introduction}
@@
 A TVS is  a  portfolio  of  risky  assets  (typically  equities)  and  a  risk-free  asset dynamically re-balanced with the aim of maintaining the overall portfolio volatility level closed to some target value.
-This products were initially offered in the Asian markets, see for instance the reports of \cite{Chew} and \cite{XUE} which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \cite{Morrison}.
+This products were initially offered in the Asian markets, see for instance the reports of \textcite{Chew} and \textcite{XUE} which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \textcite{Morrison}.
 
-In literature TVSs are tested to investigate their performances in term of realized returns, see for instance \cite{Hocquard28} and \cite{PERCHET}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim}.
-Moreover, in pricing literature derivative contracts on TVS are studied, see in particular \cite{Di_Graziano}, \cite{Grasselli}, and \cite{Albeverio}.
+In literature TVSs are tested to investigate their performances in term of realized returns, see for instance \textcite{Hocquard28} and \textcite{PERCHET}, and the soundness of the volatility targeting algorithm, as described in \textcite{Kim}.
+Moreover, in pricing literature derivative contracts on TVS are studied, see in particular \textcite{Di_Graziano}, \textcite{Grasselli}, and \textcite{Albeverio}.
 
@@
 Valuation adjustments (XVAs) is a topic widely discussed in the literature.
-We refer to \cite{Brigo} for a discussion.
-Since fund managers allocating TVS usually rely on Equity assets, here we use the results of \cite{Gabrielli} which  analyze the valuation adjustments for equity products.
+We refer to \textcite{Brigo} for a discussion.
+Since fund managers allocating TVS usually rely on Equity assets, here we use the results of \textcite{Gabrielli} which  analyze the valuation adjustments for equity products.
 We can write
@@
 where $\Bar{\sigma}$ is the target volatility value.
-
-We assume a generic continuous semi-martingale dynamics under the risk-neutral measure for the underlying securities, so that we can write
 In practice, this means that the fund manager will select a risky-portfolio choosing $\alpha_t$ equities from the universe where he can trade and after that his choices will be scaled by the automatic target volatility algorithm\footnote{We recall that the universe of assets where the manager can trade and the value of $\bar{\sigma}$ are written in the contract.} $\omega_t$.
@@
     \alpha_{T_{k-1}}:=\argmax_{\alpha} \left\{\mathbb{E}_{T_{k-1}}\left[D_{T_{k-1}}\left(T_{k} \right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}(\alpha)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right]\right\},
-\end{equation}
+\label{eq:recursion}\end{equation}
 where $V_{T_k}$ is the option value at time $T_k$ and $X$ is any Markovian state such that the drift and the diffusion coefficient of $I_t$ are a function of $\left(X_t,I_t,\alpha_t\right)$.
@@
     V_{T_{k-1}}\left(X_{T_{k-1}}, I_{T_{k-1}}\right)=\mathbb{E}_{T_{k-1}}\left[D_{T_{k-1}}\left(T_{k}\right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right],
-\label{recursion}\end{equation}
+\label{eq:recursion2}\end{equation}
 while the iteration starts from maturity date where the boundary condition is set equal to the payoff function:
@@
 \end{equation}
+
 \subsection{Black and Scholes Model}
@@
 
+\subsection{Hamilton-Jacobi-Bellman Equation for Target Volatility Options} 
+In this section we want to provide to the reader a formal description of the dynamic problem associated to options on target volatility strategies by writing the Hamilton-Jacobi-Bellman equation for the derivative value.
+We prove that from this description it is available the same closed formula \eqref{eq:BS_optimal_strategy} for the time-dependent BS model derived from the Gy\"ongy.
+
+In full generality we suppose that the risky securities follows a generic multidimensional diffusive process $X_t$ given by the following SDE
+\begin{equation}
+		dX_t = M(X_t)dt + \Sigma(X_t) \cdot dW_t.
+\end{equation}
+The TVS dynamics is given as follows 
+\begin{equation}
+	\frac{dI_t}{I_t} = \left(r_t(X_t) -  \frac{\bar{\sigma} \alpha_t}{\|\alpha_t \cdot \nu_t(X_t)\|}  \cdot \mu_t(X_t)    \right)dt +\frac{\bar{\sigma}\alpha_t }{\|\alpha_t \cdot \nu_t(X_t) \|} \cdot \nu_t(X_t) \cdot dW_t
+\end{equation}
+Given $X := X_t$ and $I := I_t$, we can write the HJB equation for $V := V(t,I,X)$ as follows
+\begin{equation}
+	\begin{aligned}
+		\frac{\partial V}{\partial t} + \max_\alpha  \left\{   \left(r_t(X) - \bar{\sigma}\frac{\alpha \cdot \mu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right) I \frac{\partial V}{\partial I} + (\nabla_X V) \cdot M(X) + \frac{1}{2} \bar{\sigma}^2I^2 \frac{\partial^2V}{\partial I^2}  \right.\\
+		\left.
++\frac{1}{2}\Tr\left(\tr{\Sigma(X)}(H_XV)\Sigma(X)\right)  + (\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(I\bar{\sigma}\frac{\alpha \cdot \nu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right)\right\} = 0,
+	\end{aligned}
+\end{equation}
+where $\Tr(A)$ is the trace operator of $A$, $\nabla_X V$ the gradient of $V$ w.r.t. $X$, $H_XV$ the Hessian matrix of $V$ w.r.t. $X$ and $\nabla_{X,I}V $ is the vector defined by:
+\begin{equation}
+	\nabla_{X,I} V= \tr{\left(\frac{\partial^2 V}{\partial X^1 \partial I},\dots, \frac{\partial^2 V}{\partial X^n \partial I} \right)}.
+\end{equation}
+We take out from the maximum operator all the elements that do not depend on the risky allocation strategy $\alpha$
+\begin{equation}
+	\begin{aligned}
+		\frac{\partial V}{\partial t} +   r_t(X) I \frac{\partial V}{\partial I} + (\nabla_X V) \cdot M(X) + \frac{1}{2} \bar{\sigma}^2I^2 \frac{\partial^2V}{\partial I^2} + \frac{1}{2}\Tr\left(\tr{\Sigma(X)}(H_XV)\Sigma(X)\right)  \\
+		+\max_\alpha  \left\{  -\frac{\partial V}{\partial I} I  \bar{\sigma}\frac{\alpha \cdot \mu_t(X)}{\|\alpha \cdot \nu_t(X)\|} + 
+		(\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(I\bar{\sigma}\frac{\alpha \cdot \nu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right)\right\} = 0.
+		\end{aligned}
+\label{eq:HJB_general}\end{equation}
+\Cref{eq:HJB_general} represent the Hamilton-Jacobi-Bellman equation describing the TVS dynamic problem for a generic dynamics of the risky securities underlying the portfolio.
+In our work we have that\footnote{We use $\circ$ for the element-wise product between two vectors and between vector and matrix.} $X_t = S_t$, $M(X_t) \coloneqq (r_t\mathds{1} - \mu_t)\circ S_t$ and $\Sigma(X_t) = S_t \circ \nu_t$.
+
+If we assume a time-dependent BS dynamics for the risky equities ($\mu_t$, $r_t$ and $\nu_t$ deterministic), then $V = V(t,I)$ and all the derivatives w.r.t $X$ are all zero. Thus the reduced HJB is
+\begin{equation}
+	\frac{\partial V}{\partial t} +  I r(t)\frac{\partial V}{\partial I}  + \frac{1}{2}\bar{\sigma}^2I^2\frac{\partial^2V}{\partial I^2} + \max_\alpha \left\{ -\frac{\partial V}{\partial I} I  \bar{\sigma}\frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t)\|}\right\} = 0,
+\end{equation}
+so that if the payoff is increasing in $I_T$ by homogeneity of the SDE we get that $V$ is increasing and so the solution is given by
+\begin{equation}
+ \alpha^*(t) = \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t)\|}
+	\end{equation}
+which is the same result expressed in \cref{eq:BS_optimal_strategy}. If the payoff is decreasing in $I_T$ then the solution will be the argmax.
+
+Conversely, if we deal with a LV model for the $S_t$-dynamics, then $\nu_t = \nu(t,S_t)$ and there are no apparent simplifications in \cref{eq:HJB_general} thus we are not able to derive a closed form solution as in BS.
+In fact in the LV dynamics we have in the maximum operator the first order term derived from the drift of $I_t$ (as in BS) and a second order term which takes into account the volatility's smile functions deriving from the $S_t$-dynamics.
+
 \section{Reinforcement Learning}\label{sec:RL}
-As we have seen previously the determination of the derivative price on the TVS requires to  solve a stochastic control problem with continuous actions.
-Among the available standard techniques, Reinforcement Learning (RL) represents a new efficient way to solve stochastic optimal control problems in high dimensions.
+As we have discussed previously in the case of general payoffs or risky securities dynamics, one must resort to numerical approaches to solve the stochastic control problem.
+The standard approach could be to use classical techniques based on backwards recursion \eqref{eq:recursion}-\eqref{eq:recursion2}.
+However their performances may degrade as the dimension $n$ of the problem increases.
+In our contribution we adopt a novel technique that is gaining popularity in many scientific branches for solving stochastic optimal control problems: Reinforcement Learning.
 
- Reinforcement Learning (RL) describes how an agent behaves in an environment so to maximize some notion of cumulative reward.
-The actions of the agent as a function of his observations of the environment are termed the agent policy.
+Reinforcement Learning is a branch of Machine Learning (ML) which allows an artificial agent to interact with an environment through actions and observations in order to maximize total rewards to achieve specific goals.
+In RL the agent is not told which actions to take but instead it must discover by trial and error which are the behaviors yielding to the greatest reward by trying them several times.
+This aspect makes RL independent  from pre-collected data as other ML techniques.
+Because of its nature, RL has been successful in quantitative finance for solving control problems; among the most important RL applications in this field, we refer to \textcite{pioneerrlfinance} as the pioneers in studying self-taught reinforcement trading problems, while to \textcite{rlintroductionfinance} and \textcite{rlfinance} for hedging derivatives with RL under market frictions.
+
+\begin{comment}
+\section{Numerical Investigation}
+Dire qui che abbiamo lavorato con l'implementazione di OpenAi
+
+The actions of the agent as a function of its observations of the environment are termed the agent policy.
 In our case the policy is the asset allocation weights $\alpha$, while the rewards are the costs generated by hedging a derivative contract on the TVS.
@@
 
- We adopt as learning strategy the PPO algoritm developed in \cite{ppo} and \cite{gae}. We have
-
+ We adopt as learning strategy the PPO algoritm developed in \textcite{ppo} and \textcite{gae}. We have 
+\end{comment}
 \cleardoublepage

### 2021-05-07 reinforcement/print_logs.py

@@ -8,13 +8,12 @@ n_sigma = 1.5
 strategy = 'free'
 
 LOG_FOLDER = './logs/TVS_LV_newreward-v0/'   #in which folder I find the file to plot
-WHICH_LOGS = [
-('ppo2_6e7_5x8_3e-4free_no_corr','RL'),
-#('ppo2_3e7_5x8_3e-4free_no_corr2','')
+WHICH_LOGS = [  #("name of the output folder", "legend label")
+#('ppo2_8e7_5x8_3e-4freestrategy_maturity2_monthgrid_noise05_beta07','RL'),
 ]
 X_AXIS_TIMESTEPS = 0  # otherwise: episodes
 WINDOW = int(3e5)  # measured in episodes  10000
-join_learning = 0   
+join_learning = 0  
 fig, ax = plt.subplots(1, 1, figsize=(8, 5))
 
 
@@ -37,10 +36,10 @@ if strategy=="only_long":
     plt.axhline(0.014784083837497124-3.2454334287712235e-05*n_sigma, color='green',lw=2.5, linestyle = '-.')
         
 elif strategy=='free':
-    plt.axhline(0.04095011290308573 +n_sigma*5.007677825957378e-05, color='red', lw=2.5,linestyle = '--',label="BS Strategy")
-    plt.axhline( 0.04095011290308573-n_sigma*5.007677825957378e-05, color='red',lw=2.5, linestyle = '--')
-    plt.axhline(0.04148428380424038+n_sigma* 5.245313523959805e-05, color='green',lw=2.5, linestyle = '-.',label="Baseline Strategy")
-    plt.axhline(0.04148428380424038-n_sigma* 5.245313523959805e-05, color='green',lw=2.5, linestyle = '-.')
+    plt.axhline(0.040947694174563704 +n_sigma*4.325646589751639e-05, color='red', lw=2.5,linestyle = '--',label="BS Strategy")
+    plt.axhline( 0.040947694174563704-n_sigma*4.325646589751639e-05, color='red',lw=2.5, linestyle = '--')
+    plt.axhline(0.04174359990484794+n_sigma*5.798528935881833e-05, color='green',lw=2.5, linestyle = '-.',label="Baseline Strategy")
+    plt.axhline(0.04174359990484794-n_sigma* 5.798528935881833e-05, color='green',lw=2.5, linestyle = '-.')
 plt.legend()
 plt.grid(True)
 plt.xlabel('Time step' if X_AXIS_TIMESTEPS else 'Episode',fontsize=font)

### 2021-05-07 reinforcement/go.py
Other Parameters Corrections

@@ -72,7 +72,7 @@ def build_args(do_train, do_test, env, alg, num_layers, num_hidden, num_env, lr,
 
     if do_train:
        agent_mode='save'
-       seed='5343011'
+       seed='43561'
        options.append('--log_path=' + log_path)
        options.append('--lr=' + lr)
     else:
@@ -124,7 +124,8 @@ if __name__ == '__main__':
     cur_args = build_args(
        do_train=1,
        do_test=0,
-       restart_training = 0,
+       restart_training = False,
+       #initial_guess='ppo2_3e7_5x8_3e-4freestrategy_maturity2_monthgrid_noise05_beta07_restarted2',
        env='TVS_LV_newreward-v0',
        alg='ppo2',
        num_layers='5',
@@ -137,7 +138,7 @@ if __name__ == '__main__':
        print_period='64',
        save_interval='200',    
        value_network='copy',
-       noise='0.5',
+       noise='2',
        beta='0.7',
        custom_suffix='freestrategy_maturity2_monthgrid_noise05_beta07'   #test on one_month 6,10,1e6
     )

### 2021-05-13 reinforcement/go.py
.

@@ -132,14 +132,13 @@ if __name__ == '__main__':
         num_hidden='8',
         num_env='27',
         lr='3e-4',
-        gamma='0.9',  
+        gamma='0.85',  
         train_timesteps='8e7',
         test_episodes='1e6',
         print_episodes='1',
         print_period='64',
         save_interval='200',    
         value_network='copy',
-        noise='7.',
         beta='0.7',
         custom_suffix='_freestrategy_displacedmarket_2assets_monthgrid_maturity2_strikeatm'   #test on one_month 6,10,1e6
         )

### 2021-05-13 reinforcement/go_test.py

@@ -132,14 +132,13 @@ if __name__ == '__main__':
         num_hidden='8',
         num_env='27',
         lr='3e-4',
-        gamma='0.9',  #in the test phase it is set to 1
+        gamma='0.85',  #in the test phase it is set to 1
         train_timesteps='8e7',
         test_episodes='1e6',
         print_episodes='1',
         print_period='64',
         save_interval='200',    
         value_network='copy',
-        noise='7.',
         beta='0.7',
         custom_suffix='_freestrategy_displacedmarket_2assets_monthgrid_maturity2_strikeatm'   #test on one_month 6,10,1e6
         )

### 2021-05-14 tex/targetvol.tex
Paper

@@
 \section{Reinforcement Learning}\label{sec:RL}
-As we have discussed previously in the case of general payoffs or risky securities dynamics, one must resort to numerical approaches to solve the stochastic control problem.
+As we have discussed previously in the case of general payoffs or risky securities dynamics, one must resort to numerical approaches to solve the stochastic control problem related to the TVS.
 The standard approach could be to use classical techniques based on backwards recursion \eqref{eq:recursion}-\eqref{eq:recursion2}.
@@
 In RL the agent is not told which actions to take but instead it must discover by trial and error which are the behaviors yielding to the greatest reward by trying them several times.
-This aspect makes RL independent  from pre-collected data as other ML techniques.
+This is obtained by updating the agent policy which is  This peculiarity makes RL independent from pre-collected data as other ML techniques.
 Because of its nature, RL has been successful in quantitative finance for solving control problems; among the most important RL applications in this field, we refer to \textcite{pioneerrlfinance} as the pioneers in studying self-taught reinforcement trading problems, while to \textcite{rlintroductionfinance} and \textcite{rlfinance} for hedging derivatives with RL under market frictions.

### 2021-05-14 reinforcement/envs/tvs_new_reward.py
Simulations with action parameterization from baseline

@@ -17,7 +17,7 @@ class TVS_LV_newreward(gym.Env):
     maturity=2., constraint = "free", action_bound=5., sum_long = None, sum_short=None):
         #Simulation parameters
         self.constraint = constraint
-        
+        self.from_baseline = 1
         self.target_vol = target_volatility
         self.N_equity = N_equity            
         self.strike_opt = strike_opt
@@ -60,9 +60,9 @@ class TVS_LV_newreward(gym.Env):
             self.spot_prices[i] = F[i].spot
         """Preparing the LV model"""
         self.model = LV_model(fixings=self.observation_grid[1:], local_vol_curve=LV, forward_curve=F, N_grid = self.N_euler_grid)
-        euler_grid = self.model.time_grid
+        self.euler_grid = self.model.time_grid
         """Discount part"""
-        self.r_t = D.r_t(np.append(0.,euler_grid[:-1]))
+        self.r_t = D.r_t(np.append(0.,self.euler_grid[:-1]))
         self.discount = D(self.T)
         self.intermidiate_discounts = D(self.T)/D(self.observation_grid)
         self.dt_vector = self.model.dt
@@ -80,8 +80,12 @@ class TVS_LV_newreward(gym.Env):
         self.integral_variance_sqrt[0,:] = 1
         self.forwards = np.insert(self.model.forward.T,0,self.spot_prices,axis=0)[self.state_index,:]
         if self.constraint == "free":
-            low_action = np.ones(self.N_equity)*(-abs(action_bound))
-            high_action = np.ones(self.N_equity)*abs(action_bound)
+            if not self.from_baseline:
+                low_action = np.ones(self.N_equity)*(-abs(action_bound))
+                high_action = np.ones(self.N_equity)*abs(action_bound)
+            elif self.from_baseline:
+                low_action = np.ones(self.N_equity)*-1.
+                high_action = np.ones(self.N_equity)
         elif self.constraint == "only_long":
             low_action = np.ones(self.N_equity)*1e-10
             high_action = np.ones(self.N_equity)
@@ -95,10 +99,11 @@ class TVS_LV_newreward(gym.Env):
 #current time start at zero
     def step(self, action): 
         assert self.action_space.contains(action)
+        s = 0.
         if self.constraint == "only_long":
             action = action/np.sum(action)
             s = 1.
-        elif self.constraint == "free":
+        elif self.constraint == "free" and not self.from_baseline:
             s = np.sum(action)
         
[…]

### 2021-05-14 reinforcement/go.py

@@ -132,7 +132,6 @@ if __name__ == '__main__':
         num_hidden='8',
         num_env='27',
         lr='3e-4',
-        gamma='0.85',  
         train_timesteps='8e7',
         test_episodes='1e6',
         print_episodes='1',
@@ -140,6 +139,6 @@ if __name__ == '__main__':
         save_interval='200',    
         value_network='copy',
         beta='0.7',
-        custom_suffix='_freestrategy_displacedmarket_2assets_monthgrid_maturity2_strikeatm'   #test on one_month 6,10,1e6
+        custom_suffix='freestrategyfrombaseline_displacedmarket_2assets_monthgrid_maturity2'   #test on one_month 6,10,1e6
         )
     main(cur_args)

### 2021-05-14 reinforcement/go_test.py

@@ -132,7 +132,6 @@ if __name__ == '__main__':
         num_hidden='8',
         num_env='27',
         lr='3e-4',
-        gamma='0.85',  #in the test phase it is set to 1
         train_timesteps='8e7',
         test_episodes='1e6',
         print_episodes='1',
@@ -140,6 +139,6 @@ if __name__ == '__main__':
         save_interval='200',    
         value_network='copy',
         beta='0.7',
-        custom_suffix='_freestrategy_displacedmarket_2assets_monthgrid_maturity2_strikeatm'   #test on one_month 6,10,1e6
+        custom_suffix='freestrategyfrombaseline_displacedmarket_2assets_monthgrid_maturity2'   #test on one_month 6,10,1e6
         )
     main(cur_args)

### 2021-05-14 tex/targetvol.tex
Paper

@@
 \usepackage[T1]{fontenc}
-\usepackage[outer=3.5cm, marginparwidth=4.5cm]{geometry}
+\usepackage[outer=4cm, marginparwidth=4.5cm]{geometry}
 %\usepackage{marginnote}
@@
 \newcommand{\lgd}{\mbox{L{\tiny GD}}}
-
+\newcommand{\virgolette}[1]{``#1''}
 \newcommand{\pvcall}{{\mathcal{C}}}
@@
 \begin{equation}
-V_0 \coloneqq \sup_{\alpha} \mathbb{E}_0\left[D_0\left(T\right)\Phi\left(I_T^{\text{MP}}\left(\alpha\right)\right)\right].
+V_0 \coloneqq \sup_{\alpha} \mathbb{E}_0\left[D\left(0,T;\zeta\right)\Phi\left(I_T^{\text{MP}}\left(\alpha\right)\right)\right].
 \end{equation}
@@
 \begin{equation}
-    \alpha_{T_{k-1}}:=\argmax_{\alpha} \left\{\mathbb{E}_{T_{k-1}}\left[D_{T_{k-1}}\left(T_{k} \right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}(\alpha)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right]\right\},
+    \alpha_{T_{k-1}}:=\argmax_{\alpha} \left\{\mathbb{E}_{T_{k-1}}\left[D\left(T_{k-1},T_{k} \right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}(\alpha)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right]\right\},
 \label{eq:recursion}\end{equation}
@@
 \begin{equation}
-    V_{T_{k-1}}\left(X_{T_{k-1}}, I_{T_{k-1}}\right)=\mathbb{E}_{T_{k-1}}\left[D_{T_{k-1}}\left(T_{k}\right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right],
+    V_{T_{k-1}}\left(X_{T_{k-1}}, I_{T_{k-1}}\right)=\mathbb{E}_{T_{k-1}}\left[D\left(T_{k-1},T_{k}\right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right],
 \label{eq:recursion2}\end{equation}
@@
 \subsection{Black and Scholes Model}
-In the Black and Scholes model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$.
+In the \textcite{BlackScholes} model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$.
 As a consequence, the local drift can be written as
@@
 \begin{proposition}
-When the underlying securities follow a Black and Scholes model with deterministic rates, the optimal strategy for a monotone payoff consists in minimizing the local drift function, indipendently of the current state $I_t$
+When the underlying securities follow a Black and Scholes (BS) model with deterministic rates, the optimal strategy for a monotone payoff consists in minimizing the local drift function, indipendently of the current state $I_t$
 \begin{equation}
@@
 The absence of stochastic elments in \Cref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori}; in fact one can solve the optimization problema once for all for each $t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities.
-Once $\alpha^*$ is known, then one can price the payoff by Monte Carlo simulation.
+Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
+\begin{equation}
+	V_0^{BS} = BS(F^{\text{TVS}}(0,T;\alpha^*), K, T, \bar{\sigma}, D(0,T;\zeta)),
+\end{equation}
+where $F^{\text{TVS}}(t,T;\alpha)$ is the TVS forward curve defined by
+\begin{equation}
+	F^{\text{TVS}}(t,T;\alpha) = I_t \exp\left[\int_t^T\left(r(u)-\bar{\sigma}\frac{\alpha(u)\cdot \mu(u)}{\|\alpha(u)\cdot \nu(u)\|}\right)du\right],
+\end{equation}
+while $BS(F,K,T,\sigma,D)$ is the standard BS formula for an European option with forward curve $F$, strike $K$, time to maturity $T$, volatility $\sigma$ and discount factor $D$.  
 \subsubsection{Black's Closed form solution}
@@
 
-\subsection{Hamilton-Jacobi-Bellman Equation for Target Volatility Options} 
+\subsection{Hamilton-Jacobi-Bellman Equation for Target Volatility Options}\label{subsec:HJB}
 In this section we want to provide to the reader a formal description of the dynamic problem associated to options on target volatility strategies by writing the Hamilton-Jacobi-Bellman equation for the derivative value.
@@
 	\end{equation}
-which is the same result expressed in \cref{eq:BS_optimal_strategy}. If the payoff is decreasing in $I_T$ then the solution will be the argmax.
+which is the same result expressed in \cref{eq:BS_optimal_strategy}. If the payoff is decreasing in $I_T$ then the solution will be the $\argmax$.
 
@@
 In RL the agent is not told which actions to take but instead it must discover by trial and error which are the behaviors yielding to the greatest reward by trying them several times.
-This is obtained by updating the agent policy which is  This peculiarity makes RL independent from pre-collected data as other ML techniques.
+This is obtained by updating the agent policy $\pi$ which is a mapping from the environment states to the set of actions.
+Thus RL is independent from pre-collected data as opposed to other ML techniques.
 Because of its nature, RL has been successful in quantitative finance for solving control problems; among the most important RL applications in this field, we refer to \textcite{pioneerrlfinance} as the pioneers in studying self-taught reinforcement trading problems, while to \textcite{rlintroductionfinance} and \textcite{rlfinance} for hedging derivatives with RL under market frictions.
+
+In our work we adopt as learning strategy the Proximal Policy Optimization (PPO) developed in \textcite{ppo} and \textcite{gae}.
+PPO is an actor-critic algorithm well-suited for continuous control problems.
+The term \virgolette{actor-critic} means that it estimates the value function  $J$, which represents the expected value of future rewards, and updates the policy in the direction suggested by $J$.
+In PPO the policy distribution and the value function are parameterized by a set of parameters $\theta$: $J^\theta$, $\pi^\theta$.
+The PPO collects a small batch of experiences interacting with the environment to update $\theta$ by computing an estimator of the policy gradient and plugging it into a stochastic gradient ascent algorithm in order to maximize an objective function $L^{\text{PPO}}$.
+The PPO objective is defined as
+\begin{equation}
+	L^{\text{PPO}} = L^\pi - c_1 L^J + c_2 L^{E}
+\end{equation}
+where $L^\pi$ is the policy surrogate, $L^J$ the value function error term and $L^E$ the entropy bonus, while $c_1$ and $c_2$ are hyper-parameters of the algorithm.
+We refer to \textcite{ppo} for a deeper description of those terms.
+ We use the PPO implementation found in \textsc{OpenAI Baselines} \citep{baselines}, where the parameterization procedure is performed by using artificial neural networks (ANNs), where the set of parameters $\theta$ are the hidden weights of the networks.
+In the following section we describe the the way we have formalized the TVS problem in the Reinforcement Learning framework.
+
+ \subsection{Episode Description: Agent-Environment Interaction}
+ We consider an episode $\tau$ of length $m+1$ that takes place on a discrete time-grid of fixing times $\mathcal{T}:=\{T_0,...,T_k,...,T_m\}$ with $T_0:=0$ and $T_m=T$ maturity of the option.
+At a given episodic time $T_k$ the RL agent interacts with the environment: it receives a representation of the environment called state $s_k$ and on the basis of that it selects an action $a_k$ sampling from the current policy $\pi^\theta_k$.
+ In our case the agent can choose the composition of the risky asset portfolio, so that the policy is the asset allocation weights $\alpha$.
+From the HJB equation we have derived in section \Cref{subsec:HJB}, we select as state of th environment the following block
+ \begin{equation}
+ 	s_{k} \coloneqq \left[T_k,I_{T_k}, S_{T_k}\right] \quad \forall T_k \in \mathcal{T}.
+ \label{eq:state}\end{equation}
+ This state contains all the informations needed by the agent to take an optimal action, leading to the maximum option price on the TVS. 
+ Once the agent has selected the action $a_k$, it receives the next time $T_{k+1}$ a reward $r_{k+1}$ generated by the environment defined by
+ \begin{equation}
+ 	r_{k+1} \coloneqq  \gamma^{k}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)],
+ \label{eq:reward}\end{equation}
+where $\gamma \in[0,1]$ is an hyper-parameter of the PPO, while $V_{\text{BS}}(T_k)$ is the BS price as proxy of the residual option price defined by
+\begin{equation}
+	V_{\text{BS}}(T_k) \coloneqq BS(F^{\text{TVS}}(T_k,T;\alpha^*_{\text{BS}}),K,T-T_k,\bar{\sigma},D(T_k,T;\zeta)) \quad \text{and} \quad V(T_0)=0
+\end{equation} 
+with $\alpha^*_{BS}$ the BS optimal strategy \eqref{eq:BS_optimal_strategy} calculated in the state $s_k$.
+The parameter $\gamma$ plays the role of a discount rate in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future.
+Choosing $\gamma<1$ introduces a bias in the final result.
+If we take the cumulative reward per episode and set $\gamma =1$ we obtain
+\begin{equation}
+	R(\tau) = \sum_{k=0}^{m-1} r_{k+1} = \sum_{k=0}^{m-1}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)] = V_{BS}(T) = (I_T-K)^+,
+\end{equation} 
+which is equal to the intrinsic value of the option.
+This result does not depend on the definition of $V_{\text{BS}}$  $\forall T_k<T$, but we think that the closer $V_{\text{BS}}$ is to the value function, the easier the agent is in learning.
+
+ Thus one can train the agent choosing $\gamma\in[0,1]$, and then run, as test phase, a Monte Carlo (MC) simulation with $\gamma=1$, where, if the agent has reached the optimal policy $\pi^* = \pi^{\theta^*}$, the average of $R(\tau$) along different episodes will match the optimal undiscounted price of the derivative contract on the TVS.
+
+\subsection{Policy and Action Parameterization}
+As discussed before, PPO parameterizes the agent policy by means of an artificial neural network.
+In our case we use a feed forward neural network (FFNN), a network with an acyclic topology and completely characterized by the following hyper-parameters: number of hidden layers $N$, number of neurons per hidden layer $M$, activation function per hidden layer and output activation function.
+The number of output neurons is fixed by the action space dimension, which is the number of assets $n$ in the risky portfolio, while the number of  input neurons is determined by the dimensionality of the state \eqref{eq:state}, $n+2$.
 

### 2021-05-18 reinforcement/go.py
New Parameters

@@ -125,8 +125,9 @@ if __name__ == '__main__':
         do_train=True,
         do_test=False,
         test_seed='114',
-        training_seed='34561',
+        training_seed='20319',
         env='TVS_LV_newreward-v0',
+        gamma='0.95',
         alg='ppo2',
         num_layers='5',
         num_hidden='8',

### 2021-05-18 reinforcement/go_test.py

@@ -125,9 +125,10 @@ if __name__ == '__main__':
         do_train=False,
         do_test=True,
         test_seed='114',
-        training_seed='34561',
+        training_seed='20319',
         env='TVS_LV_newreward-v0',
         alg='ppo2',
+        gamma = '0.95', 
         num_layers='5',
         num_hidden='8',
         num_env='27',

### 2021-05-18 tex/targetvol.tex
Paper: done RL section

@@
 \usepackage[T1]{fontenc}
-\usepackage[outer=4cm, marginparwidth=4.5cm]{geometry}
+\usepackage[outer=3.75cm, marginparwidth=4.5cm]{geometry}
 %\usepackage{marginnote}
@@
 \usepackage{comment}
+\usepackage{bm}
 
@@
 Then, in \Cref{sec:Derivative} we derive the analytical results, and in \Cref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem.
-We conclude the paper with \Cref{sec:Numerical_Results} where we present the numerical results obtained in this work.
+We conclude the paper with \Cref{sec:Numerical_results} where we present the numerical results obtained in this work.
 
@@
 \begin{equation}
-	R(\tau) = \sum_{k=0}^{m-1} r_{k+1} = \sum_{k=0}^{m-1}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)] = V_{BS}(T) = (I_T-K)^+,
+	R(\tau) = \sum_{k=0}^{m-1} r_{k+1} \underset{\gamma =1}{=} \sum_{k=0}^{m-1}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)] = V_{BS}(T) = (I_T-K)^+,
 \end{equation} 
@@
 The number of output neurons is fixed by the action space dimension, which is the number of assets $n$ in the risky portfolio, while the number of  input neurons is determined by the dimensionality of the state \eqref{eq:state}, $n+2$.
+We work with a stochastic policy whose functional form is a multivariate diagonal Gaussian distribution where the mean $\mu^\theta(s_k)$ is the output vector of the FFNN and the log-standard deviation $\log\sigma$ is an external parameter
+\begin{equation}
+	\pi^\theta(s_k) \sim \mathcal{N}(\mu^\theta(s_k),e^{\log\sigma}).
+\end{equation}
+In the \textsc{Baselines} implementation of PPO, the parameter $\log\sigma$ is state-independent, but it decreases as the number of PPO update iterations increases.
+The idea is that the log-standard deviation will be higher at the beginning of the training phase in order to guarantee a good exploration of the action space while it will be lower at the ending phase to avoid too much noise in the proximity of the optimal policy.
 
+Another important topic is how to parameterize the agent action in order to satisfy some specific constraints on the risky asset allocation weights.
+In fact, if on one hand with free $\alpha$ it is sufficient to set bounds in the action space, on the other hand the presence of constraints on the risky strategy imposes the need to parameterize the output of the ANN.
+In this work we study the constraint where the agent can invest in the risky market adopting long positions only:
+\begin{equation}
+	\alpha \cdot \mathds{1} = 1 \quad \alpha^i >0 \quad \forall i=1,\dots,n.
+\end{equation} 
+We study two different action parameterizations for this kind of constraint.
+\begin{itemize}
+	\item \textbf{$\bm{(n-1)}$-spherical angles}: we rewrite $\alpha^j = (x^i)^2$ and we use $(n-1)$-spherical coordinates with unitary radius.
+Thus the strategy is parameterized by a vector of angles $\phi = (\phi^1,\dots,\phi^{n-1})^T$ as follows
+	\begin{equation}
+		\begin{aligned}
+			\alpha^1 &\coloneqq \cos^2 (\phi^1), \\
+			\alpha^2 &\coloneqq \sin^2(\phi^1)\cos^2(\phi^2),\\
+			&\vdots \\
+			\alpha^{n-1} &\coloneqq \sin^2(\phi^1)\dots\sin^2(\phi^{n-2})\cos^2(\phi^{n-1}),\\
+			\alpha^n &\coloneqq \sin^2(\phi^1)\dots\sin^2(\phi^{n-2})\sin^2(\phi^{n-1}).
+		\end{aligned}
+		\label{eq:euler_angles}\end{equation}
+	where $\phi^i\in[0,\pi/2]$ to avoid oscillations in $\alpha$ that could slow down the training.
+Thus, with this parameterization, the ANN sample a vector of angles ($a = \phi$) and thus the action space is reduced by one dimension.
+	\item \textbf{Normalized action}: we restrict the agent actions to a subset of $(\mathbb{R}^+)^n$ and then normalize the action to obtain the corresponding risky allocation strategy through the formula
+	\begin{equation}
+		\alpha_t = \frac{a_t}{a_t \cdot \mathds{1}}.
+		\label{eq:normalized_action}\end{equation}
+\end{itemize}
+
+\section{Numerical Investigation}\label{sec:Numerical_results}
+In this section As we have seen in 
+\subsection{Black and Scholes Dynamics}
+\subsection{Local Volatility Dynamics}
+\section{Discussion}
 \begin{comment}
-\section{Numerical Investigation}
-Dire qui che abbiamo lavorato con l'implementazione di OpenAi
+
+--scrivere input normalization della rete
+--introduzione sui modelli fatti 
+--dire perchè black e del fine tuning dei parametri
+
+\ref{sec:target_volatility_option_derivative_pricing}, the bank evaluation of the most conservative call option price strictly depends on the dynamics followed by the equities underlying the \acrshort{tvs}; in particular the volatility term of the securities affects the \acrshort{tvs} price process \eqref{TVS_last}.
+In our work we will try to solve this pricing problem through the \acrshort{rl} technique, training the agent in two different environments corresponding to two different mathematical models for the dynamics of financial markets: the time-dependent \acrlong{bs} model and the \acrlong{lv} model.
+Although the first one is quite simple (its parameters depends only on time) and it is not completely able to represent the complex dynamics of real markets, we have seen that, under Black's conditions, we are able to derive the closed form solution for the optimal strategy \eqref{optimal_strategy}.
+Thus we can use the \acrlong{bs} model as a test to control if we have built correctly the \acrlong{rl} algorithm; in other words we will check if the \acrshort{rl} agent will take decisions in agreement with the closed form.
+On the other hand, the \acrfull{lv} model depends not only on time but also on the spot prices of the equities and thus a closed formula for the optimal strategy is not available; we will try to recover it by our \acrshort{rl} algorithm.
 

### 2021-05-19 reinforcement/go.py
New Parameters

@@ -140,6 +140,7 @@ if __name__ == '__main__':
         save_interval='200',    
         value_network='copy',
         beta='0.7',
-        custom_suffix='freestrategyfrombaseline_displacedmarket_2assets_monthgrid_maturity2'   #test on one_month 6,10,1e6
+        noise='1.',
+        custom_suffix='freefrombaseline_displacedmarket_2assets_monthgrid_maturity2'   #test on one_month 6,10,1e6
         )
     main(cur_args)

### 2021-05-19 reinforcement/go_test.py

@@ -127,8 +127,8 @@ if __name__ == '__main__':
         test_seed='114',
         training_seed='20319',
         env='TVS_LV_newreward-v0',
+        gamma='0.95',
         alg='ppo2',
-        gamma = '0.95', 
         num_layers='5',
         num_hidden='8',
         num_env='27',
@@ -140,6 +140,7 @@ if __name__ == '__main__':
         save_interval='200',    
         value_network='copy',
         beta='0.7',
-        custom_suffix='freestrategyfrombaseline_displacedmarket_2assets_monthgrid_maturity2'   #test on one_month 6,10,1e6
+        noise='1.',
+        custom_suffix='freefrombaseline_displacedmarket_2assets_monthgrid_maturity2'   #test on one_month 6,10,1e6
         )
     main(cur_args)

### 2021-05-19 reinforcement/print_test_results.py

@@ -3,8 +3,8 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 ### RL Test Result ####
-MC_price = 0
-MC_error =  0
+MC_price = 0.04105623158024169
+MC_error =  5.0366052653869903e-05
 
 ### BS PRICE ###
 BS_Price = 0.040947694174563704 

### 2021-05-19 tex/targetvol.tex
Paper

@@
 \usepackage{bm}
-
+\usepackage[font=small,labelfont=bf]{caption}
+\graphicspath{{Local_Volatility/}{Black_Scholes}}
 % \setbeamertemplate{bibliography item}[text]
@@
 	\and
-	Stefano Polo \thanks{Intesa Sanpaolo Milan, stefano.polo@intesasanpaolo.com}% indicates same as 2nd thanks
+	Stefano Polo \thanks{Intesa Sanpaolo Milan, stefano.polo@intesasanpaolo.com} \thanks{We wish to thank the Italian computing centre Cineca, which approved our ISCRA C project and provided us the high-computing resources of Marconi100 for the numerical simulations of this work.}% indicates same as 2nd thanks
 }
@@
 In literature TVSs are tested to investigate their performances in term of realized returns, see for instance \textcite{Hocquard28} and \textcite{PERCHET}, and the soundness of the volatility targeting algorithm, as described in \textcite{Kim}.
-Moreover, in pricing literature derivative contracts on TVS are studied, see in particular \textcite{Di_Graziano}, \textcite{Grasselli}, and \textcite{Albeverio}.
+Moreover, in pricing literature, derivative contracts on TVS, known as target volatility options (TVOs), are studied, see in particular \textcite{Di_Graziano}, \textcite{Grasselli}, and \textcite{Albeverio}.
 
-In this contribution we study a problem related to TVS.
-We deal with the funding costs coming from hedging the risky assets underlying the TVS.
-We consider the point of view of a bank selling a protection to a portfolio manager on the capital invested in a TVS.
-the portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS.
-Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario.
+In this contribution we face a new topic never dealt with in the TVS literature: we deal with the funding costs coming from hedging the risky assets underlying the TVS.
+We consider the point of view of a bank selling a call option to a portfolio manager as protection on the capital invested in a TVS.
+The portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS.
+Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario, i.e.
+the most expensive strategy from the financing costs point of view.
 Hence, the pricing problem becomes a dynamical control problem over the risky portfolio composition.
-In our contribution we describe the dynamical control problem, and we derive an analytical solution in the simple case of the risky assets driven by a Black-Scholes (BS) model.
-Then, we tackle the problem in the general case by using reinforcement learning (RL) algorithms.
+In our contribution we describe the dynamical control problem, and we derive an analytical solution in the simple case of the risky assets driven by a Black-Scholes (BS) model \citep{BlackScholes}.
+Then, we tackle the problem in the general case of a local volatility (LV) model \citep{Dupire94pricingwith,LVfirst} by using Reinforcement Learning (RL) algorithms, in particular the Proximal Policy Optimization developed in \textcite{ppo}.
 \\
@@
 In \Cref{sec:TVS} we describe the dynamics of a TVS in presence of valuation adjustments.
-Then, in \Cref{sec:Derivative} we derive the analytical results, and in \Cref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem.
-We conclude the paper with \Cref{sec:Numerical_results} where we present the numerical results obtained in this work.
+Then, in \Cref{sec:Derivative} we study the case of derivative contracts on TVSs, we describe the dynamical control problem giving the expression of the Hamilton-Jacobi-Bellman equation and we derive the analytical results for plain vanilla options under BS dynamics.
+In \Cref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a deep description of the algorithm we have built.
+We conclude the paper with \Cref{sec:Numerical_results} where we present the numerical results obtained in this work for the Black and Scholes model and the local volatility (LV) one.
 
@@
 \begin{equation}
-    \frac{dI_t^{\text{MP}}}{I_t^{\text{MP}}} \coloneqq \left(r_t - \ell \left(t,I_t^{\text{MP}}\right)\right)dt + \bar{\sigma}dW_t^{\text{MP}},
+    \frac{dI_t^{\text{MP}}}{I_t^{\text{MP}}} \coloneqq \left(r_t - \ell_{\alpha} \left(t,I_t^{\text{MP}}\right)\right)dt + \bar{\sigma}dW_t^{\text{MP}},
 \label{eq:markovian_projection}\end{equation}
@@
 \begin{equation}
-    \ell \left(t,K\right) \coloneqq \bar{\sigma} \mathbb{E}_0\left[\frac{\mu_t \cdot \alpha_t}{\|\alpha_t \cdot \nu_t \|}\bigg|I_t = K\right],
+    \ell_\alpha \left(t,K\right) \coloneqq \bar{\sigma} \mathbb{E}_0\left[\frac{\mu_t \cdot \alpha_t}{\|\alpha_t \cdot \nu_t \|}\bigg|I_t = K\right],
 \end{equation}
@@
 
-\subsection{Black and Scholes Model}
-In the \textcite{BlackScholes} model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$.
+\subsection{Black and Scholes Model}\label{subsec:BS_model}
+In the Black and Scholes model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$.
 As a consequence, the local drift can be written as
 \begin{equation}
-     \ell \left(t,K\right) = \bar{\sigma}   \frac{\mu\left(t\right) \cdot \alpha\left(t,K\right)}{\|\alpha\left(t,K\right) \cdot \nu\left(t\right)\|},  
+     \ell_\alpha \left(t,K\right) = \bar{\sigma}   \frac{\mu\left(t\right) \cdot \alpha\left(t,K\right)}{\|\alpha\left(t,K\right) \cdot \nu\left(t\right)\|},  
 \end{equation}
@@
 \begin{proposition}
-When the underlying securities follow a Black and Scholes (BS) model with deterministic rates, the optimal strategy for a monotone payoff consists in minimizing the local drift function, indipendently of the current state $I_t$
+When the underlying securities follow a Black and Scholes (BS) model with deterministic rates, the optimal strategy for a monotone payoff consists in minimizing the local drift function, independently of the current state $I_t$
 \begin{equation}
@@
 \end{proposition}
-The absence of stochastic elments in \Cref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori}; in fact one can solve the optimization problema once for all for each $t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities.
+The absence of stochastic elements in \Cref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no simulation needed; in fact one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities.
 Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
 \begin{equation}
-	V_0^{BS} = BS(F^{\text{TVS}}(0,T;\alpha^*), K, T, \bar{\sigma}, D(0,T;\zeta)),
-\end{equation}
-where $F^{\text{TVS}}(t,T;\alpha)$ is the TVS forward curve defined by
+	V_0^{\text{BS}} = BS(F^{\text{TVS}}(0,T;\alpha^*,K), K, T, \bar{\sigma}, D(0,T;\zeta)),
+\label{eq:bs_closed_solution}\end{equation}
+where $F^{\text{TVS}}(t,T;\alpha,K)$ is the TVS forward curve defined by
 \begin{equation}
-	F^{\text{TVS}}(t,T;\alpha) = I_t \exp\left[\int_t^T\left(r(u)-\bar{\sigma}\frac{\alpha(u)\cdot \mu(u)}{\|\alpha(u)\cdot \nu(u)\|}\right)du\right],
+	F^{\text{TVS}}(t,T;\alpha, K) = I_t \exp\left[\int_t^T\left(r(u)-\ell_\alpha \left(u,K\right)\right)du\right],
 \end{equation}
@@
 \begin{lemma}
-	Let be $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \nu^T$.
+	Let be $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \tr{\nu}$.
 Then the closed solution of the optimization problem \eqref{eq:BS_optimal_strategy} is
@@
 \begin{lemma}
-	Let $\mu \in \mathbb{R}^n$ be a vector with nonnegative components, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix, and $\Sigma = \nu \nu^{T}$. Then
+	Let $\mu \in \mathbb{R}^n$ be a vector with nonnegative components, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix, and $\Sigma = \nu \tr{\nu}$. Then
 	\begin{equation}\label{eq:inf}
@@
 
-In our work we adopt as learning strategy the Proximal Policy Optimization (PPO) developed in \textcite{ppo} and \textcite{gae}.
+In our work we adopt as learning strategy the proximal policy optimization (PPO) developed in \textcite{ppo} and \textcite{gae}.
 PPO is an actor-critic algorithm well-suited for continuous control problems.
@@
 In PPO the policy distribution and the value function are parameterized by a set of parameters $\theta$: $J^\theta$, $\pi^\theta$.
-The PPO collects a small batch of experiences interacting with the environment to update $\theta$ by computing an estimator of the policy gradient and plugging it into a stochastic gradient ascent algorithm in order to maximize an objective function $L^{\text{PPO}}$.
+The PPO collects a small batch of experiences interacting with the environment to update $\theta$ by computing an estimator of the policy gradient and plugging it into a stochastic gradient ascent (SGA) algorithm in order to maximize an objective function $L^{\text{PPO}}$.
 The PPO objective is defined as
@@
  \subsection{Episode Description: Agent-Environment Interaction}
- We consider an episode $\tau$ of length $m+1$ that takes place on a discrete time-grid of fixing times $\mathcal{T}:=\{T_0,...,T_k,...,T_m\}$ with $T_0:=0$ and $T_m=T$ maturity of the option.
+ We consider an episode $\tau$ of length $m+1$ that takes place on a discrete time-grid of fixing times expressed in year fractions $\mathcal{T}:=\{T_0,...,T_k,...,T_m\}$ with $T_0:=0$ and $T_m=T$ maturity of the option.
 At a given episodic time $T_k$ the RL agent interacts with the environment: it receives a representation of the environment called state $s_k$ and on the basis of that it selects an action $a_k$ sampling from the current policy $\pi^\theta_k$.
@@
 \begin{equation}
-	V_{\text{BS}}(T_k) \coloneqq BS(F^{\text{TVS}}(T_k,T;\alpha^*_{\text{BS}}),K,T-T_k,\bar{\sigma},D(T_k,T;\zeta)) \quad \text{and} \quad V(T_0)=0
+	V_{\text{BS}}(T_k) \coloneqq BS(F^{\text{TVS}}(T_k,T;\alpha^*_{\text{BS}},K),K,T-T_k,\bar{\sigma},D(T_k,T;\zeta)) \quad \text{and} \quad V(T_0)=0
 \end{equation} 
@@
 \end{equation}
-In the \textsc{Baselines} implementation of PPO, the parameter $\log\sigma$ is state-independent, but it decreases as the number of PPO update iterations increases.
+In the \textsc{OpenAI Baselines} implementation of PPO, the parameter $\log\sigma$ is state-independent, but it decreases as the number of PPO update iterations increases.
 The idea is that the log-standard deviation will be higher at the beginning of the training phase in order to guarantee a good exploration of the action space while it will be lower at the ending phase to avoid too much noise in the proximity of the optimal policy.
@@
 In fact, if on one hand with free $\alpha$ it is sufficient to set bounds in the action space, on the other hand the presence of constraints on the risky strategy imposes the need to parameterize the output of the ANN.
-In this work we study the constraint where the agent can invest in the risky market adopting long positions only:
+In this work we study the constraint where at time $T_k\in\mathcal{T}$ the agent can invest in the risky market adopting long positions only:
 \begin{equation}
-	\alpha \cdot \mathds{1} = 1 \quad \alpha^i >0 \quad \forall i=1,\dots,n.
+	\alpha_{T_k} \cdot \mathds{1} = 1 \quad \alpha_{T_k}^i >0 \quad \forall i=1,\dots,n.
 \end{equation} 
@@
 \begin{itemize}
-	\item \textbf{$\bm{(n-1)}$-spherical angles}: we rewrite $\alpha^j = (x^i)^2$ and we use $(n-1)$-spherical coordinates with unitary radius.
-Thus the strategy is parameterized by a vector of angles $\phi = (\phi^1,\dots,\phi^{n-1})^T$ as follows
+	\item \textbf{$\bm{(n-1)}$-spherical angles}: we rewrite $\alpha_{T_k}^j = (x^i_k)^2$ and we use $(n-1)$-spherical coordinates with unitary radius.
+Thus the strategy is parameterized by a vector of angles $\phi_k = \tr{(\phi_k^1,\dots,\phi_k^{n-1})}$ as follows
 	\begin{equation}
 		\begin{aligned}
-			\alpha^1 &\coloneqq \cos^2 (\phi^1), \\
-			\alpha^2 &\coloneqq \sin^2(\phi^1)\cos^2(\phi^2),\\
+			\alpha_{T_k}^1 &\coloneqq \cos^2 (\phi_k^1), \\
+			\alpha_{T_k}^2 &\coloneqq \sin^2(\phi_k^1)\cos^2(\phi_k^2),\\
 			&\vdots \\
-			\alpha^{n-1} &\coloneqq \sin^2(\phi^1)\dots\sin^2(\phi^{n-2})\cos^2(\phi^{n-1}),\\
-			\alpha^n &\coloneqq \sin^2(\phi^1)\dots\sin^2(\phi^{n-2})\sin^2(\phi^{n-1}).
+			\alpha_{T_k}^{n-1} &\coloneqq \sin^2(\phi_k^1)\dots\sin^2(\phi_k^{n-2})\cos^2(\phi_k^{n-1}),\\
+			\alpha_{T_k}^n &\coloneqq \sin^2(\phi_k^1)\dots\sin^2(\phi_k^{n-2})\sin^2(\phi_k^{n-1}).
 		\end{aligned}
 		\label{eq:euler_angles}\end{equation}
-	where $\phi^i\in[0,\pi/2]$ to avoid oscillations in $\alpha$ that could slow down the training.
-Thus, with this parameterization, the ANN sample a vector of angles ($a = \phi$) and thus the action space is reduced by one dimension.
+	where $\phi_k^i\in[0,\pi/2]$ to avoid oscillations in $\alpha$ that could slow down the training.
+Thus, with this parameterization, the ANN sample a vector of angles ($a_k = \phi_k$) and thus the action space is reduced by one dimension.
 	\item \textbf{Normalized action}: we restrict the agent actions to a subset of $(\mathbb{R}^+)^n$ and then normalize the action to obtain the corresponding risky allocation strategy through the formula
 	\begin{equation}
-		\alpha_t = \frac{a_t}{a_t \cdot \mathds{1}}.
+		\alpha_{T_k} = \frac{a_k}{a_k \cdot \mathds{1}}.
 		\label{eq:normalized_action}\end{equation}
@@
 
-\section{Numerical Investigation}\label{sec:Numerical_results}
-In this section As we have seen in 
+\clearpage
+\section{Numerical Investigations}\label{sec:Numerical_results}
+In this section we present the numerical results of our work.
+In particular we focus our analysis on an European option on the TVS with at-the-money strike considering two different models for the dynamics followed by the risky securities: Black and Scholes and local volatility.
+In \Cref{subsec:BS_model} we have proved that under the BS dynamics the solution of the control problem is given by \Cref{eq:BS_optimal_strategy}.
+We want to take advantage of this \textit{a priori} solution as a benchmark to gather evidence on the robustness of our RL approach and to check if our analytical result is correct.
+Moreover, the availability of this faster solution allows let us to identify the key parameters of the ANN, PPO and the action parameterization, so that to analyze how they impact on the final results and performances.
+Then we dedicate a full section to the LV case where the dynamical control problem must be solved by the RL agent.
+
+In our work we keep fixed to the default values of \textsc{OpenAI Baselines} some PPO hyper-parameters that will not be study of our fine tuning tests: the SGA learning rate $\rho=3\times10^{-4}$, the lambda GAE parameter $\lambda=0.95$ and the  $L^\pi$ clip threshold\footnote{We refer to \textcite{ppo} for a more accurate description of $\epsilon$ and to \textcite{gae} for $\lambda$.} $\epsilon=0.2$.
+In every training phase, PPO builds two FFNNs with same architecture and activation functions to evaluate respectively the actions and the value function at each episode step $k$.
+The update of their hidden weights $\theta$ is performed after collecting a mini-batch of 2048 episodes.
+
+The two FFNNs take in input the environment state \eqref{eq:state}.
+It is well known in literature that ANNs give better performances in the training phase if the input data are well normalized \citep{inputnormalization,normalizationinput}.
+Thus we manipulate the input state as follows:
+\begin{equation}
+	s_{T_k} = [T_k,X_{T_k}, I_{T_k}/I_0] \quad \forall T_k\in\mathcal{T},
+\end{equation}
+with 
+	\begin{equation}
+	X_{T_k} \coloneqq \frac{\log\left(S_{T_k}/F(0,T_k)\right)+ \frac{1}{2}\int_{0}^{T_k} (\nu^2(t)\cdot \mathds{1})dt}{\sqrt{\int_{0}^{T_k} (\nu^2(t)\cdot \mathds{1})dt}}, \quad X_{T_k} \in [-2.5,2.5]^n,
+	\label{eq:input_normalization}\end{equation}
+where $F(t,T)$ is the forward curve vector of the risky assets from $t$ to $T$.
+With the normalization procedure \eqref{eq:input_normalization}  allows to remap $S_{T_K}$ in a standard Gaussian variable for the BS model, while in a proxy of it in the case of a LV dynamics.
+We adopt this normalization without loss of generality as the variables involved are observables (forward curve and at-the-money volatilities), as they are values that are always accessible by the fund manager in the market.
+
+All the training experiments required on average 40 CPU cores, 1 NVIDIA Volta V100 GPU and 100 GB RAM to manage the parallel environments generated by \textsc{OpenAI Baselines}; the code is written in Python.
+
+\subsection{Black and Scholes: Hyper-Parameters Fine Tuning}
+
+\begin{comment}
+
+In the Black and Scholes environment we study some markets composed by simple 
+\begin{figure}[ht!]
+\centering
+\includegraphics[width=10cm,height=7cm]{Learning_5x8_marketdata_long_5x8_3asset_beta07_noise05_best_WINDOW5e6_EUR_900dpi}
+\caption{Learning curve for an agent policy parameterized with $5$ layers, $8$ neurons per layer and $\tanh$ as activation function.
+The agent is trained in a BS environment with three equities.
+On the horizontal axis the number of training episodes.
+The solid lines are the moving average of the realized rewards on the last $5\times 10^6$ episodes.
+The shadows represent the $98\%$ confidence intervals.
+The dash-dotted horizontal red line is the conservative TVO price computed according to the optimal risky portfolio composition through \eqref{eq:bs_closed_solution}.}
+\label{fig:Learning_curve_3asset_BS}\end{figure}
+
+ In the last section we provide a first analysis of the LV model where a closed strategy is not available.
+The BS environment is used as a toy model in order to understand which parametersof the ANN and of the PPO play key roles in our RL algorithm.
+
 \subsection{Black and Scholes Dynamics}
+--normalizzazione degli input.
 \subsection{Local Volatility Dynamics}
 \section{Discussion}
+\end{comment}
 \begin{comment}

### 2021-05-24 tex/targetvol.tex
Paper: BS numerical results

@@
 	urlcolor= mydarkblue,linkcolor= mydarkblue,citecolor= mydarkgreen,
-	pdfauthor={Daluiso, R., Pallavicini, A., Nastasi, E., Polo, S.},
+	pdfauthor={Daluiso, R., Nastasi, E., Pallavicini, A., Polo, S.},
 }
@@
 \fancyhead[R]{}
-\lhead{\footnotesize\slshape Daluiso, Pallavicini, Nastasi, Polo, Optimal Strategies for Target Volatility Contracts}
+\lhead{\footnotesize\slshape Daluiso, Nastasi, Pallavicini, Polo, Optimal Strategies for Target Volatility Contracts}
 \chead{}
@@
 
-\subsection{Policy and Action Parameterization}
+\subsection{Policy and Action Parameterization}\label{subsec:policy_action}
 As discussed before, PPO parameterizes the agent policy by means of an artificial neural network.
@@
 In every training phase, PPO builds two FFNNs with same architecture and activation functions to evaluate respectively the actions and the value function at each episode step $k$.
-The update of their hidden weights $\theta$ is performed after collecting a mini-batch of 2048 episodes.
+The library initialize the hidden weights $\theta$ using an orthogonal initialization with $\sqrt{2}$ scaling factor and  update them after collecting a mini-batch of 2048 episodes.
+The network biases are set to zero.
 
@@
 It is well known in literature that ANNs give better performances in the training phase if the input data are well normalized \citep{inputnormalization,normalizationinput}.
-Thus we manipulate the input state as follows:
-\begin{equation}
-	s_{T_k} = [T_k,X_{T_k}, I_{T_k}/I_0] \quad \forall T_k\in\mathcal{T},
-\end{equation}
-with 
+Thus we modify the state block by replacing $\forall T_k\in\mathcal{T}$ the TVS spot price $I_{T_k}$ with $I_{T_k}/I_0$ and the equity price vector $S_{T_k}$ with:
 	\begin{equation}
@@
 \subsection{Black and Scholes: Hyper-Parameters Fine Tuning}
+We use the BS environment as toy model to understand which parameters of the RL algorithm play key roles in the training and testing phase.
+We also study which action parameteriziation is most effective in the case of a long only allocation strategy $\alpha$.
+We consider a target volatility call option with the following contract details between bank and fund manager
+\begin{equation}\nonumber
+	K = 1 \, \text{EUR}, \quad T=1 \, \text{yr}, \quad \bar{\sigma} = 5\%.
+\end{equation}
+We test the action parameterizations discussed in \Cref{subsec:policy_action} increasing the number $n$ of equities in the basket; we have evidence that the $(n-1)$-sphere parameterization  \eqref{eq:euler_angles} suffers of labelling problems when $n>2$.
+In fact if we consider a spherical case ($n=3$), \cref{eq:euler_angles} introduces an asymmetry in the action space (\Cref{fig:n_sphere_parameterization} on the left) that translates into the choice by the agent of sub-optimal policies during the training phase.
+Because of that we privilege the action parameterization through normalization \eqref{eq:normalized_action}.
+ In \Cref{fig:n_sphere_parameterization} on the right we provide the learning curve of a $5\times8$ ANN\footnote{We use the notation $L\times N$ to describe a FFNN with $L$ hidden layers each made up of $N$ neurons.} with $tanh$ as activation function, $\gamma=1$, $c_1=0.7$ and $c_2=0$.
+This learning curve, just like all the ones we will show in this work, is the best result in term of reward on the last $10^7$ episodes among four simulations with different initial seed for the weights $\theta$.
+This is done since the objective function is not convex.
 
+ In \Cref{fig:n_sphere_parameterization} we can observe that the average cumulative reward converges as the number of training episode increases to the optimal option value calculated \textit{a priori} with the BS formula \eqref{eq:bs_closed_solution}, proving a successful learning of the optimal BS strategy.
+This can be observed also by fixing a never-seen MC scenario and interrogating the trained agent on this test path: we see that the ANN takes actions that are according to the closed allocation strategy (\Cref{fig:RL_actions}).
+
+ We use the above PPO parameters since are the results of our tests of fine tuning (\Cref{fig:fine_tuning}).
+In particular we train the agent in BS environments with simple market data structures building a grid search on different hyper-parameters: the ANN architecture, the activation function and the value function trade-off $c_1$.
+We first fix a $2\times3$ ANN and $c_1=0.5$ and study the activation function, where we see that the $\tanh$ outperforms the popular sigmoid and RELU that reaches suboptimal policies.
+Then we perform a search in the $c_1$ direction, where the optimal value is 0.7.
+At the end we study the number of layers and neurons in the FFNN; we can see in \Cref{fig:fine_tuning} top-right that adding hidden layers in the ANN architecture gives better results in the learning phase than adding hidden neurons per layer, since deeper ANNs seem less sensitive to get stuck in local optimal strategies.
+We do not need to turn on the entropy element $c_2$ since in the BS environment the RL agent finds successfully $\pi^*$, and adding an entropy term could deteriorate the learning phase.
+
+ All those numerical analysis will be useful to study the LV environment.
+
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[height=5cm,width=5.8cm]{actionspace3D.pdf}
+	\includegraphics[width=8.3cm,height=6.5cm]{Learning_5x8_marketdata_long_5x8_3asset_beta07_noise05_best_WINDOW5e6_EUR_900dpi}
+
+	\caption{On the left $n$-sphere strategy parameterization for 3 equities: $\alpha$ in function of the colatitude $\phi_1$ and latitude $\phi_2$.
+The domain of the angles is $[0,\pi/2]$.
+On the right the  learning curve for an agent policy parameterized with $5\times8$ and $\tanh$ as activation function.
+The agent is trained in a BS environment with three equities.
+On the horizontal axis the number of training episodes.
+The solid lines are the moving average of the realized rewards on the last $5\times 10^6$ episodes.
+The shadows represent the $98\%$ confidence intervals.
+The dash-dotted horizontal red line is the conservative TVO price computed according to the optimal risky portfolio composition through \cref{eq:bs_closed_solution}.}
+	\label{fig:n_sphere_parameterization}
+\end{figure}  
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[width=7.2cm,height=6cm]{Instant_Optimal_percentage_strategy_three_assets_real_market.pdf}
+	\includegraphics[width=7.2cm,height=6cm]{Action_RL_5x8_marketdata_seed16_long.pdf}
+	\caption{Comparison between the BS optimal risky strategy (\Cref{eq:BS_optimal_strategy}) and the RL actions taken in a test episode never seen by the agent.
+On the horizontal axis time in term of year fractions.
+The agent training session is showed in \Cref{fig:n_sphere_parameterization}.
+}\label{fig:RL_actions}\end{figure}
+
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[height=6cm,width=7.2cm]{Activation_functions_Learning_curve_1x40_network_confront_activation_strategy_EUR_8x5_900dpi.pdf}
+	\includegraphics[height=6cm,width=7.2cm]{Value_function_Learning_curve_1x40_beta_confront_varying_strategy_EUR_8x5_400dpi_compressed.pdf}
+	\includegraphics[height=6cm,width=7.2cm]{Hidden_Layers_Learning_curve_3asset_varystrategy_window1000000_EUR_5e8timestep_comparison_900dpi_compressed.pdf}
+	\caption{Fine tuning simulations: on the top-left learning curves with different activation functions, on the top-right grid search on the policy-value trade-off and on bottom-centre learning curves on ANNs architectures.
+On the horizontal axis the number of training episodes.
+The solid lines are the moving average of the realized rewards on the last $5\times 10^5$ episodes.
+The shadows represent the $98\%$ confidence intervals.
+The dash-dotted horizontal red lines are the conservative TVO prices computed according to the optimal risky portfolio composition through \cref{eq:bs_closed_solution}.}
+	\label{fig:fine_tuning}
+\end{figure}  
+\subsection{Local Volatility Dynamics}
 \begin{comment}

### 2021-05-25 reinforcement/LSMC.py
Least Square MC: implementation of vectorized operations for TVS dynamics

@@ -1,0 +1,101 @@
+import numpy as np 
+from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, LV_model, ForwardVariance, quad_piecewise
+from envs.pricing.targetvol import Drift
+from envs.pricing.loadfromtxt import LoadFromTxt
+from envs.pricing.targetvol import Markowitz_solution
+from envs.pricing.closedforms import European_option_closed_form
+
+def alpha_vector(a1, nu, norm=1):
+    beta = nu[1,0]**2+nu[1,1]**2
+    delta = (a1**2)*(nu[0,0]**2+nu[0,1]**2)
+    gamma = a1*(nu[0,0]*nu[1,0]+nu[1,1]*nu[0,1])
+    epsilon = delta-norm
+    x2 = (-gamma-np.sqrt(gamma**2-epsilon*beta))/beta
+    return np.array([a1,x2])
+
+def nu_matrix(sigma_t_vector,identity_3d, corr_chole, N_equity, Nsim):
+    S = sigma_t_vector.T*identity_3d
+    n = S.T@corr_chole
+    return n.transpose(1,2,0).reshape(N_equity,N_equity,Nsim)
+
+def alpha_dot_nu(alpha_matrix,nu_matrix, N_equity, Nsim):
+    a = np.sum(alpha_matrix*nu_matrix.transpose(1,2,0),axis=2)
+    return a.transpose(1,0)
+    
+N_equity = 2
+frequency = "month"
+target_volatility = 5./100
+I_0 = 1.
+K = I_0     #option strike 
+T = 2.      #option expiry
+Nsim = 5  #number of MC paths
+
+
+I_grid = np.linspace(0.,I_0*10,100)
+alpha_grid = np.linspace(-5,5,100)
+
+shape = (N_equity,N_equity,Nsim)
+identity_3d = np.zeros(shape)
+idx = np.arange(shape[0])
+identity_3d[idx, idx, :] = 1
+
+identity_3d
+
+Identity = np.identity(N_equity)
+Nsim = int(Nsim)
+ACT = 365.0
+if frequency == "month":
+    month_dates = np.array([31.,28.,31.,30.,31.,30.,31.,31.,30.,31.,30.,31.])
+    months = month_dates
+    if T > 1.:
+        for i in range(int(T)-1):
+            months = np.append(months, month_dates)
[…]

### 2021-05-25 reinforcement/baseline_simulation/run.py

@@ -69,6 +69,7 @@ for i in range(len(names)):
 model = LV_model(fixings=observation_grid[1:], local_vol_curve=LV, forward_curve=F, N_grid = N_euler_grid)
 euler_grid = model.time_grid
 discount = D(T)
+print(discount)
 dt_vector = model.dt
 mu_function = Drift(forward_curves=F)
 mean = np.array([])
@@ -105,8 +106,8 @@ for i in range(Nsim):
     I_t = I_0[i]
    # counter = 0
    # counter_tot = 0
-    sigma = simulations_Vola[0]
-    norm_price = dS_S[0]
+    sigma = simulations_Vola[i]
+    norm_price = dS_S[i]
     for j  in range(len(observation_grid[1:])):
         dt = dt_vector[j]
         index_plus = j*N_euler_grid
@@ -143,7 +144,7 @@ for i in range(Nsim):
   #  mean = np.append(mean,counter/counter_tot*100)
    # if i>0 and i%20==0:
     #    print("mean accuracy ",np.mean(mean))
-    simulations_Vola = np.delete(simulations_Vola,0,axis=0)
-    dS_S = np.delete(dS_S,0,axis=0)
+    #simulations_Vola = np.delete(simulations_Vola,0,axis=0)
+    #dS_S = np.delete(dS_S,0,axis=0)
 
 print("EXECUTION TIME",(time.clock()-t0)/3600)

### 2021-05-25 reinforcement/go.py

@@ -125,13 +125,13 @@ if __name__ == '__main__':
         do_train=True,
         do_test=False,
         test_seed='114',
-        training_seed='13247',
+        training_seed='45891',
         env='TVS_LV_newreward-v0',
         alg='ppo2',
         num_layers='3',
         num_hidden='5',
         num_env='27',
-        lr='8e-5',
+        lr='3e-5',
         train_timesteps='4e8',
         test_episodes='1e6',
         print_episodes='1',

### 2021-05-25 reinforcement/go_test.py

@@ -125,13 +125,13 @@ if __name__ == '__main__':
         do_train=False,
         do_test=True,
         test_seed='114',
-        training_seed='13247',
+        training_seed='45891',
         env='TVS_LV_newreward-v0',
         alg='ppo2',
         num_layers='3',
         num_hidden='5',
         num_env='27',
-        lr='8e-5',
+        lr='3e-5',
         train_timesteps='4e8',
         test_episodes='1e6',
         print_episodes='1',
@@ -141,4 +141,5 @@ if __name__ == '__main__':
         beta='0.7',
         custom_suffix='freefrombaseline_displacedmarket_2assets_monthgrid_maturity2'   #test on one_month 6,10,1e6
         )
+
     main(cur_args)

### 2021-05-25 reinforcement/print_logs.py

@@ -9,10 +9,10 @@ strategy = 'free'
 
 LOG_FOLDER = './logs/TVS_LV_newreward-v0/'   #in which folder I find the file to plot
 WHICH_LOGS = [  #("name of the output folder", "legend label")
-('ppo2_9e7_2x8_3e-4_trainingseed20319_beta0.7_copyfreefrombaseline_displacedmarket_2assets_monthgrid_maturity2','RL'),
+('ppo2_4e8_3x5_3e-5_trainingseed45891_beta0.7_copyfreefrombaseline_displacedmarket_2assets_monthgrid_maturity2','RL'),
 ]
 X_AXIS_TIMESTEPS = 0  # otherwise: episodes
-WINDOW = int(3e5)  # measured in episodes  10000
+WINDOW = int(5e5)  # measured in episodes  10000
 join_learning = 0  
 fig, ax = plt.subplots(1, 1, figsize=(8, 5))
 

### 2021-05-25 reinforcement/print_test_results.py

@@ -3,8 +3,8 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 ### RL Test Result ####
-MC_price = 0.041024303986084185
-MC_error =  5.0366052653869903e-05
+MC_price = 0.04108891194279346
+MC_error =   5.039159686727298e-05
 
 ### BS PRICE ###
 BS_Price = 0.040947694174563704 

### 2021-05-25 tex/targetvol.tex
paper

@@
     \frac{dS_t^i}{S_t^i} = \left(r_t - \mu_t^i \right)dt + \nu_t^i \cdot dW_t,
-\label{Equity_process}\end{equation}
+\label{eq:Equity_process}\end{equation}
 where $\nu_t$ is an adapted matrix process ensuring the existence of a solution for the SDE and $W_t$ is a $n$-dimensional vector of Brownian motions under $\mathbb{Q}$.
@@
 \subsection{Local Volatility Dynamics}
+In this section we study the TVS control problem assuming a local volatility model for the dynamics of the risky assets.
+Thus in this case we have a diffusive term in \Cref{eq:Equity_process} that is a deterministic function both of time and state, i.e.
+the spot price, $\nu_t=\nu(t,S_t)$.
+This additional dependency of volatility makes the problem of finding the optimal strategy non-trivial; in fact if we consider the whole equity smiles then the second order term in the HBJ equation \eqref{eq:HJB_general} is non null and thus a closed formula for $\alpha^*$ is not available anymore.
+Because of that one must resort on Reinforcement Learning to recover the problem solution.
+Unlike BS model where the strategy depends only on time, in LV dynamics there are no unnecessary information provided by the state block in \Cref{eq:state} to take the optimal action.
+
+We first apply our RL algorithm in the same market data conditions as in the case showed in \Cref{fig:n_sphere_parameterization} on the right, but in the LV dynamics.
+This choice is done to compare the optimal solution in two well-known financial models.
+We keep the same ANN architecture as in previous section: a $5\times8$ FFNN with $\tanh$ activation function.
+This enables us to perform the training using two different initial guesses for the PPO:
+\begin{itemize}
+	\item \textbf{Random starting point}: the hidden weights $\theta$ of the ANN are randomly initialized so that the training phase starts from a casual policy.
+	\item \textbf{Black \& Scholes initial guess}: the algorithm initialize the values of $\theta$ with those of the ANN trained in the BS environment. 
+\end{itemize}    
+The idea behind the second guess is that the LV dynamics differs from BS for a corrective term in the HJB equation and thus we guess that a network which has learnt the BS optimal strategy will be a good starting point for the RL algorithm as sub-optimal value.
+We present in \Cref{fig:local_vol_learning_curves} on the left the resulting learning curves for two ANNs trained with the two initialization methods just described.
+
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[height=6cm,width=7.2cm]{Learning_curve_all_5x8_localvol_3asset_compressed.pdf}
+	\includegraphics[height=6cm,width=7.2cm]{Action_RL_5x8_marketdata_seed80_long.pdf}
+	\caption{On the left, learning curve for two $5\times8$ FFNNs with $\tanh$ activation function and $c_1 = 0.7$.
+The networks are both trained in the same LV but with different starting point: random policy for the orange line and BS optimal policy for the blue one.
+The dash-dotted horizontal red line is the conservative BS call option price  computed according to the optimal risky portfolio in \Cref{fig:RL_actions} on the left.
+	On the right agent actions in a never-seen MC episode.	
+}
+	\label{fig:local_vol_learning_curves}
+\end{figure}
+
+\noindent Analyzing the training results, we observe that the two curves converges both to the same saturation value which is compatible with the optimal BS price for the TVO curve.
+The RL agent has learnt in both cases the optimal allocation strategy of the BS model.
+In fact if we plot the agent actions taken in never-seen scenarios (\Cref{fig:local_vol_learning_curves} on the right), we see that they are in accordance with those of BS.
+
+As told before, in the LV model the optimal strategy is not available \textit{a priori} as in BS and thus we can not say if the solution found by the RL agent is sub-optimal.
+However we can check if the agent is stuck in a local optimum by building a n\"aif strategy to measure the performance of the RL algorithm; we call this strategy \virgolette{baseline}.
+Since we have proved in \Cref{subsec:BS_model} that in the BS model the solution is the allocation strategy that maximizes the local drift of the TVS, we choose as baseline a manager that applies path-wise the BS solution \eqref{eq:BS_optimal_strategy}.
+In \Cref{tab:mc_results_lv} we compare the MC prices obtained applying in $10^6$ paths the baseline and the ANNs trained before.
+\begin{table}[t]
+	\centering
+	\begin{tabular}{|c|c|} \hline
+		Method & TVO price [EUR]  \\ \hline 
+		RL from random $\theta$ & $(1.912 \pm 0.003) \times 10^{-2}$ \\
+		RL from BS $\theta$ & $(1.915 \pm 0.003) \times 10^{-2}$ \\
+		Baseline & $(1.915 \pm 0.004) \times 10^{-2}$ \\ \hline
+	\end{tabular}\caption{MC prices of an atm-spot TVO with LV dynamics for the risky assets and with $K=1\,[\text{EUR}]$, $T = 1\,[\text{yr}]$ and $\bar{\sigma}=5\%$.
+The prices are obtained computing three different allocation strategies: the ANN trained starting from a random policy, the ANN trained from the BS strategy and the baseline.
+}
+	\label{tab:mc_results_lv}\end{table}
+All the prices computed with different strategies are $2\sigma$-compatible among them.
+Thus we can assert that in the market data conditions we have considered, the LV optimal strategy coincides with the BS one.
+Moreover we observe that the baseline strategy and the BS optimal one provide the same TVO price, such as the risky securities dynamics is irrelevant for the resulting price.
+
+Thus we try different configurations for the hedging costs $\mu_t$ and the implied volatility surfaces in order to find a market where the TVO prices  
 \begin{comment}
@@
 
+%%Buona simulazione local vol [ ppo2_1.5e8_5x8_3e-4frombaselinelong_displacedmarket_month_2assets_beta0.7_maturity2_seed5343061 ] su macchina remota in TVS_LV-v0
+
 \ref{sec:target_volatility_option_derivative_pricing}, the bank evaluation of the most conservative call option price strictly depends on the dynamics followed by the equities underlying the \acrshort{tvs}; in particular the volatility term of the securities affects the \acrshort{tvs} price process \eqref{TVS_last}.

### 2021-05-26 tex/targetvol.tex
Minor changes to the paper
Mostly grammar and phrasing fixes.

@@
 \usepackage[font=small,labelfont=bf]{caption}
-\graphicspath{{Local_Volatility/}{Black_Scholes}}
+\graphicspath{{Local_Volatility/}{Black_Scholes/}}
 % \setbeamertemplate{bibliography item}[text]
@@
 
-In this contribution we face a new topic never dealt with in the TVS literature: we deal with the funding costs coming from hedging the risky assets underlying the TVS.
+In this contribution we face a new topic never dealt with in the TVS literature: the funding costs coming from hedging the risky assets underlying the TVS.
 We consider the point of view of a bank selling a call option to a portfolio manager as protection on the capital invested in a TVS.
@@
 Then, in \Cref{sec:Derivative} we study the case of derivative contracts on TVSs, we describe the dynamical control problem giving the expression of the Hamilton-Jacobi-Bellman equation and we derive the analytical results for plain vanilla options under BS dynamics.
-In \Cref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a deep description of the algorithm we have built.
+In \Cref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a description of the algorithm we have built.
 We conclude the paper with \Cref{sec:Numerical_results} where we present the numerical results obtained in this work for the Black and Scholes model and the local volatility (LV) one.
@@
 where $V_t$ is the price process of the strategy, and $K$ is the guaranteed capital.
-On the other hand, the fund manager ca n replicate the payoff by means of the put-call parity by investing the capital in a low-risk asset and buying a call on the strategy
+On the other hand, the fund manager can replicate the payoff by means of the put-call parity by investing the capital in a low-risk asset and buying a call on the strategy
 \begin{equation}
@@
 with $M_t^V$ martingale under $\mathbb{Q}$.
-Notice that,as expected from non-arbitrage considerations, the coupons paid by each security appear only in the drift of the security price process, but they do not impact the drift of the strategy.
+Notice that, as expected from non-arbitrage considerations, the coupons paid by each security appear only in the drift of the security price process, but they do not impact the drift of the strategy.
 
@@
 \begin{proposition}
-When the underlying securities follow a Black and Scholes (BS) model with deterministic rates, the optimal strategy for a monotone payoff consists in minimizing the local drift function, independently of the current state $I_t$
+When the underlying securities follow a Black and Scholes (BS) model with deterministic rates, the optimal strategy for a non-decreasing European payoff consists in minimizing the local drift function, independently of the current state $I_t$
 \begin{equation}
     \alpha^*(t) \coloneqq \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|}.
-\label{eq:BS_optimal_strategy}\end{equation}
+\label{eq:BS_optimal_strategy}
+\end{equation}
+Analogously, the optimal strategy for a non-increasing European payoff consists in maximizing the local drift function:
+\begin{equation}
+    \alpha^*(t) \coloneqq \argmax_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|}.
+\label{eq:BS_optimal_strategy_put}
+\end{equation}
 \end{proposition}
@@
 In this section we want to provide to the reader a formal description of the dynamic problem associated to options on target volatility strategies by writing the Hamilton-Jacobi-Bellman equation for the derivative value.
-We prove that from this description it is available the same closed formula \eqref{eq:BS_optimal_strategy} for the time-dependent BS model derived from the Gy\"ongy.
+We prove that from this description one can recover the same closed formula \eqref{eq:BS_optimal_strategy} for the time-dependent BS model which was derived above from the Gy\"ongy Lemma.
 
@@
 \end{equation}
-where $\Tr(A)$ is the trace operator of $A$, $\nabla_X V$ the gradient of $V$ w.r.t. $X$, $H_XV$ the Hessian matrix of $V$ w.r.t. $X$ and $\nabla_{X,I}V $ is the vector defined by:
+where $\Tr(A)$ is the trace operator of $A$, $\nabla_X V$ the gradient of $V$ w.r.t.~$X$, $H_XV$ the Hessian matrix of $V$ w.r.t.~$X$ and $\nabla_{X,I}V$ is the vector defined by:
 \begin{equation}
@@
 \end{equation}
-where $L^\pi$ is the policy surrogate, $L^J$ the value function error term and $L^E$ the entropy bonus, while $c_1$ and $c_2$ are hyper-parameters of the algorithm.
+where $L^\pi$ is a surrogate of the policy's average reward, $L^J$ the value function error term and $L^E$ the entropy bonus, while $c_1$ and $c_2$ are hyper-parameters of the algorithm.
 We refer to \textcite{ppo} for a deeper description of those terms.
- We use the PPO implementation found in \textsc{OpenAI Baselines} \citep{baselines}, where the parameterization procedure is performed by using artificial neural networks (ANNs), where the set of parameters $\theta$ are the hidden weights of the networks.
-In the following section we describe the the way we have formalized the TVS problem in the Reinforcement Learning framework.
+ We use the PPO implementation found in \textsc{OpenAI Baselines} \citep{baselines}, where the parameterization procedure is performed by using artificial neural networks (ANNs), and the set of parameters $\theta$ are the hidden weights of the networks.
+In the following section we describe  the way we have formalized the TVS problem in the Reinforcement Learning framework.
 
@@
 In the \textsc{OpenAI Baselines} implementation of PPO, the parameter $\log\sigma$ is state-independent, but it decreases as the number of PPO update iterations increases.
-The idea is that the log-standard deviation will be higher at the beginning of the training phase in order to guarantee a good exploration of the action space while it will be lower at the ending phase to avoid too much noise in the proximity of the optimal policy.
+The idea is that the log-standard deviation will be higher at the beginning of the training phase in order to guarantee a good exploration of the action space while it will be lower at the end to avoid too much noise in the proximity of the optimal policy.
 
 Another important topic is how to parameterize the agent action in order to satisfy some specific constraints on the risky asset allocation weights.
-In fact, if on one hand with free $\alpha$ it is sufficient to set bounds in the action space, on the other hand the presence of constraints on the risky strategy imposes the need to parameterize the output of the ANN.
+In fact, if on the one hand with free $\alpha$ it is sufficient to set bounds in the action space, on the other hand the presence of constraints on the risky strategy imposes the need to parameterize the output of the ANN.
 In this work we study the constraint where at time $T_k\in\mathcal{T}$ the agent can invest in the risky market adopting long positions only:
 \begin{equation}
-	\alpha_{T_k} \cdot \mathds{1} = 1 \quad \alpha_{T_k}^i >0 \quad \forall i=1,\dots,n.
+	\alpha_{T_k} \cdot \mathds{1} = 1 \quad \alpha_{T_k}^i \geq 0 \quad \forall i=1,\dots,n.
 \end{equation} 
@@
 We want to take advantage of this \textit{a priori} solution as a benchmark to gather evidence on the robustness of our RL approach and to check if our analytical result is correct.
-Moreover, the availability of this faster solution allows let us to identify the key parameters of the ANN, PPO and the action parameterization, so that to analyze how they impact on the final results and performances.
-Then we dedicate a full section to the LV case where the dynamical control problem must be solved by the RL agent.
+Moreover, the availability of this faster solution allows us to identify the key parameters of the ANN, PPO and the action parameterization, and to analyze how they impact on the final results and performances.
+Then we dedicate a full section to the LV case where the dynamical control problem cannot be solved without the RL agent.
 
-In our work we keep fixed to the default values of \textsc{OpenAI Baselines} some PPO hyper-parameters that will not be study of our fine tuning tests: the SGA learning rate $\rho=3\times10^{-4}$, the lambda GAE parameter $\lambda=0.95$ and the  $L^\pi$ clip threshold\footnote{We refer to \textcite{ppo} for a more accurate description of $\epsilon$ and to \textcite{gae} for $\lambda$.} $\epsilon=0.2$.
+In our work we keep fixed to the default values of \textsc{OpenAI Baselines} some PPO hyper-parameters that will not be subject to our fine tuning tests: the SGA learning rate $\rho=3\times10^{-4}$, the lambda Generalized Advantage Estimation parameter $\lambda=0.95$ and the $L^\pi$ clip threshold\footnote{We refer to \textcite{ppo} for a more accurate description of $\epsilon$ and to \textcite{gae} for $\lambda$.} $\epsilon=0.2$.
 In every training phase, PPO builds two FFNNs with same architecture and activation functions to evaluate respectively the actions and the value function at each episode step $k$.
-The library initialize the hidden weights $\theta$ using an orthogonal initialization with $\sqrt{2}$ scaling factor and  update them after collecting a mini-batch of 2048 episodes.
+The library initializes the hidden weights $\theta$ using an orthogonal initialization with $\sqrt{2}$ scaling factor and  updates them after collecting a mini-batch of 2048 episodes.
 The network biases are set to zero.
@@
 where $F(t,T)$ is the forward curve vector of the risky assets from $t$ to $T$.
-With the normalization procedure \eqref{eq:input_normalization}  allows to remap $S_{T_K}$ in a standard Gaussian variable for the BS model, while in a proxy of it in the case of a LV dynamics.
-We adopt this normalization without loss of generality as the variables involved are observables (forward curve and at-the-money volatilities), as they are values that are always accessible by the fund manager in the market.
+The normalization procedure \eqref{eq:input_normalization}  allows to remap $S_{T_K}$ into a standard Gaussian variable for the BS model, while into a proxy of it in the case of a LV dynamics.
+We adopt this normalization without loss of generality as the variables involved are observables (forward curve and at-the-money volatilities), since they are values that are always accessible by the fund manager in the market.
 
@@
 We use the BS environment as toy model to understand which parameters of the RL algorithm play key roles in the training and testing phase.
-We also study which action parameteriziation is most effective in the case of a long only allocation strategy $\alpha$.
+We also study which action parameterization is most effective in the case of a long only allocation strategy $\alpha$.
 We consider a target volatility call option with the following contract details between bank and fund manager
@@
  In \Cref{fig:n_sphere_parameterization} we can observe that the average cumulative reward converges as the number of training episode increases to the optimal option value calculated \textit{a priori} with the BS formula \eqref{eq:bs_closed_solution}, proving a successful learning of the optimal BS strategy.
-This can be observed also by fixing a never-seen MC scenario and interrogating the trained agent on this test path: we see that the ANN takes actions that are according to the closed allocation strategy (\Cref{fig:RL_actions}).
+This can be observed also by fixing a never-seen MC scenario and interrogating the trained agent on this test path: we see that the ANN takes actions that are close to the analytical allocation strategy (\Cref{fig:RL_actions}).
 
- We use the above PPO parameters since are the results of our tests of fine tuning (\Cref{fig:fine_tuning}).
-In particular we train the agent in BS environments with simple market data structures building a grid search on different hyper-parameters: the ANN architecture, the activation function and the value function trade-off $c_1$.
+ We use the above PPO parameters since they are the results of our tests of fine tuning (\Cref{fig:fine_tuning}).
+In particular we have trained the agent in BS environments with simple market data structures building a grid search on different hyper-parameters: the ANN architecture, the activation function and the value function trade-off $c_1$.
 We first fix a $2\times3$ ANN and $c_1=0.5$ and study the activation function, where we see that the $\tanh$ outperforms the popular sigmoid and RELU that reaches suboptimal policies.
 Then we perform a search in the $c_1$ direction, where the optimal value is 0.7.
-At the end we study the number of layers and neurons in the FFNN; we can see in \Cref{fig:fine_tuning} top-right that adding hidden layers in the ANN architecture gives better results in the learning phase than adding hidden neurons per layer, since deeper ANNs seem less sensitive to get stuck in local optimal strategies.
+At the end we study the number of layers and neurons in the FFNN; we can see in \Cref{fig:fine_tuning} top-right that adding hidden layers in the ANN architecture gives better results in the learning phase than adding hidden neurons per layer, since deeper ANNs seem less sensitive to get stuck in locally optimal strategies.
 We do not need to turn on the entropy element $c_2$ since in the BS environment the RL agent finds successfully $\pi^*$, and adding an entropy term could deteriorate the learning phase.

### 2021-05-26 reinforcement/LSMC.py
LSMC

@@ -4,14 +4,17 @@ from envs.pricing.targetvol import Drift
 from envs.pricing.loadfromtxt import LoadFromTxt
 from envs.pricing.targetvol import Markowitz_solution
 from envs.pricing.closedforms import European_option_closed_form
+from scipy.interpolate import interp1d
 
-def alpha_vector(a1, nu, norm=1):
+def alpha_vector(a1, nu, Nsim=1, norm=1):
+    a_1 = np.ones(Nsim)*a1
     beta = nu[1,0]**2+nu[1,1]**2
-    delta = (a1**2)*(nu[0,0]**2+nu[0,1]**2)
-    gamma = a1*(nu[0,0]*nu[1,0]+nu[1,1]*nu[0,1])
+    delta = (a_1**2)*(nu[0,0,:]**2+nu[0,1,:]**2)
+    gamma = a_1*(nu[0,0,:]*nu[1,0,:]+nu[1,1,:]*nu[0,1,:])
     epsilon = delta-norm
     x2 = (-gamma-np.sqrt(gamma**2-epsilon*beta))/beta
-    return np.array([a1,x2])
+    a = np.array(([a_1,x2])).T
+    return a
 
 def nu_matrix(sigma_t_vector,identity_3d, corr_chole, N_equity, Nsim):
     S = sigma_t_vector.T*identity_3d
@@ -19,21 +22,32 @@ def nu_matrix(sigma_t_vector,identity_3d, corr_chole, N_equity, Nsim):
     return n.transpose(1,2,0).reshape(N_equity,N_equity,Nsim)
 
 def alpha_dot_nu(alpha_matrix,nu_matrix, N_equity, Nsim):
-    a = np.sum(alpha_matrix*nu_matrix.transpose(1,2,0),axis=2)
-    return a.transpose(1,0)
+    a = np.sum((alpha_matrix.T*nu_matrix.transpose(1,0,2)),axis=1).T
+    return a
+
+def evolve(index_start, I_t, dS_S, alpha, omega, instant_interest_rate, dt, N_euler_grid):
+    s = np.sum(alpha, axis=1)
+    I_t_plus_1 = np.ones(dS_S[:,0,0].shape)*I_t
+    for t in range(index_start, N_euler_grid):
+        I_t_plus_1 = I_t_plus_1 * (1. + np.sum(alpha*dS_S[:,t,:],axis=1)*omega + dt[index_start] * instant_interest_rate[t]*(1.-omega*s))
[…]

### 2021-05-26 tex/targetvol.tex
Paper
add second formulation of reward function, add thanks, terminated LV

@@
 \usepackage[backend=biber,style=authoryear-comp,maxcitenames=2,sorting=nyt,natbib=true,sortcites=false]{biblatex}
+\setlength\bibitemsep{1.3\itemsep}
+
 \addbibresource{Bib.bib}
@@
 	\and
-	Stefano Polo \thanks{Intesa Sanpaolo Milan, stefano.polo@intesasanpaolo.com} \thanks{We wish to thank the Italian computing centre Cineca, which approved our ISCRA C project and provided us the high-computing resources of Marconi100 for the numerical simulations of this work.}% indicates same as 2nd thanks
+	Stefano Polo\thanks{Intesa Sanpaolo Milan, stefano.polo@intesasanpaolo.com} 
 }
@@
 The topic of hedging costs is a novelty never dealt in the TVS literature and we provide a formal description of the entire control problem.
-We tackle the problem by implementing a Reinforcement Learning algorithm to determine the optimal risky portfolio composition leading to the most conservative option price.
-We investigate the problem for two models of the risky asset dynamics: time-dependent Black \& Scholes and Local Volatility.
-In the first case we prove the existence of an analytical solution of the problem; a result that we use as benchmark to perform a series of fine-tuning of the hyper-parameters of the RL algorithm.
-At the end we provide numerical results for the Local Volatility model, for which an \textit{a priori} solution is not available.
+We tackle the problem by implementing a RL algorithm to determine the optimal risky portfolio composition leading to the most conservative option price.
+We investigate the problem for two models of the risky asset dynamics: time-dependent Black and Scholes and local volatility.
+In the first case we prove the existence of an analytical solution of the problem; a result that we use as benchmark to perform a series of fine-tuning of the RL hyper-parameters.
+At the end we provide numerical results for the local volatility model, for which an \textit{a priori} solution is not available.
 \end{abstract}
@@
 \tableofcontents
+ \mbox{}
+ \vfill
+ \noindent \small The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.
+
 \newpage
@@
 In our contribution we describe the dynamical control problem, and we derive an analytical solution in the simple case of the risky assets driven by a Black-Scholes (BS) model \citep{BlackScholes}.
-Then, we tackle the problem in the general case of a local volatility (LV) model \citep{Dupire94pricingwith,LVfirst} by using Reinforcement Learning (RL) algorithms, in particular the Proximal Policy Optimization developed in \textcite{ppo}.
+Then, we tackle the problem in the general case of a local volatility (LV) model \citep{Dupire94pricingwith,LVfirst} by using Reinforcement Learning (RL) algorithms, in particular the proximal policy optimization developed in \textcite{ppo}.
 \\
@@
 In \Cref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a description of the algorithm we have built.
-We conclude the paper with \Cref{sec:Numerical_results} where we present the numerical results obtained in this work for the Black and Scholes model and the local volatility (LV) one.
+We conclude the paper with \Cref{sec:Numerical_results} where we present the numerical results obtained in this work for the Black and Scholes model and the local volatility one.
+\\
+
+A part of this work has been developed during the Master thesis of one of us (SP), where we had the opportunity to collaborate with Marco Bianchetti and Diego Pierluigi Giovannini from Intesa Sanpaolo, Milan, that we tanks.
+Moreover we wish to thank the Italian computing centre Cineca, which approved our ISCRA C project and provided us the high-computing resources of Marconi100 for the numerical simulations of this contribution.
 
@@
 We start by introducing the Markovian projection of the dynamics followed by $I_t$.
-We name it $I_t^{\text{MP}}$, and we get by applying the Gy\"ongy Lemma
+We name it $I_t^{\text{MP}}$, and we get by applying the Gy\"ongy Lemma \citep{Gyongy}
 \begin{equation}
@@
 This is a stochastic optimal control problem where by homogeneity we can suppose that $\alpha_t$ lives in a compact domain, namely (a subset of) the unit simplex.
-A least if $r_t$, $\mu^i_t$ and $\nu^i_t$ are (uniformly) bounded, and if the eigenvalues of $\nu_t$ are (uniformly) bounded away from zero, then drift and diffusion should be uniformly Lipschitz in $\alpha_t$, and classical theorems should exist.
+A least if $r_t$, $\mu^i_t$ and $\nu^i_t$ are (uniformly) bounded, and if the eigenvalues of $\nu_t$ are (uniformly) bounded away from zero, then drift and diffusion should be uniformly Lipschitz in $\alpha_t$, and classical theorems should exist \citep{StandardTheorems}.
 \end{remark}
@@
 \begin{equation}
-	V_0^{\text{BS}} = BS(F^{\text{TVS}}(0,T;\alpha^*,K), K, T, \bar{\sigma}, D(0,T;\zeta)),
+	V_0^{\text{BS}} = BS(F^{\text{TVS}}(0,T;\alpha^*), K, T, \bar{\sigma}, D(0,T;\zeta)),
 \label{eq:bs_closed_solution}\end{equation}
-where $F^{\text{TVS}}(t,T;\alpha,K)$ is the TVS forward curve defined by
+where $F^{\text{TVS}}(t,T;\alpha)$ is the TVS forward curve defined by
 \begin{equation}
-	F^{\text{TVS}}(t,T;\alpha, K) = I_t \exp\left[\int_t^T\left(r(u)-\ell_\alpha \left(u,K\right)\right)du\right],
+	F^{\text{TVS}}(t,T;\alpha^*) = I_t \exp\left[\int_t^T\left(r(u)-\ell_{\alpha^*} \left(u\right)\right)du\right] \quad \text{with} \quad \ell_{\alpha^*}(t)=\bar{\sigma}   \frac{\mu\left(t\right) \cdot \alpha^*\left(t\right)}{\|\alpha^*\left(t\right) \cdot \nu\left(t\right)\|},
 \end{equation}
 while $BS(F,K,T,\sigma,D)$ is the standard BS formula for an European option with forward curve $F$, strike $K$, time to maturity $T$, volatility $\sigma$ and discount factor $D$.  
-\subsubsection{Black's Closed form solution}
+\subsubsection{Optimal Strategy Closed Form}
 In absence of constraints on the allocation strategy, we are able to derive a closed form solution to the BS problem \eqref{eq:BS_optimal_strategy}
@@
 		\frac{\partial V}{\partial t} +   r_t(X) I \frac{\partial V}{\partial I} + (\nabla_X V) \cdot M(X) + \frac{1}{2} \bar{\sigma}^2I^2 \frac{\partial^2V}{\partial I^2} + \frac{1}{2}\Tr\left(\tr{\Sigma(X)}(H_XV)\Sigma(X)\right)  \\
-		+\max_\alpha  \left\{  -\frac{\partial V}{\partial I} I  \bar{\sigma}\frac{\alpha \cdot \mu_t(X)}{\|\alpha \cdot \nu_t(X)\|} + 
-		(\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(I\bar{\sigma}\frac{\alpha \cdot \nu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right)\right\} = 0.
+		+\bar{\sigma}I\max_\alpha  \left\{  -\frac{\partial V}{\partial I}   \frac{\alpha \cdot \mu_t(X)}{\|\alpha \cdot \nu_t(X)\|} + 
+		(\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(\frac{\alpha \cdot \nu_t(X)}{\|\alpha \cdot \nu_t(X)\|}\right)\right\} = 0.
 		\end{aligned}
@@
 \begin{equation}
-	\frac{\partial V}{\partial t} +  I r(t)\frac{\partial V}{\partial I}  + \frac{1}{2}\bar{\sigma}^2I^2\frac{\partial^2V}{\partial I^2} + \max_\alpha \left\{ -\frac{\partial V}{\partial I} I  \bar{\sigma}\frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t)\|}\right\} = 0,
+	\frac{\partial V}{\partial t} +  I r(t)\frac{\partial V}{\partial I}  + \frac{1}{2}\bar{\sigma}^2I^2\frac{\partial^2V}{\partial I^2} + \bar{\sigma}I\max_\alpha \left\{ -\frac{\partial V}{\partial I}   \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t)\|}\right\} = 0,
 \end{equation}
@@
  This state contains all the informations needed by the agent to take an optimal action, leading to the maximum option price on the TVS. 
- Once the agent has selected the action $a_k$, it receives the next time $T_{k+1}$ a reward $r_{k+1}$ generated by the environment defined by
+ Once the agent has selected the action $a_k$, it receives the next time $T_{k+1}$ a reward $r_{k+1}$ generated by the environment.
+We have defined the reward function in two different ways.
+A first definition is
+ \begin{equation}
+ 	r_{k+1}=\left\{\begin{array}{ll}
+ 		(I_T - K)^+ & \text { if } T_{k+1} = T \\
+ 		0 & \text { otherwise }
+ 	\end{array}\right.
+ \end{equation}
+Therefore, during the episode, the agent receives a nil reward except at maturity when the reward coincides with the option intrinsic value.
+This choice may seem too daring because the agent receives a real feedback of its actions only at the end of the whole episode, leading to a slow learning.
+However, once the agent has learnt the optimal policy $\pi^* = \pi^{\theta^*}$, the average cumulative reward per episodes will coincide to the optimal TVO price.
+The second definition of reward function for this problem we have developed is
  \begin{equation}
@@
 with $\alpha^*_{BS}$ the BS optimal strategy \eqref{eq:BS_optimal_strategy} calculated in the state $s_k$.
-The parameter $\gamma$ plays the role of a discount rate in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future.
-Choosing $\gamma<1$ introduces a bias in the final result.
-If we take the cumulative reward per episode and set $\gamma =1$ we obtain
+This second definition of the reward function does not suffer of nil rewards at intermediate times of the episode.
+The hyper-parameter $\gamma$ plays the role of a discount rate in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future.
+Choosing\footnote{We refer to \textcite{gae} for a more detailed description of $\lambda$.} $\gamma,\lambda<1$ and introduces a bias in the final result.
+If we take the cumulative reward per episode and set $\gamma =\lambda =1$ we obtain
 \begin{equation}
-	R(\tau) = \sum_{k=0}^{m-1} r_{k+1} \underset{\gamma =1}{=} \sum_{k=0}^{m-1}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)] = V_{BS}(T) = (I_T-K)^+,
+	R(\tau) = \sum_{k=0}^{m-1} r_{k+1} \underset{\gamma=\lambda =1}{=} \sum_{k=0}^{m-1}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)] = V_{BS}(T) = (I_T-K)^+,
 \end{equation} 
@@
 
- Thus one can train the agent choosing $\gamma\in[0,1]$, and then run, as test phase, a Monte Carlo (MC) simulation with $\gamma=1$, where, if the agent has reached the optimal policy $\pi^* = \pi^{\theta^*}$, the average of $R(\tau$) along different episodes will match the optimal undiscounted price of the derivative contract on the TVS.
+ Thus one can train the agent choosing $\gamma,\lambda\in[0,1]$, and then run, as test phase, a Monte Carlo (MC) simulation with $\gamma=\lambda=1$, where, if the agent has reached the optimal policy $\pi^* = \pi^{\theta^*}$, the average of $R(\tau$) along different episodes will match the optimal undiscounted price of the derivative contract on the TVS.
 
@@
 
-\clearpage
 \section{Numerical Investigations}\label{sec:Numerical_results}
@@
 \begin{equation}\nonumber
-	K = 1 \, \text{EUR}, \quad T=1 \, \text{yr}, \quad \bar{\sigma} = 5\%.
+	I_0 = K = 1 \, \text{EUR}, \quad T=1 \, \text{yr}, \quad \bar{\sigma} = 5\%.
 \end{equation}
@@
 Because of that we privilege the action parameterization through normalization \eqref{eq:normalized_action}.
- In \Cref{fig:n_sphere_parameterization} on the right we provide the learning curve of a $5\times8$ ANN\footnote{We use the notation $L\times N$ to describe a FFNN with $L$ hidden layers each made up of $N$ neurons.} with $tanh$ as activation function, $\gamma=1$, $c_1=0.7$ and $c_2=0$.
+ In \Cref{fig:n_sphere_parameterization} on the right we provide the learning curve of a $5\times8$ ANN\footnote{We use the notation $L\times N$ to describe a FFNN with $L$ hidden layers each made up of $N$ neurons.} with $\tanh$ as activation function, $\gamma=1$, $c_1=0.7$ and $c_2=0$.
 This learning curve, just like all the ones we will show in this work, is the best result in term of reward on the last $10^7$ episodes among four simulations with different initial seed for the weights $\theta$.
@@
 \end{figure}  
+\clearpage
 \subsection{Local Volatility Dynamics}
@@
 	\item \textbf{Random starting point}: the hidden weights $\theta$ of the ANN are randomly initialized so that the training phase starts from a casual policy.
-	\item \textbf{Black \& Scholes initial guess}: the algorithm initialize the values of $\theta$ with those of the ANN trained in the BS environment. 
+	\item \textbf{Black and Scholes initial guess}: the algorithm initialize the values of $\theta$ with those of the ANN trained in the BS environment. 
 \end{itemize}    
@@
 
-Thus we try different configurations for the hedging costs $\mu_t$ and the implied volatility surfaces in order to find a market where the TVO prices  
+Thus we try different configurations for the hedging costs $\mu_t$ and the implied volatility (IV) surfaces in order to find a market where the TVO prices reached by the baseline and by the BS optimal strategy differ between them.
+From our analyzes we have evidence that in markets where the risky equities have different hedging costs, then the baseline and the BS solution gives the same TVO prices and optimal strategies; this is due to the fact that the numerator in \Cref{eq:BS_optimal_strategy} plays an important role in the minimization and thus in the problem solution.
+For this reason we consider a market of two equities with two different shapes for the smiles curves and a $\mu_t$ vector that has equal positive entries.
+In particular we choose a symmetric smile for one equity and a displaced diffusion for the second one (\Cref{fig:displaced_diffusion} on the left).
+In this way the BS optimal solution with long constraint depends only on the diffusive term:
+\begin{equation}
+	\alpha^*(t) = \mu(t)\argmin_\alpha \|\alpha\cdot\nu(t)\|^{-1}.
+\label{eq:BS_strategy_particular_equal_repo}\end{equation}    
+In a similar way the baseline
+\begin{equation}
+	\alpha^*(t,S_t) = \mu(t)\argmin_\alpha \|\alpha\cdot\nu(t,S_t)\|^{-1}.
+\label{eq:baseline_strategy}\end{equation}
+Therefore, having a completely different smiles curves will make differ the BS solution from the baseline one since the former is computed with a diffusive term $\nu(t)$ obtained by interpolating linearly the IV curve on the atm-spot strike and thus loosing all the information about the $K$-dependence.
+We train the RL agent in this market configuration by parameterizing the actions with the baseline strategy.
+In particular at each observational time $T_K\in\mathcal{T}$ the risky allocation strategy is obtained by summing the ANN output with the \eqref{eq:baseline_strategy} and then remapped so that the long only strategy constraint is satisfied.
+We adopt this parameterization to help the agent find the optimal strategy as correction to the baseline one.
+We report in \Cref{fig:displaced_diffusion} on the right the resulting learning curve compared with the TVO price adopting the state-independent BS strategy \eqref{eq:BS_strategy_particular_equal_repo} and with the baseline \eqref{eq:baseline_strategy}.
+We observe that it learns as optimal policy the baseline strategy.
+We obtain the same result by adopting the second reward function \eqref{eq:reward} and decreasing the $\gamma$ discount factor to make the agent prefer immediate rewards.
+\begin{figure}[ht!]
+	\centering
+	\includegraphics[height=6cm,width=7.2cm]{SmileDisplacedDiff.pdf}
+	\includegraphics[height=6cm,width=7.2cm]{Learning_curve_market_displaced_compressed.pdf}
+	\caption{On the left market smiles for two equities in function of moneyness: symmetric smile at atm for equity 2 and displaced diffusion for equity 1.
+On the right learning curve for a $5\times8$ ANN trained in the LV market with the smile curves on the left.
+The learning curve is compared with the baseline and the static BS strategies.
+}
+	\label{fig:displaced_diffusion}
+\end{figure}
+
+\section{Conclusion and Further Developments}
+In this paper we described a non-trivial control problem related to derivative contracts on target volatility strategies.
+We considered a bank selling a call option to a fund manager as protection  on the capital
+invested on the TVS.
+We showed how the presence of different funding costs coming from hedging the risky assets underlying the TVS, makes the bank solving a stochastic optimal control problem to adjust the price of the protection, since the bank strategy is not self-financing.
+We derived a closed form solution of the control problem in a Black and Scholes framework in two different ways: first by applying the Gy\"ongy  Lemma and then by writing the Hamilton-Jacobi-Bellman equation.
+We studied the problem both in the BS world and in the local volatility one where the solution is not available and thus numerical investigation is needed.
+We tackled the problem by means of the novel Reinforcement Learning techniques, such as the proximal policy optimization.
+We used the BS model, where the solution is \textit{a priori} known, as benchmark to perform a series of fine-tuning of the RL algorithm hyper-parameters, such as the artificial neural network architecture.
+Then we investigated the problem under LV dynamics.
+From our simulations we had evidence that under this model the RL agent learnt as optimal strategy the application path-wise of the BS solution.
+This topic is still under study since our results for the LV model do not provide a formal demonstration of that the problem solution is the path-wise BS one.
+For this reason the natural developments of this work is to compare our RL results with other standard techniques like the least-square Monte Carlo.
+If this comparison is positive, then it will provide robustness of our LV results.
 \begin{comment}
@@
 \end{comment}
-\cleardoublepage
+\\
+
 \printbibliography

### 2021-05-27 reinforcement/LSMC.py
Training of LSMC

@@ -1,14 +1,21 @@
 import numpy as np 
+import statsmodels.api as sm
+from scipy.interpolate import interp1d
+import time
+import matplotlib.pyplot as plt
+
+
+### personal libraries
 from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, LV_model, ForwardVariance, quad_piecewise
 from envs.pricing.targetvol import Drift
 from envs.pricing.loadfromtxt import LoadFromTxt
 from envs.pricing.targetvol import Markowitz_solution
 from envs.pricing.closedforms import European_option_closed_form
-from scipy.interpolate import interp1d
+
 
 def alpha_vector(a1, nu, Nsim=1, norm=1):
     a_1 = np.ones(Nsim)*a1
-    beta = nu[1,0]**2+nu[1,1]**2
+    beta = nu[1,0,:]**2+nu[1,1,:]**2
     delta = (a_1**2)*(nu[0,0,:]**2+nu[0,1,:]**2)
     gamma = a_1*(nu[0,0,:]*nu[1,0,:]+nu[1,1,:]*nu[0,1,:])
     epsilon = delta-norm
@@ -25,12 +32,26 @@ def alpha_dot_nu(alpha_matrix,nu_matrix, N_equity, Nsim):
     a = np.sum((alpha_matrix.T*nu_matrix.transpose(1,0,2)),axis=1).T
     return a
 
-def evolve(index_start, I_t, dS_S, alpha, omega, instant_interest_rate, dt, N_euler_grid):
+def evolve(index_start, I_t, dS_S, Vola, alpha, target_vola, instant_interest_rate, dt, N_euler_grid, N_equity, Nsim, identity_3d, corr_chole):
     s = np.sum(alpha, axis=1)
     I_t_plus_1 = np.ones(dS_S[:,0,0].shape)*I_t
-    for t in range(index_start, N_euler_grid):
-        I_t_plus_1 = I_t_plus_1 * (1. + np.sum(alpha*dS_S[:,t,:],axis=1)*omega + dt[index_start] * instant_interest_rate[t]*(1.-omega*s))
+    for t in range(index_start, N_euler_grid+index_start):
+        if t==index_start:
+            omega = target_vola
+        else:
+            nu = nu_matrix(Vola[:,t,:],identity_3d, corr_chole, N_equity, Nsim)
+            omega = target_vola/np.linalg.norm(alpha_dot_nu(alpha, nu, N_equity, Nsim),axis=1)
[…]

### 2021-05-28 tex/targetvol.tex
Paper
 first completed version

@@
 
-\date{
-	First Version: March 2021. This Version: \today
-}
+%\date{
+%	First Version: March 2021. This Version: \today
+%}
 \begin{document} % goes here
@@
  	\end{array}\right.
- \end{equation}
+ \label{eq:reward2}\end{equation}
 Therefore, during the episode, the agent receives a nil reward except at maturity when the reward coincides with the option intrinsic value.
@@
 \end{equation}
+The reward function adopted in this environment is that expressed by \Cref{eq:reward2}.
 We test the action parameterizations discussed in \Cref{subsec:policy_action} increasing the number $n$ of equities in the basket; we have evidence that the $(n-1)$-sphere parameterization  \eqref{eq:euler_angles} suffers of labelling problems when $n>2$.
@@
 This additional dependency of volatility makes the problem of finding the optimal strategy non-trivial; in fact if we consider the whole equity smiles then the second order term in the HBJ equation \eqref{eq:HJB_general} is non null and thus a closed formula for $\alpha^*$ is not available anymore.
-Because of that one must resort on Reinforcement Learning to recover the problem solution.
+Because of that one must resort on numerical techniques to recover the problem solution.
 Unlike BS model where the strategy depends only on time, in LV dynamics there are no unnecessary information provided by the state block in \Cref{eq:state} to take the optimal action.
@@
 For this reason the natural developments of this work is to compare our RL results with other standard techniques like the least-square Monte Carlo.
-If this comparison is positive, then it will provide robustness of our LV results.
+If the resulting comparison is positive, then it will provide robustness of our LV results.
 \begin{comment}

### 2022-03-08 tex/reinforcement_learning_for_options_on_target_volatility_funds.tex
Code Refactoring

\documentclass[11pt]{article}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[outer=3.75cm, marginparwidth=4.5cm]{geometry}
\usepackage{amssymb,amsmath,amsthm,latexsym,amsfonts,amscd,dsfont,enumerate,bbm,mathabx,mathtools}
\usepackage{graphicx,tikz}
\usepackage{lipsum}
\usepackage{enumerate} 
\usepackage{soul}
\usepackage{comment}
\usepackage{bm}
\usepackage[font=footnotesize,labelfont=bf]{caption}
\graphicspath{{graphs/}}

\definecolor{mylightyellow}{rgb}{1,1,.8}
\definecolor{mylightgreen}{rgb}{.8,1,.8}
\definecolor{mydarkred}{RGB}{178,34,34}
\definecolor{mydarkgreen}{RGB}{34,139,34}
\definecolor{mydarkblue}{RGB}{72,61,139}
\definecolor{mydarkyellow}{RGB}{218,165,32}

\definecolor{myblueA}{RGB}{52,41,39}
\definecolor{myblueB}{RGB}{92,81,109}
\definecolor{myblueC}{RGB}{132,121,179}
\definecolor{myblueD}{RGB}{172,161,249}
\definecolor{myblueE}{RGB}{212,201,255}

\usetikzlibrary{arrows,patterns,shapes,snakes,shadows,fit,backgrounds,positioning,decorations.pathmorphing,external}
\tikzstyle{nameusd} = [circle, draw, top color=white, bottom color=mydarkblue!50, draw=mydarkblue!75!black!100, drop shadow, minimum height=4em]
\tikzstyle{nameeur} = [circle, draw, top color=white, bottom color=mydarkred!50, draw=mydarkred!75!black!100, drop shadow, minimum height=4em]
\tikzstyle{collusd} = [rectangle,fill=mydarkblue!10, inner sep=0.2cm, rounded corners=5mm]
\tikzstyle{colleur} = [rectangle,fill=mydarkred!10, inner sep=0.2cm, rounded corners=5mm]
\tikzstyle{fixto} = [draw, -latex']
\tikzstyle{fixfrom} = [draw, latex'-]
\tikzstyle{floatto} = [draw, snake=coil, segment aspect=0, line before snake=2ex, line after snake=1ex, -latex']
\tikzstyle{floatfrom} = [draw, snake=coil, segment aspect=0, line before snake=2ex, line after snake=1ex, latex'-]
\tikzstyle{investor} = [rectangle, draw, top color=white, bottom color=mydarkyellow!50, draw=mydarkyellow!75!black!100, drop shadow, rounded corners, minimum height=3em, text width=4em, text centered]
\tikzstyle{trader} = [circle, draw, top color=white, bottom color=blue!30, draw=blue!50!black!100, drop shadow, minimum height=4em]
\tikzstyle{bank} = [rectangle, draw, top color=white, bottom color=red!20, draw=red!50!black!100, drop shadow, rounded corners, minimum height=3em, text width=4em, text centered]
\tikzstyle{market} = [rectangle, draw, top color=white, bottom color=green!20, draw=green!50!black!100, drop shadow, rounded corners, minimum height=3em, text width=4em, text centered]
\tikzstyle{yellowbox} = [rectangle, draw, top color=white, bottom color=mydarkyellow!50, draw=mydarkyellow!75!black!100, drop shadow, rounded corners, text centered]
\tikzstyle{fadedyellowbox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkyellow!25, draw=mydarkyellow!50, drop shadow, rounded corners, text centered]
\tikzstyle{redbox} = [rectangle, draw, top color=white, bottom color=mydarkred!50, draw=mydarkred!75!black!100, drop shadow, rounded corners, text centered]
\tikzstyle{fadedredbox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkred!25, draw=mydarkred!50, drop shadow, rounded corners, text centered]
\tikzstyle{bluebox} = [rectangle, draw, top color=white, bottom color=mydarkblue!50, draw=mydarkblue!75!black!100, drop shadow, rounded corners, text centered]
\tikzstyle{fadedbluebox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkblue!25, draw=mydarkblue!50, drop shadow, rounded corners, text centered]
\tikzstyle{greenbox} = [rectangle, draw, top color=white, bottom color=mydarkgreen!50, draw=mydarkgreen!75!black!100, drop shadow, rounded corners, text centered]
\tikzstyle{fadedgreenbox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkgreen!25, draw=mydarkgreen!50, drop shadow, rounded corners, text centered]

\newcommand{\tikzmark}[1]{\tikz[remember picture] \node[coordinate] (#1) {#1};}

% \pgfplotstableread{data/total_consumption_example.dat}\swingtotal

\usepackage[]{hyperref}
\hypersetup{
	breaklinks, 
	baseurl = http://,
	pdfborder = 0 0 0,
	pdfpagemode = UseNone, % do not show thumbnails or bookmarks on opening
	pdfstartpage = 1,
	colorlinks=true,breaklinks=true,
	urlcolor= mydarkblue,linkcolor= mydarkblue,citecolor= mydarkgreen,
	pdfauthor={Daluiso, R., Nastasi, E., Pallavicini, A., Polo, S.},
	pdftitle = {Reinforcement Learning for Options on Target Volatility Funds},
	pdfsubject = {Reinforcement Learning for Options on Target Volatility Funds}
	pdfkeywords = {Reinforcement Learning, Hedging Costs, Funding Costs, Proximal Policy Optimization, Target Volatility, Asset Allocation, Stochastic Control Problem}
}

\usepackage[backend=biber,style=authoryear-comp,maxcitenames=2,maxbibnames=99,sorting=nyt,sortcites=false, natbib=true]{biblatex}
\addbibresource{Bib.bib}

\AtEveryBibitem{\clearfield{month}}
\AtEveryCitekey{\clearfield{month}}
\setlength\bibitemsep{1.3\itemsep}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{remark}{Remark}[section]
\newtheorem{example}[remark]{Example}

\newcommand{\Eq}[1]{{\begin{equation}{#1}\end{equation}}}

\newcommand{\Exo}[1]{\mathbb{E}\!\left[\,#1\,\right]}
\newcommand{\Ex}[2]{\mathbb{E}_{#1}\!\left[\,#2\,\right]}
\newcommand{\ExG}[2]{\mathbb{E}\!\left[\,#2\,\left|{\cal G}_{#1}\right.\right]}
\newcommand{\ExF}[2]{\mathbb{E}\!\left[\,#2\,\left|{\cal F}_{#1}\right.\right]}
\newcommand{\ExFT}[3]{\mathbb{E}^{#2}\!\left[\,#3\,\left.\right|{\cal F}_{#1}\right]}
\newcommand{\ExGT}[3]{\mathbb{E}^{#2}\!\left[\,#3\,\left.\right|{\cal G}_{#1}\right]}
\newcommand{\ExCo}[2]{\mathbb{E}\!\left[\left.\,#1\,\right|\,#2\,\right]}
\newcommand{\ExC}[3]{\mathbb{E}_{#1}\!\left[\left.\,#2\,\right|\,#3\,\right]}
\newcommand{\ExCT}[4]{\mathbb{E}_{#1}^{#2}\!\left[\left.\,#3\,\right|\,#4\,\right]}
\newcommand{\ExT}[3]{\mathbb{E}_{#1}^{#2}\!\left[\,#3\,\right]}
\newcommand{\Exf}[2]{\hat{\mathbb{E}}_{#1}\!\left[\,#2\,\right]}
\newcommand{\ExfT}[3]{\hat{\mathbb{E}}_{#1}^{#2}\!\left[\,#3\,\right]}
\newcommand{\Px}[1]{\mathbb{P}\left\{\,#1\,\right\}}
\newcommand{\PxC}[2]{\mathbb{P}\!\left\{\,#1\,\left.\right|\,#2\,\right\}}
\newcommand{\Qx}[1]{\mathbb{Q}\left\{\,#1\,\right\}}
\newcommand{\QxC}[2]{\mathbb{Q}\!\left\{\,#1\,\left.\right|\,#2\,\right\}}
\newcommand{\QxCT}[3]{\mathbb{Q}^{#1}\!\left\{\,#2\,\left.\right|\,#3\,\right\}}
\newcommand{\QxT}[2]{\mathbb{Q}^{#1}\left\{\,#2\,\right\}}
\newcommand{\Qxf}[1]{\hat{\mathbb{Q}}\left\{\,#1\,\right\}}
\newcommand{\Var}[2]{\operatorname{Var}_{#1}\!\left[\,#2\,\right]}
\newcommand{\Covar}[3]{\operatorname{Covar}_{#1}\!\left[\,#2,#3\,\right]}
\newcommand{\ind}[1]{1_{\{#1\}}}
\newcommand{\clip}[3]{\operatorname{clip}\!\left({#1},{#2},{#3}\right)}
\newcommand{\one}{\ensuremath{\mathbbm{1}}}
\newcommand{\onehalf}{\tfrac{1}{2}}

\newcommand{\tr}[1]{{#1}^{\intercal}} %transpose
\DeclareMathOperator{\Tr}{Tr}
\let\oldnorm\norm   % <-- Store original \norm as \oldnorm
\let\norm\undefined % <-- "Undefine" \norm
\DeclarePairedDelimiter\norm{\lVert}{\rVert}

\newcommand{\rec}{R}
\newcommand{\lgd}{\mbox{L{\tiny GD}}}
\newcommand{\virgolette}[1]{``#1''}
\newcommand{\pvcall}{{\mathcal{C}}}
\newcommand{\pvput}{{\mathcal{P}}}
\newcommand{\rom}[1]{\uppercase\expandafter{\romannumeral #1\relax}}

\usepackage{eurosym}
\newcommand{\euroms}[0]{\mbox{\small \euro}}
\usepackage{cleveref}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\newenvironment{eqsys}{\begin{equation}\begin{dcases}}{\end{dcases}\end{equation}}
\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[R]{}
\lhead{\footnotesize\slshape Daluiso, Nastasi, Pallavicini, Polo, Optimal Strategies for Target Volatility Contracts}
\chead{}
\rhead{\thepage}
\lfoot{}
\cfoot{}
\rfoot{}
\date{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Reinforcement Learning for Options on \\Target Volatility Funds}

\author{
	Roberto Daluiso\thanks{Intesa Sanpaolo Group Milano, \texttt{roberto.daluiso@intesasanpaolo.com}}
	\and 
	Emanuele Nastasi\thanks{Marketz S.p.A. Milano, \texttt{emanuele.nastasi@marketz.eu}}
	\and Andrea Pallavicini\thanks{Intesa Sanpaolo Group Milano, \texttt{andrea.pallavicini@intesasanpaolo.com}} 
	\and
	Stefano Polo\thanks{X-Numeris S.r.l. Milano, \texttt{sp.stefano.polo@gmail.com}} 
}

\begin{document} % goes here

\maketitle

\begin{abstract}
\footnotesize	
\noindent
In this work we deal with the funding costs rising from hedging the risky securities underlying a target volatility strategy (TVS), a portfolio of risky assets and a risk-free one dynamically rebalanced in order to keep the realized volatility of the portfolio on a certain level. The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs for each component requires to solve a control problem to evaluate the option prices. We derive an analytical solution of the problem in the Black and Scholes (BS) scenario. Then we use Reinforcement Learning (RL) techniques to determine the fund composition leading to the most conservative price under the local volatility (LV) model, for which an \textit{a priori} solution is not available. We show how the performances of the RL agents are compatible with those obtained by applying path-wise the BS analytical strategy to the TVS dynamics, which therefore appears competitive also in the LV scenario.
\end{abstract}
\smallskip
\noindent
\textbf{JEL classification codes:} C63, C45, G13.\\
\textbf{AMS classification codes:} 65C05, 68T07, 91G20.\\
\textbf{Keywords:} Reinforcement Learning, Hedging Costs, Target Volatility, Asset Allocation, Stochastic Optimal Control Problem. 
\newpage
\tableofcontents
 \mbox{}
 \vfill
 \noindent \small The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.

\newpage
\pagestyle{myheadings} \markboth{}{{\footnotesize Daluiso, Pallavicini, Nastasi, Polo, Reinforcement Learning for Options on Target Volatility Funds}}

\section{Introduction}
In the recent years portfolio managers are exposed to very low interest rates and quickly changing market volatilities. An effective solution to control risks under such an environment is given by target volatility strategies (TVSs) (also known as constant volatility targeting) which are able to preserve the portfolio at a predetermined level of volatility. A TVS is a portfolio of risky assets (typically  equities) and a risk-free asset dynamically re-balanced with the aim of maintaining the overall portfolio volatility level closed to some target value. The constant volatility approach can help investors to obtain desired risk exposures over the short and long term and it increases the risk-adjusted performance of the portfolio.

These products were initially offered in the Asian markets, see for instance the reports of \textcite{Chew2011} and \textcite{Xue2012}, which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \textcite{Morrison2013}. At the present day we can observe some new market indices based on the mechanism of the target volatility strategy such as Dow Jones Volatility Control Index, and S\&P 500 Risk Control Index.

In the recent literature TVSs are tested to investigate their performances in term of realized returns, see for instance \textcite{Hocquard2013} and \textcite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \textcite{Kim2018}. Moreover, in the present day, in the derivative market new structured financial products that see TVSs as underlying are becoming widely used by practitioners. Those derivative instruments are known as target volatility options (TVOs) and have gained great interest in the derivative pricing literature, see in particular  \textcite{DiGraziano2012}, \textcite{Grasselli2016}, \textcite{Albeverio2019}, and \textcite{DiPersio2019}.

This paper is the first attempt, to the best of our knowledge, to study the funding costs coming from hedging the risky assets underlying the target volatility strategy. In particular, we consider the point of view of a bank selling a call option to a portfolio manager as protection on the capital invested in a TVS. The portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS. Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario, i.e. the most expensive strategy in term of the financing costs. Hence, the pricing problem becomes a continuous dynamical control problem over the risky portfolio composition. In our work we provide to the reader a formal description of the stochastic control problem and most importantly we derive an analytical solution assuming that the risky asset dynamics underlying the TVS  follow a Black and Scholes (BS) model \citep{BlackScholes1973} and that the derivative contract is a European-style option.

Although from a theoretical point of view this solution represents a sub-optimal strategy when dealing with generic dynamics for the risky assets, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem. More precisely we numerically study the problem in the general case of a local volatility dynamics \citep{Dupire1994,Derman1994}  by adopting a Reinforcement Learning approach, in particular an \textit{ad hoc} vanilla direct policy algorithm and the more sophisticated proximal policy optimization technique developed in \textcite{Schulman2017}. 
\\

The paper is organized as follows. In Section \ref{sec:TVS} we introduce more in detail the manager-bank contract and provide the description of the TVS dynamics in presence of valuation adjustments such as the hedging costs. Then, in Section \ref{sec:Derivative} we introduce the structured class of derivative contracts linked to TVSs, where we describe the arising dynamical control problem for pricing those options. Moreover in this section we derive the BS closed solution for European TVOs in two different ways: one applying the Gy\"ongy Lemma and the other by writing the Hamilton-Jacobi-Bellman equation. In Section \ref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a description of the algorithm we have built. We conclude the paper with Section \ref{sec:Numerical_investigations} where we present the numerical results obtained in this work for the Black and Scholes model and the local volatility one.
\\

A part of this work has been developed during the Master thesis of one of us (SP), where we had the opportunity to collaborate with Marco Bianchetti and Diego Pierluigi Giovannini from Intesa Sanpaolo Milan, that we thank. Moreover we wish to thank the Italian computing centre Cineca, which approved our ISCRA C project and provided us with the high-computing resources of Marconi100 for the numerical simulations of this contribution.

\section{Target Volatility Strategy}\label{sec:TVS}
As mentioned before, this work aims to enrich the TVS pricing literature by studying the aspect related to the funding costs coming from hedging the risky assets underlying a target volatility portfolio. Thus we consider the following scenario: a bank selling a protection to a portfolio manager who has his capital invested in a TVS. In our case the fund manager has the freedom of changing the relative weights of the risky asset during the life of the TVS; once the allocation strategy is selected then the volatility targeting algorithm rebalances the risky component of the portfolio with the risk-free one in order to keep the overall portfolio volatility close to a target value. Clients investing in the fund pay a running fee for the service of the fund manager and their capital is protected.

The fund manager usually buys from a bank an option on the TVS to ensure capital protection. For instance, the capital can be protected by buying a put option. In this case, we can write the net asset value (NAV) $A_t$ of the strategy as given by\footnote{Here we neglect discounting factors.}
\begin{equation}
    A_t \coloneqq \max\{V_t,K\} = V_t + (K-V_t)^+ \; ,
\end{equation}
where $V_t$ is the price process of the strategy, and $K$ is the guaranteed capital. On the other hand, the fund manager can replicate the payoff by means of the put-call parity by investing the capital in a low-risk asset and buying a call on the strategy
\begin{equation}
    A_t = K + (V_t-K)^+ \; .
\end{equation}
In this way, the TVS is only defined in the two contracts client-fund and fund-bank. The fund manager is not implementing the strategy by trading in the market, and he is not subject to additional costs to access the market. On the other way, the bank is paying such costs since she is actively hedging the call option sold to the manager.  The bank trading activity implemented to actively hedge the option requires funding the collateral procedures of the hedging instruments along with any lending/borrowing fee.
We remark that the choices of the manager trading activity are stochastic processes since they will depend on the market evolution. Thus, neither the manager strategy nor the bank financing costs for hedging the option can be written in the fund-bank contract. For this reason, the price of a financial product sold by the bank is adjusted to include any valuation adjustment due to the trading activity.

Our aim is to find the most expensive investing strategy from the point of view of financing costs for the bank that the manager could choose in the market. In other words, we want to determine the worst-case scenario for the bank. In the current section, we proceed by defining the price process of the TVS so that we can highlight the impact of valuation adjustments.

\subsection{The strategy Price Process}
We work on a filtered probability space $\left(\Omega, \mathcal{F}, \{\mathcal{F}_t\}_{t\geq0}, \mathbb{P}  \right)$ satisfying the usual assumptions for a market model, where $\mathbb{P}$ is the physical probability measure representing the actual distribution of supply and demand shocks on equities prices.

We consider a fund trading a basket of $n$ risky securities with price process $S_t = (S_t^i, \, i=1,\dots,n)$ funded with a cash account $B_t$ accruing at $r_t$. Any dividend paid by the securities is re-invested in the fund. Here, we assume that the TVS is implemented in continuous time, even if in the practice we can implement the strategy only on a discrete set of dates. We introduce the deflated gain process $\Bar{G}_t^i$ associated with the risky securities as given by
\begin{equation}
    \Bar{G}_t^i \coloneqq \Bar{S}_t^i + \Bar{D}_t^i \; ,
\end{equation}
where we define the deflated price and cumulative dividend processes as 
\begin{equation}
    \bar{S}_t^i \coloneqq \frac{S_t^i}{B_t} \; , \quad \bar{D}^i_t \coloneqq \int_0^t\frac{d\pi_u^i}{B_u} + \int_0^t\frac{d\psi_u^i}{B_u} \; ,
\end{equation}
where $\pi^i_t$ represents the cumulative contractual-coupon process paid by the security, and $\psi_t^i$ represents the cumulative valuation adjustments.
Since fund managers allocating TVS usually rely on equity assets, here we use the results of \textcite{Gabrielli2020} that analyzes the valuation adjustments for equity products. We can write 
\begin{equation}
    \psi_t^i \coloneqq \int_0^t S_u^i\mu_u^i du \; ,
\label{eq:XVA_equity}\end{equation}
where we call $\mu_t^i$ cost of carry, which basically represents the hedging costs for the $i$-th security.

Then, we introduce the strategy price process $V_t$, and we define the deflated gain process $\Bar{G}_t^V$ as given by
\begin{equation}
    \Bar{G}_t^V \coloneqq \frac{V_t}{B_t} + \int_0^t\frac{V_u\phi_u}{B_u}du \; ,
\end{equation}
where $\phi_t$ are the running fees earned by the fund manager for his activity. We assume that the strategy is self-financing, so that we can write
\begin{equation}
    d\Bar{G}_t^V = q_t \cdot d\bar{G}_t \; ,
\label{eq:self_financing}\end{equation}
where $q_t^i$ is the quantity invested in the $i$-th security.\footnote{In all formulae we use dot notation for scalar product between vectors, i.e. $a \cdot b = \sum_i a_i b_i$, or between matrix and vector, i.e. $A \cdot b = \sum_j a_{ij}b_j$ or $b \cdot A = \sum_i b_i a_{ij}$.}

Now, in order to prevent arbitrages, we assume the existence of a risk-neutral measure $\mathbb{Q}$ equivalent to $\mathbb{P}$ under which the deflated gain processes of all traded securities are martingales. Under this assumption, we are able to derive the drift conditions on the security price processes, and in turn on the strategy price process.
\begin{equation}
    \forall u >t \quad \bar{G}_t^i = \mathbb{E}_t \left[\bar{G}_u^i\right] \quad\Longrightarrow \quad dS_t^i = r_tS_t^idt-d\pi_t^i-d\psi_t^i + dM_t^i \; ,
\label{eq:risk_neutral}\end{equation}
where $M_t^i$ are martingale under $\mathbb{Q}$. If we substitute this expression for the security dynamics into the definition of the strategy we can check that the price process of the strategy is accruing at a cash account rate $r_t$ compensated for the fund manager fees
\begin{equation}
    dV_t = V_t(r_t-\phi_t)dt + dM_t^V \; ,
\end{equation}
with $M_t^V$ martingale under $\mathbb{Q}$. Notice that, as expected from non-arbitrage considerations, the coupons paid by each security appear only in the drift of the security price process, but they do not impact the drift of the strategy.

Yet, the strategy priced by $V_t$ cannot be described in the contract between the parties, since Equation~\eqref{eq:self_financing} depends via the security gain processes on the valuation adjustment $\psi_t^i$, which is specific to the investor pricing the strategy. Thus, the TVS defined in the contract will be
\begin{equation}
    d\bar{I}_t \coloneqq q_t \cdot \left(d\bar{S}_t + \frac{d\pi_t}{B_t} \right) - \bar{I}_t\phi_t dt \quad \text{with} \quad I_0 = V_0 \; , 
\label{eq:TVS_first}\end{equation}
leading to the following price process dynamics
\begin{equation}
    dI_t = I_t(r_t-\phi_t)dt -q_t\cdot d\psi_t +dM_t^I \; ,
\end{equation}
with $M_t^I$ martingale under $\mathbb{Q}$. In this case we observe that $I_t$ depends explicitly both on the valuation adjustments and on the allocation strategy. Indeed, if we substitute the valuation adjustments with their explicit expression \eqref{eq:XVA_equity}, we get
\begin{equation}
    dI_t = I_t(r_t-\phi_t)dt - q_t \cdot S_t \mu_t dt + dM_t^I \; ,
\end{equation}
where we can see the dependency on the cost of carry $\mu_t$ and on the allocation strategy.

\subsection{The Volatility Targeting Constraint}
In a typical TVS, the fund manager selects a risky-asset portfolio with a specific time-dependent allocation strategy expressed by means of the vector of relative weights $\alpha_t$, along with a risk-free asset, which we can identify with the bank account $B_t$. In the following, for sake of simplicity in exposition, we consider only total-return securities, namely we set $\pi_t=0$. Thus we can write Equation~\eqref{eq:TVS_first} as given by
\begin{equation}
    \frac{dI_t}{I_t}= \omega_t \alpha_t \cdot \frac{dS_t}{S_t} + \left(1-\omega_t  \alpha_t \cdot \mathds{1} \right)\frac{dB_t}{B_t}- \phi_t dt \; ,
\label{eq:TVS_elegant}\end{equation}
where $\mathds{1}$ is a $n$-dimensional vector of ones and $\omega_t\in[0,1]$ is determined so that the strategy log-normal volatility is kept constant, namely
\begin{equation}
        \omega_t: \quad \mathrm{Var}_t[dI_t] = \bar{\sigma}^2I_t^2dt \; ,
\end{equation}
where $\Bar{\sigma}$ is the target volatility value.
In practice, this means that the fund manager will select a risky-portfolio choosing $\alpha_t$ equities from the universe where he can trade and after that, his choices will be scaled by the automatic target volatility algorithm $\omega_t$.\footnote{We recall that the universe of assets where the manager can trade and the value of $\bar{\sigma}$ are written in the contract.}

To derive the expression for $\omega_t$ we need to assume a generic continuous semi-martingales dynamics under the risk-neutral measure for the underlying securities, so that we can write Equation~\eqref{eq:risk_neutral} as 
\begin{equation}
    \frac{dS_t^i}{S_t^i} = \left(r_t - \mu_t^i \right)dt + \nu_t^i \cdot dW_t \; ,
\label{eq:Equity_process}\end{equation}
where $\nu_t$ is an adapted matrix process ensuring the existence of a solution for the stochastic differential equation (SDE) and $W_t$ is a $n$-dimensional vector of Brownian motions under $\mathbb{Q}$. Under these assumptions we can derive an expression for $\omega_t$, and we get\footnote{In all formulae the norm for a vector $a$ is defined as $\|a\|\coloneqq\sqrt{a\cdot a}$.}
\begin{equation}
    \omega_t = \frac{\Bar{\sigma}}{\norm{ \alpha_t\cdot \nu_t }} \; .
\end{equation}
Hence, putting this last result in the dynamics of $I_t$ we obtain

\begin{equation}
     \frac{dI_t}{I_t} = \left(r_t -  \phi_t     - \frac{\bar{\sigma} \alpha_t}{\norm {\alpha_t \cdot \nu_t }}  \cdot \mu_t \right)dt+ \frac{\bar{\sigma}\alpha_t }{\norm{ \alpha_t \cdot \nu_t }} \cdot \nu_t \cdot dW_t \; ,
\label{eq:TVS_last}\end{equation}
where we can see, as expected, that the strategy grows at the risk-free rate but for adjustments due to valuation adjustments and fees.

We highlight that, to derive the dynamics expressed in Equation~\eqref{eq:TVS_last}, we have not made any assumptions on the risky allocation strategy; thus all this argument is valid for any constraints on the process $\alpha_t$.

\section{Derivative Pricing}\label{sec:Derivative}
In this section, we analyze derivative contracts linked to the TVS described previously. In particular we will focus on European style options, and we will show that under appropriate assumptions it is possible to find a closed form solution for the optimal allocation strategy which maximizes the contract price.

In a general framework, a derivative contract on the TVS with maturity $T$ can be defined as
\begin{equation}
    V_0 \coloneqq \sup_\alpha \mathbb{E}_0\left[\int_0^T D(0,u;\zeta_u)d\pi_u(\alpha)\right] \; ,
\end{equation}
where $D(0,T;\zeta_t)$ is the discount factor with rate $\zeta_t$, inclusive of the derivative valuation adjustments, and $\pi_t$ is the cumulative coupon process paid by the derivative, and it depends on the allocation strategy since in turn the TVS depends on it via the valuation adjustments. We take the supremum over the strategies since we do not have any information on the future activity of the fund manager.

\subsection{European Options}
If the derivative contract depends only on the marginal distribution of $I_T$ at maturity (a European payoff), we are able to calculate an optimal strategy. We consider the following pricing problem
\begin{equation}
    V_0 \coloneqq \sup_\alpha \mathbb{E}_0\left[D(0,T;\zeta)\Phi(I_T(\alpha))\right] \; ,
\end{equation}
where $\Phi$ is the payoff function of the derivative.

We start by introducing the Markovian projection of the dynamics followed by $I_t$. We name it $I_t^{\text{MP}}$, and we get by applying the Gy\"ongy Lemma \citep{Gyongy1986}
\begin{equation}
    \frac{dI_t^{\text{MP}}}{I_t^{\text{MP}}} \coloneqq \left(r_t - \ell_{\alpha} \left(t,I_t^{\text{MP}}\right)\right)dt + \bar{\sigma}dW_t^{\text{MP}} \quad \text{with} \quad I_0^\text{MP} = I_0 \; ,
\label{eq:markovian_projection}\end{equation}
where the local drift is defined as
\begin{equation}
    \ell_\alpha \left(t,K\right) \coloneqq \bar{\sigma} \mathbb{E}_0\left[\frac{\mu_t \cdot \alpha_t}{\norm{\alpha_t \cdot \nu_t }}\bigg|I_t = K\right] \; ,
\label{eq:local_drift}\end{equation}
and $W_t^{\text{MP}}$ is a Brownian motion under the risk-neutral measure $\mathbb{Q}$. Notice that the diffusion coefficient collapses to the target volatility value $\bar{\sigma}$. Since European payoffs depend only on the marginal distribution at maturity, they can be calculated only by means of the Markovian projection $I_t^{\text{MP}}$, namely
\begin{equation}
V_0 \coloneqq \sup_{\alpha} \mathbb{E}_0\left[D\left(0,T;\zeta\right)\Phi\left(I_T^{\text{MP}}\left(\alpha\right)\right)\right] \; .
\end{equation}
Hence, we have our first result valid only if valuation adjustments can be neglected:
\begin{proposition}
A European payoff on the TVS can be calculated by assuming any allocation in the underlying risky basket if all the underlying
securities grow under the risk-neutral measure at the risk-free rate without any valuation adjustment, namely if we can write $\mu_t=0$.
\end{proposition}

\subsection{Stochastic Optimal Control Problem}
In presence of valuation adjustments, we need to solve the full optimization problem. We discretize the optimal strategy $\alpha_t$ as\footnote{We use the symbol $\mathbf{1}_A$ for the indicator function of a subset $A$.}
\begin{equation}
    \alpha_t \coloneqq \sum_k \mathbf{1}_{ \{t \in [T_{k-1}, T_k)\}}\alpha_{T_{k-1}} \; ,
\label{eq:piecewise_strategy}\end{equation}
according to a time grid $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq t$ the pricing date and $T_m \coloneqq T$ the maturity of the option. 
Therefore we can apply the dynamic programming principle to express the optimal $\alpha_t$ at time $T_{k-1}$ as
\begin{equation}
    \alpha_{T_{k-1}}\coloneqq \argmax_{\alpha} \left\{\mathbb{E}_{T_{k-1}}\left[D\left(T_{k-1},T_{k} \right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}(\alpha)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right]\right\} \; ,
\label{eq:recursion}\end{equation}
where $V_{T_k}$ is the option value at time $T_k$ and $X$ is any Markovian state such that the drift and the diffusion coefficient of $I_t$ are a function of $\left(X_t,I_t,\alpha_t\right)$. We calculate $I_{T_k}\left(\alpha_{T_{k-1}}\right)$ for any given strategy $\alpha_{T_{k-1}}$ by a suitable discretization of \eqref{eq:TVS_last} starting from $X_{T_{k-1}}$ and $I_{T_{k-1}}$.

Thus the derivative price is given by:
\begin{equation}
    V_{T_{k-1}}\left(X_{T_{k-1}}, I_{T_{k-1}}\right)=\mathbb{E}_{T_{k-1}}\left[D\left(T_{k-1},T_{k}\right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right],
\label{eq:recursion2}\end{equation}
while the iteration starts from maturity date where the boundary condition is set equal to the payoff function:
\begin{equation}
    V_{T_m} = \Phi\left(I_{T_m}\right) \; .
\end{equation}

\subsection{Black and Scholes Model}\label{subsec:BS_model}
In the time-dependent Black and Scholes model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$. As a consequence, the local drift defined in Equation~\eqref{eq:local_drift} can be written as
\begin{equation}
     \ell_\alpha \left(t,K\right) = \bar{\sigma}   \frac{\mu\left(t\right) \cdot \alpha\left(t,K\right)}{\norm{\alpha\left(t,K\right) \cdot \nu\left(t\right)}} \; ,  
\end{equation}
so that the optimization problem can be solved by looking only at the Markovian projection without simulating all the Brownian motions $W_t$. Notice that we are indicating the dependency on time in parenthesis to highlight that in this formula all the quantities are deterministic functions of time.

A direct consequence is the following proposition, which is relevant for plain vanilla options on TVS.

\begin{proposition}
When the underlying securities follow a Black and Scholes model with deterministic rates, the optimal strategy for a non-decreasing European payoff consists in minimizing the local drift function, independently of the current state $I_t$
\begin{equation}
    \alpha^*(t) \coloneqq \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|} \; .
\label{eq:BS_optimal_strategy}
\end{equation}
Analogously, the optimal strategy for a non-increasing European payoff consists in maximizing the local drift function:
\begin{equation}
    \alpha^*(t) \coloneqq \argmax_\alpha \frac{\alpha \cdot \mu(t)}{\norm{\alpha \cdot \nu(t)}} \; .
\label{eq:BS_optimal_strategy_put}
\end{equation}
\end{proposition}
The absence of stochastic elements in Equation~\eqref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no simulation needed; in fact one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities. Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
\begin{equation}
	V_0^{\text{BS}} = BS(F^{\text{TVS}}(0,T;\alpha^*), K, T, \bar{\sigma}, D(0,T;\zeta)) \; ,
\label{eq:bs_closed_solution}\end{equation}
where $F^{\text{TVS}}(t,T;\alpha)$ is the TVS forward curve defined by
\begin{equation}
	F^{\text{TVS}}(t,T;\alpha) = I_t \exp\left[\int_t^T\left(r(u)-\ell_{\alpha} \left(u\right)\right)du\right] \; ,
\label{eq:tvs_forward}\end{equation}
while $BS(F,K,T,\sigma,D)$ is the standard BS formula for a European option with forward curve $F$, strike $K$, time to maturity $T$, volatility $\sigma$ and discount factor $D$.

In Figure~\ref{fig:strategy_price_comparison} we provide a comparison among the option price obtained with the optimal strategy (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and those with other intuitive strategies (S$_A$, S$_B$, S$_C$). Here we consider the case of an at-the-money call option with spot $I_0 = 1\, \text{EUR}$, maturity $T=5 \, \text{yr}$, target volatility $\bar{\sigma}=5\%$, and a nonnegative constraint on $\alpha$. The intuitive strategies are
\begin{itemize}
\item S$_A$: invest all in the asset with the maximum forward curve at maturity;
\item S$_B$: for each market pillar before maturity invest all in the asset with minimum $\mu(t)$; 
\item S$_C$: for each market pillar before maturity invest all in the asset with minimum $\mu(t)/\|\nu(t)\|$.
\end{itemize}
As the reader can observe, our strategy outperforms any other intuitive one a practitioner might adopt.\begin{figure}[t]
	\centering
	\includegraphics[height=5cm,width=10cm]{strategy_price_comparison.pdf}
	\caption{Comparison of plain vanilla prices on the TVS by adopting different allocation strategies: the optimal Black and Scholes solution (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and other intuitive strategies (S$_A$, S$_B$, S$_C$).}
	\label{fig:strategy_price_comparison}
\end{figure}

\subsubsection{Unconstrained Allocation Strategy: Closed Form Solution}
In absence of constraints on the allocation strategy, we are able to derive a closed form solution to the BS problem \eqref{eq:BS_optimal_strategy}.
\begin{lemma}
	Let be $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \tr{\nu}$. Then the closed solution of the optimization problem \eqref{eq:BS_optimal_strategy} is 
	\begin{equation}
			\alpha^* = - \frac{\Sigma^{-1} \cdot \mu}{\norm{(\Sigma^{-1} \cdot \mu)\cdot \nu}} \; .
	\end{equation}
\end{lemma}
\begin{proof}
Since the argument of the minimum \eqref{eq:BS_optimal_strategy} is zero-homogeneous, then we can rewrite the problem as
\begin{equation}
		\begin{aligned}
			&\text { minimize } \alpha \cdot \mu\\
			&\text { subject to } \|\alpha \cdot \nu\|^2=1 \; .
	\end{aligned}
\end{equation}
By setting the Lagrangian function associated with the problem
\begin{equation}
	\mathcal{L}\left(\alpha, \lambda\right)=\alpha \cdot \mu-\lambda\left(\|\alpha \cdot \nu\|^2-1\right) \; ,
\end{equation}
we obtain the first order conditions
\begin{eqsys}
	\frac{\partial \mathcal{L}}{\partial \alpha}=\mu -2\lambda \Sigma \cdot \alpha=0 \\
	\frac{\partial \mathcal{L}}{\partial \lambda}= \norm{\alpha \cdot \nu}^2-1=0
	\; ,
\end{eqsys}
Then, by applying simple algebra, we obtain the analytical form for the free optimal strategy
\begin{equation}
	\alpha^* = \pm \frac{\Sigma^{-1} \cdot \mu}{\norm{(\Sigma^{-1} \cdot \mu)\cdot \nu}} \; .
\label{eq:analytical_solution_free_strategy}\end{equation}
We take the minus sign to get the minimum value of the TVS local drift while the plus sign for the maximum one (put payoff).
\end{proof}

\subsubsection{Active Asset (or Bang Bang) Solution}
A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry are nonnegative and the only constraint on portfolio weights is nonnegativity, which would mean a long-only strategy by the fund manager.

\begin{lemma}
	Let $\mu \in \mathbb{R}^n$ be a vector with nonnegative components, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix, and $\Sigma = \nu \tr{\nu}$. Then
	\begin{equation}\label{eq:inf}
		\inf_{\alpha \in \mathbb{R}_{+}^n \setminus \{0\}} \frac{\alpha \cdot \mu}{\norm{\alpha \cdot \nu}} = \min_{i\leq n} \frac{\mu_i}{\sqrt{\Sigma_{ii}}} \; ;
	\end{equation}
	if $\bar{\imath}$ is the index which realizes the min, then the infimimum is realized by a vector concentrated on the $\bar{\imath}$ component: $\alpha_i = \delta_{i\bar{\imath}}$. 
\end{lemma}
\begin{proof}
	Let us first consider the case in which $\mu = \mathbbm{1}$. Since the argument of the infimum is zero-homogeneous, normalizing by $\alpha \cdot \mathbbm{1} > 0$ we can restrict to the affine hyperspace $\{\alpha \cdot \mathbbm{1} = 1\}$, where the minimization \eqref{eq:inf} reduces to the maximization of its denominator: the required infimum will be the square root of the reciprocal of
	\begin{equation}
	\sup\left\{\norm{\alpha \cdot \nu}^2 \mid \alpha \in \mathbb{R}_{+}^n,\, \alpha \cdot \mathbbm{1} = 1\right\} \; .
	\end{equation}
	Now we can note that $\Sigma$ is positive definite, hence $\Sigma_{ij} < \sqrt{\Sigma_{ii}\Sigma_{jj}} \leq \Sigma_{\bar{\imath}\bar{\imath}}$, which implies
	\begin{equation}
	\norm{\alpha \cdot \nu}^2 = \sum_{i,j=1}^{n} \alpha_i \alpha_j \Sigma_{ij} \leq \sum_{i,j=1}^{n} \alpha_i \alpha_j \Sigma_{\bar{\imath}\bar{\imath}} = \Sigma_{\bar{\imath}\bar{\imath}} \; ,
	\end{equation}
	because $\sum_i \alpha_i = 1$. Since we trivially have equality for $\alpha_i = \delta_{i\bar{\imath}}$, this concludes the proof of the case $\mu = \mathbbm{1}$.

	Next, let us consider the case in which all components of $\mu$ are strictly positive, and define $M$ as the diagonal matrix with diagonal $\mu$. Then we can rewrite the infimum as a function of $\beta = M\alpha$:
	\begin{equation}
	\inf_{\beta \in \mathbb{R}_{+}^n \setminus \{0\}} \frac{\beta \cdot \mathbbm{1}}{\norm{\beta \cdot M^{-1}\nu}} \; ,
	\end{equation}
	which by the first part of the proof equals 
	\begin{equation}
	\min_{i \leq n} \frac{1}{\sqrt{\tilde{\Sigma}_{ii}}} = \min_{i \leq n} \frac{\mu_i}{\sqrt{\Sigma_{ii}}}, \quad \tilde{\Sigma} \coloneqq M^{-1}\nu\nu^{T} M^{-1} = M^{-1}\Sigma M^{-1} \; .
	\end{equation}

	Finally, let us consider the general case in which $\mu$ may have some components equal to zero. For an arbitrary $\epsilon \geq 0$ let us define
	\begin{equation}
	f_{\epsilon}(\alpha) = \frac{\alpha \cdot (\mu+\epsilon)}{\norm{\alpha \cdot \nu}} \; .
	\end{equation}
	One can easily note that as $\epsilon \to 0$, $f_{\epsilon}$ tends to $f_{0}$ uniformly on the compact set $\{\alpha \in \mathbb{R}^n_+ \mid \alpha \cdot \mathbbm{1} = 1\}$, so that the minimum converges to the minimum on that set. Since we know by homogeneity that the minimum on $\{\alpha \in \mathbb{R}^n_+ \mid \alpha \cdot \mathbbm{1} = 1\}$ equals the minimum on $\mathbb{R}^n_+ \setminus \{0\}$, we conclude
	\begin{equation}
	\inf_{\alpha \in \mathbb{R}_{+}^n \setminus \{0\}} f_0(\alpha) = 
	\lim_{\epsilon \to 0+} \inf_{\alpha \in \mathbb{R}_{+}^n \setminus \{0\}} f_{\epsilon}(\alpha) = 
	\lim_{\epsilon \to 0+} \min_{i\leq n} \frac{\mu_i+\epsilon}{\sqrt{\Sigma_{ii}}} =
	\min_{i\leq n} \frac{\mu_i}{\sqrt{\Sigma_{ii}}} \; . \; 
	\end{equation}
\end{proof}

\subsection{Hamilton-Jacobi-Bellman Equation for Target Volatility Options}\label{subsec:HJB}
In this section, we want to provide to the reader a formal description of the dynamic problem associated with options on target volatility strategies by writing the Hamilton-Jacobi-Bellman (HJB) equation for the derivative price. We prove that from this equation we can recover the same closed formula \eqref{eq:BS_optimal_strategy} derived above from the Gy\"ongy Lemma for the time-dependent BS model.

In full generality, we assume that the time evolution of the Markovian factors governing the problem dynamics is given by  a stochastic multidimensional process in $\mathbb{R}^n$, $X_t$, which is the unique strong solution to the following It$\hat{\text{o}}$ SDE 
\begin{equation}
		dX_t = M(X_t)dt + \Sigma(X_t) \cdot dW_t \; ,
\end{equation}
where $W_t$ is a $n$-dimensional Wiener process with independent components. We point out to the reader that with this notation we are including general dynamics models like those with stochastic drift $M(X_t)$ and stochastic diffusive $\Sigma(X_t)$ coefficients.

Let be the dynamics of the securities $S_t$ a generic function of the Markovian factors, namely
\begin{equation}
 S_t \coloneqq  f(X_t) \; .
\label{eq:S_f}\end{equation}

In this framework, the TVS price process dynamics is given by the SDE 
\begin{equation}
	\frac{dI_t}{I_t} = \left(r(X_t) -  \frac{\bar{\sigma} \alpha_t}{\norm{\alpha_t \cdot \nu(X_t)}}  \cdot \mu(X_t)    \right)dt +\frac{\bar{\sigma}\alpha_t }{\norm{ \alpha_t \cdot \nu(X_t) }} \cdot \nu(X_t) \cdot dW_t \; ,
\end{equation}
where the expression of $\mu(X_t)$ and $\nu(X_t)$ can be recovered by applying the It$\hat{\text{o}}$ formula to Equation~\eqref{eq:S_f}.

Given $X \coloneqq X_t$ and $I \coloneqq I_t$, we can write the HJB equation for $V \coloneqq V(t,I,X)$ as follows
\begin{equation}
	\begin{aligned}
		\frac{\partial V}{\partial t} + \max_\alpha  \left\{   \left(r(X) - \bar{\sigma}\frac{\alpha \cdot \mu(X)}{\norm{\alpha \cdot \nu(X)}}\right) I \frac{\partial V}{\partial I} + (\nabla_X V) \cdot M(X) + \frac{1}{2} \bar{\sigma}^2I^2 \frac{\partial^2V}{\partial I^2}  \right.\\ 
		\left. +\frac{1}{2}\Tr\left(\tr{\Sigma(X)}(H_XV)\Sigma(X)\right)  + (\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(I\bar{\sigma}\frac{\alpha \cdot \nu(X)}{\norm{\alpha \cdot \nu(X)}}\right)\right\} = 0 \; ,
	\end{aligned}
\end{equation}
where $\Tr(A)$ is the trace operator of $A$, $\nabla_X V$ the gradient of $V$ w.r.t. $X$, $H_XV$ the Hessian matrix of $V$ w.r.t.~$X$ and $\nabla_{X,I}V$ is the vector defined by:
\begin{equation}
	\nabla_{X,I} V \coloneqq \tr{\left(\frac{\partial^2 V}{\partial X^1 \partial I},\dots, \frac{\partial^2 V}{\partial X^n \partial I} \right)} \; .
\end{equation}
We take out from the maximum operator all the elements that do not depend on the risky allocation strategy $\alpha$
\begin{equation}
	\begin{aligned}
		\frac{\partial V}{\partial t} +   r(X) I \frac{\partial V}{\partial I} + (\nabla_X V) \cdot M(X) + \frac{1}{2} \bar{\sigma}^2I^2 \frac{\partial^2V}{\partial I^2} + \frac{1}{2}\Tr\left(\tr{\Sigma(X)}(H_XV)\Sigma(X)\right)  \\ 
		+\bar{\sigma}I\max_\alpha  \left\{  -\frac{\partial V}{\partial I}   \frac{\alpha \cdot \mu(X)}{\norm{\alpha \cdot \nu(X)}} + 
		(\nabla_{X,I}V) \cdot \Sigma(X) \cdot \left(\frac{\alpha \cdot \nu(X)}{\norm{\alpha \cdot \nu(X)}}\right)\right\} = 0 \; .
		\end{aligned}
\label{eq:HJB_general}\end{equation}
Equation~\eqref{eq:HJB_general} is the Hamilton-Jacobi-Bellman equation describing the TVS dynamic problem for a generic dynamics of the risky securities underlying the portfolio.

If we assume a time-dependent BS dynamics for the risky equities ($\mu_t$, $r_t$ and $\nu_t$ deterministic), then $V = V(t,I)$ and all the derivatives w.r.t. $X$ are zero. Therefore the reduced HJB equation is
\begin{equation}
	\frac{\partial V}{\partial t} +  r(t) I \frac{\partial V}{\partial I}  + \frac{1}{2}\bar{\sigma}^2I^2\frac{\partial^2V}{\partial I^2} + \bar{\sigma}I\max_\alpha \left\{ -\frac{\partial V}{\partial I}   \frac{\alpha \cdot \mu(t)}{\norm{\alpha \cdot \nu(t)}}\right\} = 0 \; ;
\end{equation}
if the payoff is non-decreasing in $I$ by homogeneity of the SDE we get that $V$ is non-decreasing; thus the solution is given by
\begin{equation}
 \alpha^*(t) = \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t)\|} \; ,
	\end{equation}
which is the same result expressed in Equation~\eqref{eq:BS_optimal_strategy}. On the other hand, if the payoff is non-increasing in $I$ then the solution will be the $\argmax$.

Conversely, if we deal with a dynamic model for which the derivative contract $V$ depends on $X$ then the volatility versor term, namely the second one, in the $\max$ operator of Equation~\eqref{eq:HJB_general} is no longer zero and thus one must solve the entire control problem numerically.

In this work we tackle the non-trivial case of a local volatility (LV) model for the $S_t$-dynamics, such that $\nu_t = \nu(t,S_t)$. We have chosen the LV model since it is well known in financial literature and among practitioners.

\section{Reinforcement Learning}\label{sec:RL}
As we have discussed in the previous sections, one must resort to numerical approaches to solve the stochastic control problem related to the TVS in the case of general payoffs or risky securities dynamics. The standard approach could be to use classical techniques based on backwards recursion \eqref{eq:recursion}-\eqref{eq:recursion2} such as American Monte Carlo \citep{Longstaff2001}. However, their performances degrade exponentially as the dimension $n$ of the problem increases, making prohibitively costly finding the solution to the problem. In our contribution, we adopt a novel technique which is free from the curse of dimensionality and is gaining popularity in many scientific branches for solving stochastic optimal control problems: Reinforcement Learning (RL) \citep{Sutton2018}.

Reinforcement Learning is a branch of Machine Learning which allows an artificial agent to interact with an environment through actions and observations in order to maximize some notion of cumulative reward. In RL the agent is not told which actions to take but instead, it must discover by trial and error which are the behaviours  yielding the highest reward. This is obtained by updating the agent policy $\pi$ which is a mapping from the environment states to the set of actions. Thus RL is independent of pre-collected data as opposed to other Machine Learning techniques. Because of its nature, RL has been successful in quantitative finance for solving control problems; among the most important RL applications in this field, we refer to \textcite{Deng2017} as the pioneers in studying self-taught reinforcement trading problems, while to \textcite{Kolm2019} and \textcite{Halperin2020}  for hedging derivatives with RL under market frictions.

In our work we adopt two learning strategies to compare their performances in terms of overall reward: an \textit{ad hoc} direct policy algorithm and the state of the art proximal policy optimization (PPO) developed in \textcite{Schulman2017} and \textcite{Schulman2016}.

The first method is a specific algorithm developed by us to fit the problem aim to find the optimal option price; this technique can take place as a direct policy optimization method in the wide taxonomy of Reinforcement Learning algorithms. We will go into details in Section~\ref{sec:direct_policy}.

On the other hand, the PPO is a high-level actor-critic algorithm well-suited for continuous control problems. It collects a small batch of experiences interacting with the environment to update its decision-making policy. From those interactions with the environment, PPO is able to compute the expected reward and the value function. We will not provide a complete description of this sophisticated learning strategy; for more details, we refer to the authors' papers. In our work, we adopt the PPO implementation found in \textsc{OpenAI Baselines}.\footnote{\url{https://github.com/openai/baselines}.}

In the following sections, we describe the way we have formalized the TVS problem in the Reinforcement Learning framework.

\subsection{Direct Policy Approach}\label{sec:direct_policy}
We consider an episode $\tau$ of length $m+1$ that takes place on a discrete time-grid of fixing times expressed in year fractions $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$ maturity of the option. At a given episodic time $T_k$ the RL agent interacts with the environment: it receives a representation of the environment called state $s_k$ and on the basis of that it selects an action $a_k$ sampling from the current policy $\pi^{\theta_j}$. Here with $\theta_j$ we refer to the set of parameters through which we parameterize the agent policy at the $j$-th algorithm iteration. In our case of study the agent can choose the composition of the risky asset portfolio, so that the policy is the allocation strategy $\alpha$ introduced in Equation~\eqref{eq:TVS_elegant}:
\begin{equation}
 	a_k = \alpha_{T_k} \quad \forall T_k \in \mathcal{T} \; , \quad a_k \in \mathcal{A}\subset \mathbb{R}^n \; .
\label{eq:action}\end{equation}

Since the value function of the problem depends on the Markovian state $X_t$, the portfolio level $I_t$ and time $t$, our natural choice for the observation state is the following block
\begin{equation}
 	s_{k} \coloneqq \left[X_{T_k}, I_{T_k}, T_k \right] \quad \forall T_k \in \mathcal{T} \; , \quad s_k \in \mathcal{S} \subset \mathbb{R}^{n+2} \; .
\label{eq:state}\end{equation}
In this way, the state contains all the information needed by the agent to take an optimal action, leading to the maximum plain vanilla TVO price.

In this algorithm, we parameterize the agent policy with a feed forward neural network (FFNN), such that $\theta_j$ coincides with the hidden weights, $s_k$ with the inputs, and $a_k$ with the output. In this way, we are dealing with a deterministic policy, where its functional form is given by the neural network. The parameters update is performed as follows: the agent collects a finite batch of experiences interacting with the environment in a set of episodes $\tau$, then the loss function is evaluated as
\begin{equation}
L(\theta_j) = D(0,T;\zeta)\hat{\mathbb{E}}\left[ \Phi(I_T(\alpha(\theta_j))) \right] \; ,
\label{eq:direct_policy_loss}\end{equation}  
where the expectation $\hat{\mathbb{E}}[\dots]$ indicates the empirical average on the batch. Then the parameters are updated by plugging the policy into a stochastic gradient ascent algorithm. The choice of Equation~\eqref{eq:direct_policy_loss} is justified by the fact that, when the algorithm will find the optimal set $\theta^*$, then $L(\theta^*)$ will be a good proxy of the optimal option price.

Once the training phase is ended and the agent has selected the optimal policy, we can run a Monte Carlo (MC) simulation with never seen scenarios to price the optimal target volatility option and test if the algorithm has not overfitted the data.

\subsection{Proximal Policy Optimization Approach}
As for the direct policy approach, we model the pricing problem considering an episode that takes place on the time-grid $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$. Again we choose as observation state the block defined in Equation~\eqref{eq:state} and the agent policy coincides with the risky allocation strategy $\alpha$ \eqref{eq:action}. An important difference from the previous method is that: once the agent has selected the action $a_k$ sampled from the current policy, it receives at the next time $T_{k+1}$ a reward $r_{k+1}$ generated by the environment.

In our work, we have defined two different reward functions with the purpose to analyze which could help the agent to learn more efficiently the optimal policy $\pi^* = \pi^{\theta^*}$. 
Our first definition is
 \begin{equation}
 	r_{k+1}=\left\{\begin{array}{ll}
 		 D(0,T;\zeta)(I_T(\alpha(\theta_j)) - K)^+ & \text { if } T_{k+1} = T \\
 		0 & \text { otherwise }
 	\end{array}\right.
 \label{eq:reward1} \; . \end{equation}
Therefore, during the whole episode, the agent receives a nil reward except at maturity when the reward coincides with the option intrinsic value. This choice may seem too daring because the agent receives a real feedback of its actions only at the end of the episode, increasing the probability to obtain a slow learning. However, if the agent has learnt $\pi^*$, the average cumulative reward per episode will coincide with the optimal TVO price.

The second reward function we have defined is 
\begin{equation}
 	r_{k+1} \coloneqq  \gamma^{k}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)] \; ,
 \label{eq:reward2}\end{equation}
where $\gamma \in[0,1]$ is an hyper-parameter of the PPO, while $V_{\text{BS}}(T_k)$ is a proxy of the residual option price defined by
\begin{equation}
\begin{aligned}
	V_{\text{BS}}(T_k) &\coloneqq BS(F^{\text{TVS}}(T_k,T;\alpha^*_{\text{BS}},K),K,T-T_k,\bar{\sigma},D(T_k,T;\zeta)) \; ,\\
	V_{\text{BS}}(T_0)&=0 \; , 
\end{aligned}
\end{equation} 
with $\alpha^*_{BS}$ the BS optimal strategy \eqref{eq:BS_optimal_strategy} evaluated at the state $s_k$. In this form the agent actions are hidden inside the term $I_t$ used to compute the TVS forward curve $F^{\text{TVS}}$ defined by Equation~\eqref{eq:tvs_forward}. In this case, the reward function does not suffer of nil values for $0<T_k<T$: the RL agent always gets a feedback from the environment for its choices.  The hyper-parameter $\gamma$ plays the role of a discount factor in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future.
If we take the cumulative reward per episode and set the PPO parameters\footnote{We refer to \textcite{Schulman2016} for a more detailed description for the generalized advantage estimation parameter $\lambda$.} $\gamma =\lambda =1$ we obtain
\begin{equation}
	R(\tau) = \sum_{k=0}^{m-1} r_{k+1} \underset{\gamma=\lambda =1}{=} \sum_{k=0}^{m-1}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)] = V_{BS}(T) = (I_T-K)^+ \; ,
\end{equation} 
which is equal to the intrinsic value of the option. This result does not depend on the definition of $V_{\text{BS}}$  $\forall T_k<T$, but we conjecture that the closer $V_{\text{BS}}$ is to the value function, the easier the agent is in learning.

Thus one can train the agent choosing the optimal value for $\gamma,\lambda\in[0,1]$, and then run, as test phase, a Monte Carlo simulation with $\gamma=\lambda=1$ and the optimized $\theta$ fixed, where, if the agent has learnt $\pi^*$, the average of $R(\tau$) along different episodes will match the optimal undiscounted price of the derivative contract on the TVS.

In the \textsc{OpenAI Baselines} implementation of the PPO, the agent policy is parameterized again by a neural network; as for the previous method we have chosen an FFNN. In particular with PPO we deal with a stochastic policy whose functional form is a multivariate diagonal Gaussian distribution where the mean $\mu^\theta(s_k)$ is the output vector of the FFNN and the log-standard deviation $\log\sigma$ is an external parameter
\begin{equation}
	\pi^\theta(s_k) \sim \mathcal{N}(\mu^\theta(s_k),e^{\log\sigma}) \; .
\end{equation}
As one can observe, $\log\sigma$ is state-independent, but it is reduced as the number of the PPO update iterations increases. The idea is that the log-standard deviation will be higher at the beginning of the training phase in order to guarantee a good exploration of the action space while it will be lower at the end to avoid too much noise in the proximity of the optimal policy.

The fact that the PPO implementation exploits a stochastic policy ensures us a better exploration of the action space than with the previous approach. Moreover, the algorithm tries to learn an approximator of the on-policy value function as control variate for the training phase. This approximator is an FFNN with the same architecture as the one for the policy.

\section{Numerical Investigations}\label{sec:Numerical_investigations}
Here we present the numerical results obtained with our proposed methods. We focus our analysis on a European call option on a TVS with the following product details
\begin{equation}\nonumber
	I_0 = K = 1 \, \text{EUR} \; , \quad T=2 \, \text{yr} \; , \quad \bar{\sigma} = 5\% \; .
\end{equation}
Moreover without loss of generality we consider the case of a completely free allocation strategy $\alpha$. The extension to the constrained case is easy.

It is our aim to investigate the control problem under non-trivial dynamics like the local volatility model where the volatility of the risky asset is also a function of the state. By looking at the HJB equation \eqref{eq:HJB_general}, we expect that the volatility versor will play a role in finding the optimal solution, giving rise to a non-trivial strategy.

Although in Section~\ref{subsec:BS_model} we have proved that under the Black and Scholes model the Equation~\eqref{eq:BS_optimal_strategy} solves the control problem, we want to take advantage of this \textit{a priori} solution as a benchmark to gather evidence on the robustness of our Reinforcement Learning approach and to check if our analytical result is correct. Moreover, we use the BS model as numerical laboratory to perform fine-tuning tests for the RL algorithms hyper-parameters and analyze how they impact the final results and performances.

We recall that in both the algorithms we parameterize the agent policy with an FFNN; thus this is completely characterized by the following hyper-parameters: number of hidden layers, number of neurons per hidden layer, activation function per hidden layer. This is due to the fact that in this Reinforcement Learning problem the number of the input neurons is equal to the state space dimension \eqref{eq:state}, while the number of the output ones coincides with the action space dimension \eqref{eq:action}. It is well known in the literature that neural networks give better performances in the training phase if the input data are well normalized \citep{Sola1997,Puheim2014} . Thus we choose as state $s_k$ the following normalized block  
\begin{equation}
	[\log(S_{T_k}/F(0,T_k)), I_{T_k}/I_0, T_k ] \quad \forall T_k\in\mathcal{T} \; ,
\label{eq:input_state}\end{equation}
where $F(t,T)$ is the forward curve vector of the risky assets from $t$ to $T$. In Equation~\eqref{eq:input_state} we have chosen as Markovian state $X_{T_k}$ the martingale term of the securities dynamics. In this way we have that in the input block the variables have the same order of magnitude.

\subsection{Black and Scholes: Hyper-Parameters Fine Tuning}
We use the BS environment as toy model to understand which parameters of the RL algorithm play key roles in the training and testing phase. 
Our first approach has been the direct policy one since it represents a natural way to tackle the problem: since our goal is to find the allocation 
strategy that maximizes the option price, we update the FFNN weights following the gradient direction of the loss function defined in Equation~\eqref{eq:direct_policy_loss}.

We try to investigate the following hyper-parameters: the FFNN architecture, in particular which feature between the depth and the width of the network is more important, the activation function and the learning rate of the optimizer. Moreover, we compare the performances of two well-known optimizers in Machine Learning literature: Nadam \citep{Dozat2016} and RMSprop \citep{Hiton2012}. Since we deal with a free allocation strategy $\alpha$ that can assume negative values, we have chosen among the wide variety of activation functions the tanh and the elu. In this way, we can analyze the performance of a saturating activation function and a non-saturating one. Firstly we have performed a grid search on the learning rate starting value and we have found $10^{-3}$ a good choice in terms of speed of learning and avoiding over-fitting.
 \begin{figure}[h!]
	\centering
	\includegraphics[width=7.2cm,height=6cm]{BS_fine_tuning_Nadam_tanh.pdf}
	\includegraphics[width=7.2cm,height=6cm]{BS_fine_tuning_Nadam_elu.pdf}
	\caption{Learning curves of the direct policy algorithm applied to the Black and Scholes model. The solid lines are the loss functions defined in Equation~\eqref{eq:direct_policy_loss} in function of the number of training epochs. Each continuous line represents a different neural network architecture with tanh activation function (left) and elu activation function (right). The optimizer adopted is the Nadam with a learning rate of $10^{-3}$. The horizontal red-dashed line is the conservative option price obtained by maximizing the TVS drift through Equation~\eqref{eq:analytical_solution_free_strategy}.}
	\label{fig:result_bs_nadam}
\end{figure}  
 \begin{figure}[h!]
	\centering
	\includegraphics[width=7.2cm,height=6cm]{BS_fine_tuning_RMSprop_tanh.pdf}
	\includegraphics[width=7.2cm,height=6cm]{BS_fine_tuning_RMSprop_elu.pdf}
	\caption{Learning curves of the direct policy algorithm applied to the Black and Scholes model. The solid lines are the loss values defined in Equation~\eqref{eq:direct_policy_loss} in function of the number of training epochs. Each continuous line represents a different neural network architecture with tanh activation function (left) and elu activation function (right). The optimizer adopted is the RMSprop with a learning rate of $10^{-3}$. The horizontal red-dashed line is the conservative option price obtained by maximizing the TVS drift through Equation~\eqref{eq:analytical_solution_free_strategy}.}
	\label{fig:result_bs_rmsprop}
\end{figure}

In Figures~\ref{fig:result_bs_nadam} and \ref{fig:result_bs_rmsprop} we present the learning curves of our fine-tuning tests for RMSProp and Nadam respectively. The three lines of each plot display the learning curves for different FFNN architectures: a one hidden layer network with 50 neurons, a 2 hidden layers with 30 and 20 neurons respectively, and a deeper one with three layers with 20, 15, and 5 neurons. Each learning curve is compared with the optimal option price (red-dashed line) that we have computed with the closed form solution we have derived for the BS model in Section~\ref{subsec:BS_model}. We can see that all the learning curves converge to our expected price, providing a numerical demonstration of our theoretical result. More deeply, from the plots we observe that the RMSprop optimizer outperforms the Nadam. Moreover, the tanh activation function seems to be more preferable than the non-vanishing elu. However, more importantly, we have evidence that a deeper architecture of the neural network outperforms the shallow one. All the learning curves we have presented are the best-in-sample results, in terms of performance, of four runs with different random starting guesses for the hidden weights $\theta$. This procedure is necessary since the objective function is not convex.

We take the $20\times 15 \times 5$ network with tanh from RMSprop as the best optimized network, and we run a Monte Carlo simulation with $10^6$ never-seen scenarios to check if the agent overfits the new data. 
\begin{table}
\begin{center}
\begin{tabular}{ll}
\hline\noalign{\smallskip}
Pricing method & TVO price [EUR] \\
\noalign{\smallskip}
\hline
\noalign{\smallskip}
Analytical Solution & $4.2634\times 10^{-2}$ \\
Direct policy RL & $(4.2624 \pm 0.005) \times 10^{-2}$ \\ 
\hline
\end{tabular}
\end{center}
\caption{Comparison of TVO prices under Black and Scholes model: analytical solution price and direct policy reinforcement learning. The option parameters are: $I_0=K=1\,[\text{EUR}]$, $T=2\,[\text{yr}]$ and $\bar{\sigma}=5\%$.}
\label{tab:mc_results_bs}
\end{table}
\setlength{\tabcolsep}{1.4pt}
We report the results in Table~\ref{tab:mc_results_bs}: the RL price is compatible with the closed formula price. Thus the RL agent did not overfit the data during the training phase.

We will take advantage of those fine-tuning results to tackle the local volatility problem.

\subsection{Local Volatility Dynamics}
In this section, we study the TVS control problem assuming a local volatility model for the dynamics of the risky assets. Thus in this case we have a diffusive term in Equation~\eqref{eq:Equity_process} that is a deterministic function both of time and state, i.e. the spot price, $\nu_t=\nu(t,S_t)$. This additional dependency of volatility makes the problem of finding the optimal strategy non-trivial; in fact if we consider the whole securities smiles then the second-order term in the HJB equation \eqref{eq:HJB_general} is not zero and thus a closed formula for $\alpha^*$ is not available anymore. Because of that, one must resort to numerical techniques to recover the problem solution. Unlike the BS model where the strategy depends only on time, in LV dynamics there is no unnecessary information provided by the state block in Equation~\eqref{eq:state} to take the optimal action. Here we consider the same market data ($r_t$, $\mu_t$ and $\nu_t$) as in the Black and Scholes environment to study how the optimal solution changes with the dynamics model. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=9cm,height=6cm]{LV_direct_policy_approach.pdf}
	\caption{Learning curves of the direct policy algorithm applied to the local volatility model. The solid blue line is the learning curve of an agent with free actions, while the green solid one is the learning curve of an agent whose actions are parameterized with the maximum drift strategy. The two agents are parameterized by a $20 \times 10 \times 5$ FFNN with tanh activation function. The loss function is computed according to \Cref{eq:direct_policy_loss}. The optimizer adopted is the RMSprop with a learning rate of $10^{-3}$ for the blue line while $10^{-4}$ for the green one. The horizontal red-dashed lines are the 99\% confidence interval of the MC price obtained by maximizing path-wise the TVS drift through \Cref{eq:analytical_solution_free_strategy}.}
	\label{fig:result_lv_direct_policy}
\end{figure}  
Our first way to tackle the problem is to exploit the results obtained in the BS case. Thus we train by direct policy algorithm a deep $20\times 10\times 5$ neural network with tanh activation function and a RMSprop optimizer. We report in Figure~\ref{fig:result_lv_direct_policy} the corresponding learning curve (solid blue line). From this, we observe that the network has learnt some good policy since the curve grows as the number of training epochs increases until it saturates at a certain value. To try to measure the performance of the policy selected by the agent, we can build a n\"aif strategy called \virgolette{baseline}. By looking at the Markovian projection in \eqref{eq:markovian_projection}-\eqref{eq:local_drift} and the first term in the HJB \eqref{eq:HJB_general}, a natural choice for the baseline is to maximize the TVS local drift. In other words this coincides by applying the path-wise the Black and Scholes solution of Equation~\eqref{eq:BS_optimal_strategy}; since here we deal with a free allocation strategy, we can simply use our analytical result \eqref{eq:analytical_solution_free_strategy}. In the same Figure~\ref{fig:result_lv_direct_policy} we report the 99\% confidence interval of the MC price obtained with the maximum drift baseline as two red-dashed lines. As we can see, the optimized loss function is compatible with the baseline price. Following the theoretical result of the HJB equation, we can assert that the agent has learnt a sub-optimal policy. Since we LV model differs from the BS one for a corrective term in the HJB \eqref{eq:HJB_general}, we expect that the optimal solution  in the LV framework will be in a close region of the maximum drift strategy. Thus we train another agent with the direct policy learning by parameterizing its action with the baseline strategy and choosing a smaller learning rate of $10^{-4}$. With this parameterization, at each observational time $T_k\in\mathcal{T}$ the risky allocation strategy is obtained by summing the network output with the Equation~\eqref{eq:analytical_solution_free_strategy}. Even in this case, the corresponding learning curve (solid green line in Figure~\ref{fig:result_lv_direct_policy}) is stuck in the maximum drift strategy. The first possible reason for this behaviour is that the learning strategy of the direct policy approach is too simple for the learning task. The second interpretation is that the volatility versor in the HJB does not affect significantly the optimum location since it is a second-order term and is lost in the MC error.

Because of that, we change the learning strategy by adopting a more sophisticated one: the PPO algorithm. The big advantage of PPO is that, in addition to the policy, a guess of the value function in Equation~\eqref{eq:HJB_general} is also computed by a parameterization through an FFNN. As for the direct policy approach, we use the BS environment to fine-tune the PPO hyper-parameters. Again we experienced that deeper FFNNs outperform shallow ones and tanh is the most preferable. We report for completeness the values of the other PPO hyper-parameters, for the description of which we refer to \cite{Schulman2017}: learning rate $3\times 10^{-4}$, $\lambda=0.95$, $\epsilon=0.2$, $c_1=0.7$, $c_2=0$ and mini-batch size of 2048 episodes.

We have trained a 5 layer FFNN with 8 neurons each with the PPO algorithm in two different environments: one environment generates the reward according to Equation~\eqref{eq:reward1}, while the other exploits the reward function \eqref{eq:reward2} where we set $\gamma=0.98$ as discount factor to make the agent prefer immediate rewards. Again, for both environments, we train two different agents: one whose action is given by Equation~\eqref{eq:action}, while the other implements the action parameterization with the baseline.
\begin{figure}[h!]
	\centering
	\includegraphics[height=6cm,width=7.2cm]{LV_PPO_reward_learning_curve.png}
	\includegraphics[height=6cm,width=7.2cm]{LV_PPO_new_reward_learning_curve.png}
	\caption{Learning curves of the PPO algorithm applied to the local volatility model. On the horizontal axis the number of training episodes. The solid lines are the moving average of the realized rewards on the last $10^5$ episodes. The shadows represent the $98\%$ confidence intervals. The label {\it reward} ({\it a}) indicates that the RL agent is trained in an environment where the reward function is defined by Equation~\eqref{eq:reward1}, while {\it reward} ({\it b}) refers to the reward function in Equation~\eqref{eq:reward2}. The solid blue lines are the learning curves of an agent with free actions, while the green solid ones are the learning curves of an agent whose actions are parameterized with the maximum drift strategy. In both plots, the agents are parameterized by an FFNN with 5 hidden layers with 5 neurons each with tanh activation function. The horizontal red-dashed lines delimit the 99\% confidence interval of the MC price obtained by maximizing path-wise the TVS drift through Equation~\eqref{eq:analytical_solution_free_strategy}.}\label{fig:PPO_results}
\end{figure}
We report in Figure~\ref{fig:PPO_results} the results of the learning curves. In all the PPO cases, the trained agents seem to be stuck in the sub-optimal maximum drift strategy policy. For the case of the immediate reward function (Figure~\ref{fig:PPO_results} on the right), we have that the learning curves converge to a saturation value that is not compatible with the baseline price; this is due to the fact that with $\lambda,\gamma \neq 1$ we introduce a bias in the problem. Because of that, we need to run a test MC simulation with freezed agents and $\lambda=\gamma=1$ to obtain a result to compare with the baseline price. We report the results in Table~\ref{tab:mc_results_lv}. The Reinforcement Learning agents we have trained with PPO give all prices compatible with the baseline one.
\begin{table}
\begin{center}
\begin{tabular}{ll}
\hline\noalign{\smallskip}
Pricing method & TVO price [EUR] \\
\noalign{\smallskip}
\hline
\noalign{\smallskip}
Baseline &  $(4.138 \pm 0.005)\times 10^{-2}$\\
RL free (a) & $(4.127 \pm 0.005) \times 10^{-2}$ \\
RL parameterized (a) & $(4.131 \pm 0.005) \times 10^{-2}$ \\
RL free (b) & $(4.130 \pm 0.005) \times 10^{-2}$ \\
RL parameterized (b) & $(4.135 \pm 0.005) \times 10^{-2}$ \\
\hline
\end{tabular}
\caption{Comparison of TVO prices under local volatility model. The baseline price is obtained by applying path-wise the maximum drift strategy \eqref{eq:analytical_solution_free_strategy}. The RL free agent implements the policy defined by Equation~\eqref{eq:action}, while the parameterized agent chooses its actions in terms of the baseline strategy. We label with letter (a) the agents trained in environments with the reward function \eqref{eq:reward1}, while with (b) the reward function \eqref{eq:reward2}. The option parameters are: $I_0=K=1\,[\text{EUR}]$, $T=2\,[\text{yr}]$ and $\bar{\sigma}=5\%$.}
\label{tab:mc_results_lv}
\end{center}
\end{table}
\setlength{\tabcolsep}{1.4pt}

\section{Conclusion and Further Developments}
In this paper, we described a non-trivial control problem related to derivative contracts on target volatility strategies. In particular, we have considered a bank selling a call option to a fund manager as protection on the capital invested on the TVS. We showed how the presence of different funding costs coming from hedging the risky assets underlying the TVS, obliges the bank to solve a stochastic optimal control problem to price the protection. This is due to the fact that the bank strategy is not self-financing. This kind of control problem is hard to solve because here the control process affects both drift and diffusive coefficients of the controlled process. Despite its complexity, our first contribution is the derivation of a closed form solution of the control problem in a Black and Scholes framework, which could represent a useful tool for practitioners since it outperforms intuitive trading strategies. We have derived this solution in two different ways: first by applying the Gy\"ongy  Lemma and then by writing the Hamilton-Jacobi-Bellman equation.  We numerically studied the problem in the more general local volatility model where the solution is not available and thus a numerical investigation is needed. We tackled the problem by means of the novel Reinforcement Learning techniques, by both the direct policy learning and the proximal policy optimization one. We used the BS model, where the solution is \textit{a priori} known, as benchmark to perform a series of fine-tuning of the RL algorithm hyper-parameters, such as the artificial neural network architecture. We have tested in the LV model the two RL approaches and from our simulations, we have evidence that nor the simple direct policy learning strategy nor the sophisticated PPO are able to outperform our analytical solution applied path-wise. Thus our analytical result for the Black and Scholes model seems to be a good proxy solution also for the local volatility one.

 This result seems to be a local optimum from the HJB equation of the problem, since in the LV model the volatility versor term should influence the RL agent actions. Thus natural development of this work could be to solve the HJB numerically in low dimension in order to check why such sophisticated algorithms are not able to find the global optimum of the problem, or to understand which are the key elements of the problem, such as market data or the payoff function, that can give rise to a solution far from the intuitive one.

\printbibliography
\end{document}

### 2022-03-08 local_volatility/localvol.py

@@ -1,147 +1,0 @@
-import numpy as np
-import lets_be_rational.LetsBeRational as lbr
-from numpy import log, exp,sqrt
-from scipy.interpolate import interp1d,RegularGridInterpolator
-from pricing import ACT_365, DiscountingCurve, EquityForwardCurve, Curve
-
-class LocalVolatilityCurve(Curve):
-    
-    def __init__(self, market_volatility=None, strikes=None, maturities=None):
-        self.volatilities = market_volatility
-        self.K = strikes
-        self.T = np.append(0.,maturities[:-1])     #it is fundamental this transformation for the piecewise interpolation
-        self.vola_interpolated = interp1d(self.K,self.volatilities,axis=0,fill_value="extrapolate")   #linear interpolation along strike
-    
-    def curve(self,price):
-        return self.vola_interpolated(np.array(price))
-    
-    def value_at_time(self,time,price):
-        if time in self.T:
-            return self(price).T[np.searchsorted(self.T, time, side='left')]
-        else:
-            return self(price).T[np.searchsorted(self.T, time, side='left')-1]
-    
-def A_i_matrix(alpha_i, beta_i, gamma_i):
-    b = np.diag(beta_i)
-    a = np.diag(alpha_i[1:],k=-1)
-    c = np.diag(gamma_i[:-1],k=1)
-    A = a +b +c
-    A[0] = 0
-    A[-1] = 0
-    return A
-
-
-def alpha_i_vector(eta_i_plus_square, dh):
-    """Calculates the alpha_i coefficients of the tri-diagonal matrix"""
-    return (eta_i_plus_square) * 0.5 * (-(0.5/dh)-1/(dh**2))
-
-def beta_i_vector(eta_i_plus_square, dh):
-    """Calculates the beta_i coefficients of the tri-diagonal matrix"""
-    return (eta_i_plus_square) * (1/(dh**2))
-
-
-def gamma_i_vector(eta_i_plus_square, dh):
-    """Calculates the gamma_i coefficients of the tri-diagonal matrix"""
-    return (eta_i_plus_square) * 0.5 * ((0.5/dh)-1/(dh**2))
-
[…]

### 2022-03-08 local_volatility/pricing.py

@@ -1,231 +1,0 @@
-import numpy as np
-from scipy.interpolate import interp1d
-from numpy import exp, sqrt, log
-from scipy.integrate import quad
-from numpy.linalg import cholesky
-
-"""Classes for my simulation"""
-
-class Curve:
-
-    def __init__(self, **kwargs):
-
-        raise Exception('do not instantiate this class.')
-
-    def __call__(self, date): #return the value of the curve at a defined time
-
-        return self.curve(date)
-
-class EquityForwardCurve(Curve):
-
-    def __init__(self, spot=None, reference=None, discounting_curve=None,
-                repo_rates=None, repo_dates=None, act = "No"):
-        self.spot = spot
-        self.reference = reference
-        self.discounting_curve = discounting_curve
-        if act =="360":
-            self.T = ACT_360(repo_dates,self.reference)
-            self.T = self.T*360/365
-        elif act == "365":
-            self.T = ACT_365(repo_dates,self.reference)
-        else:
-            self.T = abs(repo_dates - self.reference)
-        self.q_values = np.array([repo_rates[0]])
-        for i in range(1,len(self.T)):
-            alpha = ((self.T[i])*(repo_rates[i])-(self.T[i-1])*
-                     (repo_rates[i-1]))/(self.T[i]-self.T[i-1])
-            self.q_values = np.append(self.q_values,alpha)
-        if self.T[0] !=0:
-            self.T = np.insert(self.T[:-1],0,0)
-        self.q = interp1d(self.T, self.q_values, kind='previous',fill_value="extrapolate", assume_sorted=False) 
-        print("Forward repo time grid",self.T)
-        print("Forward repo rate: ", self.q_values)
-
-    def curve(self, date):
-        date = np.array(date)
-        if date.shape!=():
-            return  np.asarray([(self.spot/self.discounting_curve(extreme))*exp(-quad_piecewise(self.q,self.T,0,extreme)) for extreme in date])
-        else:
-            return (self.spot/self.discounting_curve(date))*exp(-quad_piecewise(self.q,self.T,0,date))
-
-
-class DiscountingCurve(Curve):
-
-    def __init__(self, reference=None, discounts=None, dates=None, act = "No"):
-        self.reference = reference
-        if act=="360":
-            self.T = ACT_360(dates,self.reference)
-        elif act == "365":
-            self.T = ACT_365(dates,self.reference)
[…]

### 2022-03-08 pricing/Tests/Asian_option_monoasset_Test.py

@@ -1,0 +1,330 @@
+# ---
+# jupyter:
+#   jupytext:
+#     formats: ipynb,py
+#     text_representation:
+#       extension: .py
+#       format_name: light
+#       format_version: '1.5'
+#       jupytext_version: 1.13.7
+#   kernelspec:
+#     display_name: Python 3 (ipykernel)
+#     language: python
+#     name: python3
+# ---
+
+# # <span style="color:orange"> Monoasset test: Pricing Geometric Average Asian Option </span>
+
+# +
+import sys
+sys.path.insert(1, '../')
+
+
+import numpy as np
+import matplotlib.pyplot as plt
+from pricing import DiscountingCurve, EquityForwardCurve, Black, Vanilla_PayOff,ForwardVariance
+from montecarlo import MC_results, MC_Data_Blocking
+import time, sys
+from matplotlib import ticker
+from scipy.stats.mstats import gmean
+from closedforms import Price_to_BS_ImpliedVolatility, GA_Asian_option_closed_form, BS_European_option_closed_form, volatility_asian, interest_rate_asian
+from numpy import log, exp
+# -
+
+# ### Parameters of Simulation
+
+t = 0
+spot_price = 150.
+N_simulations = 5e5
+N_block = 100
+N_averages = 30
+T_max = 10
+maturity = 4.
+dates = np.linspace(0.,maturity,N_averages)
+r = 1./100
+volatility = 50./100
+asian_vola = volatility_asian(N_averages,volatility)
+sampling = "antithetic"
+
+# ### Market Data
+#
+
+# +
+"""Discounut Factors"""
+zero_interest_rate = np.array([r,r,r])
+zero_interest_rate_dates = np.array([0.0,5,T_max])
+d = exp(-zero_interest_rate*zero_interest_rate_dates)        #market discount factors 
+D = DiscountingCurve(reference=t, discounts=d, discount_dates=zero_interest_rate_dates)   #discounting curve
+
+plt.figure(figsize=(15,4))
+
+plt.subplot(1,2,1)
+plt.step(zero_interest_rate_dates, D.R(zero_interest_rate_dates), color="red")
+plt.xlabel("time")
+plt.ylabel("zero interest rate")
+
+plt.subplot(1,2,2)
[…]

### 2022-03-08 pricing/Tests/European_option_monoasset_test.py

@@ -1,0 +1,321 @@
+# ---
+# jupyter:
+#   jupytext:
+#     formats: ipynb,py
+#     text_representation:
+#       extension: .py
+#       format_name: light
+#       format_version: '1.5'
+#       jupytext_version: 1.13.7
+#   kernelspec:
+#     display_name: Python 3 (ipykernel)
+#     language: python
+#     name: python3
+# ---
+
+# # <span style="color:blue"> Monoasset test: Pricing European Option </span>
+
+# +
+import sys
+sys.path.insert(1, '../')
+
+import numpy as np
+import matplotlib.pyplot as plt
+from pricing import DiscountingCurve, EquityForwardCurve, Black,ForwardVariance,Vanilla_PayOff
+from montecarlo import MC_results, MC_Data_Blocking
+from closedforms import BS_European_option_closed_form, Price_to_BS_ImpliedVolatility
+import time
+from numpy import log, exp
+from matplotlib import ticker
+# -
+
+# ### Parameters of Simulation
+
+r = 1/100
+t = 0
+volatility = 20/100
+spot_price = 150.
+T_max = 10.
+N_simulations = 5e5
+N_block = 100
+reference_date = 0
+sampling = "antithetic"
+
+# ### Market Data
+
+# +
+"""Discounut Factors"""
+zero_interest_rate = np.array([r,r,r])
+zero_interest_rate_dates = np.array([0.0,5,T_max])
+d = exp(-zero_interest_rate*zero_interest_rate_dates)        #market discount factors 
+D = DiscountingCurve(reference=t, discounts=d, discount_dates=zero_interest_rate_dates)   #discounting curve
+
+plt.figure(figsize=(15,4))
+
+plt.subplot(1,2,1)
+plt.step(zero_interest_rate_dates, D.R(zero_interest_rate_dates), color="red")
+plt.xlabel("time")
+plt.ylabel("zero interest rate")
+
+plt.subplot(1,2,2)
+x = np.linspace(0.,6,100)
+plt.plot(x,D(x))
+plt.xlabel("time")
+plt.ylabel("Discount curve")
+plt.show()
+
+# +
+"""Forward curve and repo rates"""
[…]

### 2022-03-08 pricing/Tests/Interpolation_monoasset_Test.py

@@ -1,0 +1,104 @@
+# ---
+# jupyter:
+#   jupytext:
+#     formats: ipynb,py
+#     text_representation:
+#       extension: .py
+#       format_name: light
+#       format_version: '1.5'
+#       jupytext_version: 1.13.7
+#   kernelspec:
+#     display_name: Python 3 (ipykernel)
+#     language: python
+#     name: python3
+# ---
+
+# # <span style="color:green"> Monoasset test: Plain Vanilla Pricing with Time-Dependent Volatility </span>
+
+# +
+import sys
+sys.path.insert(1, '../')
+
+from pricing import EquityForwardCurve, DiscountingCurve, Black, ForwardVariance, Vanilla_PayOff, ACT_365
+import numpy as np
+import matplotlib.pyplot as plt
+from montecarlo import MC_results
+from closedforms import Price_to_BS_ImpliedVolatility
+from matplotlib import ticker 
+from scipy.interpolate import interp1d
+from read_market import LoadFromTxt
+
+# +
+idx= 0
+
+folder = "../../market_data/Basket10Assets/"
+asset_names_list = "DJ 50 TR", "S&P 500 NET EUR","MSCI EM MKT EUR","I NKY NTR EUR","DAX 30 E", "FTSE100 NTR E","SMI TR EUR","FTSEMIBN", "CAC 40 NTR", "HSI NTR EUR"
+D, F, V, correlation_matrix = LoadFromTxt(asset_names_list,folder, "ATM_FWD", local_vol_model=False)
+F = F[idx]
+V = V[idx]
+assert V.asset_name == F.asset_name
+# -
+
+# ### Parameters of Simulation
+
+t = 0
+N_simulations = 6e5
+T_max = 10
+r = 1./100
+sampling = "antithetic"
+call_put = 1
+
+# ### Simulation
+
+maturities = np.loadtxt(folder + "maturities_vola_data_" + F.asset_name + ".txt")
+BS = Black(maturities, forward_curve=F,variance_curve=V, sampling=sampling)
+S_t = BS.simulate(n_sim=N_simulations, seed=14)
+
+X_t = S_t/F(maturities)
+strike = 1.0
+pay = Vanilla_PayOff(X_t, strike, call_put, sampling=sampling)
+result,err = MC_results(pay)
+E = np.mean(X_t,axis=0)
[…]

### 2022-03-08 pricing/Tests/Local_volatility_european_pricing_Test.py

@@ -1,0 +1,121 @@
+# ---
+# jupyter:
+#   jupytext:
+#     formats: ipynb,py
+#     text_representation:
+#       extension: .py
+#       format_name: light
+#       format_version: '1.5'
+#       jupytext_version: 1.13.7
+#   kernelspec:
+#     display_name: Python 3 (ipykernel)
+#     language: python
+#     name: python3
+# ---
+
+# # <span style="color:purple"> Local Volatility Pricing Test </span>
+
+# +
+import sys
+sys.path.insert(1, '../')
+
+import numpy as np
+from time import time
+import matplotlib.pyplot as plt
+from read_market import LoadFromTxt
+from pricing import LocalVolatilityCurve
+from montecarlo import MC_results, MC_Data_Blocking
+from closedforms import Price_to_BS_ImpliedVolatility
+from pricing import LocalVolatilityCurve, EquityForwardCurve, ForwardVariance, DiscountingCurve,Vanilla_PayOff,LV_model
+
+# +
+idx= 0
+
+
+folder = "../../market_data/Basket10Assets/"
+asset_names_list = "DJ 50 TR", "S&P 500 NET EUR","MSCI EM MKT EUR","I NKY NTR EUR","FTSE100 NTR E","SMI TR EUR","DAX 30 E","FTSEMIBN","HSI NTR EUR"
+D, F, V, LV, correlation_matrix = LoadFromTxt(asset_names_list,folder, "ATM_FWD", local_vol_model=True)
+F = F[idx]
+V = V[idx]
+LV = LV[idx]
+assert V.asset_name == F.asset_name
+assert LV.asset_name == F.asset_name
+
+# +
+spot_volatilities = V.market_volatility.T
+vola_strikes = V.K
+maturities = np.loadtxt(folder + "maturities_vola_data_" + F.asset_name + ".txt")
+money_ness = np.zeros((len(vola_strikes),len(maturities)))
+for j in range(len(vola_strikes)):
+    for i in range(len(maturities)):
+        money_ness[j,i] = vola_strikes[j]/F(maturities[i])
+
+IV_curve = LocalVolatilityCurve(spot_volatilities,money_ness,maturities,V.asset_name, "piecewise")
+# -
+
+# %%time
+N_grid = 160
+sampling = "antithetic"
+N_simulation = 1e5
+expiries = LV.T
[…]

### 2022-03-08 pricing/Tests/Multi_asset_Test.py

@@ -1,0 +1,349 @@
+# ---
+# jupyter:
+#   jupytext:
+#     formats: ipynb,py
+#     text_representation:
+#       extension: .py
+#       format_name: light
+#       format_version: '1.5'
+#       jupytext_version: 1.13.7
+#   kernelspec:
+#     display_name: Python 3 (ipykernel)
+#     language: python
+#     name: python3
+# ---
+
+# # <span style="color:red"> Multiasset test: Pricing Geometric Average Basket Option </span>
+
+# +
+import sys
+sys.path.insert(1, '../')
+
+import numpy as np
+from pricing import EquityForwardCurve, DiscountingCurve, ForwardVariance, Black, Vanilla_PayOff
+from montecarlo import MC_results, MC_Data_Blocking
+from numpy import exp, log
+import matplotlib.pyplot as plt
+from closedforms import Price_to_BS_ImpliedVolatility, GAM_Basket_option_closed_form, volatility_basket, forward_basket
+from scipy.stats.mstats import gmean
+import time
+from matplotlib import ticker
+# -
+
+# ### Parameters of the Simulation
+
+r = 1./100
+t = 0
+spot_price = np.array([110.0,120.0,97.0,133.0])
+T_max = 10
+N_simulation = 1e6
+N_block = 100
+volatility = np.array([20.,30.,25.,32.])/100
+corr = np.array(([1,0.15,0.10,0.20],[0.15,1.,-0.05,0.18],[0.1,-0.05,1.,0.13],[0.20,0.18,0.13,1]))
+sampling = "antithetic"
+N_equity = len(corr)
+sigma_basket = volatility_basket(volatility,corr)
+
+# ### Market Data
+
+# +
+"""Discounut Factors"""
+zero_interest_rate = np.array([r,r,r])
+zero_interest_rate_dates = np.array([0.0,5,T_max])
+d = exp(-zero_interest_rate*zero_interest_rate_dates)        #market discount factors 
+D = DiscountingCurve(reference=t, discounts=d, discount_dates=zero_interest_rate_dates)   #discounting curve
+plt.figure(figsize=(15,4))
+
+plt.subplot(1,2,1)
[…]

### 2022-03-08 pricing/Tests/Optimal_target_volatility_strategy_BS_model_Test.py

@@ -1,0 +1,511 @@
+# ---
+# jupyter:
+#   jupytext:
+#     formats: ipynb,py
+#     text_representation:
+#       extension: .py
+#       format_name: light
+#       format_version: '1.5'
+#       jupytext_version: 1.13.7
+#   kernelspec:
+#     display_name: Python 3 (ipykernel)
+#     language: python
+#     name: python3
+# ---
+
+# # <span style="color:brown"> Markowitz solution: Evaluation of the most constervative call option price under Black and Scholes model</span>
+
+# +
+import sys
+sys.path.insert(1, '../')
+
+import numpy as np
+import matplotlib.pyplot as plt
+from pricing import EquityForwardCurve, ForwardVariance, DiscountingCurve,Vanilla_PayOff
+from closedforms import Price_to_BS_ImpliedVolatility
+import time
+from montecarlo import MC_results, MC_Data_Blocking
+from matplotlib import ticker
+from read_market import LoadFromTxt
+from targetvol import time_grid_union, Drift, CholeskyTDependent, Strategy,TVSForwardCurve ,TargetVolatilityStrategy
+# -
+
+# ## Parsing market data
+
+idx= 0
+folder = "../../market_data/Basket10Assets"
+asset_names_list = "DJ 50 TR", "S&P 500 NET EUR","MSCI EM MKT EUR","I NKY NTR EUR","DAX 30 E", "FTSE100 NTR E","SMI TR EUR","FTSEMIBN", "CAC 40 NTR", "HSI NTR EUR"
+D, F, V, correlation_matrix = LoadFromTxt(asset_names_list,folder, strike_interpolation_rule="ATM_FWD", local_vol_model=False)
+assert len(F) == len(V) == len(correlation_matrix) == len(correlation_matrix.T)
+names = []
+spot_prices = []
+N_equity = len(F)
+for i in range(len(V)):
+    assert V[i].asset_name == F[i].asset_name
+    names.append(F[i].asset_name)
+    spot_prices.append(F[i].spot)
+
+# #### Discounting curve
+#
+
+# +
+plt.figure(figsize=(15,4))
+
+plt.subplot(1,2,1)
+x = np.linspace(0,60,1000)
+plt.step(x, D.R(x), color="red")
[…]

### 2022-03-08 pricing/closedforms.py

@@ -1,87 +1,156 @@
 import numpy as np
-import scipy.stats as si  #for gaussian cdf
+import scipy.stats as si  # for gaussian cdf
 from numpy import exp, log, sqrt
 
+""" Closed-form pricing for Black-Scholes-Merton model """
 
-"""Functions for Black & Scholes Formula"""
-def d1(forward = None, strike= None, maturity=None, reference=None, volatility=None):
-    """d_1 function for BS model"""
-    return (log(forward / strike) + 0.5 * (volatility ** 2) * (maturity-reference)) / (volatility * sqrt(maturity-reference))
-
-def d2(forward = None, strike= None, maturity=None, reference=None, volatility=None):
-    """d_1 function for BS model"""
-    return (log(forward / strike) - 0.5 * (volatility ** 2) * (maturity-reference)) / (volatility * sqrt(maturity-reference))
-
-def European_option_closed_form(forward = None, strike= None, maturity=None, reference=None, zero_interest_rate = None, volatility=None, typo = 1):
-    """Closed form of Call Option for BS model"""
-    d_1 = d1(forward, strike, maturity, reference, volatility)
-    d_2 = d2(forward, strike, maturity, reference, volatility)
-    discount = exp(-zero_interest_rate*(maturity-reference))
-    if typo ==1:
-        """Call option"""
-        return discount * (forward * si.norm.cdf(d_1, 0.0, 1.0) - strike * si.norm.cdf(d_2, 0.0, 1.0))
-    elif typo==-1:
-        """Put option"""
-        return discount * (strike * si.norm.cdf(-d_2, 0.0, 1.0) - forward * si.norm.cdf(-d_1, 0.0, 1.0))
-
-def Delta(forward = None, strike= None, maturity=None, reference=0, volatility=None):
-    return si.norm.cdf(d1(forward, strike, maturity, reference, volatility))
-
-def StrikeFromDelta(forward=None, maturity=None, reference=None, delta=None, volatility=None):
-    return forward(maturity)*np.exp(0.5*(volatility**2)*(maturity-reference)-volatility*np.sqrt(maturity-reference)*si.norm.ppf(delta))
-
-def Vega(forward=None, strike=None, maturity=None, reference=None, dividends=None, volatility=None):
-    d_1 = d1(forward, strike, maturity, reference, volatility)
-    return S * np.exp(-dividends * (maturity-reference)) * np.sqrt(maturity-reference) * si.norm.pdf(d_1)
-
[…]

### 2022-03-08 pricing/montecarlo.py

@@ -1,95 +1,125 @@
 import numpy as np
 
 
-def MC(data):
-    """mean and error of MC simulation"""
-    mean = np.mean(data,axis=0)
-    std = np.std(data,axis=0)/np.sqrt(len(data))
-    return mean, std    
+def MC_results(mc_paths: np.ndarray) -> tuple:
+    """
+    Returns mean and uncertainty of a Monte Carlo simulation.
+    :param mc_paths (np.ndarray[float]): Monte Carlo simulation. It can be a vector or a matrix.
+    The mean and uncertainty are returned for each column.
+    :return: mean, uncertainty.
+    """
+    mean = np.mean(mc_paths, axis=0)
+    std = np.std(mc_paths, axis=0) / np.sqrt(len(mc_paths))
+    return mean, std
 
 
-def MC_Analisys(data,N_block):
-    """Data blocking algorithm"""
-    if data.ndim == 1:
-        return MC_Analisys_vector(data,N_block)
-    elif data.ndim == 2:
-        return MC_Analisys_matrix(data,N_block)
-
-    
-def MC_Analisys_matrix(data,N_block): #method to check convergence of the MC result
-    """Data blocking method for a matrix"""
-    N = N_block
-    M= int(len(data))              # Total number of throws
-    L=int(M/N)           # Number of throws in each block, please use for M a multiple of N
-    K = int(len(data.T))
-    ave = np.zeros((N,K))
-    av2 = np.zeros((N,K))
-    sum_prog = np.zeros((N,K))
-    su2_prog = np.zeros((N,K))
-    err_prog = np.zeros((N,K))
-    x = np.arange(N)
-    x*=L
-    for l in range(K):
-        for i in range(N):
-            sum = 0
-            for j in range(L):
-                k = j+i*L
-                sum += data[k][l]
-            ave[i][l] = sum/L
-            av2[i][l] = (ave[i][l])**2
-    for l in range (K):
-        for i in range(N):
-            for j in range(i+1):
-                sum_prog[i][l] += ave[j][l] # SUM_{j=0,i} r_j
-                su2_prog[i][l] += av2[j][l] # SUM_{j=0,i} (r_j)^2
-            sum_prog[i][l]/=(i+1) # Cumulative average
-            su2_prog[i][l]/=(i+1) # Cumulative square average
-            err_prog[i][l] = error(sum_prog,su2_prog,i,l) # Statistical uncertainty
[…]

### 2022-03-08 pricing/pricing.py

@@ -1,374 +1,570 @@
-import numpy as np
 from scipy.interpolate import interp1d, PchipInterpolator
 from numpy import exp, sqrt, log
-from scipy.integrate import quad
-import matplotlib.pyplot as plt
+from pricing_utility import *
 
 """Classes for my simulation"""
+
 
 class Curve:
 
     def __init__(self, **kwargs):
-
         raise Exception('do not instantiate this class.')
 
-    def __call__(self, date): #return the value of the curve at a defined time
-
+    def __call__(self, date):  # return the value of the curve at a defined time
         return self.curve(date)
-
-class EquityForwardCurve(Curve):
-
-    def __init__(self, spot=None, reference=None, discounting_curve=None,
-                repo_rates=None, repo_dates=None, act = "No"):
-        self.spot = spot
-        self.reference = reference
-        self.discounting_curve = discounting_curve
-        if act =="360":
-            self.T = ACT_360(repo_dates,self.reference)
-            self.T = self.T*360/365
-        elif act == "365":
-            self.T = ACT_365(repo_dates,self.reference)
-        else:
-            self.T = abs(repo_dates - self.reference)
-        if self.T[0] == 0.:
-            self.T = np.delete(self.T,0)
-            repo_rates = np.delete(repo_rates,0)
-        self.q_values = np.array([repo_rates[0]])
-        for i in range(1,len(self.T)):
-            alpha = ((self.T[i])*(repo_rates[i])-(self.T[i-1])*
-                     (repo_rates[i-1]))/(self.T[i]-self.T[i-1])
-            self.q_values = np.append(self.q_values,alpha)
-        self.T = np.append(0.,self.T[:-1])
-        self.q = interp1d(self.T, self.q_values, kind='previous',fill_value="extrapolate", assume_sorted=False) 
-     #   print("Forward repo time grid",self.T)
-      #  print("Forward repo rate: ", self.q_values)
-
-    def curve(self, date):
-        date = np.array(date)
-        if date.shape!=():
-            return  np.asarray([(self.spot/self.discounting_curve(extreme))*exp(-quad_piecewise(self.q,self.T,0,extreme)) for extreme in date])
-        else:
-            return (self.spot/self.discounting_curve(date))*exp(-quad_piecewise(self.q,self.T,0,date))
 
 
 class DiscountingCurve(Curve):
+    """Discounting curve"""
 
-    def __init__(self, reference=None, discounts=None, dates=None, act = "No"):
[…]

### 2022-03-08 pricing/pricing_utility.py

@@ -1,0 +1,104 @@
+import numpy as np
+
+
+def quad_piecewise(piece_wise_function, time_grid_of_piece_wise_function: np.ndarray, lower_int: float,
+                   upper_int: float or np.ndarray, vectorized: bool = False) -> float or np.ndarray:
+    """
+    Function that integrates a piecewise constant function (right open).
+    :param piece_wise_function: function to be integrated
+    :param time_grid_of_piece_wise_function: time grid where the piecewise function is defined.
+    :param lower_int (float): lower integration bound.
+    :param upper_int (float): upper integration bound.
+    :param vectorized (bool, default=False): boolean that indicates if the function is vectorial (True) or scalar (False).
+    :return: the integral of the piecewise function.
+    """
+    t_in = float(lower_int)
+    t_fin = float(upper_int)
+    time_grid = np.float64(time_grid_of_piece_wise_function)
+    if t_in == t_fin:
+        return 0
+    if t_fin in time_grid:
+        time_grid = time_grid[np.where(time_grid <= t_fin)[0]]
+    if t_in in time_grid:
+        time_grid = time_grid[np.where(time_grid >= t_in)[0]]
+    if t_in not in time_grid:
+        time_grid = time_grid[np.where(time_grid > t_in)[0]]
+        time_grid = np.insert(time_grid, 0, t_in)
+    if t_fin not in time_grid:
+        time_grid = time_grid[np.where(time_grid < t_fin)[0]]
+        time_grid = np.insert(time_grid, len(time_grid), t_fin)
+    if vectorized:
+        y = np.array([])
+        for i in range(1, len(time_grid)):
+            y = np.append(y, piece_wise_function(time_grid[i]))
+    else:
+        y = piece_wise_function(time_grid[:-1])
+
+    dt = np.diff(time_grid)
+    return np.sum(y * dt)
+
+
+def get_euler_grid(fixings: np.ndarray, n_intervals: int) -> tuple:
+    """
+    Function that takes a vector of fixings (expressed in yrf) and returns a tuple of two vectors:
+    - the first vector contains the Euler grid (expressed in yrf) of the fixings (among each fixings
+    the Euler grid is the vector of the fixings divided by n_intervals)
+    tuple of the grid of the Euler method and the number of intervals
+    :param fixings (np.ndarray[float]): vector of fixings (expressed in yrf)
+    :param n_intervals (int): number of intervals between the fixings.
+    :return (tuple): tuple of the grid of the Euler method (np.ndarray with shape: len(fixings)*n_intervals)
[…]

### 2022-03-08 pricing/print_xml_market_to_txt.py

@@ -1,0 +1,103 @@
+from pyGem import GemItem
+import numpy as np
+import pandas as pd
+import warnings
+import os
+
+"""
+This script is used to convert xml files with FE market data conventions to txt files used by this library.
+The user must provide the xml file (FileName), the folder of the xml file, a list of the assets names that 
+the user wants to export (asset_names), the discounting curve nave (discount_curve) and the output folder 
+where the txt files will be saved (output_folder).
+"""
+
+############################################################ SCRIPT INPUTS #########################################################################################
+
+FileName = 'calibration_output'
+input_folder = "../../market_data"
+asset_names = "DJ 50 TR", "S&P 500 NET EUR", "MSCI EM MKT EUR", "I NKY NTR EUR", "FTSE100 NTR E", "SMI TR EUR", "DAX 30 E", "FTSEMIBN", "CAC 40 NTR", "HSI NTR EUR"
+discount_curve = "EUR"
+saving_folder = "../../market_data/Basket10Assets"
+
+######################################################################################################################################################################
+
+
+if input_folder[-1] != '/':
+    input_folder += '/'
+if saving_folder[-1] != '/':
+    saving_folder += '/'
+if not os.path.exists(saving_folder):
+    os.makedirs(saving_folder)
+
+# Read GemItem
+market = GemItem()
+market.load("", "file:///" + str(input_folder) + FileName + ".xml")
+
+# Read discounting curve
+reference = market.get(f"market.{discount_curve}.zero_curve.{discount_curve} :STD.reference_date")
+dates = np.array(market.get(f"market.{discount_curve}.zero_curve.{discount_curve} :STD.data.dates"))
+discounts = np.array(market.get(f"market.{discount_curve}.zero_curve.{discount_curve} :STD.data.discounts"))
+day_count_convention = market.get(f"market.{discount_curve}.zero_curve.{discount_curve} :STD.convention.day_count")
+
+if day_count_convention == "ACT365":
+    discount_dates = (dates - reference) / 365.
+elif day_count_convention == "ACT360":
+    discount_dates = (dates - reference) / 360.
+else:
+    raise ValueError(f" {day_count_convention} day count convention not supported")
+
+np.savetxt(saving_folder + "discount_data.txt", (discount_dates, discounts))
+
+# Read Correlation Matrix
+correlation_matrix = np.array(market.get("market.data.correlation_matrix.data.correlations"))
[…]

### 2022-03-08 pricing/read_market.py

@@ -1,135 +1,72 @@
-import xml.etree.ElementTree as ET
-from pricing import EquityForwardCurve, DiscountingCurve, ForwardVariance,LocalVolatilityCurve
-from numpy import array, delete, zeros, reshape, append, max
+from pricing import DiscountingCurve, EquityForwardCurve, ForwardVariance, LocalVolatilityCurve
 import numpy as np
-
-class MarketDataReader:
-
-    def __init__(self,filename = None):
-        tree = ET.parse(filename)
-        self.root = tree.getroot()
-
-    def get_stock_number(self):
-        return len(self.root[3][0][1][1][1][0])
-
-    def get_stock_names(self):
-        N_stocks = self.get_stock_number()
-        names = array([])
-        for i in range(N_stocks):
-            names=append(names,self.root[3][0][1][1][1][0][i].text)
-        delete_equity = [0,11,12,13,14,15]
-        names = delete(names,delete_equity)
-        wrong_names = [names[4],names[5],names[6]]
-        names[4] = wrong_names[2]
-        names[5] = wrong_names[0]
-        names[6] = wrong_names[1]
-        return names
-    def get_correlation(self):
-        N_stocks = self.get_stock_number()
-        correlation_matrix = zeros(N_stocks**2)
-        for i in range (N_stocks**2):
-            correlation_matrix[i] = float(self.root[3][0][1][1][0][0][i].text)
-        correlation_matrix = reshape(correlation_matrix, (N_stocks,N_stocks))
-        """Delete cash bank"""
-        delete_equity = [0,11,12,13,14,15]
-        correlation_matrix = delete(correlation_matrix,delete_equity, axis = 0)
-        correlation_matrix = delete(correlation_matrix,delete_equity, axis = 1)
-        return correlation_matrix
-
-    def get_discounts(self):
-        discounts = zeros(len(self.root[3][1][0][0][0][0][0]))
-        discounts_dates = zeros(len(self.root[3][1][0][0][0][0][0]))
-        for i in range(len(self.root[3][1][0][0][0][0][0])):
-            discounts_dates[i] = float(self.root[3][1][0][0][0][0][0][i].text)
-            discounts[i] = float(self.root[3][1][0][0][0][1][0][i].text)
-        return DiscountingCurve(reference=self.get_reference_date(), discounts=discounts, dates=discounts_dates, act="365")
-
-    def get_reference_date(self):
-        return float(self.root[3][1][0][0][2][0].text)
-
-    def get_spot_prices(self):
[…]

### 2022-03-08 pricing/run.py

@@ -1,74 +1,0 @@
-from mpi4py import MPI
-import numpy as np
-from scipy.interpolate import interp1d
-from time import time
-#import utilities
-from scipy.interpolate import interp1d
-import matplotlib.pyplot as plt
-#import generator
-import xml.etree.ElementTree as ET
-from pricing import LocalVolatilityCurve
-from read_market import MarketDataReader
-from montecarlo import MC_Analisys, MC
-import lets_be_rational.LetsBeRational as lbr
-from pricing import LocalVolatilityCurve, EquityForwardCurve, ForwardVariance, DiscountingCurve,piecewise_function,Vanilla_PayOff,PricingModel,LV_model
-
-
-
-comm = MPI.COMM_WORLD
-size = comm.Get_size()
-rank = comm.Get_rank()
-for i in range(9):
-    if rank == 0:
-        t_start = MPI.Wtime()
-    reader = MarketDataReader("TV_example.xml")
-    correlation_matrix = reader.get_correlation()
-    N_equity = len(correlation_matrix)
-    names = reader.get_stock_names()
-    spot_prices = reader.get_spot_prices()
-    D = reader.get_discounts()
-    F = reader.get_forward_curves()
-    V = reader.get_volatilities()
-
-
-    tree = ET.parse('calibration_output.xml')
-    root = tree.getroot()
-    N_equity = i
-    name=root[1][2][1][0][N_equity].text
-    expiry_yrf = 4
-    n_expiries = len(root[1][2][2][N_equity][expiry_yrf][0])#.attrib
-    expiries = np.array([])
-    for i in range(n_expiries):
-        expiries = np.append(expiries,float(root[1][2][2][N_equity][expiry_yrf][0][i].text))
-    print(name)
-    moneyness = 6
-    moneyness_matrix = np.array([])
-    n_matrix =len(root[1][2][2][N_equity][moneyness][0])
-    n_strikes = int(n_matrix/n_expiries)
-    for i in range(n_matrix):
-        moneyness_matrix = np.append(moneyness_matrix,float(root[1][2][2][N_equity][moneyness][0][i].text))
-    moneyness_matrix = moneyness_matrix.reshape(n_strikes,n_expiries)
-
-    vola = 7
-    vola_matrix = np.array([])
-    for i in range(n_matrix):
-        vola_matrix = np.append(vola_matrix,float(root[1][2][2][N_equity][vola][0][i].text))
[…]

### 2022-03-08 pricing/targetvol.py

@@ -1,276 +1,512 @@
 import numpy as np
-from numpy import exp, sqrt, log
-from pricing import piecewise_function, Curve, PricingModel, quad_piecewise
+from numpy import exp, sqrt
+from pricing import Curve, ForwardVariance, EquityForwardCurve, DiscountingCurve
+from pricing_utility import quad_piecewise
 from numpy.linalg import cholesky
-from scipy.optimize import minimize
 from scipy.interpolate import interp1d
+from scipy.optimize import minimize, LinearConstraint, Bounds
 
-
-"""Notation of pallavicini and daluiso draft"""
-
+"""
+Black and Scholes model for target volatility pricing.
+The notation is based on the paper in the tex directory
+"""
 
 class Drift(Curve):
-    """Time dependent Repo rates union"""
-    def __init__(self, forward_curves = None):
-        self.T = time_grid_union(curve_array_list = forward_curves)
-        Ndim = len(forward_curves)
+    """Time dependent repo rates curve"""
+
+    def __init__(self, forward_curves: list) -> None:
+        """
+        Drift curve. This function builds the mu function presented in the paper.
+        The class returns (with the curve method) the instantaneous repo rates of the asset contained in the
+        basket at a given time. The drift curve is a piecewise constant function (right open) of time (the first element
+        of the time-grid definition must be 0.0).
+        :param forward_curves [list[EquityCurve]]: list of equity curves in the basket.
+        """
+        assert type(forward_curves) == list and all(isinstance(x, EquityForwardCurve) for x in forward_curves)
+        assert len(forward_curves) > 1
+        # time grid of the piecewise constant function defined as union of the time grids of the equity curves in the basket
+        self.T = time_grid_union(curve_array_list=forward_curves)
+        n_dim = len(forward_curves)
         self.mu = forward_curves[0].q(self.T)
-        self.mu = np.stack((self.mu,forward_curves[1].q(self.T)),axis=1)
+        self.mu = np.stack((self.mu, forward_curves[1].q(self.T)), axis=1)
         """Calculating the instant repo rates"""
-        for i in range(2,Ndim):
-            self.mu = np.insert(self.mu,len(self.mu.T),forward_curves[i].q(self.T),axis=1)
-        self.m = interp1d(self.T, self.mu, axis=0, kind='previous',fill_value="extrapolate", assume_sorted=False) 
-        print("Drift time grid:",self.T)
-        print("Drift values:", self.mu)   
-    def curve(self,date):
[…]

### 2022-03-08 reinforcement/LSMC.py

@@ -1,304 +1,0 @@
-import numpy as np 
-import statsmodels.api as sm
-from scipy.interpolate import interp1d
-import time
-import matplotlib.pyplot as plt
-
-
-### personal libraries
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, LV_model, ForwardVariance, quad_piecewise, Vanilla_PayOff
-from envs.pricing.targetvol import Drift
-from envs.pricing.loadfromtxt import LoadFromTxt
-from envs.pricing.targetvol import Markowitz_solution
-from envs.pricing.closedforms import European_option_closed_form
-from envs.pricing.montecarlo import MC_Analisys
-
-
-
-def alpha_vector(a1, nu, Nsim=1, norm=1):
-    a_1 = np.ones(Nsim)*a1
-    beta = nu[1,0,:]**2 + nu[1,1,:]**2
-    gamma = a_1*(nu[0,0,:]*nu[1,0,:]+ nu[0,1,:]*nu[1,1,:])
-    delta = (a_1**2)*(nu[0,0,:]**2+nu[0,1,:]**2) - norm
-    Delta = gamma**2-beta*delta
-    a_2 = (-gamma-np.sqrt(Delta))/beta
-    return np.array(([a_1,a_2])).T
-
-def nu_matrix(sigma_t_vector,identity_3d, corr_chole, N_equity, Nsim):
-    S = sigma_t_vector.T*identity_3d
-    n = S.T@corr_chole
-    return n.transpose(1,2,0).reshape(N_equity,N_equity,Nsim)
-
-def alpha_dot_nu(alpha_matrix,nu_matrix, N_equity, Nsim):
-    a = np.sum((alpha_matrix.T*nu_matrix.transpose(1,0,2)),axis=1).T
-    return a
-
-def evolve_TVS(index_start, I_t, dS_S, Vola, alpha, target_vola, instant_interest_rate, dt, N_euler_grid, N_equity, Nsim, identity_3d, corr_chole):
-    s = np.sum(alpha, axis=1)
-    I_t_plus_1 = np.ones(dS_S[:,0,0].shape)*I_t
-    for t in range(index_start, N_euler_grid+index_start):
-        nu = nu_matrix(Vola[:,t,:],identity_3d, corr_chole, N_equity, Nsim)
-        omega = target_vola/np.linalg.norm(alpha_dot_nu(alpha, nu, N_equity, Nsim),axis=1)
[…]

### 2022-03-08 reinforcement/LSMCRegression.py

@@ -1,254 +1,0 @@
-import numpy as np 
-import statsmodels.api as sm
-from sklearn.preprocessing import PolynomialFeatures
-from sklearn import linear_model
-from scipy.interpolate import interp1d
-import time
-import matplotlib.pyplot as plt
-import math
-
-### personal libraries
-from envs.pricing.pricing import LV_model, Vanilla_PayOff
-from envs.pricing.loadfromtxt import LoadFromTxt
-from envs.pricing.montecarlo import MC_Analisys
-
-
-def nu_matrix(sigma_t_vector,identity_3d, corr_chole, N_equity, Nsim):
-    S = sigma_t_vector.T*identity_3d
-    n = S.T@corr_chole
-    return n.transpose(1,2,0).reshape(N_equity,N_equity,Nsim)
-
-def alpha_dot_nu(alpha_matrix,nu_matrix, N_equity, Nsim):
-    a = np.sum((alpha_matrix.T*nu_matrix.transpose(1,0,2)),axis=1).T
-    return a
-
-def find_nearest(array,value):
-    idx = np.searchsorted(array, value, side="left")
-    if idx > 0 and (idx == len(array) or math.fabs(value - array[idx-1]) < math.fabs(value - array[idx])):
-        return idx-1
-    else:
-        return idx
-
-
-def evolve_TVS(index_start, I_t, dS_S, Vola, alpha, target_vola, instant_interest_rate, dt, N_euler_grid, N_equity, Nsim, identity_3d, corr_chole):
-    s = 1.0
-    I_t_p = np.ones(dS_S[:,0,0].shape)*I_t
-    for t in range(index_start, N_euler_grid+index_start):
-        nu = nu_matrix(Vola[:,t,:],identity_3d, corr_chole, N_equity, Nsim)
-        omega = target_vola/np.linalg.norm(alpha_dot_nu(alpha, nu, N_equity, Nsim),axis=1)
-        I_t_p = I_t_p * (1. + np.sum(alpha*dS_S[:,t,:],axis=1)*omega + dt * instant_interest_rate[t]*(1.-omega*s))
-    return I_t_p
-
-
-def plot_regression(states, V_realized, V_alpha, I_t, alpha, time, number_state):
-    plt.scatter(states,V_realized, label="Realized")
-    plt.scatter(states,V_alpha, c='red', label='Regression')
[…]

### 2022-03-08 reinforcement/LSMC_long.py

@@ -1,376 +1,0 @@
-import numpy as np 
-import statsmodels.api as sm
-from scipy.interpolate import interp1d
-import time
-import matplotlib.pyplot as plt
-import math
-from sklearn.preprocessing import PolynomialFeatures
-from sklearn import linear_model
-
-### personal libraries
-from envs.pricing.pricing import LV_model, Vanilla_PayOff
-from envs.pricing.loadfromtxt import LoadFromTxt
-from envs.pricing.montecarlo import MC_Analisys
-
-
-def nu_matrix(sigma_t_vector,identity_3d, corr_chole, N_equity, Nsim):
-    S = sigma_t_vector.T*identity_3d
-    n = S.T@corr_chole
-    return n.transpose(1,2,0).reshape(N_equity,N_equity,Nsim)
-
-def alpha_dot_nu(alpha_matrix,nu_matrix, N_equity, Nsim):
-    a = np.sum((alpha_matrix.T*nu_matrix.transpose(1,0,2)),axis=1).T
-    return a
-
-def find_nearest(array,value):
-    idx = np.searchsorted(array, value, side="left")
-    if idx > 0 and (idx == len(array) or math.fabs(value - array[idx-1]) < math.fabs(value - array[idx])):
-        return idx-1
-    else:
-        return idx
-
-
-def evolve_TVS(index_start, I_t, dS_S, Vola, alpha, target_vola, instant_interest_rate, dt, N_euler_grid, N_equity, Nsim, identity_3d, corr_chole):
-    s = 1.0
-    I_t_p = np.ones(dS_S[:,0,0].shape)*I_t
-    for t in range(index_start, N_euler_grid+index_start):
-        nu = nu_matrix(Vola[:,t,:],identity_3d, corr_chole, N_equity, Nsim)
-        omega = target_vola/np.linalg.norm(alpha_dot_nu(alpha, nu, N_equity, Nsim),axis=1)
-        I_t_p = I_t_p * (1. + np.sum(alpha*dS_S[:,t,:],axis=1)*omega + dt * instant_interest_rate[t]*(1.-omega*s))
-    return I_t_p
-
-
-def plot_regression(states, V_realized, V_alpha, I_t, alpha, time, number_state):
-    plt.scatter(states,V_realized, label="Realized")
-    plt.scatter(states,V_alpha, c='red', label='Regression')
[…]

### 2022-03-08 reinforcement/LSMC_long_simulated_I.py

@@ -1,341 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Tue Jul 13 14:32:18 2021
-
-@author: u453878
-"""
-
-
-import numpy as np 
-import statsmodels.api as sm
-from scipy.interpolate import interp1d
-import time
-import matplotlib.pyplot as plt
-import math
-
-### personal libraries
-from envs.pricing.pricing import LV_model, Vanilla_PayOff
-from envs.pricing.loadfromtxt import LoadFromTxt
-from envs.pricing.montecarlo import MC_Analisys
-
-
-def nu_matrix(sigma_t_vector,identity_3d, corr_chole, N_equity, Nsim):
-    S = sigma_t_vector.T*identity_3d
-    n = S.T@corr_chole
-    return n.transpose(1,2,0).reshape(N_equity,N_equity,Nsim)
-
-def alpha_dot_nu(alpha_matrix,nu_matrix, N_equity, Nsim):
-    a = np.sum((alpha_matrix.T*nu_matrix.transpose(1,0,2)),axis=1).T
-    return a
-
-def find_nearest(array,value):
-    idx = np.searchsorted(array, value, side="left")
-    if idx > 0 and (idx == len(array) or math.fabs(value - array[idx-1]) < math.fabs(value - array[idx])):
-        return idx-1
-    else:
-        return idx
-
-
-def evolve_TVS(index_start, I_t, dS_S, Vola, alpha, target_vola, instant_interest_rate, dt, N_euler_grid, N_equity, Nsim, identity_3d, corr_chole):
-    s = 1.0
-    I_t_p = np.ones(dS_S[:,0,0].shape)*I_t
-    for t in range(index_start, N_euler_grid+index_start):
-        nu = nu_matrix(Vola[:,t,:],identity_3d, corr_chole, N_equity, Nsim)
-        omega = target_vola/np.linalg.norm(alpha_dot_nu(alpha, nu, N_equity, Nsim),axis=1)
-        I_t_p = I_t_p * (1. + np.sum(alpha*dS_S[:,t,:],axis=1)*omega + dt * instant_interest_rate[t]*(1.-omega*s))
-    return I_t_p
-
-
-def plot_regression(states, V_realized, V_alpha, I_t, alpha, time, number_state):
-    plt.scatter(states,V_realized, label="Realized")
-    plt.scatter(states,V_alpha, c='red', label='Regression')
[…]

### 2022-03-08 reinforcement/Simulazione_I_t.py

@@ -1,144 +1,0 @@
-import numpy as np 
-import matplotlib.pyplot as plt
-
-### personal libraries
-from envs.pricing.pricing import LV_model, Vanilla_PayOff
-from envs.pricing.loadfromtxt import LoadFromTxt
-from envs.pricing.montecarlo import MC_Analisys
-
-
-def nu_matrix(sigma_t_vector,identity_3d, corr_chole, N_equity, Nsim):
-    S = sigma_t_vector.T*identity_3d
-    n = S.T@corr_chole
-    return n.transpose(1,2,0).reshape(N_equity,N_equity,Nsim)
-
-def alpha_dot_nu(alpha_matrix,nu_matrix, N_equity, Nsim):
-    a = np.sum((alpha_matrix.T*nu_matrix.transpose(1,0,2)),axis=1).T
-    return a
-
-
-def evolve_TVS(index_start, I_t, dS_S, Vola, alpha, target_vola, instant_interest_rate, dt, N_euler_grid, N_equity, Nsim, identity_3d, corr_chole):
-    s = 1.0
-    I_t_p = np.ones(dS_S[:,0,0].shape)*I_t
-    for t in range(index_start, N_euler_grid+index_start):
-        nu = nu_matrix(Vola[:,t,:],identity_3d, corr_chole, N_equity, Nsim)
-        omega = target_vola/np.linalg.norm(alpha_dot_nu(alpha, nu, N_equity, Nsim),axis=1)
-        I_t_p = I_t_p * (1. + np.sum(alpha*dS_S[:,t,:],axis=1)*omega + dt * instant_interest_rate[t]*(1.-omega*s))
-    return I_t_p
-
-
-def plot_regression(states, V_realized, V_alpha, I_t, alpha, time, number_state):
-    plt.scatter(states,V_realized, label="Realized")
-    plt.scatter(states,V_alpha, c='red', label='Regression')
-    plt.title(r"$I_t$ = "+str(round(I_t,4))+" --- alpha = "+str(round(alpha,4))+" --- time = "+str(round(time,3)))
-    plt.xlabel("State "+str(number_state))
-    plt.ylabel("$V_t$")
-    plt.legend()
-    plt.show()
-    
-def plot_regression3d(state1, state2,V_realized, V_alpha, I_t, alpha, time):
-    fig = plt.figure()
-    ax = fig.add_subplot(projection='3d')
[…]

### 2022-03-08 reinforcement/baseline_simulation/job_script.sh

@@ -1,21 +1,0 @@
-#!/bin/bash
-
-#SBATCH --account=IscrC_REAL-COP
-#SBATCH --partition=m100_usr_prod                #m100_all_serial
-#SBATCH --qos=normal
-#SBATCH -J mpi4py_test
-#SBATCH -o mpi4py.out
-#SBATCH -e mpi4py.err
-#SBATCH -N 1
-#SBATCH -n 64
-#SBATCH --time=02:00:00
-#SBATCH --mail-type=ALL
-#SBATCH --mail-user=stefano.polo@studenti.unimi.it
-
-module load anaconda/2020.02
-module load autoload lapack/3.9.0--gnu--8.4.0
-module load openblas/0.3.9--gnu--8.4.0
-source activate reinforcement_env
-
-
-srun -n $SLURM_NTASKS --mpi=pmi2 python run.py

### 2022-03-08 reinforcement/baseline_simulation/pricing/fake_market_complex.py

@@ -1,65 +1,0 @@
-import numpy as np
-from numpy import log, exp, sqrt
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, Black, ForwardVariance
-
-def load_fake_market(N_equity, r, maturity):
-    reference =0
-    if N_equity ==3:
-        correlation = np.array(([1,0.5,0.3],[0.5,1,0.3],[0.3,0.3,1]))       #correlation matrix
-        spot_prices = np.array([100,200,89])
-        T_repo3 = np.array([2/12.,5/12.,maturity])
-        repo_rate3 = (-1)*np.array([0.03,0.022,0.04])/100
-    else:
-        correlation = np.array(([1,0.5],[0.5,1]))       #correlation matrix
-        spot_prices = np.array([100,200])
-
-    r_t = np.array([r,r,r,r])
-    T_discounts = np.array([0.,3/12,4/12.,maturity])      #data observation of the market discounts factor
-    market_discounts = exp(-r_t*T_discounts)       #market discounts factor
-    repo_rate1 = (-1)*np.array([0.03,0.03,0.03])/100
-    T_repo1 = np.array([1/12,4./12,maturity])       #data observation of the market repo rates for equity 1
-    repo_rate2 = (-1)*np.array([0.04,0.03,0.035])/100  #market repo rates for equity 1 0.52
-    T_repo2 = np.array([1/12.,4/12.,maturity])
-
-    sigma1 = np.array([20,24.,23.])/100
-    T_sigma1 = np.array([2/12,5./12,maturity])
-    K1 = np.array([spot_prices[0],500])
-    spot_vola1 = np.array((sigma1,sigma1))                                      #market implied volatility matrix
-    sigma2 = np.array([22,21,20])/100
-    T_sigma2 =  np.array([2/12.,6/12,maturity])
-    K2 = np.array([spot_prices[1],600])
-    spot_vola2 = np.array((sigma2,sigma2))
-    if N_equity==3:
-        sigma3 = np.array([14,17,18])/100
-        T_sigma3 =  np.array([2/12.,6/12,maturity])
-        K3 = np.array([spot_prices[2],600])
-        spot_vola3 = np.array((sigma3,sigma3))
-
-    D = DiscountingCurve(reference=reference, discounts=market_discounts,dates=T_discounts)
-    F = []
-    V = []
-    q = repo_rate1
-    T_q = T_repo1
[…]

### 2022-03-08 reinforcement/baseline_simulation/pricing/loadfromtxt.py

@@ -1,23 +1,0 @@
-from pricing.pricing import DiscountingCurve, EquityForwardCurve, ForwardVariance, LocalVolatilityCurve
-import numpy as np
-
-def LoadFromTxt(asset_name, folder = None):
-    dates_discounts, discounts = np.loadtxt(folder+"/discount_data.txt")
-    D = DiscountingCurve(reference=0.,discounts=discounts,dates=dates_discounts)
-    F, V, LV = [], [], []
-    for name in asset_name:
-        spot = np.loadtxt(folder+"/spot_data"+name+".txt")[0]
-        repo_dates, repo_rates = np.loadtxt(folder+"/repo_data_"+str(name)+".txt")
-        F.append(EquityForwardCurve(reference=0., discounting_curve = D, spot=spot, repo_rates=repo_rates,repo_dates=repo_dates,name=name,act="No"))
-        
-        spot_vola = np.loadtxt(folder+"/vola_data_"+str(name)+".txt")
-        vola_strikes = np.loadtxt(folder+"/strikes_vola_data_"+str(name)+".txt")
-        vola_dates = np.loadtxt(folder+"/maturities_vola_data_"+name+".txt")
-        V.append(ForwardVariance(reference=0.,maturities=vola_dates,strikes=vola_strikes,spot_volatility=spot_vola,strike_interp=spot,act="No", name=name))
-        
-        lv_pars = np.loadtxt(folder+"/LV_param_data_"+str(name)+".txt")
-        lv_money = np.loadtxt(folder+"/LV_money_vola_data_"+str(name)+".txt")
-        lv_dates = np.loadtxt(folder+"/LV_maturities_vola_data_"+name+".txt")
-        LV.append(LocalVolatilityCurve(lv_pars,lv_money,lv_dates,name))
-
-    return D, F, V, LV

### 2022-03-08 reinforcement/baseline_simulation/pricing/n_sphere.py

@@ -1,27 +1,0 @@
-import numpy as np
-
-def n_sphere_to_cartesian(radius, angles):
-    """Function that takes n-spherical coordinates and convert into n-cartesian coordinates
-       angles: the n-2 values between [0,\pi) and last one between [0,2\pi)
-    """
-    a = np.concatenate((np.array([2*np.pi]), angles))
-    si = np.sin(a)
-    si[0] = 1
-    si = np.cumprod(si)
-    co = np.cos(a)
-    co = np.roll(co, -1)
-    return si*co*radius
-
-
-def sign_renormalization(vec, sum_plus, sum_minus):
-    """renormalize positive and negative elements of an array separately if necessary such 
-    that sum of positive is equal to sum_plus and the abs sum of negative is equal to sum_neg"""
-    pos = vec>=0
-    neg = np.invert(pos)
-    p_s = np.sum(vec[pos])
-    n_s = abs(np.sum(vec[neg]))
-    if p_s > sum_plus:
-        vec[pos] = (vec[pos]/p_s)*sum_plus
-    if n_s > sum_minus:
-        vec[neg] = (vec[neg]/n_s)*sum_minus
-    return vec

### 2022-03-08 reinforcement/baseline_simulation/pricing/pricing.py

@@ -1,383 +1,0 @@
-import numpy as np
-from scipy.interpolate import interp1d, PchipInterpolator
-from numpy import exp, sqrt, log
-from scipy.integrate import quad
-"""Classes for my simulation"""
-
-class Curve:
-
-    def __init__(self, **kwargs):
-
-        raise Exception('do not instantiate this class.')
-
-    def __call__(self, date): #return the value of the curve at a defined time
-
-        return self.curve(date)
-
-class EquityForwardCurve(Curve):
-
-    def __init__(self, spot=None, reference=None, discounting_curve=None,
-                repo_rates=None, repo_dates=None, name = None, act = "No"):
-        self.spot = spot
-        self.asset_name = name
-        self.reference = reference
-        self.discounting_curve = discounting_curve
-        if act =="ACT360":
-            self.T = ACT_360(repo_dates,self.reference)
-            self.T = self.T*360/365
-        elif act == "ACT365":
-            self.T = ACT_365(repo_dates,self.reference)
-        else:
-            self.T = abs(repo_dates - self.reference)
-        if self.T[0] == 0.:
-            self.T = np.delete(self.T,0)
-            repo_rates = np.delete(repo_rates,0)
-        self.q_values = np.array([repo_rates[0]])
-        for i in range(1,len(self.T)):
-            alpha = ((self.T[i])*(repo_rates[i])-(self.T[i-1])*
-                     (repo_rates[i-1]))/(self.T[i]-self.T[i-1])
-            self.q_values = np.append(self.q_values,alpha)
-        if self.T[0] != 0:
-            self.T = np.append(0.,self.T[:-1])
-        self.q = interp1d(self.T, self.q_values, kind='previous',fill_value="extrapolate", assume_sorted=False) 
-     #   print("Forward repo time grid",self.T)
-      #  print("Forward repo rate: ", self.q_values)
-
-    def curve(self, date):
-        date = np.array(date)
-        if date.shape!=():
-            return  np.asarray([(self.spot/self.discounting_curve(extreme))*exp(-quad_piecewise(self.q,self.T,0,extreme)) for extreme in date])
-        else:
-            return (self.spot/self.discounting_curve(date))*exp(-quad_piecewise(self.q,self.T,0,date))
-
-
-class DiscountingCurve(Curve):
-
-    def __init__(self, reference=None, discounts=None, dates=None, act = "No"):
-        self.reference = reference
-        if act=="ACT360":
-            self.T = ACT_360(dates,self.reference)
[…]

### 2022-03-08 reinforcement/baseline_simulation/pricing/read_market.py

@@ -1,135 +1,0 @@
-import xml.etree.ElementTree as ET
-from pricing.pricing import EquityForwardCurve, DiscountingCurve, ForwardVariance,LocalVolatilityCurve
-from numpy import array, delete, zeros, reshape, append, max
-import numpy as np
-
-class MarketDataReader:
-
-    def __init__(self,filename = None):
-        tree = ET.parse(filename)
-        self.root = tree.getroot()
-
-    def get_stock_number(self):
-        return len(self.root[3][0][1][1][1][0])
-
-    def get_stock_names(self):
-        N_stocks = self.get_stock_number()
-        names = array([])
-        for i in range(N_stocks):
-            names=append(names,self.root[3][0][1][1][1][0][i].text)
-        delete_equity = [0,11,12,13,14,15]
-        names = delete(names,delete_equity)
-        wrong_names = [names[4],names[5],names[6]]
-        names[4] = wrong_names[2]
-        names[5] = wrong_names[0]
-        names[6] = wrong_names[1]
-        return names
-    def get_correlation(self):
-        N_stocks = self.get_stock_number()
-        correlation_matrix = zeros(N_stocks**2)
-        for i in range (N_stocks**2):
-            correlation_matrix[i] = float(self.root[3][0][1][1][0][0][i].text)
-        correlation_matrix = reshape(correlation_matrix, (N_stocks,N_stocks))
-        """Delete cash bank"""
-        delete_equity = [0,11,12,13,14,15]
-        correlation_matrix = delete(correlation_matrix,delete_equity, axis = 0)
-        correlation_matrix = delete(correlation_matrix,delete_equity, axis = 1)
-        return correlation_matrix
-
-    def get_discounts(self):
-        discounts = zeros(len(self.root[3][1][0][0][0][0][0]))
-        discounts_dates = zeros(len(self.root[3][1][0][0][0][0][0]))
-        for i in range(len(self.root[3][1][0][0][0][0][0])):
-            discounts_dates[i] = float(self.root[3][1][0][0][0][0][0][i].text)
-            discounts[i] = float(self.root[3][1][0][0][0][1][0][i].text)
-        return DiscountingCurve(reference=self.get_reference_date(), discounts=discounts, dates=discounts_dates, act="365")
-
-    def get_reference_date(self):
-        return float(self.root[3][1][0][0][2][0].text)
-
-    def get_spot_prices(self):
-        index_equity = [1,2,3,4,5,6,7,8,9,10]
[…]

### 2022-03-08 reinforcement/baseline_simulation/pricing/targetvol.py

@@ -1,302 +1,0 @@
-import numpy as np
-from numpy import exp, sqrt, log
-from pricing.pricing import piecewise_function, Curve, PricingModel, quad_piecewise
-from numpy.linalg import cholesky
-from scipy.optimize import minimize, LinearConstraint, Bounds
-from scipy.interpolate import interp1d
-
-
-"""Notation of pallavicini and daluiso draft"""
-
-
-class Drift(Curve):
-    """Time dependent Repo rates union"""
-    def __init__(self, forward_curves = None):
-        self.T = time_grid_union(curve_array_list = forward_curves)
-        Ndim = len(forward_curves)
-        self.mu = forward_curves[0].q(self.T)
-        self.mu = np.stack((self.mu,forward_curves[1].q(self.T)),axis=1)
-        """Calculating the instant repo rates"""
-        for i in range(2,Ndim):
-            self.mu = np.insert(self.mu,len(self.mu.T),forward_curves[i].q(self.T),axis=1)
-        self.m = interp1d(self.T, self.mu, axis=0, kind='previous',fill_value="extrapolate", assume_sorted=False) 
-       # print("Drift time grid:",self.T)
-        #print("Drift values:", self.mu)   
-    def curve(self,date):
-        return self.m(date)
-
-
-class CholeskyTDependent(Curve):
-    """Time dependent cholesky variance-covariance matrix union"""
-    def __init__(self, variance_curves = None, correlation_chole = None):
-        self.T = time_grid_union(curve_array_list = variance_curves)
-        N_times = len(self.T)
-        Ndim = len(variance_curves)
-        self.nu = np.zeros((Ndim,Ndim,N_times))
-        I = np.identity(Ndim)
-        for i in range(N_times):
-            vol = np.zeros(Ndim)
-            for j in range(Ndim):
-                vol[j] = sqrt(variance_curves[j](self.T[i]))
-            vol = I*vol
-            self.nu[:,:,i] = vol@correlation_chole
-        self.n = interp1d(self.T, self.nu, axis=2, kind='previous',fill_value="extrapolate", assume_sorted=False) 
-      #  print("Cholesky covariance-variance time grid:",self.T)
-     #   print("Cholesky covariance-variance matrix values:", self.nu)
-   
-    def curve(self,date):
-        return self.n(date)
-
-        
-class Strategy(Curve):
[…]

### 2022-03-08 reinforcement/baseline_simulation/run.py

@@ -1,150 +1,0 @@
-from pricing.pricing import LV_model
-from pricing.loadfromtxt import LoadFromTxt
-from pricing.targetvol import optimization_only_long, CholeskyTDependent, Strategy,Drift,  Markowitz_solution
-import numpy as np
-import time
-try:
-    from mpi4py import MPI
-except ImportError:
-    MPI = None
-
-if MPI is None or MPI.COMM_WORLD.Get_rank()==0:
-    rank = 0
-else:
-    rank = MPI.COMM_WORLD.Get_rank()
-
-t0 = time.clock()
-Nsim = int(1e5)
-run = 0
-Seed = run+rank#3000
-strategy = "free"   #free
-#print("Seme ",Seed)
-I_0 = 1.
-restart =0 
-Black = 0
-daily = 0
-N_equity = 2                                #number of equities
-target_vol = 5./100.
-T = 2.
-title = 'new/d'+str(int(T))+'/final_price_lvinvestigation_2_asset_equalrepo_long_Nsim_'+str(Nsim)+'_vola'+str(target_vol)+'_maturity'+str(T)+'_Black'+str(Black)+'_daily'+str(daily)+strategy
-restart_file = 'day_new/d'+str(int(T-1))+'/final_price_lvinvestigation_2_asset_equalrepo_long_Nsim_8000_vola0.05_maturity'+str(T-1)+'_Black'+str(Black)+'_daily'+str(daily)+strategy
-
-if restart:
-    I_0 = np.loadtxt(restart_file+"_rank"+str(rank+run)+'.txt')
-    Nsim =len(I_0)
-else:
-    I_0 = np.ones(Nsim)*I_0
-"""Time grid creation for the simulation"""
-if daily:
-    days = 365
-    n_days =days * int(T)
-    observation_grid = np.linspace(1./days,T,n_days)
-    observation_grid = np.insert(observation_grid,0,0.)
-    time_index = 0
-    current_time = 0.
-    N_euler_grid = 2
-
-else:
-    month_dates = np.array([31.,28.,31.,30.,31.,30.,31.,31.,30.,31.,30.,31.])
-    months = month_dates
-    N_euler_grid = 60
-    if T > 1.:
-        for i in range(int(T)-1):
-            months = np.append(months, month_dates)
-    observation_grid = np.cumsum(months)/365.
-    observation_grid = np.insert(observation_grid,0,0.)
- 
-simulation_index = 0
[…]

### 2022-03-08 reinforcement/envs/call_option.py

@@ -1,70 +1,0 @@
-import gym
-from gym import spaces
-from gym.utils import seeding
-from numpy import array, exp, linspace, float32
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, ForwardVariance, Black
-from envs.pricing.closedforms import European_option_closed_form
-
-class PlainVanillaOption(gym.Env):
-    """Plain Vanilla Option environment"""
-    def __init__(self, spot = 100, drift = 1/100, volatility = 20/100, strike = 100, maturity = 1.):
-        self.sigma = volatility
-        self.r = drift
-        self.T = maturity
-        self.s0 = spot
-        self.strike_opt = strike
-        self.asset_history = []
-        T_max = 10
-        zero_interest_rate = array([self.r,self.r,self.r])
-        zero_interest_rate_dates = array([0.0,5,T_max])
-        d = exp(-zero_interest_rate*zero_interest_rate_dates)
-        D = DiscountingCurve(reference = 0, discounts= d, dates=zero_interest_rate_dates)
-        F = EquityForwardCurve(reference=0, discounting_curve=D, spot=self.s0,repo_dates=array([0.,T_max]), repo_rates=array([0.,0.]))
-        K_spot_vola = array([self.s0,200])
-        spot_vol = array(([self.sigma,self.sigma],[0.3,0.3]))
-        spot_vol_dates = array([0.1,T_max])
-        V = ForwardVariance(reference=0,maturities=spot_vol_dates,strikes=K_spot_vola,spot_volatility=spot_vol,strike_interp=self.s0)
-        self.BS = Black(variance=V, forward_curve=F)
-        # possible actions are exercise(1) or not (0)
-        self.action_space = spaces.Discrete(2)
-        self.observation_space = spaces.Box(low=float32(array([0.,0.])),high=float32(array([self.s0*10000.,self.T+1./365.])))
-        self.seed()
-        self.reset()
-
-    def step(self, action):
-        assert self.action_space.contains(action)
-
-        if self.current_time < self.T:
-            done = False
-            self.current_asset = self.BS.simulate(fixings=array([self.T]), random_gen = self.np_random)[0,0]
-            self.current_time = self.T
-            self.asset_history.append(self.current_asset)
-            reward = 0
-            return array((self.current_asset,self.current_time)), reward, done, {}
-
-        if self.current_time == self.T:
-            done = True
-            if action:
-                reward = (self.current_asset - self.strike_opt)*exp(-self.r*self.T)
-            else:
[…]

### 2022-03-08 reinforcement/envs/fe_envs.py

@@ -5,42 +5,23 @@ from gym.envs.registration import register
 # FE agents
 # ----------------------------------------
 register(
-    id='CompoundOption-v0',                    #name of the enviroment
-    entry_point='envs.compound_option:CompoundOption',         #location of the file
+    id='VanillaOption-v0',                                      # name of the environment
+    entry_point='envs.plain_vanilla_env:PlainVanillaOption',    #location of the file
 )
 
 register(
-    id='VanillaOption-v0',
-    entry_point='envs.call_option:PlainVanillaOption',
-)
-
-register(
-    id='TVS_simple-v0',
-    entry_point='envs.tvs_simple_env:TVS_simple',
-)
-
-register(
-    id='TVS-v0',
-    entry_point='envs.tvs_env:TVS_environment',
-)
-
-register(
-    id='TVS_2assets-v0',
-    entry_point='envs.tvs_2market_env:TVS_environment2',
-)
-
-register(
-    id='TVS_3assets-v0',
-    entry_point='envs.tvs_3market_env:TVS_envirnoment3',
+    id="TVS_BS-v0",
+    entry_point='envs.tvs_bs_env:TVS_BS_ENV',
 )
 
 register(
     id='TVS_LV-v0',
-    entry_point='envs.tvs_lv_env:TVS_LV',
+    entry_point='envs.tvs_lv_env:TVS_LV_ENV_reward1',
 
 )
+
 register(
-    id='TVS_LV_newreward-v0',
-    entry_point='envs.tvs_new_reward:TVS_LV_newreward',
+    id='TVS_LV-v2',
+    entry_point='envs.tvs_lv_env_new_reward:TVS_LV_ENV_reward2',
 
 )

### 2022-03-08 reinforcement/envs/plain_vanilla.py

@@ -1,70 +1,0 @@
-import gym
-from gym import spaces
-from gym.utils import seeding
-from numpy import array, exp, linspace
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, ForwardVariance, Black
-from envs.pricing.closedforms import European_option_closed_form
-
-class PlainVanillaOption(gym.Env):
-    """Plain Vanilla Option environment"""
-    def __init__(self, spot = 100, drift = 1/100, volatility = 20/100, strike = 100, maturity = 1.):
-        self.sigma = volatility
-        self.r = drift
-        self.T = maturity
-        self.s0 = spot
-        self.strike_opt = strike
-        self.asset_history = []
-        T_max = 10
-        zero_interest_rate = array([self.r,self.r,self.r])
-        zero_interest_rate_dates = array([0.0,5,T_max])
-        d = exp(-zero_interest_rate*zero_interest_rate_dates)
-        D = DiscountingCurve(reference = 0, discounts= d, dates=zero_interest_rate_dates)
-        F = EquityForwardCurve(reference=0, discounting_curve=D, spot=self.s0,repo_dates=array([0.,T_max]), repo_rates=array([0.,0.]))
-        K_spot_vola = array([self.s0,200])
-        spot_vol = array(([self.sigma,self.sigma],[0.3,0.3]))
-        spot_vol_dates = array([0.1,T_max])
-        V = ForwardVariance(reference=0,maturities=spot_vol_dates,strikes=K_spot_vola,spot_volatility=spot_vol,strike_interp=self.s0)
-        self.BS = Black(variance=V, forward_curve=F)
-        # possible actions are exercise(1) or not (0)
-        self.action_space = spaces.Discrete(2)
-        self.observation_space = spaces.Box(low=array([0.,0.]),high=array([self.s0*10000.,self.T+1./365.]))
-        self.seed()
-        self.reset()
-
-    def step(self, action):
-        assert self.action_space.contains(action)
-
-        if self.current_time < self.T:
-            done = False
-            self.current_asset = self.BS.simulate(fixings=array([self.T]), random_gen = self.np_random)[0,0]
-            self.current_time = self.T
-            self.asset_history.append(self.current_asset)
-            reward = 0
-            return array((self.current_asset,self.current_time)), reward, done, {}
-
-        if self.current_time == self.T:
-            done = True
-            if action:
-                reward = self.current_asset - self.strike_opt
-            else:
-                reward = 0
[…]

### 2022-03-08 reinforcement/envs/plain_vanilla_env.py

@@ -1,0 +1,97 @@
+import gym
+from gym import spaces
+from gym.utils import seeding
+from numpy import array, exp
+
+
+import sys
+sys.path.insert(1, '../pricing')
+from pricing import EquityForwardCurve, DiscountingCurve, ForwardVariance, Black
+from closedforms import BS_European_option_closed_form
+
+class PlainVanillaOption(gym.Env):
+    """Plain Vanilla Option environment under single-asset Black and Scholes model"""
+    def __init__(self, spot_price: float = 100.0, zero_interest_rate: float = 1/100, volatility: float = 20/100,
+                 option_strike: float = 100.0, option_maturity: float = 1., call_put: int = 1,
+                 n_sim_for_cache: int = int(1e4)) -> None:
+        """
+        Plain Vanilla Option environment under single-asset Black and Scholes model. The agent actions are dicrete: 0 not
+        exercise, 1 exercise the right on the option.
+        :param spot_price (float): initial spot price of the underlying asset
+        :param zero_interest_rate (float): zero interest rate
+        :param volatility (float): volatility of the underlying asset
+        :param option_strike (float): strike price of the option
+        :param option_maturity (float): maturity of the option
+        :param call_put (int): 1 for call, -1 for put
+        :param n_sim_for_cache (int, default=int(2e3)): number of simulated paths cached in memory
+        (the cache speed-up the computation because it exploits the vectorization provided by numpy)
+        """
+        self.T = option_maturity
+        self.strike_opt = option_strike
+        self.asset_history = []
+        self.cp = call_put
+        self.Nsim = n_sim_for_cache
+        self.simulation_index = 0
+        t_max = 10.
+        zero_interest_rate = array([zero_interest_rate, zero_interest_rate, zero_interest_rate])
+        zero_interest_rate_dates = array([0.0, 5, t_max])
+        d = exp(-zero_interest_rate*zero_interest_rate_dates)
+        self.discounting_curve = DiscountingCurve(reference=0, discounts=d, discount_dates=zero_interest_rate_dates)
+        self.forward_curve = EquityForwardCurve(reference=0, spot=spot_price, discounting_curve=self.discounting_curve,
+                                                repo_dates=array([0., t_max]), repo_rates=array([0., 0.]))
+        strike_spot_vola = array([spot_price, 200])
+        spot_vol = array(([volatility, volatility], [volatility, volatility]))
+        spot_vol_dates = array([0.1, t_max])
[…]

### 2022-03-08 reinforcement/envs/pricing/closedforms.py

@@ -1,91 +1,0 @@
-import numpy as np
-import scipy.stats as si  #for gaussian cdf
-from numpy import exp, log, sqrt
-
-
-"""Functions for Black & Scholes Formula"""
-def d1(forward = None, strike= None, time_to_maturity=None, reference=None, volatility=None):
-    """d_1 function for BS model"""
-    return (log(forward / strike) + 0.5 * (volatility ** 2) * time_to_maturity) / (volatility * sqrt(time_to_maturity))
-
-def d2(forward = None, strike= None, maturity=None, reference=None, volatility=None):
-    """d_1 function for BS model"""
-    return (log(forward / strike) - 0.5 * (volatility ** 2) * (maturity-reference)) / (volatility * sqrt(maturity-reference))
-
-def European_option_closed_form(forward = None, strike= None, time_to_maturity=None, discount = None, volatility=None, typo = 1):
-    """Closed form of Call Option for BS model"""
-    if time_to_maturity>1e-14:
-        variance = (volatility ** 2) * time_to_maturity
-        std = sqrt(time_to_maturity)*volatility
-        d_1 = (log(forward / strike) + 0.5*variance)/std
-        d_2 = d_1 - std
-        if typo ==1:
-            """Call option"""
-            return discount * (forward * si.norm.cdf(d_1, 0.0, 1.0) - strike * si.norm.cdf(d_2, 0.0, 1.0))
-        elif typo==-1:
-            """Put option"""
-            return discount * (strike * si.norm.cdf(-d_2, 0.0, 1.0) - forward * si.norm.cdf(-d_1, 0.0, 1.0))
-    else:
-        return np.maximum(typo*(forward - strike),0.)
-
-def Delta(forward = None, strike= None, maturity=None, reference=0, volatility=None):
-    return si.norm.cdf(d1(forward, strike, maturity, reference, volatility))
-
-def StrikeFromDelta(forward=None, maturity=None, reference=None, delta=None, volatility=None):
-    return forward(maturity)*np.exp(0.5*(volatility**2)*(maturity-reference)-volatility*np.sqrt(maturity-reference)*si.norm.ppf(delta))
-
-def Vega(forward=None, strike=None, maturity=None, reference=None, dividends=None, volatility=None):
-    d_1 = d1(forward, strike, maturity, reference, volatility)
-    return S * np.exp(-dividends * (maturity-reference)) * np.sqrt(maturity-reference) * si.norm.pdf(d_1)
-
[…]

### 2022-03-08 reinforcement/envs/pricing/fake_market.py

@@ -1,65 +1,0 @@
-import numpy as np
-from numpy import log, exp, sqrt
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, Black, ForwardVariance
-
-def load_fake_market(N_equity, r, maturity):
-    reference =0
-    if N_equity ==3:
-        correlation = np.array(([1,0.5,0.3],[0.5,1,0.3],[0.3,0.3,1]))       #correlation matrix
-        spot_prices = np.array([100,200,89])
-        T_repo3 = np.array([2/12.,5/12.,maturity])
-        repo_rate3 = np.array([0.02,0.02,0.02])/10
-    else:
-        correlation = np.array(([1,0.5],[0.5,1]))       #correlation matrix
-        spot_prices = np.array([100,200])
-
-    r_t = np.array([r,r,r,r])
-    T_discounts = np.array([0.,3/12,4/12.,maturity])      #data observation of the market discounts factor
-    market_discounts = exp(-r_t*T_discounts)       #market discounts factor
-    repo_rate1 = np.array([0.22,0.22,0.22])/10
-    T_repo1 = np.array([1/12,4./12,maturity])       #data observation of the market repo rates for equity 1
-    repo_rate2 = np.array([0.72,0.42,0.02])/10  #market repo rates for equity 1 0.52
-    T_repo2 = np.array([1/12.,4/12.,maturity])
-
-    sigma1 = np.array([20,20.,20.])/100
-    T_sigma1 = np.array([2/12,5./12,maturity])
-    K1 = np.array([spot_prices[0],500])
-    spot_vola1 = np.array((sigma1,sigma1))                                      #market implied volatility matrix
-    sigma2 = np.array([20,20,20])/100
-    T_sigma2 =  np.array([2/12.,6/12,maturity])
-    K2 = np.array([spot_prices[1],600])
-    spot_vola2 = np.array((sigma2,sigma2))
-    if N_equity==3:
-        sigma3 = np.array([10,10,10])/100
-        T_sigma3 =  np.array([2/12.,6/12,maturity])
-        K3 = np.array([spot_prices[2],600])
-        spot_vola3 = np.array((sigma3,sigma3))
-
-    D = DiscountingCurve(reference=reference, discounts=market_discounts,dates=T_discounts)
-    F = []
-    V = []
-    q = repo_rate1
-    T_q = T_repo1
-    s_vola = spot_vola1
-    T_vola = T_sigma1
[…]

### 2022-03-08 reinforcement/envs/pricing/fake_market_lv.py

@@ -1,73 +1,0 @@
-import numpy as np
-from numpy import log, exp, sqrt
-import xml.etree.ElementTree as ET
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, Black, ForwardVariance, ACT_365, LocalVolatilityCurve
-
-def load_fake_market_lv():
-    correlation = np.array(([1,0.5],[0.5,1]))       #correlation matrix
-    index_equity = [5]
-
-    tree = ET.parse('../TV_example.xml')
-    root = tree.getroot()
-    N_stocks = len(root[3][0][1][1][1][0])
-    names = np.array([])
-    for i in range(N_stocks):
-        names=np.append(names,root[3][0][1][1][1][0][i].text)
-    delete_equity = [0,11,12,13,14,15]
-    names = np.delete(names,delete_equity)
-    reference_date = float(root[3][1][0][0][2][0].text)
-
-    spot_prices = np.zeros(len(index_equity))
-    j = 0
-    for i in index_equity:
-        if i>=6:
-            spot_prices[j] = float(root[3][i+3][0][1][0][0].text)
-        else:
-            spot_prices[j] = float(root[3][3+i][0][0][1][0].text)
-        j = j+1
-        
-    discounts = np.zeros(len(root[3][1][0][0][0][0][0]))
-    discounts_dates = np.zeros(len(root[3][1][0][0][0][0][0]))
-    for i in range(len(root[3][1][0][0][0][0][0])):
-        discounts_dates[i] = float(root[3][1][0][0][0][0][0][i].text)
-        discounts[i] = float(root[3][1][0][0][0][1][0][i].text)
-    D = DiscountingCurve(reference=reference_date, discounts=discounts, dates=discounts_dates, act="365")
-
-    F = []
-    max_dates = np.array([10])
-    index = 0
-    for i in index_equity:
-        if i>=6:
-            repo_dates = np.array([reference_date+1, max(max_dates)])
-            repo_rates = np.zeros(2)
-            F.append(EquityForwardCurve(reference=reference_date, discounting_curve=D, repo_dates=repo_dates,repo_rates=repo_rates, spot=spot_prices[index],act="360"))
-        else:
-            repo_dates = np.zeros(len(root[3][3+i][0][0][0][0][0]))
-            repo_rates = np.zeros(len(root[3][3+i][0][0][0][1][0]))
[…]

### 2022-03-08 reinforcement/envs/pricing/loadfromtxt.py

@@ -1,23 +1,0 @@
-from envs.pricing.pricing import DiscountingCurve, EquityForwardCurve, ForwardVariance, LocalVolatilityCurve
-import numpy as np
-
-def LoadFromTxt(asset_name, folder = None):
-    dates_discounts, discounts = np.loadtxt(folder+"/discount_data.txt")
-    D = DiscountingCurve(reference=0.,discounts=discounts,dates=dates_discounts)
-    F, V, LV = [], [], []
-    for name in asset_name:
-        spot = np.loadtxt(folder+"/spot_data"+name+".txt")[0]
-        repo_dates, repo_rates = np.loadtxt(folder+"/repo_data_"+str(name)+".txt")
-        F.append(EquityForwardCurve(reference=0., discounting_curve = D, spot=spot, repo_rates=repo_rates,repo_dates=repo_dates,name=name,act="No"))
-        
-        spot_vola = np.loadtxt(folder+"/vola_data_"+str(name)+".txt")
-        vola_strikes = np.loadtxt(folder+"/strikes_vola_data_"+str(name)+".txt")
-        vola_dates = np.loadtxt(folder+"/maturities_vola_data_"+name+".txt")
-        V.append(ForwardVariance(reference=0.,maturities=vola_dates,strikes=vola_strikes,spot_volatility=spot_vola,strike_interp=spot,act="No", name=name))
-        
-        lv_pars = np.loadtxt(folder+"/LV_param_data_"+str(name)+".txt")
-        lv_money = np.loadtxt(folder+"/LV_money_vola_data_"+str(name)+".txt")
-        lv_dates = np.loadtxt(folder+"/LV_maturities_vola_data_"+name+".txt")
-        LV.append(LocalVolatilityCurve(lv_pars,lv_money,lv_dates,name))
-
-    return D, F, V, LV

### 2022-03-08 reinforcement/envs/pricing/montecarlo.py

@@ -1,95 +1,0 @@
-import numpy as np
-
-
-def MC(data):
-    """mean and error of MC simulation"""
-    mean = np.mean(data,axis=0)
-    std = np.std(data,axis=0)/np.sqrt(len(data))
-    return mean, std    
-
-
-def MC_Analisys(data,N_block):
-    """Data blocking algorithm"""
-    if data.ndim == 1:
-        return MC_Analisys_vector(data,N_block)
-    elif data.ndim == 2:
-        return MC_Analisys_matrix(data,N_block)
-
-    
-def MC_Analisys_matrix(data,N_block): #method to check convergence of the MC result
-    """Data blocking method for a matrix"""
-    N = N_block
-    M= int(len(data))              # Total number of throws
-    L=int(M/N)           # Number of throws in each block, please use for M a multiple of N
-    K = int(len(data.T))
-    ave = np.zeros((N,K))
-    av2 = np.zeros((N,K))
-    sum_prog = np.zeros((N,K))
-    su2_prog = np.zeros((N,K))
-    err_prog = np.zeros((N,K))
-    x = np.arange(N)
-    x*=L
-    for l in range(K):
-        for i in range(N):
-            sum = 0
-            for j in range(L):
-                k = j+i*L
-                sum += data[k][l]
-            ave[i][l] = sum/L
-            av2[i][l] = (ave[i][l])**2
-    for l in range (K):
-        for i in range(N):
-            for j in range(i+1):
-                sum_prog[i][l] += ave[j][l] # SUM_{j=0,i} r_j
-                su2_prog[i][l] += av2[j][l] # SUM_{j=0,i} (r_j)^2
-            sum_prog[i][l]/=(i+1) # Cumulative average
-            su2_prog[i][l]/=(i+1) # Cumulative square average
-            err_prog[i][l] = error(sum_prog,su2_prog,i,l) # Statistical uncertainty
-
-
-    return x,sum_prog,err_prog
-
-
-
-def MC_Analisys_vector(data,N_block): #method to check convergence of the MC result
-    """Data blocking method vector"""
-    M=int(len(data))              # Total number of throws
-    N=N_block                 # Number of blocks
-    L=int(M/N)            # Number of throws in each block, please use for M a multiple of N
-    x = np.arange(N)      # [0,1,2,...,N-1]
-    ave = np.zeros(N)
[…]

### 2022-03-08 reinforcement/envs/pricing/pricing.py

@@ -1,383 +1,0 @@
-import numpy as np
-from scipy.interpolate import interp1d, PchipInterpolator
-from numpy import exp, sqrt, log
-from scipy.integrate import quad
-"""Classes for my simulation"""
-
-class Curve:
-
-    def __init__(self, **kwargs):
-
-        raise Exception('do not instantiate this class.')
-
-    def __call__(self, date): #return the value of the curve at a defined time
-
-        return self.curve(date)
-
-class EquityForwardCurve(Curve):
-
-    def __init__(self, spot=None, reference=None, discounting_curve=None,
-                repo_rates=None, repo_dates=None, name = None, act = "No"):
-        self.spot = spot
-        self.asset_name = name
-        self.reference = reference
-        self.discounting_curve = discounting_curve
-        if act =="ACT360":
-            self.T = ACT_360(repo_dates,self.reference)
-            self.T = self.T*360/365
-        elif act == "ACT365":
-            self.T = ACT_365(repo_dates,self.reference)
-        else:
-            self.T = abs(repo_dates - self.reference)
-        if self.T[0] == 0.:
-            self.T = np.delete(self.T,0)
-            repo_rates = np.delete(repo_rates,0)
-        self.q_values = np.array([repo_rates[0]])
-        for i in range(1,len(self.T)):
-            alpha = ((self.T[i])*(repo_rates[i])-(self.T[i-1])*
-                     (repo_rates[i-1]))/(self.T[i]-self.T[i-1])
-            self.q_values = np.append(self.q_values,alpha)
-        if self.T[0] != 0:
-            self.T = np.append(0.,self.T[:-1])
-        self.q = interp1d(self.T, self.q_values, kind='previous',fill_value="extrapolate", assume_sorted=False) 
-     #   print("Forward repo time grid",self.T)
-      #  print("Forward repo rate: ", self.q_values)
-
-    def curve(self, date):
-        date = np.array(date)
-        if date.shape!=():
-            return  np.asarray([(self.spot/self.discounting_curve(extreme))*exp(-quad_piecewise(self.q,self.T,0,extreme)) for extreme in date])
-        else:
-            return (self.spot/self.discounting_curve(date))*exp(-quad_piecewise(self.q,self.T,0,date))
-
-
-class DiscountingCurve(Curve):
-
-    def __init__(self, reference=None, discounts=None, dates=None, act = "No"):
-        self.reference = reference
-        if act=="ACT360":
-            self.T = ACT_360(dates,self.reference)
[…]

### 2022-03-08 reinforcement/envs/pricing/read_market.py

@@ -1,135 +1,0 @@
-import xml.etree.ElementTree as ET
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, ForwardVariance,LocalVolatilityCurve
-from numpy import array, delete, zeros, reshape, append, max
-import numpy as np
-
-class MarketDataReader:
-
-    def __init__(self,filename = None):
-        tree = ET.parse(filename)
-        self.root = tree.getroot()
-
-    def get_stock_number(self):
-        return len(self.root[3][0][1][1][1][0])
-
-    def get_stock_names(self):
-        N_stocks = self.get_stock_number()
-        names = array([])
-        for i in range(N_stocks):
-            names=append(names,self.root[3][0][1][1][1][0][i].text)
-        delete_equity = [0,11,12,13,14,15]
-        names = delete(names,delete_equity)
-        wrong_names = [names[4],names[5],names[6]]
-        names[4] = wrong_names[2]
-        names[5] = wrong_names[0]
-        names[6] = wrong_names[1]
-        return names
-    def get_correlation(self):
-        N_stocks = self.get_stock_number()
-        correlation_matrix = zeros(N_stocks**2)
-        for i in range (N_stocks**2):
-            correlation_matrix[i] = float(self.root[3][0][1][1][0][0][i].text)
-        correlation_matrix = reshape(correlation_matrix, (N_stocks,N_stocks))
-        """Delete cash bank"""
-        delete_equity = [0,11,12,13,14,15]
-        correlation_matrix = delete(correlation_matrix,delete_equity, axis = 0)
-        correlation_matrix = delete(correlation_matrix,delete_equity, axis = 1)
-        return correlation_matrix
-
-    def get_discounts(self):
-        discounts = zeros(len(self.root[3][1][0][0][0][0][0]))
-        discounts_dates = zeros(len(self.root[3][1][0][0][0][0][0]))
-        for i in range(len(self.root[3][1][0][0][0][0][0])):
-            discounts_dates[i] = float(self.root[3][1][0][0][0][0][0][i].text)
-            discounts[i] = float(self.root[3][1][0][0][0][1][0][i].text)
-        return DiscountingCurve(reference=self.get_reference_date(), discounts=discounts, dates=discounts_dates, act="365")
-
-    def get_reference_date(self):
-        return float(self.root[3][1][0][0][2][0].text)
-
-    def get_spot_prices(self):
-        index_equity = [1,2,3,4,5,6,7,8,9,10]
[…]

### 2022-03-08 reinforcement/envs/pricing/targetvol.py

@@ -1,302 +1,0 @@
-import numpy as np
-from numpy import exp, sqrt, log
-from envs.pricing.pricing import piecewise_function, Curve, PricingModel, quad_piecewise
-from numpy.linalg import cholesky
-from scipy.optimize import minimize, LinearConstraint, Bounds
-from scipy.interpolate import interp1d
-
-
-"""Notation of pallavicini and daluiso draft"""
-
-
-class Drift(Curve):
-    """Time dependent Repo rates union"""
-    def __init__(self, forward_curves = None):
-        self.T = time_grid_union(curve_array_list = forward_curves)
-        Ndim = len(forward_curves)
-        self.mu = forward_curves[0].q(self.T)
-        self.mu = np.stack((self.mu,forward_curves[1].q(self.T)),axis=1)
-        """Calculating the instant repo rates"""
-        for i in range(2,Ndim):
-            self.mu = np.insert(self.mu,len(self.mu.T),forward_curves[i].q(self.T),axis=1)
-        self.m = interp1d(self.T, self.mu, axis=0, kind='previous',fill_value="extrapolate", assume_sorted=False) 
-       # print("Drift time grid:",self.T)
-        #print("Drift values:", self.mu)   
-    def curve(self,date):
-        return self.m(date)
-
-
-class CholeskyTDependent(Curve):
-    """Time dependent cholesky variance-covariance matrix union"""
-    def __init__(self, variance_curves = None, correlation_chole = None):
-        self.T = time_grid_union(curve_array_list = variance_curves)
-        N_times = len(self.T)
-        Ndim = len(variance_curves)
-        self.nu = np.zeros((Ndim,Ndim,N_times))
-        I = np.identity(Ndim)
-        for i in range(N_times):
-            vol = np.zeros(Ndim)
-            for j in range(Ndim):
-                vol[j] = sqrt(variance_curves[j](self.T[i]))
-            vol = I*vol
-            self.nu[:,:,i] = vol@correlation_chole
-        self.n = interp1d(self.T, self.nu, axis=2, kind='previous',fill_value="extrapolate", assume_sorted=False) 
-      #  print("Cholesky covariance-variance time grid:",self.T)
-     #   print("Cholesky covariance-variance matrix values:", self.nu)
-   
-    def curve(self,date):
-        return self.n(date)
-
-        
-class Strategy(Curve):
[…]

### 2022-03-08 reinforcement/envs/tvs_2market_env.py

@@ -1,141 +1,0 @@
-import gym
-from gym import spaces
-from gym.utils import seeding
-from numpy import exp, log, sqrt
-from numpy.linalg import cholesky
-import numpy as np
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, Black, ForwardVariance
-from envs.pricing.closedforms import European_option_closed_form
-from envs.pricing.targetvol import Drift, CholeskyTDependent, Strategy, TVSForwardCurve, TargetVolatilityStrategy
-from envs.pricing.read_market import MarketDataReader
-from envs.pricing.n_sphere import sign_renormalization
-
-class TVS_environment2(gym.Env):
-    """Target volatility strategy Option environment for equity I NKY NTR EUR and FTSE100 NTR E"""
-    def __init__(self, filename= "TVS_example.xml", spot_I = 1., target_volatility=5./100,strike_opt=1., maturity=1., constraint = "only_long", action_bound = 25/100, sum_long=None, sum_short=None):
-        #Preparing Time grid for the RL agent
-        self.constraint = constraint
-        self.I_0 = spot_I
-        self.strike_option = strike_opt
-        self.target_vol = target_volatility
-        self.T = maturity
-        self.Nsim = 1e4
-        self.current_time = 0.
-        self.time_index = 0
-        self.simulation_index = 0
-        months = np.array([31,28,31,30,31,30,31,31,30,31,30,31])
-        self.time_grid = np.cumsum(months)/365.
-        self.time_grid = np.insert(self.time_grid,0,0)
-        #Loading Market data and preparing the BS model"""
-        reader = MarketDataReader(filename)
-        self.correlation = np.array(([1.,0.86],[0.86,1.]))
-        self.correlation_chole = cholesky(self.correlation)
-        self.N_equity = 2
-        self.D = reader.get_discounts()
-        self.discount = self.D(maturity)
-        self.F = reader.get_forward_curves()
-        self.V = reader.get_volatilities()
-        self.F = [self.F[0],self.F[3]]
-        self.V = [self.V[0],self.V[3]]
-        names = reader.get_stock_names()
-        names = [names[0],names[3],names[4]]
-        print("Reinforcement Learning simulation for the market with equities: ",names)
-        #Creating the objects for the TVS
-        self.mu = Drift(forward_curves = self.F)
-        self.nu = CholeskyTDependent(variance_curves = self.V, correlation_chole = self.correlation_chole)
-        print('Observations along the time grid', self.time_grid)
[…]

### 2022-03-08 reinforcement/envs/tvs_3market_env.py

@@ -1,145 +1,0 @@
-import gym
-from gym import spaces
-from gym.utils import seeding
-from numpy import exp, log, sqrt
-from numpy.linalg import cholesky
-import numpy as np
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, Black, ForwardVariance
-from envs.pricing.closedforms import European_option_closed_form
-from envs.pricing.targetvol import Drift, CholeskyTDependent, Strategy, TVSForwardCurve, TargetVolatilityStrategy
-from envs.pricing.read_market import MarketDataReader
-from envs.pricing.n_sphere import sign_renormalization
-
-class TVS_environment3(gym.Env):
-    """Target volatility strategy Option environment for equity I NKY NTR EUR and FTSE100 NTR E"""
-    def __init__(self, filename= "TVS_example.xml", spot_I = 1., target_volatility=5./100,strike_opt=1., maturity=1., constraint = "only_long", action_bound = 25/100, sum_long=None, sum_short=None):
-        #Preparing Time grid for the RL agent
-        self.constraint = constraint
-        self.I_0 = spot_I
-        self.strike_option = strike_opt
-        self.target_vol = target_volatility
-        self.T = maturity
-        self.Nsim = 1e4
-        self.current_time = 0.
-        n_observations = 12
-        self.time_index = 0
-        self.simulation_index = 0
-        #self.time_grid = np.linspace(self.T/n_observations,self.T,n_observations)
-        #self.time_grid = np.insert(self.time_grid,0,0)
-        #Loading Market data and preparing the BS model"""
-        reader = MarketDataReader(filename)
-        #self.spot_prices = reader.get_spot_prices()
-        #self.spot_prices = np.array([self.spot_prices[3],self.spot_prices[4]])
-        self.correlation = np.array(([1.,0.86,0.],[0.86,1.,0.],[0.,0.,1.]))
-        self.correlation_chole = cholesky(self.correlation)
-        self.N_equity = 3
-        self.D = reader.get_discounts()
-        self.discount = self.D(maturity)
-        self.F = reader.get_forward_curves()
-        self.V = reader.get_volatilities()
-        self.F = [self.F[0],self.F[3],self.F[4]]
-        self.V = [self.V[0],self.V[3],self.V[4]]
-        names = reader.get_stock_names()
-        names = [names[0],names[3],names[4]]
-        print("Reinforcement Learning simulation for the market with equities: ",names)
-        #Creating the objects for the TVS
-        self.mu = Drift(forward_curves = self.F)
[…]

### 2022-03-08 reinforcement/envs/tvs_bs_env.py

@@ -1,0 +1,177 @@
+import gym
+from gym import spaces
+from gym.utils import seeding
+import numpy as np
+
+import sys
+sys.path.insert(1, '../pricing')
+from pricing import EquityForwardCurve, DiscountingCurve, Black, ForwardVariance
+from closedforms import BS_European_option_closed_form
+from read_market import LoadFromTxt
+from targetvol import Drift, CholeskyTDependent, Strategy, TVSForwardCurve, TargetVolatilityStrategy
+
+from envs.utils import build_allocation_time_grid, sign_renormalization
+
+class TVS_BS_ENV(gym.Env):
+    """Target volatility strategy Option environment for the Black-Scholes model."""
+
+    def __init__(self, market_folder: str = "../market_data/FakeSmilesDisplacedDiffusion",
+                 asset_names: list = ["DJ 50 EURO E", "S&P 500 NET EUR"],
+                 allocation_frequency: str = "monthly", target_volatility: float = 5 / 100, tvs_spot_value: float = 1.0,
+                 option_strike: float = 1.0, option_maturity: float = 2.0, action_constraints: str = None,
+                 action_bound: float = 5.0, overall_long_position: float = None, overall_short_position: float = None,
+                 n_sim_for_cache: int = int(2e3)) -> None:
+        """
+        Target volatility strategy Option RL environment with a Black and Scholes model for the assets. In this environment the action space and the observation
+        space are continuous. The observation state is a vector of size (n_assets + 2) with the following structure: [stock price, tvs level, time]
+        and the state space is the following block: lower_bound = [-2.5, 0., 0.]; upper_bound = [2.5, 10 * tvs_spot_value, maturity].
+        :param market_folder (str, default="../market_data/FakeSmilesDisplacedDiffusion"): folder where the market data is stored
+        :param asset_names (list or tuple, default=["DJ 50 EURO E", "S&P 500 NET EUR"]): list of asset names to be loaded
+        :param allocation_frequency (str, default="monthly"): allocation frequency (available options: "monthly" and "daily"). The
+        allocation frequency determines the time grid on which the RL agent changes the risky portfolio composition (alpha).
+        :param target_volatility (float, default=5/100): target volatility level of the portfolio.
+        :param tvs_spot_value (float, default=1.0): spot value of the target volatility strategy.
+        :param option_strike (float, default=1.0): strike price of the option.
+        :param option_maturity (float, default=2.0): maturity of the option expressed in yrf (it must be an integer number of years).
[…]

### 2022-03-08 reinforcement/envs/tvs_env.py

@@ -1,138 +1,0 @@
-import gym
-from gym import spaces
-from gym.utils import seeding
-from numpy import exp, log, sqrt
-from numpy.linalg import cholesky
-import numpy as np
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, Black, ForwardVariance
-from envs.pricing.closedforms import European_option_closed_form
-from envs.pricing.targetvol import Drift, CholeskyTDependent, Strategy, TVSForwardCurve, TargetVolatilityStrategy
-from envs.pricing.read_market import MarketDataReader
-from envs.pricing.n_sphere import sign_renormalization
-
-class TVS_environment(gym.Env):
-    """Target volatility strategy Option environment"""
-    def __init__(self, filename= "TVS_example.xml", spot_I = 100, target_volatility=0.1,strike_opt=100., maturity=1., constraint = "only_long", action_bound = 25/100, sum_long=None, sum_short=None):
-        #Preparing Time grid for the RL agent
-        self.constraint = constraint
-        self.I_0 = spot_I
-        self.strike_option = strike_opt
-        self.target_vol = target_volatility
-        self.T = maturity
-        self.Nsim = 1e3
-        self.current_time = 0.
-        n_observations = 365
-        self.time_index = 0
-        self.simulation_index = 0
-        months = np.array([31,28,31,30,31,30,31,31,30,31,30,31])
-        self.time_grid = np.cumsum(months)/n_observations
-        self.time_grid = np.insert(self.time_grid,0,0.)
-        #Loading Market data and preparing the BS model"""
-        reader = MarketDataReader(filename)
-        self.spot_prices = reader.get_spot_prices()
-        self.correlation = reader.get_correlation()
-        self.correlation_chole = cholesky(self.correlation)
-        self.N_equity = len(self.correlation)
-        self.D = reader.get_discounts()
-        self.discount = self.D(maturity)
-        self.F = reader.get_forward_curves()
-        self.V = reader.get_volatilities()
-        #Creating the objects for the TVS
-        self.mu = Drift(forward_curves = self.F)
-        self.nu = CholeskyTDependent(variance_curves = self.V, correlation = self.correlation)
-        self.TVSF = TVSForwardCurve(reference=0.,vola_target = self.target_vol, spot_price = self.I_0, mu = self.mu, nu = self.nu, discounting_curve = self.D)
-        self.model = Black(fixings=self.time_grid, variance_curve=self.V, forward_curve=self.F)
[…]

### 2022-03-08 reinforcement/envs/tvs_lv_env.py

@@ -1,204 +1,246 @@
 import gym
 from gym import spaces
 from gym.utils import seeding
-from numpy import log, sqrt, exp
+from numpy import log, sqrt
 import numpy as np
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, LV_model, ForwardVariance, quad_piecewise
-from envs.pricing.closedforms import European_option_closed_form
-from envs.pricing.targetvol import Drift, Strategy, TVSForwardCurve, CholeskyTDependent, optimization_only_long
-from envs.pricing.fake_market_lv import load_fake_market_lv
-from envs.pricing.loadfromtxt import LoadFromTxt
-from envs.pricing.targetvol import optimization_only_long, Markowitz_solution
-from envs.pricing.n_sphere import sign_renormalization
 
-class TVS_LV(gym.Env):
-    """Target volatility strategy Option environment with a Local volatility model for the assets
+import sys
+sys.path.insert(1, '../pricing')
+
+from pricing import EquityForwardCurve, DiscountingCurve, LV_model, ForwardVariance, quad_piecewise
+from targetvol import Drift, Strategy, TVSForwardCurve, CholeskyTDependent, optimization_only_long
+from read_market import LoadFromTxt
+from targetvol import optimization_only_long, Markowitz_solution
+
+from envs.utils import build_allocation_time_grid, sign_renormalization
+
+
+class TVS_LV_ENV_reward1(gym.Env):
     """
-    def __init__(self, N_equity= 2, frequency = "month", target_volatility=5/100, I_0 = 1., strike_opt=1., 
-    maturity=2., constraint = "free", action_bound=5., sum_long = None, sum_short=None):
-        """Pricing parameters for the option"""
-        self.bang_bang_action = 0
-        self.frombaseline = 0
-        self.fromblack = 0
-        self.constraint = constraint
+    Target volatility strategy Option environment with a Local volatility model for the assets
+    """
+
+    def __init__(self, market_folder: str = "../market_data/FakeSmilesDisplacedDiffusion", asset_names: list = ["DJ 50 EURO E", "S&P 500 NET EUR"],
+                 allocation_frequency: str = "monthly", target_volatility: float = 5 / 100, tvs_spot_value: float = 1.0,
+                 option_strike: float = 1.0, option_maturity: float = 2.0,
+                 neural_network_action_parameterization: str = None, action_constraints: str = None,
[…]

### 2022-03-08 reinforcement/envs/tvs_lv_env_new_reward.py

@@ -1,0 +1,279 @@
+import gym
+from gym import spaces
+from gym.utils import seeding
+from numpy import log, sqrt
+import numpy as np
+
+import sys
+sys.path.insert(1, '../pricing')
+
+from pricing import EquityForwardCurve, DiscountingCurve, LV_model, ForwardVariance, quad_piecewise
+from targetvol import Drift, Strategy, TVSForwardCurve, CholeskyTDependent, optimization_only_long
+from read_market import LoadFromTxt
+from targetvol import optimization_only_long, Markowitz_solution
+from closedforms import BS_European_option_closed_form
+from envs.utils import build_allocation_time_grid, sign_renormalization
+
+
+class TVS_LV_ENV_reward2(gym.Env):
+    """Target volatility strategy Option environment with a Local volatility model for the assets"""
+
+    def __init__(self, market_folder: str = "../market_data/FakeSmilesDisplacedDiffusion", asset_names: list or tuple = ["DJ 50 EURO E", "S&P 500 NET EUR"],
+                 allocation_frequency: str = "monthly", target_volatility: float = 5 / 100, tvs_spot_value: float = 1.0,
+                 option_strike: float = 1.0, option_maturity: float = 2.0,
+                 neural_network_action_parameterization: str = None, action_constraints: str = None,
+                 action_bound: float = 5.0, overall_long_position: float = None, overall_short_position: float = None,
+                 n_sim_for_cache: int = int(2e3)) -> None:
+        """
+        Target volatility strategy Option RL environment with a Local volatility model for the assets. The environment implements
+        the second reward function experimented in the paper. In this environment the action space and the observation space are
+        continuous. The observation state is a vector of size (n_assets + 2) with the following structure: [stock price, tvs level, time]
+        and the state space is the following block: lower_bound = [-2.5, 0., 0.]; upper_bound = [2.5, 10 * tvs_spot_value, maturity].
+        :param market_folder (str, default="../market_data/FakeSmilesDisplacedDiffusion"): folder where the market data is stored
+        :param asset_names (list or tuple, default=["DJ 50 EURO E", "S&P 500 NET EUR"]): list of asset names to be loaded
+        :param allocation_frequency (str, default="monthly"): allocation frequency (available options: "monthly" and "daily"). The
+        allocation frequency determines the time grid on which the RL agent changes the risky portfolio composition (alpha).
+        :param target_volatility (float, default=5/100): target volatility level of the portfolio.
+        :param tvs_spot_value (float, default=1.0): spot value of the target volatility strategy.
[…]

### 2022-03-08 reinforcement/envs/tvs_new_reward.py

@@ -1,195 +1,0 @@
-import gym
-from gym import spaces
-from gym.utils import seeding
-from numpy import log, sqrt, exp
-import numpy as np
-from scipy.integrate import quad
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, LV_model, ForwardVariance, quad_piecewise
-from envs.pricing.targetvol import Drift
-from envs.pricing.loadfromtxt import LoadFromTxt
-from envs.pricing.targetvol import Markowitz_solution
-from envs.pricing.closedforms import European_option_closed_form
-
-class TVS_LV_newreward(gym.Env):
-    """Target volatility strategy Option environment with a Local volatility model for the assets
-    """
-    def __init__(self, N_equity= 2, frequency = "month", target_volatility=5/100, I_0 = 1., strike_opt=1., 
-    maturity=2., constraint = "free", action_bound=5., sum_long = None, sum_short=None):
-        #Simulation parameters
-        self.constraint = constraint
-        self.from_baseline = 1
-        self.target_vol = target_volatility
-        self.N_equity = N_equity            
-        self.strike_opt = strike_opt
-        self.I_0 = I_0
-        self.I_t = I_0
-        self.T = maturity
-        self.V_t = 0. #derivative value at time t
-        self.V_t_plus = 0. #derivative value at time t+1
-
-        names = ["DJ 50 EURO E","S&P 500 NET EUR"]
-        correlation = np.array(([1.,0.],[0.,1.]))
-        folder_name = "FakeSmilesDisplacedDiffusion"#"FakeSmilesDisplacedDiffusion"
-        ACT = 365.
-        """Time grid creation for the simulation"""
-        self.Identity = np.identity(self.N_equity)
-        if frequency == "month":
-            month_dates = np.array([31.,28.,31.,30.,31.,30.,31.,31.,30.,31.,30.,31.])
-            months = month_dates
-            if self.T > 1.:
-                for i in range(int(self.T)-1):
-                    months = np.append(months, month_dates)
-            self.observation_grid = np.cumsum(months)/ACT
-            self.N_euler_grid = 60       
-            self.state_index = np.arange(int(12*self.T)+1)*self.N_euler_grid
-        elif frequency == "day":
-            self.observation_grid = np.linspace(1./ACT, self.T, int(self.T*ACT))
-            self.N_euler_grid = 2
-            self.state_index = np.arange(int(365*self.T)+1)*self.N_euler_grid
-        
[…]

### 2022-03-08 reinforcement/envs/tvs_simple_env.py

@@ -1,118 +1,0 @@
-import gym
-from gym import spaces
-from gym.utils import seeding
-from numpy import log, sqrt, exp
-import numpy as np
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, Black, ForwardVariance, quad_piecewise
-from envs.pricing.closedforms import European_option_closed_form
-from envs.pricing.targetvol import Drift, CholeskyTDependent, Strategy, TVSForwardCurve, TargetVolatilityStrategy
-from envs.pricing.fake_market import load_fake_market
-from envs.pricing.n_sphere import sign_renormalization
-
-class TVS_simple(gym.Env):
-    """Target volatility strategy Option environment with a simple market
-    """
-    def __init__(self, N_equity= 3, target_volatility=5/100, I_0 = 1., r=1/100., strike_opt=1., maturity=1., constraint = "only_long", action_bound=20./100., sum_long = None, sum_short=None):
-        self.constraint = constraint
-        self.target_vol = target_volatility
-        self.I_0 = I_0
-        self.N_equity = N_equity                                #number of equities
-        self.asset_history = np.array([])
-        self.T = maturity
-        n_observations = 12
-        self.time_grid = np.linspace(self.T/n_observations,self.T,n_observations)
-        self.time_grid = np.insert(self.time_grid,0,0)
-        self.time_index = 0
-        self.current_time = 0.
-        self.asset_history = np.array([])
-        self.alpha_t = np.array([])
-        self.strike_opt = strike_opt
-        self.simulation_index = 0
-        self.Nsim = 1e4
-        self.D, self.F, self.V, self.correlation, self.spot_prices = load_fake_market(N_equity, r, self.T)
-        self.discount = self.D(self.T)
-        self.correlation_chole = np.linalg.cholesky(self.correlation)
-        self.model = Black(fixings=self.time_grid,variance_curve=self.V,forward_curve = self.F)
-        self.mu = Drift(self.F)
-        self.nu = CholeskyTDependent(self.V,self.correlation_chole)
-        self.TVSF = TVSForwardCurve(reference=0.,vola_target = self.target_vol, spot_price = self.I_0, mu = self.mu, nu = self.nu, discounting_curve = self.D)
-        self.integral_variance = np.cumsum(self.model.variance[:,1:],axis=1).T
-        self.integral_variance_sqrt = sqrt(self.integral_variance)
-        if self.constraint == 'long_short_limit' and (sum_long is None or sum_short is None):
[…]

### 2022-03-08 reinforcement/envs/utils.py

@@ -1,0 +1,72 @@
+import numpy as np
+
+def n_sphere_to_cartesian(radius: float, angles: np.ndarray) -> np.ndarray:
+    """
+    Function that takes n-spherical coordinates and convert into n-cartesian coordinates
+    angles: the n-2 values between [0,\pi) and last one between [0,2\pi)
+    :param radius (float): the radius of the n-sphere
+    :param angles (np.ndarray): the angles of the n-sphere
+    :return: the n-cartesian coordinates (shape: (len(angles),))
+    """
+    a = np.concatenate((np.array([2*np.pi]), angles))
+    si = np.sin(a)
+    si[0] = 1
+    si = np.cumprod(si)
+    co = np.cos(a)
+    co = np.roll(co, -1)
+    return si*co*radius
+
+
+def sign_renormalization(vec: np.ndarray, sum_plus: float, sum_minus: float) -> np.ndarray:
+    """
+    Renormalize positive and negative elements of an array separately if necessary such
+    that sum of positive is equal to sum_plus and the abs sum of negative is equal to sum_neg
+    :param vec: the array to be renormalized
+    :param sum_plus: the sum of positive elements
+    :param sum_minus: the sum of negative elements
+    :return: the renormalized array (shape: (len(vec),))
+    """
+    pos = vec >= 0
+    neg = np.invert(pos)
+    p_s = np.sum(vec[pos])
+    n_s = abs(np.sum(vec[neg]))
+    if p_s > sum_plus:
+        vec[pos] = (vec[pos]/p_s)*sum_plus
+    if n_s > sum_minus:
+        vec[neg] = (vec[neg]/n_s)*sum_minus
+    return vec
+
+def build_allocation_time_grid(maturity: float, allocation_frequency: str = 'monthly', day_count_convention: str = "ACT_365"):
+    """
+    Builds a time grid for the allocation of a portfolio
+    :param maturity: the maturity of the portfolio in years (it must be an integer number of years).
+    :param allocation_frequency: the frequency of the allocation
+    :param day_count_convention: the day count convention
+    :return: a numpy array of the time grid
+    """
+    if day_count_convention == "ACT_365":
+        number_days_in_year = 365.0
+    elif day_count_convention == "ACT_360":
+        number_days_in_year = 360.0
+    else:
+        raise ValueError("Unknown day count convention (available: ACT_365, ACT_360)")
+
+    if maturity % 1.0 != 0:
[…]

### 2022-03-08 reinforcement/go.py

@@ -1,41 +1,46 @@
 #! python
-
 from baselines.run import main
 import envs.fe_envs
 
-############################################ MANDATORY PARAMETERS #################################################
-# do_train (boolean) = if you want to train the agent or to do the MC to obtain the price of the strategy
-# do_test (boolean) = if you want to do MC
-# env (string) = enviroment name
-# alg (string) = rl algorithm (usually ppo2)
-# num_layers (string) = number of hidden layers
-# num_hidden (string) = number of neurons for each hidden layer
-# num_env (string) = number for multiprocessing python (parallelism)
-# lr (string) = learning rate (about 3e-4)
-# train_timesteps (string) = how much time it spends to train the agent (approximately how many episodes) (about 1e8)
-# test_episodes (string) = number of episodes for the monte carlo calculation (you need do_test=True)
-# print_episodes (string: 0 or 1) = print some information
-# print_period (string: number) = every how many steps it prints a log about the trainign
-###################################################################################################################
-
-# value_network (string) = "None/shared": creates one NN for policy and value function; "copy": creates two different NN for policy and value function
-# lamb (double) = (default 0.95) parameter that controls the bias and variance of the expected reward estimator (useful for convergence)
-# noise (double) = add noise over the output of the NN (action is not deterministic)
-# beta (double) = importance of value function
-# custom_suffix (string) = add a string of characters at the end of the folder name
-
-
-
-def build_args(do_train,do_test, env, alg, num_layers, num_hidden, num_env, lr,
+def build_args(do_train, do_test, env, alg, num_layers, num_hidden, num_env, lr,
                train_timesteps, test_episodes, print_episodes, print_period, activation=None, save_interval=None,
                value_network=None, lam=None, gamma=None, noise=None, beta=None, ent=None, training_seed=None,
                test_seed=None, restart_training=None, initial_guess=None,
                batch=None, custom_options=None, custom_suffix=None):
-    """Utility function for the most common argument configurations."""
-
-    #do_test = not do_train
+    """
+    Utility function for the most common argument configurations.
+    :param do_train (bool): whether to train the agent.
+    :param do_test (bool): whether to test the agent by means of a Monte Carlo simulation. The script uses the saved model built with the same arguments provided in build_args function in the do_train fase.
[…]

### 2022-03-08 reinforcement/go_test.py

@@ -1,145 +1,0 @@
-#! python
-
-from baselines.run import main
-import envs.fe_envs
-
-############################################ MANDATORY PARAMETERS #################################################
-# do_train (boolean) = if you want to train the agent or to do the MC to obtain the price of the strategy
-# do_test (boolean) = if you want to do MC
-# env (string) = enviroment name
-# alg (string) = rl algorithm (usually ppo2)
-# num_layers (string) = number of hidden layers
-# num_hidden (string) = number of neurons for each hidden layer
-# num_env (string) = number for multiprocessing python (parallelism)
-# lr (string) = learning rate (about 3e-4)
-# train_timesteps (string) = how much time it spends to train the agent (approximately how many episodes) (about 1e8)
-# test_episodes (string) = number of episodes for the monte carlo calculation (you need do_test=True)
-# print_episodes (string: 0 or 1) = print some information
-# print_period (string: number) = every how many steps it prints a log about the trainign
-###################################################################################################################
-
-# value_network (string) = "None/shared": creates one NN for policy and value function; "copy": creates two different NN for policy and value function
-# lamb (double) = (default 0.95) parameter that controls the bias and variance of the expected reward estimator (useful for convergence)
-# noise (double) = add noise over the output of the NN (action is not deterministic)
-# beta (double) = importance of value function
-# custom_suffix (string) = add a string of characters at the end of the folder name
-
-
-
-def build_args(do_train,do_test, env, alg, num_layers, num_hidden, num_env, lr,
-               train_timesteps, test_episodes, print_episodes, print_period, activation=None, save_interval=None,
-               value_network=None, lam=None, gamma=None, noise=None, beta=None, ent=None, training_seed=None,
-               test_seed=None, restart_training=None, initial_guess=None,
-               batch=None, custom_options=None, custom_suffix=None):
-    """Utility function for the most common argument configurations."""
-
-    #do_test = not do_train
-    suffix = "" if custom_suffix is None else custom_suffix
-    options = [] if custom_options is None else custom_options
-
-    if activation is not None:
-        suffix = '_' + activation + suffix
-        options.append('--activation=tf.nn.' + activation)
-
-    if value_network is not None:
-        suffix = '_' + value_network + suffix
[…]

### 2022-03-08 reinforcement/plot_agent.py

@@ -1,300 +1,0 @@
-from baselines.common.cmd_util import common_arg_parser
-from baselines.run import parse_cmdline_kwargs, train, remove_train_noise
-import numpy as np
-from numpy import exp, log, sqrt
-from matplotlib import pyplot as plt
-from envs.pricing.pricing import EquityForwardCurve, Black, ForwardVariance, DiscountingCurve, LV_model
-from envs.pricing.fake_market import load_fake_market
-from envs.pricing.read_market import MarketDataReader, Market_Local_volatility
-from envs.pricing.n_sphere import sign_renormalization
-from envs.pricing.targetvol import CholeskyTDependent
-from gym import spaces
-import envs.fe_envs
-from matplotlib import cm
-from mpl_toolkits.mplot3d import Axes3D
-
-def model_creation(seed=None, fixings=None, n=None, normalized=None, market = 0, pricing_model='LV', return_vola=0):
-    if pricing_model == 'LV':
-        local_vol = Market_Local_volatility()
-        LV = [local_vol[0],local_vol[1],local_vol[2],local_vol[3],local_vol[6],local_vol[4],local_vol[5],local_vol[7],local_vol[8]]
-    if not market:
-        N_equity = n
-        print('Loading fake market with number of equity: ',n)
-        r = 1/100
-        maturity=1
-        load_fake_market(N_equity, r, maturity)
-        D, F, V, correlation, spot_prices = load_fake_market(N_equity, r, maturity)
-        names = []
-        for i in range(N_equity):
-            names.append("Equity "+str(i+1))
-        if pricing_model == 'LV':
-            raise Exception("I cannot implement LV model with a fake market")
-    elif market:
-        N_equity = n
-        print('Loading real market with number of equity: ',n)
-        reader = MarketDataReader("TVS_example.xml")
-        correlation = reader.get_correlation()
-        names = reader.get_stock_names()
-        D = reader.get_discounts()
-        F = reader.get_forward_curves()
-        spot_prices = reader.get_spot_prices()
-        V = reader.get_volatilities()
-        if n==2:
-            print('Cutting real data for '+str(n)+' equities')
-            F = [F[0],F[3]]
-            V = [V[0],V[3]]
-            spot_prices = np.array([spot_prices[3],spot_prices[4]])
-            correlation = np.array(([1.,0.86],[0.86,1.]))
-            names = [names[0],names[3]]
-            if pricing_model == 'LV':
[…]

### 2022-03-08 reinforcement/plot_agent_actions_and_values.py

@@ -1,0 +1,34 @@
+from plot_utilities import plot_one_episode_actions, plot_action_space
+import numpy as np
+
+
+######################################################## SCRIPT INPUT ##############################################################################################
+
+plot_one_episode_action = False # if True, plot one episode action else plot a section of the action space or the value function
+env_id = "TVS_LV-v2"
+rl_model_parameters_dict = {"rl_algo": "ppo2",
+                 "training_timesteps": "3e4",
+                 "num_layers": "5",
+                 "num_hidden": "8",
+                 "learning_rate": "3e-4",
+                 "activation": "tanh",
+                 "value_network": "copy",
+                 "beta": "0.7",
+                 "custom_suffix": "_trainingseed45891_beta0.7_copy_tanhpricing",
+                 "seed": "14"}
+
+# parameters for plotting the action space or the value function
+if not plot_one_episode_action:
+    plot_value_function = False  # if false then plot the action space of the agent
+    reference_state = np.array([np.log(1), np.log(1), 1.0, 1./12.])  # reference state where display the action/value space
+                                                                     # block = [log(S_1(t)/F(t)), log(S_2(t)/F(t)), I/I_0, t]
+    variable_indexes_to_vary = [0, 1]  # which state variable to vary in the plot (available 0: log(S_1(t)/F(t)), 1: log(S_2(t)/F(t)), 2: I/I_0, 3: t)
+                                       # variable_indexes_to_vary must be a tuple or list of length 2 (3d plot) or 1 (2d plot)
+    number_of_points = 20  # number of points to plot in the action/value space
+
+################################################################################################################################################################
+
+if plot_one_episode_action:
+    plot_one_episode_actions(env_id, rl_model_parameters_dict)
+else:
+    plot_action_space(env_id, rl_model_parameters_dict, plot_value_function, reference_state, variable_indexes_to_vary, number_of_points)

### 2022-03-08 reinforcement/plot_learning_curves.py

@@ -1,0 +1,46 @@
+from plot_utilities import plot_rolling, pu, join_curves
+import matplotlib.pyplot as plt
+import numpy as np
+from matplotlib import ticker
+
+######################################################## SCRIPT INPUT ##############################################################################################
+
+discount_factor = 1.0097815717231273  # discount factor used to discount the learning curves
+n_sigma = 2.55  # number of standard deviations to plot
+join_learning = True  # if True then the learning curves provided in WHICH_LOGS are joined together. If False then each learning curve is plotted separately.
+labels_fontsize = 12  # font size for the labels of the learning curves plots
+X_AXIS_TIMESTEPS = False  # if True then the x-axis is in time-steps, if False then the x-axis is in episodes
+MOVING_AVERAGE_WINDOW = int(1000)  # window of the moving average used to smooth the learning curves (measured in episodes)
+LOG_FOLDER = './logs/TVS_LV-v2/'   # root folder of the environment log files
+WHICH_LOGS = [  # ("name of the output folder", "legend label"). If join_learning is True then the legend label is ignored.
+    ('ppo2_1e6_5x8_3e-4_trainingseed45891_beta0.7_copy_tanhpricing_prova', 'learning_curve_1'),
+    ('ppo2_1e5_5x8_3e-4_trainingseed45891_beta0.7_copy_tanhpricing_prova', 'learning_curve_2'),
+]
+
+######################################################################################################################################################################
+
+fig, ax = plt.subplots(1, 1, figsize=(8, 5))
+if join_learning:
+    steps_joined, rewards_joined = join_curves(LOG_FOLDER, WHICH_LOGS, X_AXIS_TIMESTEPS)
+    plot_rolling(steps_joined, rewards_joined * discount_factor, MOVING_AVERAGE_WINDOW, WHICH_LOGS[0][1], n_sigma)
+else:
+    for log, title in WHICH_LOGS:
+        results = pu.load_results(LOG_FOLDER + log)
+        r = results[0]
+        steps = np.cumsum(r.monitor.l) if X_AXIS_TIMESTEPS else np.array(r.monitor.l.index)
+        rewards = r.monitor.r           # episode reward
+        plot_rolling(steps, rewards * discount_factor, MOVING_AVERAGE_WINDOW, title, n_sigma)
+
+
[…]

### 2022-03-08 reinforcement/plot_logs.py

@@ -1,42 +1,0 @@
-from baselines.common import plot_util as pu
-import matplotlib.pyplot as plt
-import numpy as np
-from math import sqrt
-import pandas as pd
-
-def plot_rolling(x, y, window, title='reward', std_mul=2.33):
-    """Plot moving average with error bands."""
-
-    y_roll = y.rolling(window)
-    y_mean = y_roll.mean()
-    y_std = y_roll.std() / sqrt(window)
-    y_lo = y_mean - std_mul * y_std
-    y_hi = y_mean + std_mul * y_std
-    plt.plot(x, y_mean, label=title)
-    plt.fill_between(x, y_lo, y_hi, alpha=0.1)
-    print("max y = {} at x = {}".format(y_mean.max(), x[np.argmax(y_mean)]))
-
-
-
-def join_curves(LOG_FOLDER, WHICH_LOGS, X_AXIS_TIMESTEPS):
-    i = 0
-    for log, title in WHICH_LOGS:
-        results = pu.load_results(LOG_FOLDER + log)
-        r = results[0]
-        steps = np.cumsum(r.monitor.l) if X_AXIS_TIMESTEPS else np.array(r.monitor.l.index)
-        rewards = r.monitor.r
-        if i == 0:
-            if X_AXIS_TIMESTEPS:
-                steps_joined = steps.values
-            else:
-                steps_joined = steps
-            rewards_joined = rewards
-        else:
-            rewards.index = rewards.index+len(rewards_joined)
-            rewards_joined = pd.concat([rewards_joined,rewards])
-            if X_AXIS_TIMESTEPS:
-                steps_joined = np.concatenate((steps_joined,steps.values+steps_joined[-1]))
-            else:
-                steps_joined = np.concatenate((steps_joined,steps+steps_joined[-1]))
-        i+=1
-    return steps_joined,rewards_joined

### 2022-03-08 reinforcement/plot_utilities.py

@@ -1,0 +1,349 @@
+from baselines.common import plot_util as pu
+from baselines.common.cmd_util import common_arg_parser
+from baselines.run import parse_cmdline_kwargs, train, remove_train_noise
+import matplotlib.pyplot as plt
+import numpy as np
+from math import sqrt
+import pandas as pd
+import envs.fe_envs
+
+
+def plot_rolling(x: pd.Series, y: pd.Series, window: int, title: str = 'reward', std_mul: float = 2.33) -> None:
+    """
+    Plot moving average with error bands of a pandas Series.
+    :param x (pd.Series): x-axis values.
+    :param y (pd.Series): y-axis values.
+    :param window (int): window size for rolling mean.
+    :param title (str): plot title.
+    :param std_mul (float): multiplier for standard deviation error bars.
+    """
+    y_roll = y.rolling(window)
+    y_mean = y_roll.mean()
+    y_std = y_roll.std() / sqrt(window)
+    y_lo = y_mean - std_mul * y_std
+    y_hi = y_mean + std_mul * y_std
+    plt.plot(x, y_mean, label=title)
+    plt.fill_between(x, y_lo, y_hi, alpha=0.1)
+    print("max y = {} at x = {}".format(y_mean.max(), x[np.argmax(y_mean)]))
+
+
+def join_curves(LOG_FOLDER: str, WHICH_LOGS: tuple, X_AXIS_TIMESTEPS: bool) -> tuple:
+    """
+    Join multiple learning curves into one plot.
+    :param LOG_FOLDER (str): folder containing log files of a given baselines environment.
+    :param WHICH_LOGS (tuple): tuple of strings containing the names of the log files to be joined and the legend label.
+    :param X_AXIS_TIMESTEPS (bool): if True, x-axis is timesteps, else x-axis is episodes.
+    :return (tuple): tuple containing the x-axis values (timesteps or episodes) and the y-axis values (rewards).
+    """
+    i = 0
+    for log, title in WHICH_LOGS:
+        results = pu.load_results(LOG_FOLDER + log)
+        r = results[0]
+        steps = np.cumsum(r.monitor.l) if X_AXIS_TIMESTEPS else np.array(r.monitor.l.index)
+        rewards = r.monitor.r
+        if i == 0:
+            if X_AXIS_TIMESTEPS:
+                steps_joined = steps.values
+            else:
+                steps_joined = steps
+            rewards_joined = rewards
+        else:
[…]

### 2022-03-08 reinforcement/print_agent.py

@@ -1,51 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-from numpy import exp, sqrt, log
-import tensorflow as tf
-from plot_agent import plot, build_args,model_creation, plot3d
-strategylong = True
-
-Seed = 19
-
-n_equity = 2
-time_dep = True
-constraint = "only_long"  #only_long, free, long_short_limit
-model = 'LV'
-sum_short=50./100.
-sum_long=50./100.
-input_normalize = 2
-market_data=1
-# configurations: which environment and agents
-PLOT_VALUE = False  # otherwise plots actions
-VARIABLE_INDEXES = (2,)  # which dimension to vary
-VARIABLE_POINTS = 12
-ENV = 'TVS_LV-v0'
-REFERENCE_STATE = np.array([np.log(1), np.log(1),1/12])
-X_MAX = 1
-ACTION_SPACE_DESCRIPTION = '[log(S/S0), t]'
-AGENTS = [
-    #(build_args(ENV, 'ppo2', '3e7', '2', '3', '3e-4', custom_suffix='long_change_2long_two_asset_change_beta0.7'), {'action_grid_size': 0}, 'PPO')
-(build_args(ENV, 'ppo2', '1e8', '1', '6', '3e-4', value_network='copy',beta='0.7',custom_suffix='long_month_observation_seed1364798666_'), {'action_grid_size': 0}, 'PPO')
-
-]
-
-reference_str = list(map(str, REFERENCE_STATE))
-for ivar in VARIABLE_INDEXES:
-    reference_str[ivar] = ':'
-z_name = 'Value function' if PLOT_VALUE else 'Actions'
-title = '{} at {} = ['.format(z_name, ACTION_SPACE_DESCRIPTION) + ', '.join(reference_str) + ']'
-
-fig, ax = plt.subplots(1, 1, figsize=(8, 5))
-
-
-for (arg, env_args, lbl) in AGENTS:
-    g = tf.Graph()
-    sess = tf.InteractiveSession(graph=g)
-    with g.as_default():
-        plot(arg, PLOT_VALUE, env_args, REFERENCE_STATE, VARIABLE_INDEXES, VARIABLE_POINTS, 
-        title, lbl, X_MAX, all_time_dep = time_dep, seed=Seed, strategy_constraint= constraint,
-        how_long=sum_long, how_short=sum_short, N_equity = n_equity, normalized=input_normalize, 
[…]

### 2022-03-08 reinforcement/print_logs.py

@@ -1,55 +1,0 @@
-from plot_logs import plot_rolling, pu, join_curves
-import matplotlib.pyplot as plt
-import numpy as np
-from matplotlib import ticker
-
-font = 13
-n_sigma = 1.5
-strategy = 'free'
-
-LOG_FOLDER = './logs/TVS_LV_newreward-v0/'   #in which folder I find the file to plot
-WHICH_LOGS = [  #("name of the output folder", "legend label")
-('ddpg_2e8_3x5_None_trainingseed45891_gamma0.98freefrombaseline_displacedmarket_2assets_monthgrid_maturity2','RL'),
-]
-X_AXIS_TIMESTEPS = 0  # otherwise: episodes
-WINDOW = int(5e5)  # measured in episodes  10000
-join_learning = 0  
-fig, ax = plt.subplots(1, 1, figsize=(8, 5))
-
-discount = 1.0097815717231273
-
-if join_learning:
-    steps_joined, rewards_joined = join_curves(LOG_FOLDER, WHICH_LOGS, X_AXIS_TIMESTEPS)
-    plot_rolling(steps_joined, rewards_joined*discount, WINDOW, r"RL")  #discount factor
-else:
-    for log, title in WHICH_LOGS:
-        results = pu.load_results(LOG_FOLDER + log)
-        r = results[0]
-        steps = np.cumsum(r.monitor.l) if X_AXIS_TIMESTEPS else np.array(r.monitor.l.index)
-        rewards = r.monitor.r           # episode reward
-        plot_rolling(steps, rewards, WINDOW, title)
-
-
-if strategy=="only_long":
-    plt.axhline( 0.014129190929724296+2.5447894383989902e-05*n_sigma, color='red', lw=2.5,linestyle = '--',label="BS Strategy")
-    plt.axhline( 0.014129190929724296-2.5447894383989902e-05*n_sigma, color='red',lw=2.5, linestyle = '--')
-    plt.axhline(0.014784083837497124+3.2454334287712235e-05*n_sigma, color='green',lw=2.5, linestyle = '-.',label="Baseline Strategy")
-    plt.axhline(0.014784083837497124-3.2454334287712235e-05*n_sigma, color='green',lw=2.5, linestyle = '-.')
-        
-elif strategy=='free':
-    plt.axhline(0.040947694174563704 +n_sigma*4.325646589751639e-05, color='red', lw=2.5,linestyle = '--',label="BS Strategy")
[…]

### 2022-03-08 reinforcement/print_multiple_joined_curves.py

@@ -1,63 +1,0 @@
-from plot_logs import plot_rolling, pu, join_curves
-import matplotlib.pyplot as plt
-import numpy as np
-from matplotlib import ticker
-
-font = 13
-LOG_FOLDER = './logs/TVS_simple-v0/'   #in which folder I find the file to plot
-
-X_AXIS_TIMESTEPS = False  # otherwise: episodes
-WINDOW = int(8e5)  # measured in episodes  10000
-
-WHICH_LOGS = [
-  ('ppo2_3e8_1x6_3e-4long_month_observation_beta0.7_2.5variance_seed1733884390','1x4 beta'),
-    ('ppo2_2e8_1x6_3e-4long_month_observation_beta0.7_2.5variance_seed1733884390_restarted','1x4 beta'),
-]
-
-fig, ax = plt.subplots(1, 1, figsize=(8, 5))
-
-steps_joined, rewards_joined = join_curves(LOG_FOLDER, WHICH_LOGS, X_AXIS_TIMESTEPS)
-plot_rolling(steps_joined, rewards_joined, WINDOW, r"1x6")
-
-
-WHICH_LOGS = [
-  ('ppo2_3e8_2x6_3e-4long_month_observation_beta0.7_2.5variance_seed1733884390','1x4 beta'),
-    ('ppo2_1e8_2x6_3e-4long_month_observation_beta0.7_2.5variance_seed1733884390_restarted','1x4 beta'),
-    ('ppo2_1e8_2x6_3e-4long_month_observation_beta0.7_2.5variance_seed1733884390_restarted2','1x4') 
-]
-
-steps_joined, rewards_joined = join_curves(LOG_FOLDER, WHICH_LOGS, X_AXIS_TIMESTEPS)
-plot_rolling(steps_joined, rewards_joined, WINDOW, r"2x6")
-
-WHICH_LOGS = [
-  ('ppo2_2e8_3x6_3e-4long_month_observation_beta0.7_2.5variance_seed3769704067', '0.45'),
-    ('ppo2_1e8_3x6_3e-4long_month_observation_beta0.7_2.5variance_seed3769704067_restarted','0.45'),
-   
-]
-
-steps_joined, rewards_joined = join_curves(LOG_FOLDER, WHICH_LOGS, X_AXIS_TIMESTEPS)
-plot_rolling(steps_joined, rewards_joined, WINDOW, r"3x6")
-
-WHICH_LOGS = [
[…]

### 2022-03-08 reinforcement/print_test_results.py

@@ -1,36 +1,0 @@
-from baselines.common.math_util import discount
-import matplotlib.pyplot as plt 
-import numpy as np
-
-### RL Test Result ####
-MC_price = 0.04108891194279346
-MC_error =   5.039159686727298e-05
-
-### BS PRICE ###
-BS_Price = 0.040947694174563704 
-BS_Price_error = 4.325646589751639e-05
-
-### Baseline Price ###
-Baseline_price = 0.04108567123304307
-Baseline_price_error = 5.03916718605177e-05
-
-discount_factor = 1.0097815717231273
-MC_price = MC_price * discount_factor
-MC_error = MC_error * discount_factor
-
-Baseline_price = Baseline_price*discount_factor
-Baseline_price_error = Baseline_price_error*discount_factor
-
-n_sigma = 2.55
-MC_error = MC_error * n_sigma
-BS_Price_error = BS_Price_error * n_sigma
-Baseline_price_error = Baseline_price_error * n_sigma
-
-errors = np.array([BS_Price_error,Baseline_price_error,MC_error])
-prices = err = np.array([BS_Price,Baseline_price,MC_price])
-fig, ax = plt.subplots(1, 1, figsize=(7, 5))
-plt.xticks([1, 2, 3], ["Black ","Baseline", "RL"],rotation=20)  # Set text labels and properties.
-plt.errorbar([1, 2, 3],prices,errors,fmt='o')
-plt.title("Pricing Results")
-#plt.savefig("strategy_comparison.pdf")
-plt.show()

### 2022-03-08 reinforcement/random_seed.py

@@ -1,4 +1,0 @@
-from numpy.random import Generator, MT19937, SeedSequence
-import sys
-sg = SeedSequence(1234)
-print("random seeds ",MT19937(sg).random_raw(int(sys.argv[1])))

### 2022-03-08 reinforcement/run.py

@@ -1,6 +1,6 @@
 from baselines.run import main   #import the main function of run.py of baselines library
 import sys
-import envs.fe_envs   #import the obgject enviroment of my problem
+import envs.fe_envs   #import the object enviroment of my problem
 
 if __name__ == '__main__':
     main(sys.argv)

### 2022-03-08 reinforcement/testTVS.py

@@ -1,117 +1,0 @@
-# -*- coding: utf-8 -*-
-"""
-Created on Thu Jun 24 17:28:46 2021
-
-@author: u453878
-"""
-
-
-import numpy as np 
-import statsmodels.api as sm
-from scipy.interpolate import interp1d
-import time
-import matplotlib.pyplot as plt
-
-### personal libraries
-from envs.pricing.pricing import EquityForwardCurve, DiscountingCurve, LV_model, ForwardVariance, quad_piecewise, Vanilla_PayOff
-from envs.pricing.targetvol import Drift
-from envs.pricing.loadfromtxt import LoadFromTxt
-from envs.pricing.targetvol import Markowitz_solution
-from envs.pricing.closedforms import European_option_closed_form
-from envs.pricing.montecarlo import MC_Analisys
-
-def nu_matrix(sigma_t_vector,identity_3d, corr_chole, N_equity, Nsim):
-    S = sigma_t_vector.T*identity_3d
-    n = S.T@corr_chole
-    return n.transpose(1,2,0).reshape(N_equity,N_equity,Nsim)
-
-def alpha_dot_nu(alpha_matrix,nu_matrix, N_equity, Nsim):
-    a = np.sum((alpha_matrix.T*nu_matrix.transpose(1,0,2)),axis=1).T
-    return a
-
-
-def evolve_martingale(S_t, LV, random_gen, t_in, t_fin, cholesky, F, Nsim, N_equity):
-    logX_t = np.zeros(S_t.shape) 
-    S_t_plus_1 = np.zeros(S_t.shape)
-    Z = random_gen.randn(Nsim,N_equity)
-    ep = cholesky@Z.T
-    dt = abs(t_fin - t_in)
-   # print("Dt in the martingale ",dt)
-    sigma = np.zeros(S_t.shape)
-    for i in range(N_equity):
-        logX_t[:,i] = np.log(S_t[:,i]/F[i](t_in))
-        sigma[:,i] = LV[i](t_in, logX_t[:,i])
-        logX_t[:,i] = logX_t[:,i]-0.5*dt*(sigma[:,i]**2)+sigma[:,i]*np.sqrt(dt)*ep[i]
-        S_t_plus_1[:,i] = np.exp(logX_t[:,i])*F[i](t_fin)
-    return S_t_plus_1, sigma
-
-Seed = 13
-N_equity = 2
-frequency = "month"
-target_volatility = 5./100
-I_0 = 1.
-K = I_0     #option strike 
[…]

### 2023-02-10 tex/m2ef/dnpp_referee_response.tex
Add files via upload

\documentclass{article}
\usepackage[english]{babel}
\usepackage{xcolor}

\usepackage[normalem]{ulem}
\newcommand\soutpars[1]{\let\helpcmd\sout\parhelp#1\par\relax\relax}
\long\def\parhelp#1\par#2\relax{%
  \helpcmd{#1}\ifx\relax#2\else\par\parhelp#2\relax\fi%
}
\newcommand{\change}[1]{{\color{red} {#1}}}%{{#1}}
\newcommand{\remove}[1]{{\color{red} \soutpars{{#1}}}}%{}

\begin{document}

\noindent {\bf ``Reinforcement learning for options on target volatility funds''.\smallskip\\ Response note to referee's report.}\\

\noindent
We thank the Referee for her interesting comments and all the appreciated suggestions to improve our work. We add few remarks to their comments in the following sections and attach a revised version of the paper. We believe that these extensions address all the points highlighted in the reviews and we hope that they will be considered sufficient to recommend our work for the publication.\\

{\bf $\bullet$ Answers to referee's report}\\

We list here a detailed list of responses Referee's major remarks. We address the minor remarks directly in the text.

\begin{enumerate}

 \item \textit{[Both in Abtract and Introduction, the presentation of the problem to deal with is not well structured. As a matter of fact, first the TVS is briefly introduces, then options and TVO are only hinted, and lastly the problem is shortly described, without all these topics are well linked each other. I suggest the authors to improve this presentation.]}

 ...

 \item \textit{[The Authors address the investigated continuous stochastic control problem through two different approaches: an exact one providing the analytical solution of the problem, and an approximate one based on RL. As for the previous point, even in this one the connections between the two solution approaches are not entirely intelligible. In other terms, it is not entirely clear how the two approaches relate each other. Is the approximate one necessary given that the exact approach provides an analytical solution to the problem? What is the "utility" to use an approximate approach given the applicability of an exact one? I suggest the Authors to clarify this issue.]}

 ...

 \item \textit{[In Subsection 3.3, three (so-called intuitive) strategies are introduced, $S_A$, $S_B$ and $S_C$. No rationales are provided concerning their choices. I suggest the Author to give some details about this point.]}

 ...

 \item \textit{[In Section 4, the Authors state that ``[t]he first [RL] method is a specific algorithm developed by us to fit the problem [...]''. As know, an important feature of any RL method is its convergence to the (unknown) best policy, if any. The Authors do not spend words about this characteristic of the specific algorithm they developed and applied, so I ask them to give some detail about this topic.]}

 ...

 \item \textit{[In Subsection 5.1, the Author claim that an appropriate choice of the learning rate starting value constitutes ``a good choice in terms of [...] avoiding over-fitting''. As known, when training a FFNN, standard methods for avoiding over-fitting are based on error metrics calculated over validation sets or through $K$-fold validation approaches. So, it is not clear as the Authors can guarantee no over-fitting without exploiting such methods or similar. I suggest the Authors to carefully clarify this issue.]}

 ...

 \end{enumerate}

\end{document}

### 2023-02-10 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
Add files via upload

--- 2022-03-08 tex/reinforcement_learning_for_options_on_target_volatility_funds.tex
@@
-\documentclass[11pt]{article}
-\usepackage[english]{babel}
-\usepackage[latin1]{inputenc}
-\usepackage{lmodern}
-\usepackage[T1]{fontenc}
-\usepackage[outer=3.75cm, marginparwidth=4.5cm]{geometry}
-\usepackage{amssymb,amsmath,amsthm,latexsym,amsfonts,amscd,dsfont,enumerate,bbm,mathabx,mathtools}
-\usepackage{graphicx,tikz}
-\usepackage{lipsum}
-\usepackage{enumerate} 
-\usepackage{soul}
-\usepackage{comment}
-\usepackage{bm}
-\usepackage[font=footnotesize,labelfont=bf]{caption}
-\graphicspath{{graphs/}}
+%June 2019
 
-\definecolor{mylightyellow}{rgb}{1,1,.8}
-\definecolor{mylightgreen}{rgb}{.8,1,.8}
-\definecolor{mydarkred}{RGB}{178,34,34}
-\definecolor{mydarkgreen}{RGB}{34,139,34}
-\definecolor{mydarkblue}{RGB}{72,61,139}
-\definecolor{mydarkyellow}{RGB}{218,165,32}
+\documentclass[runningheads]{m2ef}
 
-\definecolor{myblueA}{RGB}{52,41,39}
-\definecolor{myblueB}{RGB}{92,81,109}
-\definecolor{myblueC}{RGB}{132,121,179}
-\definecolor{myblueD}{RGB}{172,161,249}
-\definecolor{myblueE}{RGB}{212,201,255}
+\usepackage[]{latexsym}
+\usepackage{graphicx}
 
-\usetikzlibrary{arrows,patterns,shapes,snakes,shadows,fit,backgrounds,positioning,decorations.pathmorphing,external}
-\tikzstyle{nameusd} = [circle, draw, top color=white, bottom color=mydarkblue!50, draw=mydarkblue!75!black!100, drop shadow, minimum height=4em]
-\tikzstyle{nameeur} = [circle, draw, top color=white, bottom color=mydarkred!50, draw=mydarkred!75!black!100, drop shadow, minimum height=4em]
-\tikzstyle{collusd} = [rectangle,fill=mydarkblue!10, inner sep=0.2cm, rounded corners=5mm]
-\tikzstyle{colleur} = [rectangle,fill=mydarkred!10, inner sep=0.2cm, rounded corners=5mm]
-\tikzstyle{fixto} = [draw, -latex']
-\tikzstyle{fixfrom} = [draw, latex'-]
-\tikzstyle{floatto} = [draw, snake=coil, segment aspect=0, line before snake=2ex, line after snake=1ex, -latex']
-\tikzstyle{floatfrom} = [draw, snake=coil, segment aspect=0, line before snake=2ex, line after snake=1ex, latex'-]
-\tikzstyle{investor} = [rectangle, draw, top color=white, bottom color=mydarkyellow!50, draw=mydarkyellow!75!black!100, drop shadow, rounded corners, minimum height=3em, text width=4em, text centered]
-\tikzstyle{trader} = [circle, draw, top color=white, bottom color=blue!30, draw=blue!50!black!100, drop shadow, minimum height=4em]
-\tikzstyle{bank} = [rectangle, draw, top color=white, bottom color=red!20, draw=red!50!black!100, drop shadow, rounded corners, minimum height=3em, text width=4em, text centered]
-\tikzstyle{market} = [rectangle, draw, top color=white, bottom color=green!20, draw=green!50!black!100, drop shadow, rounded corners, minimum height=3em, text width=4em, text centered]
-\tikzstyle{yellowbox} = [rectangle, draw, top color=white, bottom color=mydarkyellow!50, draw=mydarkyellow!75!black!100, drop shadow, rounded corners, text centered]
-\tikzstyle{fadedyellowbox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkyellow!25, draw=mydarkyellow!50, drop shadow, rounded corners, text centered]
-\tikzstyle{redbox} = [rectangle, draw, top color=white, bottom color=mydarkred!50, draw=mydarkred!75!black!100, drop shadow, rounded corners, text centered]
-\tikzstyle{fadedredbox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkred!25, draw=mydarkred!50, drop shadow, rounded corners, text centered]
-\tikzstyle{bluebox} = [rectangle, draw, top color=white, bottom color=mydarkblue!50, draw=mydarkblue!75!black!100, drop shadow, rounded corners, text centered]
-\tikzstyle{fadedbluebox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkblue!25, draw=mydarkblue!50, drop shadow, rounded corners, text centered]
-\tikzstyle{greenbox} = [rectangle, draw, top color=white, bottom color=mydarkgreen!50, draw=mydarkgreen!75!black!100, drop shadow, rounded corners, text centered]
-\tikzstyle{fadedgreenbox} = [rectangle, draw=gray, text=gray, top color=white, bottom color=mydarkgreen!25, draw=mydarkgreen!50, drop shadow, rounded corners, text centered]
-
-\newcommand{\tikzmark}[1]{\tikz[remember picture] \node[coordinate] (#1) {#1};}
-
-% \pgfplotstableread{data/total_consumption_example.dat}\swingtotal
-
-\usepackage[]{hyperref}
-\hypersetup{
-	breaklinks, 
-	baseurl = http://,
-	pdfborder = 0 0 0,
-	pdfpagemode = UseNone, % do not show thumbnails or bookmarks on opening
-	pdfstartpage = 1,
-	colorlinks=true,breaklinks=true,
-	urlcolor= mydarkblue,linkcolor= mydarkblue,citecolor= mydarkgreen,
-	pdfauthor={Daluiso, R., Nastasi, E., Pallavicini, A., Polo, S.},
-	pdftitle = {Reinforcement Learning for Options on Target Volatility Funds},
-	pdfsubject = {Reinforcement Learning for Options on Target Volatility Funds}
-	pdfkeywords = {Reinforcement Learning, Hedging Costs, Funding Costs, Proximal Policy Optimization, Target Volatility, Asset Allocation, Stochastic Control Problem}
-}
-
-\usepackage[backend=biber,style=authoryear-comp,maxcitenames=2,maxbibnames=99,sorting=nyt,sortcites=false, natbib=true]{biblatex}
-\addbibresource{Bib.bib}
-
-\AtEveryBibitem{\clearfield{month}}
-\AtEveryCitekey{\clearfield{month}}
-\setlength\bibitemsep{1.3\itemsep}
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{remark}{Remark}[section]
-\newtheorem{example}[remark]{Example}
-
-\newcommand{\Eq}[1]{{\begin{equation}{#1}\end{equation}}}
-
-\newcommand{\Exo}[1]{\mathbb{E}\!\left[\,#1\,\right]}
-\newcommand{\Ex}[2]{\mathbb{E}_{#1}\!\left[\,#2\,\right]}
-\newcommand{\ExG}[2]{\mathbb{E}\!\left[\,#2\,\left|{\cal G}_{#1}\right.\right]}
-\newcommand{\ExF}[2]{\mathbb{E}\!\left[\,#2\,\left|{\cal F}_{#1}\right.\right]}
-\newcommand{\ExFT}[3]{\mathbb{E}^{#2}\!\left[\,#3\,\left.\right|{\cal F}_{#1}\right]}
-\newcommand{\ExGT}[3]{\mathbb{E}^{#2}\!\left[\,#3\,\left.\right|{\cal G}_{#1}\right]}
-\newcommand{\ExCo}[2]{\mathbb{E}\!\left[\left.\,#1\,\right|\,#2\,\right]}
-\newcommand{\ExC}[3]{\mathbb{E}_{#1}\!\left[\left.\,#2\,\right|\,#3\,\right]}
-\newcommand{\ExCT}[4]{\mathbb{E}_{#1}^{#2}\!\left[\left.\,#3\,\right|\,#4\,\right]}
-\newcommand{\ExT}[3]{\mathbb{E}_{#1}^{#2}\!\left[\,#3\,\right]}
-\newcommand{\Exf}[2]{\hat{\mathbb{E}}_{#1}\!\left[\,#2\,\right]}
-\newcommand{\ExfT}[3]{\hat{\mathbb{E}}_{#1}^{#2}\!\left[\,#3\,\right]}
-\newcommand{\Px}[1]{\mathbb{P}\left\{\,#1\,\right\}}
-\newcommand{\PxC}[2]{\mathbb{P}\!\left\{\,#1\,\left.\right|\,#2\,\right\}}
-\newcommand{\Qx}[1]{\mathbb{Q}\left\{\,#1\,\right\}}
-\newcommand{\QxC}[2]{\mathbb{Q}\!\left\{\,#1\,\left.\right|\,#2\,\right\}}
-\newcommand{\QxCT}[3]{\mathbb{Q}^{#1}\!\left\{\,#2\,\left.\right|\,#3\,\right\}}
-\newcommand{\QxT}[2]{\mathbb{Q}^{#1}\left\{\,#2\,\right\}}
-\newcommand{\Qxf}[1]{\hat{\mathbb{Q}}\left\{\,#1\,\right\}}
-\newcommand{\Var}[2]{\operatorname{Var}_{#1}\!\left[\,#2\,\right]}
-\newcommand{\Covar}[3]{\operatorname{Covar}_{#1}\!\left[\,#2,#3\,\right]}
-\newcommand{\ind}[1]{1_{\{#1\}}}
-\newcommand{\clip}[3]{\operatorname{clip}\!\left({#1},{#2},{#3}\right)}
-\newcommand{\one}{\ensuremath{\mathbbm{1}}}
-\newcommand{\onehalf}{\tfrac{1}{2}}
-
+\usepackage{amssymb,amsmath,latexsym,amscd,dsfont,enumerate,bbm,mathabx,mathtools}
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator{\Tr}{Tr}
 \newcommand{\tr}[1]{{#1}^{\intercal}} %transpose
-\DeclareMathOperator{\Tr}{Tr}
+\newcommand{\quotations}[1]{``#1''}
+\newenvironment{eqsys}{\begin{equation}\begin{dcases}}{\end{dcases}\end{equation}}
 \let\oldnorm\norm   % <-- Store original \norm as \oldnorm
@@
 
-\newcommand{\rec}{R}
-\newcommand{\lgd}{\mbox{L{\tiny GD}}}
-\newcommand{\virgolette}[1]{``#1''}
-\newcommand{\pvcall}{{\mathcal{C}}}
-\newcommand{\pvput}{{\mathcal{P}}}
-\newcommand{\rom}[1]{\uppercase\expandafter{\romannumeral #1\relax}}
+\usepackage[normalem]{ulem}
+\newcommand\soutpars[1]{\let\helpcmd\sout\parhelp#1\par\relax\relax}
+\long\def\parhelp#1\par#2\relax{%
+  \helpcmd{#1}\ifx\relax#2\else\par\parhelp#2\relax\fi%
+}
+\newcommand{\change}[1]{{\color{red} {#1}}}%{{#1}}
+\newcommand{\remove}[1]{{\color{red} \soutpars{{#1}}}}%{}
+\newcommand{\changeB}[1]{{\color{blue} {#1}}}%{{#1}}
+\newcommand{\removeB}[1]{{\color{blue} \soutpars{{#1}}}}%{}
 
-\usepackage{eurosym}
-\newcommand{\euroms}[0]{\mbox{\small \euro}}
-\usepackage{cleveref}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\DeclareMathOperator*{\argmin}{arg\,min}
+\begin{document}
 
-\newenvironment{eqsys}{\begin{equation}\begin{dcases}}{\end{dcases}\end{equation}}
-\usepackage{fancyhdr}
-\renewcommand{\headrulewidth}{0pt}
-\fancyhead[R]{}
-\lhead{\footnotesize\slshape Daluiso, Nastasi, Pallavicini, Polo, Optimal Strategies for Target Volatility Contracts}
-\chead{}
-\rhead{\thepage}
-\lfoot{}
-\cfoot{}
-\rfoot{}
-\date{}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\pagestyle{myheadings} \setcounter{page}{1}
 
-\title{Reinforcement Learning for Options on \\Target Volatility Funds}
+\mainmatter
 
-\author{
-	Roberto Daluiso\thanks{Intesa Sanpaolo Group Milano, \texttt{roberto.daluiso@intesasanpaolo.com}}
-	\and 
-	Emanuele Nastasi\thanks{Marketz S.p.A. Milano, \texttt{emanuele.nastasi@marketz.eu}}
-	\and Andrea Pallavicini\thanks{Intesa Sanpaolo Group Milano, \texttt{andrea.pallavicini@intesasanpaolo.com}} 
-	\and
-	Stefano Polo\thanks{X-Numeris S.r.l. Milano, \texttt{sp.stefano.polo@gmail.com}} 
-}
+\title{Reinforcement learning for options\\on target volatility funds\thanks{Supported by ISCRA from Cineca, \email{iscra@cineca.it}.}}
 
-\begin{document} % goes here
+\titlerunning{Reinforcement learning for options on target volatility funds}
+
+\author{Roberto Daluiso\inst{1}
+\and Emanuele Nastasi\inst{2} \and
+Andrea Pallavicini\inst{1} \and Stefano Polo\inst{3}}
+
+\authorrunning{R. Daluiso, E. Nastasi, A. Pallavicini, S. Polo}
+
+\institute{Financial Engineering, Intesa Sanpaolo Group,\\
+Largo Mattioli 3, 20121 Milan, Italy\\
+\email{\{Roberto.Daluiso, Andrea.Pallavicini\}@intesasanpaolo.com} \and
+Financial Engineering, Marketz S.p.A.,\\
+Corso Europa, 2, 20122 Milan, Italy\\
+\email{Emanuele.Nastasi@marketz.eu} \and Financial Engineering,
+X-Numeris S.r.l.,\\ Via Turro, 7 , 20127 Milan, Italy\\
+\email{Stefano.Polo@xnumeris.com}}
 
@@
 
+{\let\thefootnote\relax\footnote{\vskip1mm$\!\!\!\!\!\!\!\!\!\!\!\!\!${\fbox{{\parbox[c][1.0cm]{12cm}{\center{\textsc{\vskip-4.5mm \textbf{Mathematical Methods in Economics and Finance}}} \textbf{-- m$\mathbf{^2}$ef} \\ Vol.
+$\ldots$, No.
+$\ldots$, $\ldots$\\}}}}}} \setcounter{footnote}{0}
+
 \begin{abstract}
-\footnotesize	
-\noindent
-In this work we deal with the funding costs rising from hedging the risky securities underlying a target volatility strategy (TVS), a portfolio of risky assets and a risk-free one dynamically rebalanced in order to keep the realized volatility of the portfolio on a certain level.
+We deal with the funding costs rising from hedging the risky securities underlying a target volatility strategy (TVS), a portfolio of risky assets and a risk-free one dynamically rebalanced in order to keep the realized volatility of the portfolio on a certain level.
 The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs for each component requires to solve a control problem to evaluate the option prices.
 We derive an analytical solution of the problem in the Black and Scholes (BS) scenario.
-Then we use Reinforcement Learning (RL) techniques to determine the fund composition leading to the most conservative price under the local volatility (LV) model, for which an \textit{a priori} solution is not available.
+Then we use Reinforcement Learning (RL) techniques to determine the fund composition leading to the most conservative price under the local volatility (LV) model, for which an {\it a priori} solution is not available.
 We show how the performances of the RL agents are compatible with those obtained by applying path-wise the BS analytical strategy to the TVS dynamics, which therefore appears competitive also in the LV scenario.
+
+\smallskip
+
+\noindent {\bf Keywords.} Reinforcement learning, target volatility strategy, hedging costs.
+
+\smallskip
+
+\noindent {\bf M.S.C. classification.} \textsc{\small 65C05, 68T07, 91G20.}
+
+\noindent {\bf J.E.L. classification.} \textsc{\small C63, C45, G13.}
 \end{abstract}
-\smallskip
-\noindent
-\textbf{JEL classification codes:} C63, C45, G13.\\
-\textbf{AMS classification codes:} 65C05, 68T07, 91G20.\\
-\textbf{Keywords:} Reinforcement Learning, Hedging Costs, Target Volatility, Asset Allocation, Stochastic Optimal Control Problem. 
-\newpage
-\tableofcontents
- \mbox{}
- \vfill
- \noindent \small The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.
 
-\newpage
-\pagestyle{myheadings} \markboth{}{{\footnotesize Daluiso, Pallavicini, Nastasi, Polo, Reinforcement Learning for Options on Target Volatility Funds}}
+\bigskip
 
@@
 
-These products were initially offered in the Asian markets, see for instance the reports of \textcite{Chew2011} and \textcite{Xue2012}, which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \textcite{Morrison2013}.
+These products were initially offered in the Asian markets, see for instance the reports of \cite{Chew2011} and \cite{Xue2012}, which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \cite{Morrison2013}.
 At the present day we can observe some new market indices based on the mechanism of the target volatility strategy such as Dow Jones Volatility Control Index, and S\&P 500 Risk Control Index.
 
-In the recent literature TVSs are tested to investigate their performances in term of realized returns, see for instance \textcite{Hocquard2013} and \textcite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \textcite{Kim2018}.
+In the recent literature TVSs are tested to investigate their performances in term of realized returns, see for instance \cite{Hocquard2013} and \cite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim2018}.
 Moreover, in the present day, in the derivative market new structured financial products that see TVSs as underlying are becoming widely used by practitioners.
-Those derivative instruments are known as target volatility options (TVOs) and have gained great interest in the derivative pricing literature, see in particular  \textcite{DiGraziano2012}, \textcite{Grasselli2016}, \textcite{Albeverio2019}, and \textcite{DiPersio2019}.
+Those derivative instruments are known as target volatility options (TVOs) and have gained great interest in the derivative pricing literature, see in particular \cite{Albeverio2019}, \cite{DiGraziano2012}, \cite{DiPersio2019}, and \cite{Grasselli2016}.
 
@@
 Hence, the pricing problem becomes a continuous dynamical control problem over the risky portfolio composition.
-In our work we provide to the reader a formal description of the stochastic control problem and most importantly we derive an analytical solution assuming that the risky asset dynamics underlying the TVS  follow a Black and Scholes (BS) model \citep{BlackScholes1973} and that the derivative contract is a European-style option.
+In our work we provide to the reader a formal description of the stochastic control problem and most importantly we derive an analytical solution assuming that the risky asset dynamics underlying the TVS  follow a Black and Scholes (BS) model \cite{BlackScholes1973} and that the derivative contract is a European-style option.
 
 Although from a theoretical point of view this solution represents a sub-optimal strategy when dealing with generic dynamics for the risky assets, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem.
-More precisely we numerically study the problem in the general case of a local volatility dynamics \citep{Dupire1994,Derman1994}  by adopting a Reinforcement Learning approach, in particular an \textit{ad hoc} vanilla direct policy algorithm and the more sophisticated proximal policy optimization technique developed in \textcite{Schulman2017}.
+More precisely we numerically study the problem in the general case of a local volatility dynamics \cite{Derman1994}, \cite{Dupire1994}  by adopting a Reinforcement Learning approach, in particular an \textit{ad hoc} vanilla direct policy algorithm and the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}.
 \\
@@
 Moreover we wish to thank the Italian computing centre Cineca, which approved our ISCRA C project and provided us with the high-computing resources of Marconi100 for the numerical simulations of this contribution.
+\\
 
-\section{Target Volatility Strategy}\label{sec:TVS}
+The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.
+
+\section{Target volatility strategy}\label{sec:TVS}
 As mentioned before, this work aims to enrich the TVS pricing literature by studying the aspect related to the funding costs coming from hedging the risky assets underlying a target volatility portfolio.
@@
 \end{equation}
+
 In this way, the TVS is only defined in the two contracts client-fund and fund-bank.
@@
 where $\pi^i_t$ represents the cumulative contractual-coupon process paid by the security, and $\psi_t^i$ represents the cumulative valuation adjustments.
-Since fund managers allocating TVS usually rely on equity assets, here we use the results of \textcite{Gabrielli2020} that analyzes the valuation adjustments for equity products. We can write 
+Since fund managers allocating TVS usually rely on equity assets, here we use the results of \cite{Gabrielli2020} that analyzes the valuation adjustments for equity products. We can write 
 \begin{equation}
@@
 
-\section{Derivative Pricing}\label{sec:Derivative}
+\section{Derivative pricing}\label{sec:Derivative}
 In this section, we analyze derivative contracts linked to the TVS described previously.
@@
 
-We start by introducing the Markovian projection of the dynamics followed by $I_t$. We name it $I_t^{\text{MP}}$, and we get by applying the Gy\"ongy Lemma \citep{Gyongy1986}
+We start by introducing the Markovian projection of the dynamics followed by $I_t$. We name it $I_t^{\text{MP}}$, and we get by applying the Gy\"ongy Lemma \cite{Gyongy1986}
 \begin{equation}
@@
 \begin{equation}
-    V_{T_{k-1}}\left(X_{T_{k-1}}, I_{T_{k-1}}\right)=\mathbb{E}_{T_{k-1}}\left[D\left(T_{k-1},T_{k}\right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right],
+\begin{aligned}
+	    V_{T_{k-1}}\left(X_{T_{k-1}}, I_{T_{k-1}}\right)= \\ 
+	     \mathbb{E}_{T_{k-1}}\left[D\left(T_{k-1},T_{k}\right) V_{T_{k}}\left(X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\right) \mid X_{T_{k-1}}, I_{T_{k-1}}\right] \; ,
+\end{aligned}
 \label{eq:recursion2}\end{equation}
@@
 
-\subsection{Black and Scholes Model}\label{subsec:BS_model}
+\subsection{Black and Scholes model}\label{subsec:BS_model}
 In the time-dependent Black and Scholes model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$.
@@
 \end{itemize}
-As the reader can observe, our strategy outperforms any other intuitive one a practitioner might adopt.\begin{figure}[t]
+As the reader can observe, our strategy outperforms any other intuitive one a practitioner might adopt.
+\begin{figure}[t]
 	\centering
-	\includegraphics[height=5cm,width=10cm]{strategy_price_comparison.pdf}
-	\caption{Comparison of plain vanilla prices on the TVS by adopting different allocation strategies: the optimal Black and Scholes solution (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and other intuitive strategies (S$_A$, S$_B$, S$_C$).}
+	\includegraphics[width=4in]{strategy_price_comparison_grey.eps}
+	\caption{Comparison of plain vanilla prices on the TVS by adopting different allocation strategies: the optimal Black and Scholes solution (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and other intuitive strategies (S$_A$, S$_B$, S$_C$)}
 	\label{fig:strategy_price_comparison}
@@
 \label{eq:analytical_solution_free_strategy}\end{equation}
-We take the minus sign to get the minimum value of the TVS local drift while the plus sign for the maximum one (put payoff).
+We take the minus sign to get the minimum value of the TVS local drift while the plus sign for the maximum one (put payoff). $\Box$
 \end{proof}
@@
 	\min_{i\leq n} \frac{\mu_i}{\sqrt{\Sigma_{ii}}} \; . \; 
-	\end{equation}
+	\end{equation} $\Box$
 \end{proof}
 
-\subsection{Hamilton-Jacobi-Bellman Equation for Target Volatility Options}\label{subsec:HJB}
+\subsection{Hamilton-Jacobi-Bellman equation for target volatility options}\label{subsec:HJB}
 In this section, we want to provide to the reader a formal description of the dynamic problem associated with options on target volatility strategies by writing the Hamilton-Jacobi-Bellman (HJB) equation for the derivative price.
@@
 
-\section{Reinforcement Learning}\label{sec:RL}
+\section{Reinforcement learning}\label{sec:RL}
 As we have discussed in the previous sections, one must resort to numerical approaches to solve the stochastic control problem related to the TVS in the case of general payoffs or risky securities dynamics.
-The standard approach could be to use classical techniques based on backwards recursion \eqref{eq:recursion}-\eqref{eq:recursion2} such as American Monte Carlo \citep{Longstaff2001}.
+The standard approach could be to use classical techniques based on backwards recursion \eqref{eq:recursion}-\eqref{eq:recursion2} such as American Monte Carlo \cite{Longstaff2001}.
 However, their performances degrade exponentially as the dimension $n$ of the problem increases, making prohibitively costly finding the solution to the problem.
-In our contribution, we adopt a novel technique which is free from the curse of dimensionality and is gaining popularity in many scientific branches for solving stochastic optimal control problems: Reinforcement Learning (RL) \citep{Sutton2018}.
+In our contribution, we adopt a novel technique which is free from the curse of dimensionality and is gaining popularity in many scientific branches for solving stochastic optimal control problems: Reinforcement Learning (RL) \cite{Sutton2018}.
 
@@
 Thus RL is independent of pre-collected data as opposed to other Machine Learning techniques.
-Because of its nature, RL has been successful in quantitative finance for solving control problems; among the most important RL applications in this field, we refer to \textcite{Deng2017} as the pioneers in studying self-taught reinforcement trading problems, while to \textcite{Kolm2019} and \textcite{Halperin2020}  for hedging derivatives with RL under market frictions.
+Because of its nature, RL has been successful in quantitative finance for solving control problems; among the most important RL applications in this field, we refer to \cite{Deng2017} as the pioneers in studying self-taught reinforcement trading problems, while to \cite{Halperin2020} and \cite{Kolm2019} for hedging derivatives with RL under market frictions.
 
-In our work we adopt two learning strategies to compare their performances in terms of overall reward: an \textit{ad hoc} direct policy algorithm and the state of the art proximal policy optimization (PPO) developed in \textcite{Schulman2017} and \textcite{Schulman2016}.
+In our work we adopt two learning strategies to compare their performances in terms of overall reward: an \textit{ad hoc} direct policy algorithm and the state of the art proximal policy optimization (PPO) developed in \cite{Schulman2017} and \cite{Schulman2016}.
 
@@
 We will not provide a complete description of this sophisticated learning strategy; for more details, we refer to the authors' papers.
-In our work, we adopt the PPO implementation found in \textsc{OpenAI Baselines}.\footnote{\url{https://github.com/openai/baselines}.}
+In our work, we adopt the PPO implementation found in \textsc{OpenAI Baselines}.\footnote{\url{https://github.com/opernai/baselines}.}
 
@@
 
-\subsection{Direct Policy Approach}\label{sec:direct_policy}
+\subsection{Direct policy approach}\label{sec:direct_policy}
 We consider an episode $\tau$ of length $m+1$ that takes place on a discrete time-grid of fixing times expressed in year fractions $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$ maturity of the option.
@@
 
-\subsection{Proximal Policy Optimization Approach}
+\subsection{Proximal policy optimization approach}
 As for the direct policy approach, we model the pricing problem considering an episode that takes place on the time-grid $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$.
@@
 The hyper-parameter $\gamma$ plays the role of a discount factor in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future.
-If we take the cumulative reward per episode and set the PPO parameters\footnote{We refer to \textcite{Schulman2016} for a more detailed description for the generalized advantage estimation parameter $\lambda$.} $\gamma =\lambda =1$ we obtain
+If we take the cumulative reward per episode and set the PPO parameters\footnote{We refer to \cite{Schulman2016} for a more detailed description for the generalized advantage estimation parameter $\lambda$.} $\gamma =\lambda =1$ we obtain
 \begin{equation}
@@
 
-\section{Numerical Investigations}\label{sec:Numerical_investigations}
+\section{Numerical investigations}\label{sec:Numerical_investigations}
 Here we present the numerical results obtained with our proposed methods. We focus our analysis on a European call option on a TVS with the following product details
@@
 This is due to the fact that in this Reinforcement Learning problem the number of the input neurons is equal to the state space dimension \eqref{eq:state}, while the number of the output ones coincides with the action space dimension \eqref{eq:action}.
-It is well known in the literature that neural networks give better performances in the training phase if the input data are well normalized \citep{Sola1997,Puheim2014} .
+It is well known in the literature that neural networks give better performances in the training phase if the input data are well normalized \cite{Puheim2014}, \cite{Sola1997} .
 Thus we choose as state $s_k$ the following normalized block
@@
 
-\subsection{Black and Scholes: Hyper-Parameters Fine Tuning}
+\subsection{Black and Scholes: hyper-parameters fine tuning}
 We use the BS environment as toy model to understand which parameters of the RL algorithm play key roles in the training and testing phase. 
@@
 We try to investigate the following hyper-parameters: the FFNN architecture, in particular which feature between the depth and the width of the network is more important, the activation function and the learning rate of the optimizer.
-Moreover, we compare the performances of two well-known optimizers in Machine Learning literature: Nadam \citep{Dozat2016} and RMSprop \citep{Hiton2012}.
+Moreover, we compare the performances of two well-known optimizers in Machine Learning literature: Nadam \cite{Dozat2016} and RMSprop \cite{Hiton2012}.
 Since we deal with a free allocation strategy $\alpha$ that can assume negative values, we have chosen among the wide variety of activation functions the tanh and the elu.
@@
 	\centering
-	\includegraphics[width=7.2cm,height=6cm]{BS_fine_tuning_Nadam_tanh.pdf}
-	\includegraphics[width=7.2cm,height=6cm]{BS_fine_tuning_Nadam_elu.pdf}
+	\includegraphics[width=3.4in]{BS_fine_tuning_Nadam_tanh_grey.eps}
+	\includegraphics[width=3.4in]{BS_fine_tuning_Nadam_elu_grey.eps}
 	\caption{Learning curves of the direct policy algorithm applied to the Black and Scholes model.
 The solid lines are the loss functions defined in Equation~\eqref{eq:direct_policy_loss} in function of the number of training epochs.
-Each continuous line represents a different neural network architecture with tanh activation function (left) and elu activation function (right).
+Each continuous line represents a different neural network architecture with tanh activation function (top) and elu activation function (bottom).
 The optimizer adopted is the Nadam with a learning rate of $10^{-3}$.
-The horizontal red-dashed line is the conservative option price obtained by maximizing the TVS drift through Equation~\eqref{eq:analytical_solution_free_strategy}.}
+The horizontal grey-dashed line is the conservative option price obtained by maximizing the TVS drift through Equation~\eqref{eq:analytical_solution_free_strategy}}
 	\label{fig:result_bs_nadam}
@@
 	\centering
-	\includegraphics[width=7.2cm,height=6cm]{BS_fine_tuning_RMSprop_tanh.pdf}
-	\includegraphics[width=7.2cm,height=6cm]{BS_fine_tuning_RMSprop_elu.pdf}
+	\includegraphics[width=3.4in]{BS_fine_tuning_RMSprop_tanh_grey.eps}
+	\includegraphics[width=3.4in]{BS_fine_tuning_RMSprop_elu_grey.eps}
 	\caption{Learning curves of the direct policy algorithm applied to the Black and Scholes model.
 The solid lines are the loss values defined in Equation~\eqref{eq:direct_policy_loss} in function of the number of training epochs.
-Each continuous line represents a different neural network architecture with tanh activation function (left) and elu activation function (right).
+Each continuous line represents a different neural network architecture with tanh activation function (top) and elu activation function (bottom).
 The optimizer adopted is the RMSprop with a learning rate of $10^{-3}$.
-The horizontal red-dashed line is the conservative option price obtained by maximizing the TVS drift through Equation~\eqref{eq:analytical_solution_free_strategy}.}
+The horizontal grey-dashed line is the conservative option price obtained by maximizing the TVS drift through Equation~\eqref{eq:analytical_solution_free_strategy}}
 	\label{fig:result_bs_rmsprop}
@@
 The three lines of each plot display the learning curves for different FFNN architectures: a one hidden layer network with 50 neurons, a 2 hidden layers with 30 and 20 neurons respectively, and a deeper one with three layers with 20, 15, and 5 neurons.
-Each learning curve is compared with the optimal option price (red-dashed line) that we have computed with the closed form solution we have derived for the BS model in Section~\ref{subsec:BS_model}.
+Each learning curve is compared with the optimal option price (grey-dashed line) that we have computed with the closed form solution we have derived for the BS model in Section~\ref{subsec:BS_model}.
 We can see that all the learning curves converge to our expected price, providing a numerical demonstration of our theoretical result.
@@
 \begin{center}
+\caption{Comparison of TVO prices under Black and Scholes model: analytical solution price and direct policy reinforcement learning.
+The option parameters are: $I_0=K=1\,[\text{EUR}]$, $T=2\,[\text{yr}]$ and $\bar{\sigma}=5\%$}
+\label{tab:mc_results_bs}
 \begin{tabular}{ll}
@@
 \end{center}
-\caption{Comparison of TVO prices under Black and Scholes model: analytical solution price and direct policy reinforcement learning.
-The option parameters are: $I_0=K=1\,[\text{EUR}]$, $T=2\,[\text{yr}]$ and $\bar{\sigma}=5\%$.}
-\label{tab:mc_results_bs}
 \end{table}
@@
 
-\subsection{Local Volatility Dynamics}
+\subsection{Local volatility dynamics}
 In this section, we study the TVS control problem assuming a local volatility model for the dynamics of the risky assets.
@@
 Here we consider the same market data ($r_t$, $\mu_t$ and $\nu_t$) as in the Black and Scholes environment to study how the optimal solution changes with the dynamics model.
+
 \begin{figure}[h!]
 	\centering
-	\includegraphics[width=9cm,height=6cm]{LV_direct_policy_approach.pdf}
+	\includegraphics[width=4in]{LV_direct_policy_approach_grey.eps}
 	\caption{Learning curves of the direct policy algorithm applied to the local volatility model.
-The solid blue line is the learning curve of an agent with free actions, while the green solid one is the learning curve of an agent whose actions are parameterized with the maximum drift strategy.
+The solid line with triangles is the learning curve of an agent with free actions, while the solid one is the learning curve of an agent whose actions are parameterized with the maximum drift strategy.
 The two agents are parameterized by a $20 \times 10 \times 5$ FFNN with tanh activation function.
-The loss function is computed according to \Cref{eq:direct_policy_loss}.
-The optimizer adopted is the RMSprop with a learning rate of $10^{-3}$ for the blue line while $10^{-4}$ for the green one.
-The horizontal red-dashed lines are the 99\% confidence interval of the MC price obtained by maximizing path-wise the TVS drift through \Cref{eq:analytical_solution_free_strategy}.}
+The loss function is computed according to Equation~\eqref{eq:direct_policy_loss}.
+The optimizer adopted is the RMSprop with a learning rate of $10^{-3}$ for the solid line with triangles while $10^{-4}$ for the solid one.
+The horizontal grey-dashed lines delimit the 99\% confidence interval of the MC price obtained by maximizing path-wise the TVS drift through Equation~\eqref{eq:analytical_solution_free_strategy}}
 	\label{fig:result_lv_direct_policy}
@@
 Thus we train by direct policy algorithm a deep $20\times 10\times 5$ neural network with tanh activation function and a RMSprop optimizer.
-We report in Figure~\ref{fig:result_lv_direct_policy} the corresponding learning curve (solid blue line).
+We report in Figure~\ref{fig:result_lv_direct_policy} the corresponding learning curve (solid line with triangles).
 From this, we observe that the network has learnt some good policy since the curve grows as the number of training epochs increases until it saturates at a certain value.
-To try to measure the performance of the policy selected by the agent, we can build a n\"aif strategy called \virgolette{baseline}.
+To try to measure the performance of the policy selected by the agent, we can build a n\"aif strategy called \quotations{baseline}.
 By looking at the Markovian projection in \eqref{eq:markovian_projection}-\eqref{eq:local_drift} and the first term in the HJB \eqref{eq:HJB_general}, a natural choice for the baseline is to maximize the TVS local drift.
@@
 With this parameterization, at each observational time $T_k\in\mathcal{T}$ the risky allocation strategy is obtained by summing the network output with the Equation~\eqref{eq:analytical_solution_free_strategy}.
-Even in this case, the corresponding learning curve (solid green line in Figure~\ref{fig:result_lv_direct_policy}) is stuck in the maximum drift strategy.
+Even in this case, the corresponding learning curve (solid black line in Figure~\ref{fig:result_lv_direct_policy}) is stuck in the maximum drift strategy.
 The first possible reason for this behaviour is that the learning strategy of the direct policy approach is too simple for the learning task.
@@
 	\centering
-	\includegraphics[height=6cm,width=7.2cm]{LV_PPO_reward_learning_curve.png}
-	\includegraphics[height=6cm,width=7.2cm]{LV_PPO_new_reward_learning_curve.png}
+	\includegraphics[width=3.4in]{LV_PPO_reward_a_learning_curve.eps}
+	\includegraphics[width=3.4in]{LV_PPO_reward_b_learning_curve.eps}
 	\caption{Learning curves of the PPO algorithm applied to the local volatility model.
@@
 The label {\it reward} ({\it a}) indicates that the RL agent is trained in an environment where the reward function is defined by Equation~\eqref{eq:reward1}, while {\it reward} ({\it b}) refers to the reward function in Equation~\eqref{eq:reward2}.
-The solid blue lines are the learning curves of an agent with free actions, while the green solid ones are the learning curves of an agent whose actions are parameterized with the maximum drift strategy.
+The solid black lines are the learning curves of an agent with free actions, while the grey solid ones are the learning curves of an agent whose actions are parameterized with the maximum drift strategy.
 In both plots, the agents are parameterized by an FFNN with 5 hidden layers with 5 neurons each with tanh activation function.
-The horizontal red-dashed lines delimit the 99\% confidence interval of the MC price obtained by maximizing path-wise the TVS drift through Equation~\eqref{eq:analytical_solution_free_strategy}.}\label{fig:PPO_results}
+The horizontal grey-dashed lines delimit the 99\% confidence interval of the MC price obtained by maximizing path-wise the TVS drift through Equation~\eqref{eq:analytical_solution_free_strategy}}\label{fig:PPO_results}
 \end{figure}
@@
 \begin{center}
+\caption{Comparison of TVO prices under local volatility model.
+The baseline price is obtained by applying path-wise the maximum drift strategy \eqref{eq:analytical_solution_free_strategy}.
+The RL free agent implements the policy defined by Equation~\eqref{eq:action}, while the parameterized agent chooses its actions in terms of the baseline strategy.
+We label with letter (a) the agents trained in environments with the reward function \eqref{eq:reward1}, while with (b) the reward function \eqref{eq:reward2}.
+The option parameters are: $I_0=K=1\,[\text{EUR}]$, $T=2\,[\text{yr}]$ and $\bar{\sigma}=5\%$}
+\label{tab:mc_results_lv}
 \begin{tabular}{ll}
@@
 \end{tabular}
-\caption{Comparison of TVO prices under local volatility model.
-The baseline price is obtained by applying path-wise the maximum drift strategy \eqref{eq:analytical_solution_free_strategy}.
-The RL free agent implements the policy defined by Equation~\eqref{eq:action}, while the parameterized agent chooses its actions in terms of the baseline strategy.
-We label with letter (a) the agents trained in environments with the reward function \eqref{eq:reward1}, while with (b) the reward function \eqref{eq:reward2}.
-The option parameters are: $I_0=K=1\,[\text{EUR}]$, $T=2\,[\text{yr}]$ and $\bar{\sigma}=5\%$.}
-\label{tab:mc_results_lv}
 \end{center}
@@
 
-\section{Conclusion and Further Developments}
+\section{Conclusion and further developments}
 In this paper, we described a non-trivial control problem related to derivative contracts on target volatility strategies.
@@
 
-\printbibliography
+\begin{thebibliography}{99}
+%
+%
+\bibitem{Albeverio2019}
+Alberverio, S., Victoria, S., and Wallbaum K.: The volatility target effect in investment-linked products with embedded American-type derivatives.
+Investment Management and Financial Innovations {\bf 16-3} (2019) 18-28
+%
+\bibitem{BlackScholes1973}
+Black, F. and Scholes, M.: The Pricing of Options and Corporate Liabilities. The Journal of Political Economy {\bf 81-3} (1973) 637-654
+%
+\bibitem{Chew2011}
+Chew, L.: Target volatility asset allocation strategy. Society of Actuaries, International News (2011)
+%
+\bibitem{Deng2017}
+Deng, Y., Bao, F., Kong, Y., Ren, Z., and Dai, Q.: Deep direct reinforcement learning for financial signal representation and trading.
+IEEE Transactions on Neural Networks and Learning Systems {\bf 28-3} (2017) 653-664
+%
+\bibitem{Derman1994}
+Derman, E. and Kani, I.: Riding on a Smile. Risk {\bf 7} (1994) 32-39
+%
+\bibitem{DiGraziano2012}
+Di Graziano, G., and Torricelli, L.: Target volatility option pricing. International Journal of Theoretical and Applied Finance {\bf 15-1} (2012) 1-17
+%
+\bibitem{DiPersio2019}
+Di Persio, L.,  Prezioso, L., and Wallbaum, K.: Closed-end formula for options linked to target volatility strategies. Arxiv preprint (2019)
+%
+\bibitem{Dozat2016}
+Dozat, T.: Incorporating Nesterov momentum into Adam. International Conference on Learning Representations (2016)
+%
+\bibitem{Dupire1994}
+Dupire, B.: Pricing with a smile. Risk magazine {\bf 7-1} (1994) 18-20
+%
+\bibitem{Gabrielli2020}
+Gabrielli, S., Pallavicini, A., and Scoleri, S.: Funding adjustments in equity linear products. Risk (2020)
+%
+\bibitem{Grasselli2016}
+Grasselli, M., and Romo, J.M.: Stochastic skew and target volatility options. Journal of Futures Markets {\bf 36-2} (2016) 174-193
+%
+\bibitem{Gyongy1986}
+Gy\"ongy, I.: Mimicking the one-dimensional marginal distributions of processes having an ito differential. Probability Theory and Related Fields {\bf 4-71} (1986) 501-516
+%
+\bibitem{Halperin2020}
+Halperin, I.: Q-Learner in the Black-Scholes(-Merton) worlds. The Journal of Derivatives {\bf 28-1} (2020) 99-122
+%
+\bibitem{Hiton2012}
+Hinton, G., Srivastava, N., and Swersky, K.: Neural networks for machine learning lecture 6a overview of mini-batch gradient descent. Cited on {\bf 8-14} (2012)
+%
+\bibitem{Hocquard2013}
+Hocquard, A., Ng, S., and Papageorgiou, N.: A constant volatility framework for managing tail risk. The Journal of Portfolio Management {\bf 2-39} (2013) 28-40
+%
+\bibitem{Kim2018}
+Kim, Y., and Enke, D.: A dynamic target volatility strategy for asset allocation using artificial neural networks. The Engineering Economist {\bf 4-63} (2018) 273-290
+%
+\bibitem{Kolm2019}
+Kolm, P.N., and Ritter, G.: Dynamic replication and hedging: a reinforcement learning approach. The Journal of Financial Data Science {\bf 1-1} (2019) 159-171
+%
+\bibitem{Longstaff2001}
+Longstaff, F., and Schwartz, E.: Valuing American options by simulation: a simple least-squares approach. Review of Financial Studies {\bf 14-1} (2001) 113-147
+%
+\bibitem{Morrison2013}
+Morrison, S., and Tadrowski, L.: Guarantees and target volatility funds. Moody's Analytics (2013)
+%
+\bibitem{Perchet2016}
+Perchet, R., De Carvalho, R.L., Heckel, T., and Moulin, P.: Predicting the success of volatility targeting strategies: application to equities and other asset classes.
+The Journal of Alternative Investements {\bf 3-18} (2016) 21-38
+%
+\bibitem{Puheim2014}
+Puheim, M., and Madar\'{a}sz, L.: Normalization of inputs and outputs of neural network based robotic arm controller in role of inverse kinematic model.
+2014 IEEE 12th International Symposium on Applied Machine Intelligence and Informatics (2014) 85-89
+%
+\bibitem{Schulman2016}
+Schulman, J., Moritz, P., Levine, S., Jordan, M.I., and Abbeel, P.: High-dimensional continuous control using generalized advantage estimation.
+4th International Conference on Learning Representations (2016)
+%
+\bibitem{Schulman2017}
+Schulman, J.,  Wolski, P., Dhariwal, P., Radford, A., and Klimov, O.: Proximal policy optimization. Arxiv preprint (2017)
+%
+\bibitem{Sola1997}
+Sola, J., and Sevilla, J.: Importance of input data normalization for the application of neural networks to complex industrial problems.
+IEEE Transactions on Nuclear Science {\bf 44-3} (1997) 1464-1468
+%
+\bibitem{Sutton2018}
+Sutton, R.S., and Barto, A.G.: Reinforcement learning: an introduction. MIT Press, Cambridge (2018)
+%
+\bibitem{Xue2012}
+Xue, Y.: Target Volatility fund: an effective risk management tool for VA?. Society of Actuaries, International News (2012)
+\end{thebibliography}
+
+\vfill
+
+\footnotesize{\vskip1mm$\!\!\!\!\!\!\!\!\!\!${\fbox{{\parbox[c][2.0cm]{12cm}{\center{\textsc{\vskip-4.5mm \textbf{Mathematical Methods in Economics and Finance}}} \textbf{-- m$\mathbf{^2}$ef} \\ \vskip.5mm
+					Vol.
+$\ldots$, No.
+$\ldots$, $\ldots$ \\ \vskip.5mm {ISSN print edition: 1971-6419 -- ISSN online edition: 1971-3878} \\ \vskip.5mm {Web page: \email{http://www.unive.it/m2ef/} -- E-mail: \email{m2ef@unive.it}} \\}}}}}
+
+% \newpage
+% \mbox{}
+
 \end{document}

### 2023-02-11 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
Applying referee's simple corrections

@@
 \usepackage{graphicx}
+\usepackage{xcolor}
 
@@
 
-\title{Reinforcement learning for options\\on target volatility funds\thanks{Supported by ISCRA from Cineca, \email{iscra@cineca.it}.}}
+\title{Reinforcement \change{L}earning for \change{O}ptions\\on \change{T}arget \change{V}olatility \change{F}unds\thanks{Supported by ISCRA from Cineca, \email{iscra@cineca.it}.}}
 
@@
 Corso Europa, 2, 20122 Milan, Italy\\
-\email{Emanuele.Nastasi@marketz.eu} \and Financial Engineering,
+\email{Emanuele.Nastasi@marketz.eu} \and \change{Financial Risk Methodologies, Unicredit S.p.A.,\\ Piazza Gae Aulenti, 3, 20154 Milan, Italy \\ \email{Stefano.Polo@unicredit.eu}}
+\remove{Financial Engineering,
 X-Numeris S.r.l.,\\ Via Turro, 7 , 20127 Milan, Italy\\
-\email{Stefano.Polo@xnumeris.com}}
+\email{Stefano.Polo@xnumeris.com}}}
 
@@
 We deal with the funding costs rising from hedging the risky securities underlying a target volatility strategy (TVS), a portfolio of risky assets and a risk-free one dynamically rebalanced in order to keep the realized volatility of the portfolio on a certain level.
-The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs for each component requires to solve a control problem to evaluate the option prices.
+The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs \change{of its} \remove{for each} component requires to solve a control problem to evaluate the option prices.
 We derive an analytical solution of the problem in the Black and Scholes (BS) scenario.
@@
 In the recent years portfolio managers are exposed to very low interest rates and quickly changing market volatilities.
-An effective solution to control risks under such an environment is given by target volatility strategies (TVSs) (also known as constant volatility targeting) which are able to preserve the portfolio at a predetermined level of volatility.
-A TVS is a portfolio of risky assets (typically  equities) and a risk-free asset dynamically re-balanced with the aim of maintaining the overall portfolio volatility level closed to some target value.
-The constant volatility approach can help investors to obtain desired risk exposures over the short and long term and it increases the risk-adjusted performance of the portfolio.
+An effective solution to control risks under such an environment is given by target volatility strategies (TVSs)\change{,} \remove{(}also known as constant volatility targeting\remove{)}\change{,} which are able to preserve the portfolio at a predetermined level of volatility.
+A TVS is \change{a strategy for managing} a portfolio of risky assets (typically  equities) and a risk-free \change{one} \remove{asset} dynamically re-balanced with the aim of maintaining the overall portfolio volatility level closed to some target value.
+The constant volatility approach can help investors to obtain desired risk exposures over the short and long term and \change{possibly} it increases the risk-adjusted performance of the portfolio.
 
-These products were initially offered in the Asian markets, see for instance the reports of \cite{Chew2011} and \cite{Xue2012}, which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \cite{Morrison2013}.
+These products were initially offered in the Asian markets, see for instance \remove{the reports of} \cite{Chew2011} and \cite{Xue2012}, which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \cite{Morrison2013}.
 At the present day we can observe some new market indices based on the mechanism of the target volatility strategy such as Dow Jones Volatility Control Index, and S\&P 500 Risk Control Index.
 
-In the recent literature TVSs are tested to investigate their performances in term of realized returns, see for instance \cite{Hocquard2013} and \cite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim2018}.
+In the recent literature\change{,} \remove{TVSs} \change{TVS-based portfolios} are tested to investigate their performances in term of realized returns, see for instance \cite{Hocquard2013} and \cite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim2018}.
 Moreover, in the present day, in the derivative market new structured financial products that see TVSs as underlying are becoming widely used by practitioners.
@@
 Although from a theoretical point of view this solution represents a sub-optimal strategy when dealing with generic dynamics for the risky assets, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem.
-More precisely we numerically study the problem in the general case of a local volatility dynamics \cite{Derman1994}, \cite{Dupire1994}  by adopting a Reinforcement Learning approach, in particular an \textit{ad hoc} vanilla direct policy algorithm and the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}.
+More precisely we numerically study the problem in the general case of a local volatility dynamics \cite{Derman1994}, \cite{Dupire1994}  by adopting a Reinforcement Learning approach, in particular an \textit{ad hoc} \remove{vanilla} direct policy \change{search} algorithm and the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}.
 \\
@@
 Then, in Section \ref{sec:Derivative} we introduce the structured class of derivative contracts linked to TVSs, where we describe the arising dynamical control problem for pricing those options.
-Moreover in this section we derive the BS closed solution for European TVOs in two different ways: one applying the Gy\"ongy Lemma and the other by writing the Hamilton-Jacobi-Bellman equation.
+Moreover in this section we derive the BS closed solution for European TVOs in two different ways: one applying the Gy\"ongy Lemma and the other \change{through} \remove{by writing} the Hamilton-Jacobi-Bellman equation.
 In Section \ref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a description of the algorithm we have built.
@@
 
-\section{Target volatility strategy}\label{sec:TVS}
+\section{Target \change{V}olatility \change{S}trategy}\label{sec:TVS}
 As mentioned before, this work aims to enrich the TVS pricing literature by studying the aspect related to the funding costs coming from hedging the risky assets underlying a target volatility portfolio.
-Thus we consider the following scenario: a bank selling a protection to a portfolio manager who has his capital invested in a TVS.
+Thus we consider the following scenario: a bank selling a protection to a portfolio manager who has his capital invested \change{according to} \remove{in} a TVS.
 In our case the fund manager has the freedom of changing the relative weights of the risky asset during the life of the TVS; once the allocation strategy is selected then the volatility targeting algorithm rebalances the risky component of the portfolio with the risk-free one in order to keep the overall portfolio volatility close to a target value.
@@
 For instance, the capital can be protected by buying a put option.
-In this case, we can write the net asset value (NAV) $A_t$ of the strategy as given by\footnote{Here we neglect discounting factors.}
+In this case, we can write the \change{undiscounted} net asset value (NAV) $A_t$ of the strategy as given by\footnote{\remove{Here we neglect discounting factors.}}
 \begin{equation}
@@
 The fund manager is not implementing the strategy by trading in the market, and he is not subject to additional costs to access the market.
-On the other way, the bank is paying such costs since she is actively hedging the call option sold to the manager.
+On the other way, the bank is paying such costs since \change{it} \remove{she} is actively hedging the call option sold to the manager.
 The bank trading activity implemented to actively hedge the option requires funding the collateral procedures of the hedging instruments along with any lending/borrowing fee.
@@
 
-\subsection{The strategy Price Process}
-We work on a filtered probability space $\left(\Omega, \mathcal{F}, \{\mathcal{F}_t\}_{t\geq0}, \mathbb{P}  \right)$ satisfying the usual assumptions for a market model, where $\mathbb{P}$ is the physical probability measure representing the actual distribution of supply and demand shocks on equities prices.
+\subsection{The \change{S}trategy Price Process}
+We work on a filtered probability space $\left(\Omega, \mathcal{F}, \{\mathcal{F}_t\}_{t\geq0}, \mathbb{P}  \right)$ satisfying the usual assumptions for a market model, where $\mathbb{P}$ is the physical probability measure \remove{representing the actual distribution of supply and demand shocks on equities prices}.
 
-We consider a fund trading a basket of $n$ risky securities with price process $S_t = (S_t^i, \, i=1,\dots,n)$ funded with a cash account $B_t$ accruing at $r_t$.
+We consider a fund trading a basket of $n$ risky securities with price process \change{$S_t = \{S_t^i, \, i=1,\dots,n\}$} \remove{$S_t = (S_t^i, \, i=1,\dots,n)$} funded with a cash account $B_t$ accruing at $r_t$.
 Any dividend paid by the securities is re-invested in the fund.
@@
 Now, in order to prevent arbitrages, we assume the existence of a risk-neutral measure $\mathbb{Q}$ equivalent to $\mathbb{P}$ under which the deflated gain processes of all traded securities are martingales.
-Under this assumption, we are able to derive the drift conditions on the security price processes, and in turn on the strategy price process.
+Under this assumption, we are able to derive the drift conditions on the security price processes, and in turn on the strategy price process\change{, that is}
 \begin{equation}
@@
 where $M_t^i$ are martingale under $\mathbb{Q}$.
-If we substitute this expression for the security dynamics into the definition of the strategy we can check that the price process of the strategy is accruing at a cash account rate $r_t$ compensated for the fund manager fees
+If we substitute this expression for the security dynamics into the definition of the strategy we can check that the price process of the strategy is accruing at \change{the following} \remove{a} cash account rate $r_t$ compensated for the fund manager fees
 \begin{equation}
@@
 In a typical TVS, the fund manager selects a risky-asset portfolio with a specific time-dependent allocation strategy expressed by means of the vector of relative weights $\alpha_t$, along with a risk-free asset, which we can identify with the bank account $B_t$.
-In the following, for sake of simplicity in exposition, we consider only total-return securities, namely we set $\pi_t=0$.
+In the following, for sake of simplicity in exposition, we consider only total-return securities, namely we set $\pi_t=0$ \change{since no dividends are paid}.
 Thus we can write Equation~\eqref{eq:TVS_first} as given by
@@
 where $\Bar{\sigma}$ is the target volatility value.
-In practice, this means that the fund manager will select a risky-portfolio choosing $\alpha_t$ equities from the universe where he can trade and after that, his choices will be scaled by the automatic target volatility algorithm $\omega_t$.\footnote{We recall that the universe of assets where the manager can trade and the value of $\bar{\sigma}$ are written in the contract.}
+In practice, this means that the fund manager will select a risky-portfolio choosing $\alpha_t$ equities from the universe where \change{s}he can trade and after that, \change{her} \remove{his} choices will be scaled by the automatic target volatility algorithm $\omega_t$.\footnote{We recall that the universe of assets where the manager can trade and the value of $\bar{\sigma}$ are written in the contract.}
 
-To derive the expression for $\omega_t$ we need to assume a generic continuous semi-martingales dynamics under the risk-neutral measure for the underlying securities, so that we can write Equation~\eqref{eq:risk_neutral} as
+\remove{To derive the expression for $\omega_t$ we need to assume} \change{It is possible to derive the expression for $\omega_t$ assuming} a generic continuous semi-martingales dynamics under the risk-neutral measure for the underlying securities, so that \remove{we can write} Equation~\eqref{eq:risk_neutral} \change{can be written} as
 \begin{equation}
@@
 \label{eq:Equity_process}\end{equation}
-where $\nu_t$ is an adapted matrix process ensuring the existence of a solution for the stochastic differential equation (SDE) and $W_t$ is a $n$-dimensional vector of Brownian motions under $\mathbb{Q}$.
+where $\nu_t$ is an adapted matrix process ensuring the existence of a solution for the stochastic differential equation \remove{(SDE)} and $W_t$ is a $n$-dimensional vector of Brownian motions under $\mathbb{Q}$.
 Under these assumptions we can derive an expression for $\omega_t$, and we get\footnote{In all formulae the norm for a vector $a$ is defined as $\|a\|\coloneqq\sqrt{a\cdot a}$.}
@@
 \subsection{European Options}
-If the derivative contract depends only on the marginal distribution of $I_T$ at maturity (a European payoff), we are able to calculate an optimal strategy. We consider the following pricing problem
+If the derivative contract depends only on the marginal distribution of $I_T$ at maturity (\change{i.e.,} a European payoff), we are able to calculate an optimal strategy.
+We consider the following pricing problem
 \begin{equation}
@@
 
-\subsection{Black and Scholes model}\label{subsec:BS_model}
+\subsection{Black and Scholes \change{Scenario}}\label{subsec:BS_model}
 In the time-dependent Black and Scholes model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$.
@@
 \begin{proposition}
-When the underlying securities follow a Black and Scholes model with deterministic rates, the optimal strategy for a non-decreasing European payoff consists in minimizing the local drift function, independently of the current state $I_t$
+When the underlying securities follow a Black and Scholes \change{QUA IO LASCEREI MODEL (REFEREE PROPONE SCENARIO)} model with deterministic rates, the optimal strategy for a non-decreasing European payoff consists in minimizing the local drift function, independently of the current state $I_t$
 \begin{equation}
@@
 
-In Figure~\ref{fig:strategy_price_comparison} we provide a comparison among the option price obtained with the optimal strategy (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and those with other intuitive strategies (S$_A$, S$_B$, S$_C$).
+In Figure~\ref{fig:strategy_price_comparison} we provide a comparison among the option price obtained with the optimal strategy (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and those with \change{the below} \remove{other} intuitive strategies (S$_A$, S$_B$, S$_C$).
 Here we consider the case of an at-the-money call option with spot $I_0 = 1\, \text{EUR}$, maturity $T=5 \, \text{yr}$, target volatility $\bar{\sigma}=5\%$, and a nonnegative constraint on $\alpha$.
@@
 \begin{lemma}
-	Let be $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \tr{\nu}$.
+	Let \remove{be} $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \tr{\nu}$.
 Then the closed solution of the optimization problem \eqref{eq:BS_optimal_strategy} is
@@
 \subsubsection{Active Asset (or Bang Bang) Solution}
-A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry are nonnegative and the only constraint on portfolio weights is nonnegativity, which would mean a long-only strategy by the fund manager.
+A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry \change{$\mu$} are nonnegative and the only constraint on portfolio weights is nonnegativity, which would mean a long-only strategy by the fund manager.
 
@@
 
-\subsection{Hamilton-Jacobi-Bellman equation for target volatility options}\label{subsec:HJB}
-In this section, we want to provide to the reader a formal description of the dynamic problem associated with options on target volatility strategies by writing the Hamilton-Jacobi-Bellman (HJB) equation for the derivative price.
+\subsection{Hamilton-Jacobi-Bellman \change{E}quation for \change{T}arget \change{V}olatility \change{O}ptions}\label{subsec:HJB}
+In this section, we \remove{want to} provide to the reader a formal description of the dynamic problem associated with options on target volatility strategies by writing the Hamilton-Jacobi-Bellman (HJB) equation for the derivative price.
 We prove that from this equation we can recover the same closed formula \eqref{eq:BS_optimal_strategy} derived above from the Gy\"ongy Lemma for the time-dependent BS model.
@@
 \label{eq:HJB_general}\end{equation}
-Equation~\eqref{eq:HJB_general} is the Hamilton-Jacobi-Bellman equation describing the TVS dynamic problem for a generic dynamics of the risky securities underlying the portfolio.
+Equation~\eqref{eq:HJB_general} is the \change{HJB} \remove{Hamilton-Jacobi-Bellman}     equation describing the TVS dynamic problem for a generic dynamics of the risky securities underlying the portfolio.
 
@@
 \end{equation}
-if the payoff is non-decreasing in $I$ by homogeneity of the SDE we get that $V$ is non-decreasing; thus the solution is given by
+if the payoff is non-decreasing in $I$ by homogeneity of the SDE we get that $V$ is non-decreasing\change{.}\remove{; t}\change{T}hus the solution is given by
 \begin{equation}
@@
 In this work we tackle the non-trivial case of a local volatility (LV) model for the $S_t$-dynamics, such that $\nu_t = \nu(t,S_t)$.
-We have chosen the LV model since it is well known in financial literature and among practitioners.
+We have chosen the LV model since it is well known in financial literature and among practitioners \change{\cite{Gatheral2006}}.
 
-\section{Reinforcement learning}\label{sec:RL}
+\section{Reinforcement \change{L}earning}\label{sec:RL}
 As we have discussed in the previous sections, one must resort to numerical approaches to solve the stochastic control problem related to the TVS in the case of general payoffs or risky securities dynamics.
@@
 
-Reinforcement Learning is a branch of Machine Learning which allows an artificial agent to interact with an environment through actions and observations in order to maximize some notion of cumulative reward.
-In RL the agent is not told which actions to take but instead, it must discover by trial and error which are the behaviours  yielding the highest reward.
+\remove{Reinforcement Learning} \change{RL} is a branch of Machine Learning which allows an artificial agent to interact with an environment through actions and observations in order to maximize some notion of cumulative reward \change{called value function}.
+In RL\change{,} the agent is not told which actions to take but instead, it \remove{must} \change{has to} discover by trial and error which are the behaviours  yielding the highest reward.
 This is obtained by updating the agent policy $\pi$ which is a mapping from the environment states to the set of actions.
@@
 
-\subsection{Direct policy approach}\label{sec:direct_policy}
+\subsection{Direct \change{P}olicy \change{A}pproach}\label{sec:direct_policy}
 We consider an episode $\tau$ of length $m+1$ that takes place on a discrete time-grid of fixing times expressed in year fractions $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$ maturity of the option.
@@
 
-\subsection{Proximal policy optimization approach}
+\subsection{Proximal \change{P}olicy \change{O}ptimization \change{A}pproach}
 As for the direct policy approach, we model the pricing problem considering an episode that takes place on the time-grid $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$.
@@
 In this case, the reward function does not suffer of nil values for $0<T_k<T$: the RL agent always gets a feedback from the environment for its choices.
-The hyper-parameter $\gamma$ plays the role of a discount factor in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future.
+\remove{The} \change{Note that the} hyper-parameter $\gamma$ plays the role of a discount factor in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future.
 If we take the cumulative reward per episode and set the PPO parameters\footnote{We refer to \cite{Schulman2016} for a more detailed description for the generalized advantage estimation parameter $\lambda$.} $\gamma =\lambda =1$ we obtain
@@
 
-\section{Numerical investigations}\label{sec:Numerical_investigations}
+\section{Numerical \change{I}nvestigations}\label{sec:Numerical_investigations}
 Here we present the numerical results obtained with our proposed methods. We focus our analysis on a European call option on a TVS with the following product details
 \begin{equation}\nonumber
-	I_0 = K = 1 \, \text{EUR} \; , \quad T=2 \, \text{yr} \; , \quad \bar{\sigma} = 5\% \; .
+	I_0 = K = 1 \, \text{EUR} \; , \quad T=2 \, \text{yr} \; , \quad \bar{\sigma} = 5\% \; \change{,}
 \end{equation}
-Moreover without loss of generality we consider the case of a completely free allocation strategy $\alpha$. The extension to the constrained case is easy.
+\change{where $I_0$ is the starting value of the TVS-based portfolio.}
+Moreover without loss of generality we consider the case of a\change{n unconstrained} \remove{completely free} allocation strategy $\alpha$. The extension to the constrained case is easy.
 
@@
 
-Although in Section~\ref{subsec:BS_model} we have proved that under the Black and Scholes model the Equation~\eqref{eq:BS_optimal_strategy} solves the control problem, we want to take advantage of this \textit{a priori} solution as a benchmark to gather evidence on the robustness of our Reinforcement Learning approach and to check if our analytical result is correct.
+Although in Section~\ref{subsec:BS_model} we have proved that under the Black and Scholes model the Equation~\eqref{eq:BS_optimal_strategy} solves the control problem, we want to take advantage of this \textit{a priori} solution as a benchmark to gather evidence on the robustness of our \remove{Reinforcement Learning} \change{RL} approach and to check if our analytical result is correct.
 Moreover, we use the BS model as numerical laboratory to perform fine-tuning tests for the RL algorithms hyper-parameters and analyze how they impact the final results and performances.
@@
 We recall that in both the algorithms we parameterize the agent policy with an FFNN; thus this is completely characterized by the following hyper-parameters: number of hidden layers, number of neurons per hidden layer, activation function per hidden layer.
-This is due to the fact that in this Reinforcement Learning problem the number of the input neurons is equal to the state space dimension \eqref{eq:state}, while the number of the output ones coincides with the action space dimension \eqref{eq:action}.
+This is due to the fact that in this \remove{Reinforcement Learning} \change{RL} problem the number of the input neurons is equal to the state space dimension \eqref{eq:state}, while the number of the output ones coincides with the action space dimension \eqref{eq:action}.
 It is well known in the literature that neural networks give better performances in the training phase if the input data are well normalized \cite{Puheim2014}, \cite{Sola1997} .
@@
 
-\subsection{Black and Scholes: hyper-parameters fine tuning}
+\subsection{Black and Scholes: \change{H}yper-parameters \change{F}ine \change{T}uning}
 We use the BS environment as toy model to understand which parameters of the RL algorithm play key roles in the training and testing phase. 
@@
 	\includegraphics[width=3.4in]{BS_fine_tuning_Nadam_elu_grey.eps}
-	\caption{Learning curves of the direct policy algorithm applied to the Black and Scholes model.
+	\caption{Learning curves of the direct policy algorithm applied to the Black and Scholes \change{scenario}.
 The solid lines are the loss functions defined in Equation~\eqref{eq:direct_policy_loss} in function of the number of training epochs.
@@
 	\includegraphics[width=3.4in]{BS_fine_tuning_RMSprop_elu_grey.eps}
-	\caption{Learning curves of the direct policy algorithm applied to the Black and Scholes model.
+	\caption{Learning curves of the direct policy algorithm applied to the Black and Scholes \change{scenario}.
 The solid lines are the loss values defined in Equation~\eqref{eq:direct_policy_loss} in function of the number of training epochs.
@@
 Each learning curve is compared with the optimal option price (grey-dashed line) that we have computed with the closed form solution we have derived for the BS model in Section~\ref{subsec:BS_model}.
-We can see that all the learning curves converge to our expected price, providing a numerical demonstration of our theoretical result.
+We can see that all the learning curves converge to our expected price, providing a numerical \remove{demonstration} \change{finding} of our theoretical result.
 More deeply, from the plots we observe that the RMSprop optimizer outperforms the Nadam.
@@
 
-\subsection{Local volatility dynamics}
+\subsection{Local \change{V}olatility \change{D}ynamics}
 In this section, we study the TVS control problem assuming a local volatility model for the dynamics of the risky assets.
@@
 We report the results in Table~\ref{tab:mc_results_lv}.
-The Reinforcement Learning agents we have trained with PPO give all prices compatible with the baseline one.
+The \remove{Reinforcement Learning} \change{RL} agents we have trained with PPO give all prices compatible with the baseline one.
 \begin{table}
@@
 
-\section{Conclusion and further developments}
+\section{Conclusion and \change{F}urther \change{D}evelopments}
 In this paper, we described a non-trivial control problem related to derivative contracts on target volatility strategies.
@@
 Despite its complexity, our first contribution is the derivation of a closed form solution of the control problem in a Black and Scholes framework, which could represent a useful tool for practitioners since it outperforms intuitive trading strategies.
-We have derived this solution in two different ways: first by applying the Gy\"ongy  Lemma and then by writing the Hamilton-Jacobi-Bellman equation.
+We have derived this solution in two different ways: first by applying the Gy\"ongy  Lemma and then by writing the \remove{Hamilton-Jacobi-Bellman} \change{HJB} equation.
 We numerically studied the problem in the more general local volatility model where the solution is not available and thus a numerical investigation is needed.
-We tackled the problem by means of the novel Reinforcement Learning techniques, by both the direct policy learning and the proximal policy optimization one.
+We tackled the problem by means of the novel \remove{Reinforcement Learning} \change{RL} techniques, by both the direct policy learning and the proximal policy optimization one.
 We used the BS model, where the solution is \textit{a priori} known, as benchmark to perform a series of fine-tuning of the RL algorithm hyper-parameters, such as the artificial neural network architecture.
@@
 Gabrielli, S., Pallavicini, A., and Scoleri, S.: Funding adjustments in equity linear products. Risk (2020)
+%
+\bibitem{Gatheral2006}
+Gatheral, J.: The volatility surface: a practitioner's guide. Wiley Finance (2006)
 %

### 2023-02-18 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
implement other referee's comments

@@
 
-\title{Reinforcement \change{L}earning for \change{O}ptions\\on \change{T}arget \change{V}olatility \change{F}unds\thanks{Supported by ISCRA from Cineca, \email{iscra@cineca.it}.}}
+\title{Reinforcement \change{L}earning for \change{O}ptions\\on \change{T}arget \change{V}olatility \change{F}unds\thanks{\change{Our sincere thanks go to Marco Bianchetti and Diego Pierluigi Giovannini of Intesa Sanpaolo Milan, as well as to the Italian computing center Cineca, who generously supported our numerical simulations through an ISCRA C grant.
+The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.}}}
 
@@
 
-A part of this work has been developed during the Master thesis of one of us (SP), where we had the opportunity to collaborate with Marco Bianchetti and Diego Pierluigi Giovannini from Intesa Sanpaolo Milan, that we thank.
+\remove{A part of this work has been developed during the Master thesis of one of us (SP), where we had the opportunity to collaborate with Marco Bianchetti and Diego Pierluigi Giovannini from Intesa Sanpaolo Milan, that we thank.
 Moreover we wish to thank the Italian computing centre Cineca, which approved our ISCRA C project and provided us with the high-computing resources of Marconi100 for the numerical simulations of this contribution.
@@
 
-The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.
+The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.}
 
@@
 
-Then, we introduce the strategy price process $V_t$, and we define the deflated gain process $\Bar{G}_t^V$ as given by
+Then, we introduce the strategy price process $V_t$, and we define the deflated gain process $\Bar{G}_t^V$ as \remove{given by}
 \begin{equation}
@@
 
-We highlight that, to derive the dynamics expressed in Equation~\eqref{eq:TVS_last}, we have not made any assumptions on the risky allocation strategy; thus all this argument is valid for any constraints on the process $\alpha_t$.
+We highlight that, to derive the dynamics expressed in Equation~\eqref{eq:TVS_last}, we have not made any assumptions on the risky allocation strategy; thus all this argument is valid for any constraints on the process $\alpha_t$\change{ as, for instance, the case of contracts in which the fund manager is restricted to holding only long positions meaning that $\alpha_t^i>0$ $\forall i = 1,\dots,n$, or situations where the combined total of long and short trades must be capped at a specific contractual amount}.
 
@@
 In this section, we analyze derivative contracts linked to the TVS described previously.
-In particular we will focus on European style options, and we will show that under appropriate assumptions it is possible to find a closed form solution for the optimal allocation strategy which maximizes the contract price.
+In particular\change{,} we will focus on European style options, and we will show that under appropriate assumptions it is possible to find a closed form solution for the optimal allocation strategy which maximizes the contract price.
 
@@
 
-In Figure~\ref{fig:strategy_price_comparison} we provide a comparison among the option price obtained with the optimal strategy (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and those with \change{the below} \remove{other} intuitive strategies (S$_A$, S$_B$, S$_C$).
+In Figure~\ref{fig:strategy_price_comparison} we provide a comparison among the option price obtained with the optimal strategy (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and those with \change{the below expert-based single-asset} \remove{other intuitive} strategies (S$_A$, S$_B$, S$_C$).
 Here we consider the case of an at-the-money call option with spot $I_0 = 1\, \text{EUR}$, maturity $T=5 \, \text{yr}$, target volatility $\bar{\sigma}=5\%$, and a nonnegative constraint on $\alpha$.
@@
 \begin{itemize}
-\item S$_A$: invest all in the asset with the maximum forward curve at maturity;
-\item S$_B$: for each market pillar before maturity invest all in the asset with minimum $\mu(t)$; 
-\item S$_C$: for each market pillar before maturity invest all in the asset with minimum $\mu(t)/\|\nu(t)\|$.
+\item S$_A$: invest all in the asset with the maximum forward curve at maturity; \change{this strategy constists of selecting a static portfolio that is never rebalanced, such that the TVO moneyness is maximized;}
+\item S$_B$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)$; \change{in this case we consider a fund manager who rebalances the weights $\alpha(t)$ over the life of the portfolio by selecting the asset with the highest expected return at a given point in time;}
+\item S$_C$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)/\|\nu(t)\|$\remove{.}\change{; the strategy is similar to S$_B$, except in this case, the manager maximizes the expected return weighted for the target volatility mechanism.}
 \end{itemize}
@@
 	\includegraphics[width=4in]{strategy_price_comparison_grey.eps}
-	\caption{Comparison of plain vanilla prices on the TVS by adopting different allocation strategies: the optimal Black and Scholes solution (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and other intuitive strategies (S$_A$, S$_B$, S$_C$)}
+	\caption{Comparison of plain vanilla prices on the TVS by adopting different allocation strategies: the optimal Black and Scholes solution (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and \change{three expert-based single-asset} strategies (S$_A$, S$_B$, S$_C$)}
 	\label{fig:strategy_price_comparison}
 \end{figure}
+
+\change{In the next subsection we derive the analytical solution of the problem \eqref{eq:BS_optimal_strategy} under the assumption of unconstrained allocation strategy $\alpha_t$}
 
@@
 \end{proof}
+
+\change{In the following subsection we prove that an analytical solution of \eqref{eq:BS_optimal_strategy} exists also assuming a nonnegative constrained allocation strategy.
+Under this assumption we show that the optimal strategy is a single-asset one called bang bang solution.}
 
@@
 
-In full generality, we assume that the time evolution of the Markovian factors governing the problem dynamics is given by  a stochastic multidimensional process in $\mathbb{R}^n$, $X_t$, which is the unique strong solution to the following It$\hat{\text{o}}$ SDE
+In full generality, we assume that the time evolution of the Markovian factors governing the problem dynamics is given by  a stochastic multidimensional process in $\mathbb{R}^n$, $X_t$, which is the unique strong solution to the following It$\hat{\text{o}}$ \remove{SDE} \change{stochastic differential equation}
 \begin{equation}
@@
 
-In this framework, the TVS price process dynamics is given by the SDE 
+In this framework, the TVS price process dynamics is given by the \remove{SDE} \change{stochastic differential equation} 
 \begin{equation}
@@
 \end{equation}
-if the payoff is non-decreasing in $I$ by homogeneity of the SDE we get that $V$ is non-decreasing\change{.}\remove{; t}\change{T}hus the solution is given by
+if the payoff is non-decreasing in $I$\change{,} by homogeneity of the \remove{SDE} \change{stochastic differential equation} we get that $V$ is non-decreasing\change{.}\remove{; t}\change{T}hus the solution is given by
 \begin{equation}
@@
 This is obtained by updating the agent policy $\pi$ which is a mapping from the environment states to the set of actions.
-Thus RL is independent of pre-collected data as opposed to other Machine Learning techniques.
+Thus RL is independent of pre-collected data as opposed to other Machine Learning techniques\change{, that have to be fed a pre-existing dataset to learn from}.
 Because of its nature, RL has been successful in quantitative finance for solving control problems; among the most important RL applications in this field, we refer to \cite{Deng2017} as the pioneers in studying self-taught reinforcement trading problems, while to \cite{Halperin2020} and \cite{Kolm2019} for hedging derivatives with RL under market frictions.
@@
 \subsection{Direct \change{P}olicy \change{A}pproach}\label{sec:direct_policy}
-We consider an episode $\tau$ of length $m+1$ that takes place on a discrete time-grid of fixing times expressed in year fractions $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$ maturity of the option.
+\change{In RL, the learning phase takes place within a set of so-called episodes, which refer to a single run or instance of interaction between the RL agent and an environment, where the agent takes a sequence of actions to achieve a specific goal.}
+
+\change{For our problem, we} \remove{We} consider an episode $\tau$ of length $m+1$ that \remove{takes place on} \change{unfolds over} a discrete time-grid of fixing times expressed in year fractions $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$ maturity of the option.
 At a given episodic time $T_k$ the RL agent interacts with the environment: it receives a representation of the environment called state $s_k$ and on the basis of that it selects an action $a_k$ sampling from the current policy $\pi^{\theta_j}$.
 Here with $\theta_j$ we refer to the set of parameters through which we parameterize the agent policy at the $j$-th algorithm iteration.
-In our case of study the agent can choose the composition of the risky asset portfolio, so that the policy is the allocation strategy $\alpha$ introduced in Equation~\eqref{eq:TVS_elegant}:
+In our case of study\change{,} the agent\change{, representing the fund manager, has the right to select} \remove{can choose} the composition of the risky asset portfolio, so that the policy is the allocation strategy $\alpha$ introduced in Equation~\eqref{eq:TVS_elegant}:
 \begin{equation}
@@
  \begin{equation}
- 	r_{k+1}=\left\{\begin{array}{ll}
+ 	r_{k+1}\change{\coloneqq}\left\{\begin{array}{ll}
  		 D(0,T;\zeta)(I_T(\alpha(\theta_j)) - K)^+ & \text { if } T_{k+1} = T \\
@@
 Moreover, we compare the performances of two well-known optimizers in Machine Learning literature: Nadam \cite{Dozat2016} and RMSprop \cite{Hiton2012}.
-Since we deal with a free allocation strategy $\alpha$ that can assume negative values, we have chosen among the wide variety of activation functions the tanh and the elu.
+Since we deal with a free allocation strategy $\alpha$ that can assume negative values, we have chosen among the wide variety of activation functions \change{\cite{Goodfellow2016}} the tanh and the elu.
 In this way, we can analyze the performance of a saturating activation function and a non-saturating one.
@@
 \begin{center}
-\caption{Comparison of TVO prices under Black and Scholes model: analytical solution price and direct policy reinforcement learning.
+\caption{Comparison of TVO prices under Black and Scholes \change{scenario}: analytical solution price and direct policy reinforcement learning.
 The option parameters are: $I_0=K=1\,[\text{EUR}]$, $T=2\,[\text{yr}]$ and $\bar{\sigma}=5\%$}
@@
 	\includegraphics[width=4in]{LV_direct_policy_approach_grey.eps}
-	\caption{Learning curves of the direct policy algorithm applied to the local volatility model.
+	\caption{Learning curves of the direct policy algorithm applied to the local volatility \change{scenario}.
 The solid line with triangles is the learning curve of an agent with free actions, while the solid one is the learning curve of an agent whose actions are parameterized with the maximum drift strategy.
@@
 	\includegraphics[width=3.4in]{LV_PPO_reward_b_learning_curve.eps}
-	\caption{Learning curves of the PPO algorithm applied to the local volatility model.
+	\caption{Learning curves of the PPO algorithm applied to the local volatility \change{scenario}.
 On the horizontal axis the number of training episodes.
@@
 \begin{center}
-\caption{Comparison of TVO prices under local volatility model.
+\caption{Comparison of TVO prices under local volatility \change{scenario}.
 The baseline price is obtained by applying path-wise the maximum drift strategy \eqref{eq:analytical_solution_free_strategy}.
@@
 \bibitem{Gatheral2006}
-Gatheral, J.: The volatility surface: a practitioner's guide. Wiley Finance (2006)
+\change{Gatheral, J.: The volatility surface: a practitioner's guide. Wiley Finance (2006)}
+%
+\bibitem{Goodfellow2016}
+\change{Goodfellow, I., Bengio, Y., Courville, A.: Deep learning. MIT Press (2016)}
 %

### 2023-02-25 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
Add introduction and abstract corrections

@@
 
-\title{Reinforcement \change{L}earning for \change{O}ptions\\on \change{T}arget \change{V}olatility \change{F}unds\thanks{\change{Our sincere thanks go to Marco Bianchetti and Diego Pierluigi Giovannini of Intesa Sanpaolo Milan, as well as to the Italian computing center Cineca, who generously supported our numerical simulations through an ISCRA C grant.
+\title{Reinforcement \change{L}earning for \change{O}ptions\\on \change{T}arget \change{V}olatility \change{F}unds\thanks{\change{Our sincere thanks go to Marco Bianchetti and Diego Pierluigi Giovannini of Intesa Sanpaolo Milan, as well as to the Italian computing center Cineca for its support on our numerical simulations.
 The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.}}}
@@
 \begin{abstract}
-We deal with the funding costs rising from hedging the risky securities underlying a target volatility strategy (TVS), a portfolio of risky assets and a risk-free one dynamically rebalanced in order to keep the realized volatility of the portfolio on a certain level.
-The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs \change{of its} \remove{for each} component requires to solve a control problem to evaluate the option prices.
+We deal with \remove{the funding costs} \change{a stochastic optimal control problem} rising from hedging the risky securities underlying a target volatility strategy (TVS), a portfolio \remove{of risky assets and a risk-free one dynamically rebalanced in order to keep the} \change{whose asset-allocation is adjusted to mantain the} realized volatility of the portfolio on a certain level.
+\change{We consider the point of view of a derivative writer selling an option contract as protection to a portfolio manager on the capital invested in a TVS.} The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs \change{of its} \remove{for each} component requires to \remove{solve a control problem to evaluate the option} \change{adjust the protection} price\remove{s} \change{to include these costs in the worst-case scenario for the seller}.
 We derive an analytical solution of the problem in the Black and Scholes (BS) scenario.
-Then we use Reinforcement Learning (RL) techniques to determine the fund composition leading to the most conservative price under the local volatility (LV) model, for which an {\it a priori} solution is not available.
+Then we use Reinforcement Learning (RL) techniques to determine the fund composition leading to the \remove{most conservative price} \change{optimal policy} under the local volatility (LV) model, for which an {\it a priori} solution is not available.
 We show how the performances of the RL agents are compatible with those obtained by applying path-wise the BS analytical strategy to the TVS dynamics, which therefore appears competitive also in the LV scenario.
@@
 \section{Introduction}
-In the recent years portfolio managers are exposed to very low interest rates and quickly changing market volatilities.
+In the recent years portfolio managers \remove{are} \change{were} exposed to very low interest rates and quickly changing market volatilities.
 An effective solution to control risks under such an environment is given by target volatility strategies (TVSs)\change{,} \remove{(}also known as constant volatility targeting\remove{)}\change{,} which are able to preserve the portfolio at a predetermined level of volatility.
 A TVS is \change{a strategy for managing} a portfolio of risky assets (typically  equities) and a risk-free \change{one} \remove{asset} dynamically re-balanced with the aim of maintaining the overall portfolio volatility level closed to some target value.
-The constant volatility approach can help investors to obtain desired risk exposures over the short and long term and \change{possibly} it increases the risk-adjusted performance of the portfolio.
+\change{The strategy's concept involves adjusting the allocation of high-risk assets to low-risk assets during periods of high volatility in the high-risk assets, and vice versa.
+Thus the dynamic allocation is based on both the actual volatility of the high-risk asset and a predetermined target level of volatility.
+}The constant volatility approach can help investors to obtain desired risk exposures over the short and long term and \change{possibly} it increases the risk-adjusted performance of the portfolio.
 
@@
 At the present day we can observe some new market indices based on the mechanism of the target volatility strategy such as Dow Jones Volatility Control Index, and S\&P 500 Risk Control Index.
+In the recent literature\change{,} \remove{TVSs} \change{TVS-based portfolios} are tested to investigate their performances in term of realized returns, see for instance \cite{Hocquard2013} and \cite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim2018}\change{, where the authors propose the use of artificial neural networks for volatility forecasting volatility to enhance the performance of an asset allocation strategy}.
 
-In the recent literature\change{,} \remove{TVSs} \change{TVS-based portfolios} are tested to investigate their performances in term of realized returns, see for instance \cite{Hocquard2013} and \cite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim2018}.
-Moreover, in the present day, in the derivative market new structured financial products that see TVSs as underlying are becoming widely used by practitioners.
-Those derivative instruments are known as target volatility options (TVOs) and have gained great interest in the derivative pricing literature, see in particular \cite{Albeverio2019}, \cite{DiGraziano2012}, \cite{DiPersio2019}, and \cite{Grasselli2016}.
+\remove{Moreover, in the present day, in the derivative market new structured financial products that see TVSs as underlying are becoming widely used by practitioners.} \change{The success of TVS-based portfolios among investors has led to the emergence of new structured products with such portfolios as underlyings}.
+Those derivative instruments are known as target volatility options (TVOs) \change{and have become increasingly important in the derivative market.} \change{For these reasons TVOs}  have gained great interest in the derivative pricing literature \change{and research}, see in particular \cite{Albeverio2019}, \cite{DiGraziano2012}, \cite{DiPersio2019}, and \cite{Grasselli2016}.
 
-This paper is the first attempt, to the best of our knowledge, to study the funding costs coming from hedging the risky assets underlying the target volatility strategy.
-In particular, we consider the point of view of a bank selling a call option to a portfolio manager as protection on the capital invested in a TVS.
-The portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS.
-Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario, i.e.
+This paper \change{aims to enrich the pricing framework of the TVOs by studying for the first time,} \remove{is the first attempt} to the best of our knowledge, \remove{to study} the funding costs coming from hedging the risky assets underlying the target volatility strategy.
+In particular, we consider the point of view of \remove{a bank} \change{an option writer (for example a bank)} selling a call option to a portfolio manager as protection on the capital invested in a TVS.
+\change{During her activity the} \remove{The} portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS \change{depending on how the market evolves.
+This mechanism is no regulated by the contract, as the manager has decision making power on which asset invest}.
+Since \change{the fund manager's strategy cannot be known {\it a priori} and since} the risky assets have different hedging costs, the \remove{bank} \change{option writer} shall adjust the price of the protection to include \remove{them} \change{these costs} in the worst-case scenario, i.e.
 the most expensive strategy in term of the financing costs.
-Hence, the pricing problem becomes a continuous dynamical control problem over the risky portfolio composition.
-In our work we provide to the reader a formal description of the stochastic control problem and most importantly we derive an analytical solution assuming that the risky asset dynamics underlying the TVS  follow a Black and Scholes (BS) model \cite{BlackScholes1973} and that the derivative contract is a European-style option.
+Hence, the pricing problem becomes a continuous dynamical control problem over the risky portfolio composition\change{: determine the manager's optimal strategy that leads to the maximum protection price}.
 
-Although from a theoretical point of view this solution represents a sub-optimal strategy when dealing with generic dynamics for the risky assets, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem.
-More precisely we numerically study the problem in the general case of a local volatility dynamics \cite{Derman1994}, \cite{Dupire1994}  by adopting a Reinforcement Learning approach, in particular an \textit{ad hoc} \remove{vanilla} direct policy \change{search} algorithm and the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}.
+In our work we provide to the reader a formal description of the stochastic control problem \change{and we show that it can be solved only numerically considering a general state-dependent dynamics for the risky assets.} \change{However, we prove that} \remove{and most importantly we derive} an analytical solution to the problem \change{exists,} assuming that the risky asset dynamics underlying the TVS  follow a Black and Scholes (BS) model \cite{BlackScholes1973} and that the derivative contract is a European-style option.
+\change{As our first contribution, we derive the BS closed solution into two different ways: applying the Gy\"ongy Lemma and the other through the Hamilton-Jacobi-Bellman equation.}
+
+\change{In the second part of our work we tackle the problem in the general case of a state-dependent dynamics for the risky assets for which only the numerical approach is admissible.}
+\remove{Although from a theoretical point of view this solution represents a sub-optimal strategy when dealing with generic dynamics for the risky assets, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem}.
+More precisely we numerically study the problem \remove{in the general case of  a local volatility dynamics \cite{Derman1994}, \cite{Dupire1994} by adopting a Reinforcement Learning approach, in particular an \textit{ad hoc} vanilla direct policy algorithm and the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}} \change{assuming a local volatility (LV) dynamics \cite{Derman1994}, \cite{Dupire1994} \change{for the risky assets}.
+The reason for selecting the LV model is its widespread adoption in the industry to value derivatives, especially those related to asset baskets \cite{Gatheral2006}.
+The LV investigation is based on a Reinforcement Learning approach, in particular through an \textit{ad hoc} direct policy search algorithm and the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}.
+From our study, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem under the LV dynamics.}
 \\
@@
 The paper is organized as follows.
-In Section \ref{sec:TVS} we introduce more in detail the manager-bank contract and provide the description of the TVS dynamics in presence of valuation adjustments such as the hedging costs.
+In Section \ref{sec:TVS} we introduce more in detail the \remove{manager-bank} contract \change{between the option-writer and the manager} and provide the description of the TVS dynamics in presence of valuation adjustments such as the hedging costs.
 Then, in Section \ref{sec:Derivative} we introduce the structured class of derivative contracts linked to TVSs, where we describe the arising dynamical control problem for pricing those options.
@@
 As mentioned before, this work aims to enrich the TVS pricing literature by studying the aspect related to the funding costs coming from hedging the risky assets underlying a target volatility portfolio.
-Thus we consider the following scenario: a bank selling a protection to a portfolio manager who has his capital invested \change{according to} \remove{in} a TVS.
+Thus we consider the following scenario: a \remove{bank} \change{option writer} selling a protection to a portfolio manager who has \remove{his} \change{her} capital invested \change{according to} \remove{in} a TVS.
 In our case the fund manager has the freedom of changing the relative weights of the risky asset during the life of the TVS; once the allocation strategy is selected then the volatility targeting algorithm rebalances the risky component of the portfolio with the risk-free one in order to keep the overall portfolio volatility close to a target value.
@@
 
-The fund manager usually buys from a bank an option on the TVS to ensure capital protection.
+The fund manager usually buys from a \remove{bank} \change{counterpart} an option on the TVS to ensure capital protection.
 For instance, the capital can be protected by buying a put option.
-In this case, we can write the \change{undiscounted} net asset value (NAV) $A_t$ of the strategy as given by\footnote{\remove{Here we neglect discounting factors.}}
+In this case, we can write the \change{undiscounted} net asset value (NAV) $A_t$ of the strategy as given by\footnote{Here we neglect discounting factors \change{for sake of clarity}.}
 \begin{equation}
@@
 
-In this way, the TVS is only defined in the two contracts client-fund and fund-bank.
-The fund manager is not implementing the strategy by trading in the market, and he is not subject to additional costs to access the market.
-On the other way, the bank is paying such costs since \change{it} \remove{she} is actively hedging the call option sold to the manager.
-The bank trading activity implemented to actively hedge the option requires funding the collateral procedures of the hedging instruments along with any lending/borrowing fee.
-We remark that the choices of the manager trading activity are stochastic processes since they will depend on the market evolution.
-Thus, neither the manager strategy nor the bank financing costs for hedging the option can be written in the fund-bank contract.
-For this reason, the price of a financial product sold by the bank is adjusted to include any valuation adjustment due to the trading activity.
+In this way, the TVS is only defined in the two contracts client-fund and fund-\change{option writer} \remove{fund-bank}.
+The fund manager is not implementing the strategy by trading in the market, and \remove{he} \change{she} is not subject to additional costs to access the market.
+On the other way, the \remove{bank} \change{writer} is paying such costs since she is actively hedging the call option sold to the manager.
+The \remove{bank} \change{writer} trading activity implemented to actively hedge the option requires funding the collateral procedures of the hedging instruments along with any lending/borrowing fee.
+We remark that\change{, upon entering into the contract,}  the choices of the manager trading activity \remove{are} \change{can be seen as} stochastic processes since they \change{are not known {\it a priori} and }will depend on the market evolution.
+Thus, neither the manager strategy nor the \remove{bank} \change{writer} financing costs for hedging the option can be written in the fund-\remove{bank}\change{writer} contract\change{: the only contractual elements are the protection strike and maturity, the trading market of the manager and the target volatility of the TVS portfolio}.
+For this reason, the price of a financial product sold by the \remove{bank} \change{writer} is adjusted to include any valuation adjustment due to the trading activity.
 
-Our aim is to find the most expensive investing strategy from the point of view of financing costs for the bank that the manager could choose in the market.
+Our aim is to find the most expensive investing strategy from the point of view of financing costs for the \remove{bank} \change{writer} that the manager could choose in the market.
 In other words, we want to determine the worst-case scenario for the bank.
@@
 In a typical TVS, the fund manager selects a risky-asset portfolio with a specific time-dependent allocation strategy expressed by means of the vector of relative weights $\alpha_t$, along with a risk-free asset, which we can identify with the bank account $B_t$.
-In the following, for sake of simplicity in exposition, we consider only total-return securities, namely we set $\pi_t=0$ \change{since no dividends are paid}.
+In the following, for sake of simplicity in exposition, we consider only total-return securities, namely we set $\pi_t=0$ \change{, which means that the fees received by the stock lender are null since no dividends are paid}.
 Thus we can write Equation~\eqref{eq:TVS_first} as given by

### 2023-02-25 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
add minibatch definition

@@
 \begin{abstract}
-We deal with \remove{the funding costs} \change{a stochastic optimal control problem} rising from hedging the risky securities underlying a target volatility strategy (TVS), a portfolio \remove{of risky assets and a risk-free one dynamically rebalanced in order to keep the} \change{whose asset-allocation is adjusted to mantain the} realized volatility of the portfolio on a certain level.
+We deal with \remove{the funding costs} \change{a stochastic optimal control problem} rising from hedging the risky securities underlying a target volatility strategy (TVS), a portfolio \remove{of risky assets and a risk-free one dynamically rebalanced in order to keep the} \change{whose asset-allocation is adjusted to maintain the} realized volatility of the portfolio on a certain level.
 \change{We consider the point of view of a derivative writer selling an option contract as protection to a portfolio manager on the capital invested in a TVS.} The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs \change{of its} \remove{for each} component requires to \remove{solve a control problem to evaluate the option} \change{adjust the protection} price\remove{s} \change{to include these costs in the worst-case scenario for the seller}.
@@
 
-In this work we tackle the non-trivial case of a local volatility (LV) model for the $S_t$-dynamics, such that $\nu_t = \nu(t,S_t)$.
+In this work we tackle the non-trivial case of a local volatility \remove{(LV)} model for the $S_t$-dynamics, such that $\nu_t = \nu(t,S_t)$.
 We have chosen the LV model since it is well known in financial literature and among practitioners \change{\cite{Gatheral2006}}.
@@
 Again we experienced that deeper FFNNs outperform shallow ones and tanh is the most preferable.
-We report for completeness the values of the other PPO hyper-parameters, for the description of which we refer to \cite{Schulman2017}: learning rate $3\times 10^{-4}$, $\lambda=0.95$, $\epsilon=0.2$, $c_1=0.7$, $c_2=0$ and mini-batch size of 2048 episodes.
+We report for completeness the values of the other PPO hyper-parameters, for the description of which we refer to \cite{Schulman2017}: learning rate $3\times 10^{-4}$, $\lambda=0.95$, $\epsilon=0.2$, $c_1=0.7$, $c_2=0$ and mini-batch\footnote{\change{A minibatch is a subset of a larger dataset, which is used to train a machine learning model.
+The model updates its parameters based on the error calculated from the predictions it makes on each minibatch.
+This allows for more efficient memory use, faster convergence of the model, and better generalization to new data.}} size of 2048 episodes.
 

### 2023-02-26 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
corrections of RL sections

@@
 \usepackage{graphicx}
-\usepackage{xcolor}
+\usepackage{xcolor, cancel}
 
@@
 \newcommand{\remove}[1]{{\color{red} \soutpars{{#1}}}}%{}
+\newcommand{\removeeq}[1]{{\color{red} \sout{{#1}}}}%{}
 \newcommand{\changeB}[1]{{\color{blue} {#1}}}%{{#1}}
 \newcommand{\removeB}[1]{{\color{blue} \soutpars{{#1}}}}%{}
+\newcommand\hcancel[2][red]{\setbox0=\hbox{$#2$}%
+\rlap{\raisebox{.45\ht0}{\textcolor{#1}{\rule{\wd0}{1pt}}}}#2}
 
@@
 In this way, the TVS is only defined in the two contracts client-fund and fund-\change{option writer} \remove{fund-bank}.
-The fund manager is not implementing the strategy by trading in the market, and \remove{he} \change{she} is not subject to additional costs to access the market.
+The fund manager is not implementing the strategy by trading in the market, and \change{s}he  is not subject to additional costs to access the market.
 On the other way, the \remove{bank} \change{writer} is paying such costs since she is actively hedging the call option sold to the manager.
@@
 Our aim is to find the most expensive investing strategy from the point of view of financing costs for the \remove{bank} \change{writer} that the manager could choose in the market.
-In other words, we want to determine the worst-case scenario for the bank.
+In other words, we want to determine the worst-case scenario for the \remove{bank} \change{option-seller}.
 In the current section, we proceed by defining the price process of the TVS so that we can highlight the impact of valuation adjustments.
@@
 \end{equation}
-where $\phi_t$ are the running fees earned by the fund manager for his activity. We assume that the strategy is self-financing, so that we can write
+where $\phi_t$ are the running fees earned by the fund manager for \remove{his} \change{her} activity. We assume that the strategy is self-financing, so that we can write
 \begin{equation}
@@
 \end{proposition}
-The absence of stochastic elements in Equation~\eqref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no simulation needed; in fact one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities.
+The absence of stochastic elements in Equation~\eqref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no \change{numerical} simulation needed \change{to solve the control problem}; in fact \remove{one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities} \change{, once the risky-assets market data $\mu(t)$ and $\nu(t)$, it is sufficient to apply Equation~\eqref{eq:BS_optimal_strategy} $\forall t \in \mathcal{T}$  to retrieve the deterministic optimal strategy}.
 Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
@@
 
-In our work we adopt two learning strategies to compare their performances in terms of overall reward: an \textit{ad hoc} direct policy algorithm and the state of the art proximal policy optimization (PPO) developed in \cite{Schulman2017} and \cite{Schulman2016}.
+In our work we adopt two learning strategies to compare their performances in terms of overall reward: an \textit{ad hoc} direct policy \change{search} algorithm and the state of the art proximal policy optimization (PPO) developed in \cite{Schulman2017} and \cite{Schulman2016}.
 
@@
 
-\subsection{Direct \change{P}olicy \change{A}pproach}\label{sec:direct_policy}
+\subsection{Direct \change{P}olicy \change{Search} \change{A}pproach}\label{sec:direct_policy}
 \change{In RL, the learning phase takes place within a set of so-called episodes, which refer to a single run or instance of interaction between the RL agent and an environment, where the agent takes a sequence of actions to achieve a specific goal.}
@@
 Here with $\theta_j$ we refer to the set of parameters through which we parameterize the agent policy at the $j$-th algorithm iteration.
-In our case of study\change{,} the agent\change{, representing the fund manager, has the right to select} \remove{can choose} the composition of the risky asset portfolio, so that the policy is the allocation strategy $\alpha$ introduced in Equation~\eqref{eq:TVS_elegant}:
+In our case of study\change{,} the \change{RL} agent\change{, representing the fund manager, has the right to select} \remove{can choose} the composition of the risky asset portfolio, so that the policy is the allocation strategy $\alpha$ introduced in Equation~\eqref{eq:TVS_elegant}:
 \begin{equation}
@@
 \label{eq:state}\end{equation}
-In this way, the state contains all the information needed by the agent to take an optimal action, leading to the maximum plain vanilla TVO price.
+In this way, the state contains all the information needed by the agent to \remove{take an} \change{select the} optimal action, leading to the maximum plain vanilla TVO price.
 
-In this algorithm, we parameterize the agent policy with a feed forward neural network (FFNN), such that $\theta_j$ coincides with the hidden weights, $s_k$ with the inputs, and $a_k$ with the output.
-In this way, we are dealing with a deterministic policy, where its functional form is given by the neural network.
-The parameters update is performed as follows: the agent collects a finite batch of experiences interacting with the environment in a set of episodes $\tau$, then the loss function is evaluated as
+\change{Once the agent has selected the action $a_k$ sampled from the current policy, it receives at the next time $T_{k+1}$ a reward $r_{k+1}$ generated by the environment.
+In this algorithm we have defined the reward function as follows}
  \begin{equation}
-L(\theta_j) = D(0,T;\zeta)\hat{\mathbb{E}}\left[ \Phi(I_T(\alpha(\theta_j))) \right] \; ,
+ 	\change{r_{k+1}\change{\coloneqq}\left\{\begin{array}{ll}
+ 		 D(0,T;\zeta)(I_T(\alpha(\theta_j)) - K)^+ & \text { if } T_{k+1} = T \\
+ 		0 & \text { otherwise }
+ 	\end{array}\right.
+ \label{eq:reward1} \; .} \end{equation}
+
+ \change{Therefore, during the whole episode, the agent receives a nil reward except at maturity when the reward coincides with the option intrinsic value.
+This choice may seem too daring because the agent receives a real feedback of its actions only at the end of the episode, increasing the probability to obtain a slow learning.
+However, if the agent has learnt the optimal policy $\pi^* = \pi^{\theta^*}$, the average cumulative reward per episode will coincide with the optimal TVO price.}
+
+In this algorithm, we parameterize the agent policy with a feed forward neural network (FFNN), such that $\theta_j$ coincides with the hidden weights, $s_k$ with the \remove{inputs} \change{input neurons}, and $a_k$ with the output \change{ones}.
+In this way, we are dealing with a deterministic policy, \change{that maps each state directly to a single action without any randomness or probability distribution;} \remove{where its} \change{the map} functional form is given by the neural network.
+The parameters update is performed as follows: the agent collects a finite batch of experiences interacting with the environment in a set of episodes $\tau$, then the loss-\change{value} function is evaluated as
+\begin{equation}
+\change{L(\theta_j) =\hat{\mathbb{E}}\left[ \sum_{k=0}^{m-1} r_{k+1} | \theta_j \right]} \; ,
 \label{eq:direct_policy_loss}\end{equation}
+\begin{equation}
+\hcancel[red]{\change{L(\theta_j) = D(0,T;\zeta)\hat{\mathbb{E}}\left[ \Phi(I_T(\alpha(\theta_j))) \right]}} \; ,
+\end{equation}  
 where the expectation $\hat{\mathbb{E}}[\dots]$ indicates the empirical average on the batch.
-Then the parameters are updated by plugging the policy into a stochastic gradient ascent algorithm.
-The choice of Equation~\eqref{eq:direct_policy_loss} is justified by the fact that, when the algorithm will find the optimal set $\theta^*$, then $L(\theta^*)$ will be a good proxy of the optimal option price.
+Then the parameters are updated by plugging the policy into a stochastic gradient ascent algorithm \change{with the aim of maximizing Equation~\eqref{eq:direct_policy_loss}}.
+\remove{The choice of Equation~\eqref{eq:direct_policy_loss} is justified by the fact that, when the algorithm will find the optimal set $\theta^*$, then $L(\theta^*)$ will be a good proxy of the optimal option price.}
 
@@
 \subsection{Proximal \change{P}olicy \change{O}ptimization \change{A}pproach}
-As for the direct policy approach, we model the pricing problem considering an episode that takes place on the time-grid $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$.
+As for the direct policy \change{search} approach, we model the pricing problem considering an episode that takes place on the time-grid $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$.
 Again we choose as observation state the block defined in Equation~\eqref{eq:state} and the agent policy coincides with the risky allocation strategy $\alpha$ \eqref{eq:action}.
-An important difference from the previous method is that: once the agent has selected the action $a_k$ sampled from the current policy, it receives at the next time $T_{k+1}$ a reward $r_{k+1}$ generated by the environment.
+\remove{An important difference from the previous method is that: once the agent has selected the action $a_k$ sampled from the current policy, it receives at the next time $T_{k+1}$ a reward $r_{k+1}$ generated by the environment.}
 
-In our work, we have defined two different reward functions with the purpose to analyze which could help the agent to learn more efficiently the optimal policy $\pi^* = \pi^{\theta^*}$. 
-Our first definition is
+\remove{In our work, we have defined two different reward functions with the purpose to analyze which could help the agent to learn more efficiently the optimal policy $\pi^* = \pi^{\theta^*}$. 
+Our first definition is} 
  \begin{equation}
- 	r_{k+1}\change{\coloneqq}\left\{\begin{array}{ll}
+ 	\hcancel[red]{\change{r_{k+1} =\left\{\begin{array}{ll}
  		 D(0,T;\zeta)(I_T(\alpha(\theta_j)) - K)^+ & \text { if } T_{k+1} = T \\
  		0 & \text { otherwise }
- 	\end{array}\right.
- \label{eq:reward1} \; . \end{equation}
-Therefore, during the whole episode, the agent receives a nil reward except at maturity when the reward coincides with the option intrinsic value.
+ 	\end{array}\right.}}
+  \; . \end{equation}
+\remove{Therefore, during the whole episode, the agent receives a nil reward except at maturity when the reward coincides with the option intrinsic value.
 This choice may seem too daring because the agent receives a real feedback of its actions only at the end of the episode, increasing the probability to obtain a slow learning.
-However, if the agent has learnt $\pi^*$, the average cumulative reward per episode will coincide with the optimal TVO price.
+However, if the agent has learnt $\pi^*$, the average cumulative reward per episode will coincide with the optimal TVO price.}
 
-The second reward function we have defined is 
+\change{As the PPO algorithm is more sophisticated than the previous one, we have decided to test its performance by means of two different reward functions.
+The former is the same as defined by Equation~\eqref{eq:reward1}, while the latter is given by}
+\remove{The second reward function we have defined is}
 \begin{equation}
@@
 the spot price, $\nu_t=\nu(t,S_t)$.
-This additional dependency of volatility makes the problem of finding the optimal strategy non-trivial; in fact if we consider the whole securities smiles then the second-order term in the HJB equation \eqref{eq:HJB_general} is not zero and thus a closed formula for $\alpha^*$ is not available anymore.
-Because of that, one must resort to numerical techniques to recover the problem solution.
-Unlike the BS model where the strategy depends only on time, in LV dynamics there is no unnecessary information provided by the state block in Equation~\eqref{eq:state} to take the optimal action.
+This additional dependency of volatility makes the problem of finding the optimal strategy non-trivial\change{, as a closed form formula cannot be derived under this scenario}; in fact if we consider the whole \remove{securities} \change{volatility} smiles \change{of the securities,} then the second-order term in the HJB equation \eqref{eq:HJB_general} \remove{is not zero and thus a closed formula for $\alpha^*$ is not available anymore} \change{does not elide and consequently the only way to derive $\alpha^*$ is through a numerical method}.
+\remove{Because of that, one must resort to numerical techniques to recover the problem solution.} \remove{Unlike the BS model where the strategy depends only on time, in LV dynamics there is no unnecessary information provided by the state block in Equation~\eqref{eq:state} to take the optimal action} \change{The additional dependence on the state $S_t$ under LV dynamics will cause the agent to consider also the $X_t$ component value in the  state block \eqref{eq:state}, unlike under the BS scenario where this information is redundant as the optimal strategy was function only of time \eqref{eq:TVS_last}}.
 Here we consider the same market data ($r_t$, $\mu_t$ and $\nu_t$) as in the Black and Scholes environment to study how the optimal solution changes with the dynamics model.
@@
 To try to measure the performance of the policy selected by the agent, we can build a n\"aif strategy called \quotations{baseline}.
-By looking at the Markovian projection in \eqref{eq:markovian_projection}-\eqref{eq:local_drift} and the first term in the HJB \eqref{eq:HJB_general}, a natural choice for the baseline is to maximize the TVS local drift.
-In other words this coincides by applying the path-wise the Black and Scholes solution of Equation~\eqref{eq:BS_optimal_strategy}; since here we deal with a free allocation strategy, we can simply use our analytical result \eqref{eq:analytical_solution_free_strategy}.
+By looking at the Markovian projection in \eqref{eq:markovian_projection}-\eqref{eq:local_drift} and the first term in the HJB \change{equation} \eqref{eq:HJB_general}, \change{we see that the control problem is similar to the BS one with an additional second order term.
+Thus} a natural choice for the baseline is to maximize the TVS local drift \remove{.
+In other words this coincides} by applying the path-wise the Black and Scholes solution of Equation~\eqref{eq:BS_optimal_strategy}.
+\change{S}ince here we deal with a free allocation strategy, we can simply use our analytical result \eqref{eq:analytical_solution_free_strategy}.
 In the same Figure~\ref{fig:result_lv_direct_policy} we report the 99\% confidence interval of the MC price obtained with the maximum drift baseline as two red-dashed lines.
@@
 Following the theoretical result of the HJB equation, we can assert that the agent has learnt a sub-optimal policy.
-Since we LV model differs from the BS one for a corrective term in the HJB \eqref{eq:HJB_general}, we expect that the optimal solution  in the LV framework will be in a close region of the maximum drift strategy.
+Since we LV model differs from the BS one for a corrective term in the HJB \change{equation} \eqref{eq:HJB_general}, we expect that the optimal solution  in the LV framework will be in a close region of the maximum drift strategy.
 Thus we train another agent with the direct policy learning by parameterizing its action with the baseline strategy and choosing a smaller learning rate of $10^{-4}$.
-With this parameterization, at each observational time $T_k\in\mathcal{T}$ the risky allocation strategy is obtained by summing the network output with the Equation~\eqref{eq:analytical_solution_free_strategy}.
+With this parameterization, \change{the RL agent actions are computed by summing} at each observational time $T_k\in\mathcal{T}$ \remove{the risky allocation strategy is obtained by summing} the FFNN output with the Equation~\eqref{eq:analytical_solution_free_strategy}\change{, giving rise to the manager's risky allocation strategy}.
 Even in this case, the corresponding learning curve (solid black line in Figure~\ref{fig:result_lv_direct_policy}) is stuck in the maximum drift strategy.
@@
 
-We have trained a 5 layer FFNN with 8 neurons each with the PPO algorithm in two different environments: one environment generates the reward according to Equation~\eqref{eq:reward1}, while the other exploits the reward function \eqref{eq:reward2} where we set $\gamma=0.98$ as discount factor to make the agent prefer immediate rewards.
+We have trained a 5 layer FFNN with 8 neurons each with the PPO algorithm in two different environments: one environment generates the reward according to Equation~\eqref{eq:reward1}, while the other exploits the reward function \eqref{eq:reward2} where we set $\gamma=0.98$ as discount factor to make the agent prefer immediate rewards \change{more than in the previous simulations}.
 Again, for both environments, we train two different agents: one whose action is given by Equation~\eqref{eq:action}, while the other implements the action parameterization with the baseline.
@@
 In this paper, we described a non-trivial control problem related to derivative contracts on target volatility strategies.
-In particular, we have considered a bank selling a call option to a fund manager as protection on the capital invested on the TVS.
-We showed how the presence of different funding costs coming from hedging the risky assets underlying the TVS, obliges the bank to solve a stochastic optimal control problem to price the protection.
-This is due to the fact that the bank strategy is not self-financing.
+In particular, we have considered a\change{n option writer} \remove{bank} selling a call option to a fund manager as protection on the capital invested on the TVS.
+We showed how the presence of different funding costs coming from hedging the risky assets underlying the TVS, obliges the \remove{bank} \change{writer} to solve a stochastic optimal control problem to price the protection.
+This is due to the fact that the \remove{bank} \change{option-seller's} strategy is not self-financing.
 This kind of control problem is hard to solve because here the control process affects both drift and diffusive coefficients of the controlled process.

### 2023-02-27 tex/m2ef/dnpp_referee_response.tex
Answers to major remarks 2 and 5
This commit also addresses the "make it clearer" at page 17.

@@
 
- ...
+The analytical solution applies only for a specific choice of the model for the dynamics of the underlying and under suitable hypotheses on the payoff.
+We added equation (52) and a sentence in the first paragraph of section 4 to stress this point.
+STEFANO, FORSE HAI ANCHE AGGIUNTO QUALCOSA IN INTRODUZIONE?
 
@@
 
- ...
+We added to the second paragraph of section 5.1 a sentence which explains how over-fitting is detected.
 

### 2023-02-27 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
Answers to major remarks 2 and 5
This commit also addresses the "make it clearer" at page 17.

@@
 
-\begin{proposition}
+\begin{proposition}\label{prop:bs}
 When the underlying securities follow a Black and Scholes \change{QUA IO LASCEREI MODEL (REFEREE PROPONE SCENARIO)} model with deterministic rates, the optimal strategy for a non-decreasing European payoff consists in minimizing the local drift function, independently of the current state $I_t$
@@
 if the payoff is non-decreasing in $I$\change{,} by homogeneity of the \remove{SDE} \change{stochastic differential equation} we get that $V$ is non-decreasing\change{.}\remove{; t}\change{T}hus the solution is given by
-\begin{equation}
+\begin{equation}\label{eq:HJB_solution}
  \alpha^*(t) = \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t)\|} \; ,
@@
 Conversely, if we deal with a dynamic model for which the derivative contract $V$ depends on $X$ then the volatility versor term, namely the second one, in the $\max$ operator of Equation~\eqref{eq:HJB_general} is no longer zero and thus one must solve the entire control problem numerically.
+\changeB{Indeed, one can check by direct substitution that the natural generalization of \eqref{eq:HJB_solution}
+\begin{equation}
+ \alpha^*(X_t) = \argmin_\alpha \frac{\alpha \cdot \mu(X_t)}{\|\alpha \cdot \nu(X_t)\|}
+\end{equation}
+fails to satisfy \eqref{eq:HJB_general}.}
 
-In this work we tackle the non-trivial case of a local volatility \remove{(LV)} model for the $S_t$-dynamics, such that $\nu_t = \nu(t,S_t)$.
+In \removeB{this work}\changeB{what follows} we tackle the non-trivial case of a local volatility \remove{(LV)} model for the $S_t$-dynamics, such that $\nu_t = \nu(t,S_t)$.
 We have chosen the LV model since it is well known in financial literature and among practitioners \change{\cite{Gatheral2006}}.
@@
 \section{Reinforcement \change{L}earning}\label{sec:RL}
-As we have discussed in the previous sections, one must resort to numerical approaches to solve the stochastic control problem related to the TVS in the case of general payoffs or risky securities dynamics.
+As we have discussed in the previous sections, \changeB{the analytical solution described in Proposition \ref{prop:bs} crucially relies on the assumption that the underlyings evolve according to a Black and Scholes model, and on a monotonicity assumption on the payoff.
+Therefore,}
+one must resort to numerical approaches to solve the stochastic control problem related to the TVS in the case of general payoffs or risky securities dynamics.
 The standard approach could be to use classical techniques based on backwards recursion \eqref{eq:recursion}-\eqref{eq:recursion2} such as American Monte Carlo \cite{Longstaff2001}.
@@
 In the \textsc{OpenAI Baselines} implementation of the PPO, the agent policy is parameterized again by a neural network; as for the previous method we have chosen an FFNN.
-In particular with PPO we deal with a stochastic policy whose functional form is a multivariate diagonal Gaussian distribution where the mean $\mu^\theta(s_k)$ is the output vector of the FFNN and the log-standard deviation $\log\sigma$ is an external parameter
+\removeB{In particular with PPO we deal with a stochastic policy, whose functional form is} \changeB{More precisely, training is based on a stochastic policy, i.e.
+the action is drawn from a probability law, and the neural network is used to determine its parameters.
+In settings like ours where the action has a continuous domain, such law is} a multivariate diagonal Gaussian distribution where the mean $\mu^\theta(s_k)$ is the output vector of the FFNN and the log-standard deviation $\log\sigma$ is an \removeB{external} \changeB{additional trainable} parameter:
 \begin{equation}
@@
 The idea is that the log-standard deviation will be higher at the beginning of the training phase in order to guarantee a good exploration of the action space while it will be lower at the end to avoid too much noise in the proximity of the optimal policy.
+\changeB{Eventually at the end of training, the standard deviation is put to zero, so that the action becomes deterministically equal to the output of the FFNN.}
 
@@
 Firstly we have performed a grid search on the learning rate starting value and we have found $10^{-3}$ a good choice in terms of speed of learning and avoiding over-fitting.
+\changeB{Note that since our environment is simulated, over-fitting is easily detected by cross-validation on a set of freshly-generated scenarios which were not seen during optimization.}
  \begin{figure}[h!]

### 2023-02-27 tex/m2ef/dnpp_referee_response.tex
complete dnpp_referee_response + some paper little corrections

@@
 \newcommand{\remove}[1]{{\color{red} \soutpars{{#1}}}}%{}
+\newcommand{\quotations}[1]{``#1''}
 
@@
 
- ...
+ We have conducted a comprehensive review of the crucial components of the abstract and introduction sections to ensure that the reader is presented with a lucid and concise overview of the problem being addressed.
+In the introduction, our focus was primarily on establishing a stronger connection between TVSs and the derivatives market, which are based on these underlying assets.
+Furthermore, we have expounded upon and elucidated the description of the problem addressed in this paper, with a particular emphasis on the methods employed and the resultant findings.
+It is hoped that this will enable the reader to contextualize our work more effectively within the pertinent literature.
 
@@
 We added equation (52) and a sentence in the first paragraph of section 4 to stress this point.
-STEFANO, FORSE HAI ANCHE AGGIUNTO QUALCOSA IN INTRODUZIONE?
+Furthermore, we have enhanced the clarity of this point in the introduction section where we explicate the contribution of this work.
 
@@
 
- ...
+ The three \quotations{intuitive strategies} presented in this paper represent plausible expert-based approaches that a practitioner may choose to adopt when addressing the control problem.
+We have provided further elucidation on this matter in Subsection 3.3, where we have also incorporated the financial implications of these strategies.
 
@@
 
- ...
+ The intention behind the sentence was to apprise the reader that the first algorithm was developed by the authors themselves to tackle the problem at hand.
+In formal terms, it can be classified as a direct policy search algorithm that relies on gradient-based optimization of the Monte Carlo outcome obtained from numerical simulations.
+This kind of algorithms is well-known in the Reinforcement Learning literature.
+We have revised this statement to ensure that its meaning is conveyed with greater clarity.
 

### 2023-02-27 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
complete dnpp_referee_response + some paper little corrections

@@
 The reason for selecting the LV model is its widespread adoption in the industry to value derivatives, especially those related to asset baskets \cite{Gatheral2006}.
-The LV investigation is based on a Reinforcement Learning approach, in particular through an \textit{ad hoc} direct policy search algorithm and the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}.
-From our study, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem under the LV dynamics.}
+The LV investigation is based on a Reinforcement Learning approach due to the high-dimensionality problema we deal with.
+In particular we compare the performances of a plain direct policy search gradient-based algorithm with those of the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}.
+At the end of our study, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem under the LV dynamics.}
 \\
@@
 
-In our work we adopt two learning strategies to compare their performances in terms of overall reward: an \textit{ad hoc} direct policy \change{search} algorithm and the state of the art proximal policy optimization (PPO) developed in \cite{Schulman2017} and \cite{Schulman2016}.
+In our work we adopt two learning strategies to compare their performances in terms of overall reward: \remove{an \textit{ad hoc}} \change{a plain} direct policy \change{search} algorithm and the state of the art proximal policy optimization (PPO) developed in \cite{Schulman2017} and \cite{Schulman2016}.
 
-The first method is a specific algorithm developed by us to fit the problem aim to find the optimal option price; this technique can take place as a direct policy optimization method in the wide taxonomy of Reinforcement Learning algorithms.
+The first method is \change{well-known in RL literature \cite{Sutton2018}: it is based on the direct update of the agent policy via a gradient-based optimization of the value function.}\remove{a specific algorithm developed by us to fit the problem aim to find the optimal option price; this technique can take place as a direct policy optimization method in the wide taxonomy of Reinforcement Learning algorithms}.
 We will go into details in Section~\ref{sec:direct_policy}.

### 2023-02-27 tex/m2ef/dnpp_referee_response.tex
small corrections

@@
 
- We have conducted a comprehensive review of the crucial components of the abstract and introduction sections to ensure that the reader is presented with a lucid and concise overview of the problem being addressed.
-In the introduction, our focus was primarily on establishing a stronger connection between TVSs and the derivatives market, which are based on these underlying assets.
+ We have conducted a comprehensive review of the crucial components of the Abstract and Introduction sections to ensure that the reader is presented with a lucid and concise overview of the problem being addressed.
+In the Introduction, our focus was primarily on establishing a stronger connection between TVSs and the derivatives market, which are based on these underlying assets.
 Furthermore, we have expounded upon and elucidated the description of the problem addressed in this paper, with a particular emphasis on the methods employed and the resultant findings.
@@
 The analytical solution applies only for a specific choice of the model for the dynamics of the underlying and under suitable hypotheses on the payoff.
-We added equation (52) and a sentence in the first paragraph of section 4 to stress this point.
-Furthermore, we have enhanced the clarity of this point in the introduction section where we explicate the contribution of this work.
+We added equation (52) and a sentence in the first paragraph of Section 4 to stress this point.
+Furthermore, we have enhanced the clarity of this point in the Introduction section where we explicate the contribution of this work.
 
@@
 
-We added to the second paragraph of section 5.1 a sentence which explains how over-fitting is detected.
+We added to the second paragraph of Section 5.1 a sentence which explains how over-fitting is detected.
 

### 2023-02-27 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
small corrections

@@
 \item S$_B$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)$; \change{in this case we consider a fund manager who rebalances the weights $\alpha(t)$ over the life of the portfolio by selecting the asset with the highest expected return at a given point in time;}
-\item S$_C$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)/\|\nu(t)\|$\remove{.}\change{; the strategy is similar to S$_B$, except in this case, the manager maximizes the expected return weighted for the target volatility mechanism.}
+\item S$_C$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)/\|\nu(t)\|$\remove{.}\change{; the strategy is akin to S$_B$, with the exception being that the manager endeavors to maximize the expected return that has been adjusted to account for the target volatility mechanism..}
 \end{itemize}
-As the reader can observe, our strategy outperforms any other intuitive one a practitioner might adopt.
+As the reader can observe, our strategy outperforms any other \remove{intuitive} \change{expert-based} one a practitioner might adopt \change{to face the control problem}.
 \begin{figure}[t]

### 2023-02-27 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
small corrections

@@
 \begin{itemize}
-\item S$_A$: invest all in the asset with the maximum forward curve at maturity; \change{this strategy constists of selecting a static portfolio that is never rebalanced, such that the TVO moneyness is maximized;}
+\item S$_A$: invest all in the asset with the maximum forward curve at maturity; \change{this strategy consists of selecting a static portfolio that is never rebalanced, such that the TVO moneyness is maximized;}
 \item S$_B$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)$; \change{in this case we consider a fund manager who rebalances the weights $\alpha(t)$ over the life of the portfolio by selecting the asset with the highest expected return at a given point in time;}

### 2023-03-04 tex/m2ef/dnpp_referee_response.tex
Add Andrea corrections

@@
 In the Introduction, our focus was primarily on establishing a stronger connection between TVSs and the derivatives market, which are based on these underlying assets.
-Furthermore, we have expounded upon and elucidated the description of the problem addressed in this paper, with a particular emphasis on the methods employed and the resultant findings.
+Furthermore, we have improved the description of the problem addressed in this paper, with a particular emphasis on the methods employed and the resultant findings.
 It is hoped that this will enable the reader to contextualize our work more effectively within the pertinent literature.
@@
 
-The analytical solution applies only for a specific choice of the model for the dynamics of the underlying and under suitable hypotheses on the payoff.
+The analytical solution applies only for a specific choice of the model for the dynamics of the underlying  assets and under suitable hypotheses on the payoff.
 We added equation (52) and a sentence in the first paragraph of Section 4 to stress this point.
-Furthermore, we have enhanced the clarity of this point in the Introduction section where we explicate the contribution of this work.
+Furthermore, we have enhanced the clarity of this point in the Introduction section where we discuss the contribution of this work.
 

### 2023-03-04 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
Add Andrea corrections

@@
 
-\title{Reinforcement \change{L}earning for \change{O}ptions\\on \change{T}arget \change{V}olatility \change{F}unds\thanks{\change{Our sincere thanks go to Marco Bianchetti and Diego Pierluigi Giovannini of Intesa Sanpaolo Milan, as well as to the Italian computing center Cineca for its support on our numerical simulations.
+\title{\change{Optimal strategies for options on target volatility funds}\thanks{\change{Our sincere thanks go to Marco Bianchetti and Diego Pierluigi Giovannini of Intesa Sanpaolo Milan, as well as to the Italian computing center Cineca for its support on our numerical simulations.
 The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.}}}
 
-\titlerunning{Reinforcement learning for options on target volatility funds}
+\titlerunning{\change{Optimal strategies for options on target volatility funds}}
 
@@
 \email{Emanuele.Nastasi@marketz.eu} \and \change{Financial Risk Methodologies, Unicredit S.p.A.,\\ Piazza Gae Aulenti, 3, 20154 Milan, Italy \\ \email{Stefano.Polo@unicredit.eu}}
-\remove{Financial Engineering,
-X-Numeris S.r.l.,\\ Via Turro, 7 , 20127 Milan, Italy\\
-\email{Stefano.Polo@xnumeris.com}}}
+}
 
@@
 \begin{abstract}
-We deal with \remove{the funding costs} \change{a stochastic optimal control problem} rising from hedging the risky securities underlying a target volatility strategy (TVS), a portfolio \remove{of risky assets and a risk-free one dynamically rebalanced in order to keep the} \change{whose asset-allocation is adjusted to maintain the} realized volatility of the portfolio on a certain level.
-\change{We consider the point of view of a derivative writer selling an option contract as protection to a portfolio manager on the capital invested in a TVS.} The uncertainty in the TVS risky portfolio composition along with the difference in hedging costs \change{of its} \remove{for each} component requires to \remove{solve a control problem to evaluate the option} \change{adjust the protection} price\remove{s} \change{to include these costs in the worst-case scenario for the seller}.
-We derive an analytical solution of the problem in the Black and Scholes (BS) scenario.
-Then we use Reinforcement Learning (RL) techniques to determine the fund composition leading to the \remove{most conservative price} \change{optimal policy} under the local volatility (LV) model, for which an {\it a priori} solution is not available.
-We show how the performances of the RL agents are compatible with those obtained by applying path-wise the BS analytical strategy to the TVS dynamics, which therefore appears competitive also in the LV scenario.
+We deal with \remove{the funding costs} \change{a stochastic optimal control problem} rising from hedging the risky securities underlying a target volatility strategy \remove{(TVS)}, a portfolio \remove{of risky assets and a risk-free one dynamically rebalanced in order to keep the} \change{whose asset-allocation is adjusted to maintain the} realized volatility of the portfolio \remove{on} \change{at} a certain level.
+\change{We consider the point of view of a derivative writer selling an option contract as protection to a portfolio manager on the invested capital.} The uncertainty in the \remove{TVS} risky portfolio composition along with the difference in hedging costs \change{of its} \remove{for each} component\change{s} requires to \remove{solve a control problem to evaluate the option} \change{adjust the protection} price\remove{s} \change{to include these costs in the worst-case scenario for the seller}.
+We derive an analytical solution of the problem in \remove{the} \change{a} Black and Scholes \remove{(BS)} scenario.
+Then\change{,} we use Reinforcement Learning \remove{(RL)} techniques to determine the fund composition leading to the \remove{most conservative price} \change{optimal policy} under the local volatility \remove{(LV)} model, for which an {\it a priori} solution is not available.
+We show how the performances of the \remove{RL agents} \change{numerical solution}  are compatible with those obtained by applying path-wise the \remove{BS} analytical \change{solution previously derived.} \remove{strategy to the TVS dynamics, which therefore appears competitive also in the LV scenario}
 
@@
 
-\noindent {\bf Keywords.} Reinforcement learning, target volatility strategy, hedging costs.
+\noindent {\bf Keywords.} \change{Stochastic control problem, reinforcement learning, target volatility, option pricing, hedging costs.}
 
@@
 An effective solution to control risks under such an environment is given by target volatility strategies (TVSs)\change{,} \remove{(}also known as constant volatility targeting\remove{)}\change{,} which are able to preserve the portfolio at a predetermined level of volatility.
-A TVS is \change{a strategy for managing} a portfolio of risky assets (typically  equities) and a risk-free \change{one} \remove{asset} dynamically re-balanced with the aim of maintaining the overall portfolio volatility level closed to some target value.
-\change{The strategy's concept involves adjusting the allocation of high-risk assets to low-risk assets during periods of high volatility in the high-risk assets, and vice versa.
-Thus the dynamic allocation is based on both the actual volatility of the high-risk asset and a predetermined target level of volatility.
+A TVS is \change{a strategy for managing} a portfolio of risky assets (typically  equities) and a risk-free \change{one} \remove{asset} dynamically re-balanced with the aim of maintaining the overall portfolio volatility level \remove{closed} \change{close} to some target value.
+\change{The strategy is designed to achieve a stable level of volatility for the underlying portfolio in any market scenario, and to allow a free selection of the relative allocation weights among the risky assets.
 }The constant volatility approach can help investors to obtain desired risk exposures over the short and long term and \change{possibly} it increases the risk-adjusted performance of the portfolio.
 
-These products were initially offered in the Asian markets, see for instance \remove{the reports of} \cite{Chew2011} and \cite{Xue2012}, which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \cite{Morrison2013}.
-At the present day we can observe some new market indices based on the mechanism of the target volatility strategy such as Dow Jones Volatility Control Index, and S\&P 500 Risk Control Index.
-In the recent literature\change{,} \remove{TVSs} \change{TVS-based portfolios} are tested to investigate their performances in term of realized returns, see for instance \cite{Hocquard2013} and \cite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim2018}\change{, where the authors propose the use of artificial neural networks for volatility forecasting volatility to enhance the performance of an asset allocation strategy}.
+\remove{These products} \change{Financial products embedding this strategy} were initially offered in the Asian markets, see for instance \remove{the reports of} \cite{Chew2011} and \cite{Xue2012}, which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \cite{Morrison2013}.
+At the present day\change{,} we can observe some new market indices based on the mechanism of the target volatility strategy such as Dow Jones Volatility Control Index, and S\&P 500 Risk Control Index.
+In the recent literature\change{,} \remove{TVSs are tested to investigate their performances} \change{TVS-based portfolios are investigated with respect to} their performances in term\change{s} of realized returns, see for instance \cite{Hocquard2013} and \cite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim2018}\change{, where the authors propose the use of artificial neural networks for volatility forecasting to enhance the performance of an asset allocation strategy}.
 
-\remove{Moreover, in the present day, in the derivative market new structured financial products that see TVSs as underlying are becoming widely used by practitioners.} \change{The success of TVS-based portfolios among investors has led to the emergence of new structured products with such portfolios as underlyings}.
-Those derivative instruments are known as target volatility options (TVOs) \change{and have become increasingly important in the derivative market.} \change{For these reasons TVOs}  have gained great interest in the derivative pricing literature \change{and research}, see in particular \cite{Albeverio2019}, \cite{DiGraziano2012}, \cite{DiPersio2019}, and \cite{Grasselli2016}.
+\remove{Moreover, in the present day, in the derivative market new structured financial products that see TVSs as underlying are becoming widely used by practitioners.
+Those derivative instruments are known as target volatility options (TVOs) have gained great interest in the derivative pricing literature, see in particular} \change{The success of TVS-based products has led to the emergence of derivative contracts, known as target volatility options (TVO), where the TVS itself is used as underlying asset of the contract.
+In the derivative pricing literature TVOs are reviewed in } \cite{Albeverio2019}, \cite{DiGraziano2012}, \cite{DiPersio2019}, and \cite{Grasselli2016}.
 
-This paper \change{aims to enrich the pricing framework of the TVOs by studying for the first time,} \remove{is the first attempt} to the best of our knowledge, \remove{to study} the funding costs coming from hedging the risky assets underlying the target volatility strategy.
+This paper \change{aims to enrich the pricing framework of the TVOs by studying for the first time,} \remove{is the first attempt} to the best of our knowledge, \remove{to study} the funding costs coming from hedging \change{(or simply hedging costs)} the risky assets underlying the target volatility strategy.
 In particular, we consider the point of view of \remove{a bank} \change{an option writer (for example a bank)} selling a call option to a portfolio manager as protection on the capital invested in a TVS.
-\change{During her activity the} \remove{The} portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS \change{depending on how the market evolves.
-This mechanism is no regulated by the contract, as the manager has decision making power on which asset invest}.
-Since \change{the fund manager's strategy cannot be known {\it a priori} and since} the risky assets have different hedging costs, the \remove{bank} \change{option writer} shall adjust the price of the protection to include \remove{them} \change{these costs} in the worst-case scenario, i.e.
-the most expensive strategy in term of the financing costs.
-Hence, the pricing problem becomes a continuous dynamical control problem over the risky portfolio composition\change{: determine the manager's optimal strategy that leads to the maximum protection price}.
+\change{During her activity the} \remove{The} portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS \change{depending on how the market evolves}.
+Since \change{the fund manager's decisions cannot be known {\it a priori} and since} the risky assets have different hedging costs, the \remove{bank} \change{option writer} shall adjust the price of the protection to include \remove{them} \change{these costs} in the worst-case scenario, i.e.
+the most expensive strategy in term of the \remove{financing} \change{hedging} costs.
+Hence, the pricing problem becomes a continuous dynamical control problem over the risky portfolio composition\change{: the problem of finding the manager's optimal strategy that leads to the maximum protection price}.
 
-In our work we provide to the reader a formal description of the stochastic control problem \change{and we show that it can be solved only numerically considering a general state-dependent dynamics for the risky assets.} \change{However, we prove that} \remove{and most importantly we derive} an analytical solution to the problem \change{exists,} assuming that the risky asset dynamics underlying the TVS  follow a Black and Scholes (BS) model \cite{BlackScholes1973} and that the derivative contract is a European-style option.
-\change{As our first contribution, we derive the BS closed solution into two different ways: applying the Gy\"ongy Lemma and the other through the Hamilton-Jacobi-Bellman equation.}
+In our work\change{,} we provide \remove{to} the reader \change{with} a formal description of the stochastic control problem \change{and we show that it can be solved only numerically considering a general state-dependent dynamics for the risky assets.} \change{However, we prove that} \remove{and most importantly we derive} an analytical solution to the problem \change{exists,} assuming that the risky asset\change{,} dynamics underlying the TVS \change{portfolio}  follow a Black and Scholes (BS) model \cite{BlackScholes1973} and that the derivative contract is a European-style option.
+\change{As our first contribution, we derive the BS closed solution in two different ways: either applying the Gy\"ongy Lemma or the Hamilton-Jacobi-Bellman equation.}
 
@@
 The reason for selecting the LV model is its widespread adoption in the industry to value derivatives, especially those related to asset baskets \cite{Gatheral2006}.
-The LV investigation is based on a Reinforcement Learning approach due to the high-dimensionality problema we deal with.
-In particular we compare the performances of a plain direct policy search gradient-based algorithm with those of the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}.
-At the end of our study, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem under the LV dynamics.}
+The LV investigation is based on a Reinforcement Learning approach due to the high dimensionality of the problem we deal with.
+In particular, we compare the performances of a plain direct policy search gradient-based algorithm \cite{Sutton2018} with those of the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}.
+At the end of our study, we show that the solution found in the BS scenario provides a good approximation of the LV solution, which could be favored by practitioners if computational time is valued.}
 \\
@@
 Then, in Section \ref{sec:Derivative} we introduce the structured class of derivative contracts linked to TVSs, where we describe the arising dynamical control problem for pricing those options.
-Moreover in this section we derive the BS closed solution for European TVOs in two different ways: one applying the Gy\"ongy Lemma and the other \change{through} \remove{by writing} the Hamilton-Jacobi-Bellman equation.
+Moreover\change{,} in this section we derive the BS closed solution for European TVOs in two different ways: one applying the Gy\"ongy Lemma and the other \change{through} \remove{by writing} the Hamilton-Jacobi-Bellman equation.
 In Section \ref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a description of the algorithm we have built.
@@
 
-\section{Target \change{V}olatility \change{S}trategy}\label{sec:TVS}
-As mentioned before, this work aims to enrich the TVS pricing literature by studying the aspect related to the funding costs coming from hedging the risky assets underlying a target volatility portfolio.
-Thus we consider the following scenario: a \remove{bank} \change{option writer} selling a protection to a portfolio manager who has \remove{his} \change{her} capital invested \change{according to} \remove{in} a TVS.
+\section{Target volatility strategy}\label{sec:TVS}
+As mentioned before, this work aims to enrich the TVS pricing literature by studying the aspect related to the funding costs coming from hedging \remove{the risky assets underlying a target volatility portfolio} \change{the underlying risky assets.
+In the following, we will refer to these costs simply as hedging costs}.
+\remove{Thus we} \change{We} consider the following scenario: \remove{a bank} \change{an option writer} selling a protection to a portfolio manager who has \remove{his} \change{her} capital invested \change{according to} \remove{in} a TVS.
 In our case the fund manager has the freedom of changing the relative weights of the risky asset during the life of the TVS; once the allocation strategy is selected then the volatility targeting algorithm rebalances the risky component of the portfolio with the risk-free one in order to keep the overall portfolio volatility close to a target value.
@@
 On the other way, the \remove{bank} \change{writer} is paying such costs since she is actively hedging the call option sold to the manager.
-The \remove{bank} \change{writer} trading activity implemented to actively hedge the option requires funding the collateral procedures of the hedging instruments along with any lending/borrowing fee.
+The \remove{bank} \change{writer} trading activity implemented to \remove{actively} hedge the option requires funding the collateral procedures of the hedging instruments along with any lending/borrowing fee.
+\change{We refer to \cite{Gabrielli2020} for a discussion of how to explicitly calculate hedging costs in equity trading.}
+
 We remark that\change{, upon entering into the contract,}  the choices of the manager trading activity \remove{are} \change{can be seen as} stochastic processes since they \change{are not known {\it a priori} and }will depend on the market evolution.
-Thus, neither the manager strategy nor the \remove{bank} \change{writer} financing costs for hedging the option can be written in the fund-\remove{bank}\change{writer} contract\change{: the only contractual elements are the protection strike and maturity, the trading market of the manager and the target volatility of the TVS portfolio}.
-For this reason, the price of a financial product sold by the \remove{bank} \change{writer} is adjusted to include any valuation adjustment due to the trading activity.
+Thus, neither the manager strategy nor the \remove{bank financing costs for hedging the option} \change{writer hedging costs} can be written in the fund-\change{writer} \remove{bank} contract\change{: the only contractual elements are the protection strike and maturity, the market in which the manager can trade and the target volatility of the TVS portfolio}.
+For this reason, the price of a financial product sold by the \remove{bank} \change{writer} is adjusted to include any valuation adjustment due to the trading activity\change{, such as the hedging costs}.
 
-Our aim is to find the most expensive investing strategy from the point of view of financing costs for the \remove{bank} \change{writer} that the manager could choose in the market.
+Our aim is to find the most expensive investing strategy from the point of view of \remove{financing} \change{hedging} costs for the \remove{bank} \change{writer} that the manager could choose in the market.
 In other words, we want to determine the worst-case scenario for the \remove{bank} \change{option-seller}.
@@
 
-\subsection{The \change{S}trategy Price Process}
+\subsection{The strategy \change{p}rice \change{p}rocess}
 We work on a filtered probability space $\left(\Omega, \mathcal{F}, \{\mathcal{F}_t\}_{t\geq0}, \mathbb{P}  \right)$ satisfying the usual assumptions for a market model, where $\mathbb{P}$ is the physical probability measure \remove{representing the actual distribution of supply and demand shocks on equities prices}.
@@
 where $\pi^i_t$ represents the cumulative contractual-coupon process paid by the security, and $\psi_t^i$ represents the cumulative valuation adjustments.
-Since fund managers allocating TVS usually rely on equity assets, here we use the results of \cite{Gabrielli2020} that analyzes the valuation adjustments for equity products. We can write 
+Since fund managers allocating TVS usually rely on equity assets, here we use the results of \cite{Gabrielli2020} that \remove{analyzes} \change{analyze} the valuation adjustments for equity products.
+We can write
 \begin{equation}
@@
 \label{eq:risk_neutral}\end{equation}
-where $M_t^i$ are martingale under $\mathbb{Q}$.
+where $M_t^i$ are martingale\change{s} under $\mathbb{Q}$.
 If we substitute this expression for the security dynamics into the definition of the strategy we can check that the price process of the strategy is accruing at \change{the following} \remove{a} cash account rate $r_t$ compensated for the fund manager fees
@@
 with $M_t^I$ martingale under $\mathbb{Q}$.
-In this case we observe that $I_t$ depends explicitly both on the valuation adjustments and on the allocation strategy.
+In this case\change{,} we observe that $I_t$ depends explicitly both on the valuation adjustments and on the allocation strategy.
 Indeed, if we substitute the valuation adjustments with their explicit expression \eqref{eq:XVA_equity}, we get
@@
 
-\subsection{The Volatility Targeting Constraint}
+\subsection{The \change{v}olatility \change{t}argeting \change{c}onstraint}
 In a typical TVS, the fund manager selects a risky-asset portfolio with a specific time-dependent allocation strategy expressed by means of the vector of relative weights $\alpha_t$, along with a risk-free asset, which we can identify with the bank account $B_t$.
-In the following, for sake of simplicity in exposition, we consider only total-return securities, namely we set $\pi_t=0$ \change{, which means that the fees received by the stock lender are null since no dividends are paid}.
+In the following, for sake of simplicity in exposition, we consider only total-return securities, namely we set $\pi_t=0$ \change{, which means that the security is not paying dividends}.
 Thus we can write Equation~\eqref{eq:TVS_first} as given by
@@
         \omega_t: \quad \mathrm{Var}_t[dI_t] = \bar{\sigma}^2I_t^2dt \; ,
-\end{equation}
+\label{eq:general_tvs_wght}\end{equation}
 where $\Bar{\sigma}$ is the target volatility value.
-In practice, this means that the fund manager will select a risky-portfolio choosing $\alpha_t$ equities from the universe where \change{s}he can trade and after that, \change{her} \remove{his} choices will be scaled by the automatic target volatility algorithm $\omega_t$.\footnote{We recall that the universe of assets where the manager can trade and the value of $\bar{\sigma}$ are written in the contract.}
+In practice, this means that the fund manager will select a risky-portfolio choosing $\alpha_t$ equities from the universe where \change{s}he can trade\change{,} and after that, \change{her} \remove{his} choices will be scaled by the automatic target volatility algorithm $\omega_t$ \change{\eqref{eq:general_tvs_wght}}.\footnote{We recall that the universe of assets where the manager can trade and the value of $\bar{\sigma}$ are written in the contract.}
 
@@
 \label{eq:TVS_last}\end{equation}
-where we can see, as expected, that the strategy grows at the risk-free rate but for adjustments due to valuation adjustments and fees.
+where we can see, as expected, that the strategy grows at the risk-free rate but \remove{for adjustments due to valuation adjustments and fees} \change{for the valuation adjustments, given by the hedging costs expressed by the term proportional to $\mu_t$ in \eqref{eq:TVS_last}, and the fees, given by the term $\phi_t$ in \eqref{eq:TVS_last}}.
 
@@
 
-\section{Derivative pricing}\label{sec:Derivative}
-In this section, we analyze derivative contracts linked to the TVS described previously.
+\section{Derivative pricing\change{: target volatility options}}\label{sec:Derivative}
+In this section, we analyze \remove{derivative} \change{TVO contracts, namely} contracts linked to the TVS described previously.
 In particular\change{,} we will focus on European style options, and we will show that under appropriate assumptions it is possible to find a closed form solution for the optimal allocation strategy which maximizes the contract price.
@@
 where $D(0,T;\zeta_t)$ is the discount factor with rate $\zeta_t$, inclusive of the derivative valuation adjustments, and $\pi_t$ is the cumulative coupon process paid by the derivative, and it depends on the allocation strategy since in turn the TVS depends on it via the valuation adjustments.
-We take the supremum over the strategies since we do not have any information on the future activity of the fund manager.
+We take the supremum over the strategies since we do not have any information on the future activity of the fund manager \change{and we wish to calculate the worst-case scenario for the option seller, as discussed in the introduction of Section \ref{sec:TVS}}.
 
-\subsection{European Options}
-If the derivative contract depends only on the marginal distribution of $I_T$ at maturity (\change{i.e.,} a European payoff), we are able to calculate an optimal strategy.
+\subsection{European \change{o}ptions}
+If the derivative contract depends only on the marginal distribution of $I_T$ at maturity (\change{i.e.,} a European payoff), we are able to \remove{calculate an optimal strategy} \change{simplify the pricing problem}.
 We consider the following pricing problem
@@
 \begin{proposition}
-A European payoff on the TVS can be calculated by assuming any allocation in the underlying risky basket if all the underlying
+A European \remove{payoff on the TVS} \change{TVO} can be calculated by assuming any allocation in the underlying risky basket if all the underlying
 securities grow under the risk-neutral measure at the risk-free rate without any valuation adjustment, namely if we can write $\mu_t=0$.
@@
 
-\subsection{Stochastic Optimal Control Problem}
+\subsection{Stochastic \change{o}ptimal \change{c}ontrol \change{p}roblem}
 In presence of valuation adjustments, we need to solve the full optimization problem.
@@
 
-\subsection{Black and Scholes \change{Scenario}}\label{subsec:BS_model}
+\subsection{Black and Scholes \change{scenario}}\label{subsec:BS_model}
 In the time-dependent Black and Scholes model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$.
@@
 \begin{proposition}\label{prop:bs}
-When the underlying securities follow a Black and Scholes \change{QUA IO LASCEREI MODEL (REFEREE PROPONE SCENARIO)} model with deterministic rates, the optimal strategy for a non-decreasing European payoff consists in minimizing the local drift function, independently of the current state $I_t$
+When the underlying securities follow a Black and Scholes model with deterministic rates, the optimal strategy for a non-decreasing European payoff consists in minimizing the local drift function, independently of the current state $I_t$
 \begin{equation}
@@
 \end{proposition}
-The absence of stochastic elements in Equation~\eqref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no \change{numerical} simulation needed \change{to solve the control problem}; in fact \remove{one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities} \change{, once the risky-assets market data $\mu(t)$ and $\nu(t)$, it is sufficient to apply Equation~\eqref{eq:BS_optimal_strategy} $\forall t \in \mathcal{T}$  to retrieve the deterministic optimal strategy}.
+The absence of stochastic elements in Equation~\eqref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no \change{numerical} simulation needed \change{to solve the control problem}; in fact\change{,} \remove{one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities} \change{the functions $\mu(t)$ and $\nu(t)$ can be directly derived from the market quotes of the risky assets and of its derivative contracts, then we can apply Equation~\eqref{eq:BS_optimal_strategy} $\forall t \in \mathcal{T}$  to retrieve the deterministic optimal strategy}.
 Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
@@
 
-\change{In the next subsection we derive the analytical solution of the problem \eqref{eq:BS_optimal_strategy} under the assumption of unconstrained allocation strategy $\alpha_t$}
+\change{In the next subsection we derive the analytical solution of the problem \eqref{eq:BS_optimal_strategy} under the assumption of unconstrained allocation strategy $\alpha_t$.}
 
-\subsubsection{Unconstrained Allocation Strategy: Closed Form Solution}
+\subsubsection{Unconstrained \change{a}llocation \change{s}trategy: \change{c}losed \change{f}orm \change{s}olution}
 In absence of constraints on the allocation strategy, we are able to derive a closed form solution to the BS problem \eqref{eq:BS_optimal_strategy}.
@@
 \change{In the following subsection we prove that an analytical solution of \eqref{eq:BS_optimal_strategy} exists also assuming a nonnegative constrained allocation strategy.
-Under this assumption we show that the optimal strategy is a single-asset one called bang bang solution.}
+Under this assumption, we show that the optimal strategy consists in investing in a single asset.
+We name this strategy the bang-bang solution.}
 
-\subsubsection{Active Asset (or Bang Bang) Solution}
-A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry \change{$\mu$} are nonnegative and the only constraint on portfolio weights is nonnegativity, which would mean a long-only strategy by the fund manager.
+\subsubsection{Active \change{a}sset (or \change{b}ang\change{-b}ang) \change{s}olution}
+A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry \change{$\mu$} are nonnegative\change{, as they are costs,} and the only constraint on portfolio weights is nonnegativity, which would mean a long-only strategy by the fund manager.
 
 \begin{lemma}
-	Let $\mu \in \mathbb{R}^n$ be a vector with nonnegative components, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix, and $\Sigma = \nu \tr{\nu}$. Then
+	Let \remove{$\mu \in \mathbb{R}^n$} \change{$\mu \in \mathbb{R}^n_+$} be a vector with nonnegative components, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix, and $\Sigma = \nu \tr{\nu}$.
+Then
 	\begin{equation}\label{eq:inf}
@@
 	\end{equation}
-	if $\bar{\imath}$ is the index which realizes the min, then the infimimum is realized by a vector concentrated on the $\bar{\imath}$ component: $\alpha_i = \delta_{i\bar{\imath}}$. 
+	if $\bar{\imath}$ is the index which realizes the min, then the infimum is realized by a vector concentrated on the $\bar{\imath}$ component: $\alpha_i = \delta_{i\bar{\imath}}$. 
 \end{lemma}
@@
 
-\subsection{Hamilton-Jacobi-Bellman \change{E}quation for \change{T}arget \change{V}olatility \change{O}ptions}\label{subsec:HJB}
+\subsection{Hamilton-Jacobi-Bellman equation for target volatility options}\label{subsec:HJB}
 In this section, we \remove{want to} provide to the reader a formal description of the dynamic problem associated with options on target volatility strategies by writing the Hamilton-Jacobi-Bellman (HJB) equation for the derivative price.
@@
 
-\section{Reinforcement \change{L}earning}\label{sec:RL}
+\section{Reinforcement learning}\label{sec:RL}
 As we have discussed in the previous sections, \changeB{the analytical solution described in Proposition \ref{prop:bs} crucially relies on the assumption that the underlyings evolve according to a Black and Scholes model, and on a monotonicity assumption on the payoff.
@@
 one must resort to numerical approaches to solve the stochastic control problem related to the TVS in the case of general payoffs or risky securities dynamics.
-The standard approach could be to use classical techniques based on backwards recursion \eqref{eq:recursion}-\eqref{eq:recursion2} such as American Monte Carlo \cite{Longstaff2001}.
+The standard approach could be to use classical techniques based on \remove{backwards} \change{backward} recursion \eqref{eq:recursion}-\eqref{eq:recursion2} such as American Monte Carlo \cite{Longstaff2001}.
 However, their performances degrade exponentially as the dimension $n$ of the problem increases, making prohibitively costly finding the solution to the problem.
@@
 \remove{Reinforcement Learning} \change{RL} is a branch of Machine Learning which allows an artificial agent to interact with an environment through actions and observations in order to maximize some notion of cumulative reward \change{called value function}.
-In RL\change{,} the agent is not told which actions to take but instead, it \remove{must} \change{has to} discover by trial and error which are the behaviours  yielding the highest reward.
+In RL\change{,} the agent is not told which actions to take but instead, it \remove{must} \change{has to} discover by trial and error which are the behaviors  yielding the highest reward.
 This is obtained by updating the agent policy $\pi$ which is a mapping from the environment states to the set of actions.
@@
 
-In our work we adopt two learning strategies to compare their performances in terms of overall reward: \remove{an \textit{ad hoc}} \change{a plain} direct policy \change{search} algorithm and the state of the art proximal policy optimization (PPO) developed in \cite{Schulman2017} and \cite{Schulman2016}.
+In our work we adopt two learning strategies to compare their performances in terms of overall reward: \remove{an \textit{ad hoc}} \change{a plain} direct policy \change{search} algorithm \change{\cite{Sutton2018}} and the state of the art proximal policy optimization (PPO) developed in \cite{Schulman2017} and \cite{Schulman2016}.
 
-The first method is \change{well-known in RL literature \cite{Sutton2018}: it is based on the direct update of the agent policy via a gradient-based optimization of the value function.}\remove{a specific algorithm developed by us to fit the problem aim to find the optimal option price; this technique can take place as a direct policy optimization method in the wide taxonomy of Reinforcement Learning algorithms}.
+The first method is \change{well-known in RL literature: it is based on the direct update of the agent policy via a gradient-based optimization of the value function.}\remove{a specific algorithm developed by us to fit the problem aim to find the optimal option price; this technique can take place as a direct policy optimization method in the wide taxonomy of Reinforcement Learning algorithms}.
 We will go into details in Section~\ref{sec:direct_policy}.
@@
 
-\subsection{Direct \change{P}olicy \change{Search} \change{A}pproach}\label{sec:direct_policy}
+\subsection{Direct policy \change{search} approach}\label{sec:direct_policy}
 \change{In RL, the learning phase takes place within a set of so-called episodes, which refer to a single run or instance of interaction between the RL agent and an environment, where the agent takes a sequence of actions to achieve a specific goal.}
@@
  \change{Therefore, during the whole episode, the agent receives a nil reward except at maturity when the reward coincides with the option intrinsic value.
-This choice may seem too daring because the agent receives a real feedback of its actions only at the end of the episode, increasing the probability to obtain a slow learning.
+This choice may seem too daring because the agent receives a real feedback of its actions only at the end of the episode, which could result in a slower learning.
 However, if the agent has learnt the optimal policy $\pi^* = \pi^{\theta^*}$, the average cumulative reward per episode will coincide with the optimal TVO price.}
@@
 In this algorithm, we parameterize the agent policy with a feed forward neural network (FFNN), such that $\theta_j$ coincides with the hidden weights, $s_k$ with the \remove{inputs} \change{input neurons}, and $a_k$ with the output \change{ones}.
-In this way, we are dealing with a deterministic policy, \change{that maps each state directly to a single action without any randomness or probability distribution;} \remove{where its} \change{the map} functional form is given by the neural network.
+In this way, we are dealing with a deterministic policy, \change{that maps each state directly to a single action without any addition randomness or probability distribution;} \remove{where its} \change{the map} functional form is given by the neural network.
 The parameters update is performed as follows: the agent collects a finite batch of experiences interacting with the environment in a set of episodes $\tau$, then the loss-\change{value} function is evaluated as
@@
 where the expectation $\hat{\mathbb{E}}[\dots]$ indicates the empirical average on the batch.
-Then the parameters are updated by plugging the policy into a stochastic gradient ascent algorithm \change{with the aim of maximizing Equation~\eqref{eq:direct_policy_loss}}.
+Then the parameters are updated by plugging the policy into a stochastic gradient ascent algorithm \change{\cite{Sutton2018} with the aim of maximizing Equation~\eqref{eq:direct_policy_loss}}.
 \remove{The choice of Equation~\eqref{eq:direct_policy_loss} is justified by the fact that, when the algorithm will find the optimal set $\theta^*$, then $L(\theta^*)$ will be a good proxy of the optimal option price.}
@@
 
-\subsection{Proximal \change{P}olicy \change{O}ptimization \change{A}pproach}
+\subsection{Proximal policy optimization approach}
 As for the direct policy \change{search} approach, we model the pricing problem considering an episode that takes place on the time-grid $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$.
@@
 
-\section{Numerical \change{I}nvestigations}\label{sec:Numerical_investigations}
+\section{Numerical investigations}\label{sec:Numerical_investigations}
 Here we present the numerical results obtained with our proposed methods. We focus our analysis on a European call option on a TVS with the following product details
@@
 \change{where $I_0$ is the starting value of the TVS-based portfolio.}
-Moreover without loss of generality we consider the case of a\change{n unconstrained} \remove{completely free} allocation strategy $\alpha$. The extension to the constrained case is easy.
+Moreover\change{,} without loss of generality\change{,} we consider the case of a\change{n unconstrained} \remove{completely free} allocation strategy $\alpha$.
+The extension to the constrained case is easy.
 
@@
 
-We recall that in both the algorithms we parameterize the agent policy with an FFNN; thus this is completely characterized by the following hyper-parameters: number of hidden layers, number of neurons per hidden layer, activation function per hidden layer.
+We recall that in both the algorithms we parameterize the agent policy with an FFNN; thus this is completely characterized by the following hyper-parameters: number of hidden layers, number of neurons per hidden layer, \change{and} activation function per hidden layer.
 This is due to the fact that in this \remove{Reinforcement Learning} \change{RL} problem the number of the input neurons is equal to the state space dimension \eqref{eq:state}, while the number of the output ones coincides with the action space dimension \eqref{eq:action}.
-It is well known in the literature that neural networks give better performances in the training phase if the input data are well normalized \cite{Puheim2014}, \cite{Sola1997} .
+It is well known in the literature that neural networks give better performances in the training phase if the input data are well normalized \cite{Puheim2014}, \cite{Sola1997}.
 Thus we choose as state $s_k$ the following normalized block
@@
 In Equation~\eqref{eq:input_state} we have chosen as Markovian state $X_{T_k}$ the martingale term of the securities dynamics.
-In this way we have that in the input block the variables have the same order of magnitude.
+In this way\change{,} we have that in the input block the variables have the same order of magnitude.
 
-\subsection{Black and Scholes: \change{H}yper-parameters \change{F}ine \change{T}uning}
-We use the BS environment as toy model to understand which parameters of the RL algorithm play key roles in the training and testing phase. 
+\subsection{Black and Scholes: hyper-parameters fine tuning}
+We use the BS environment as \change{a} toy model to understand which parameters of the RL algorithm play key roles in the training and testing phase. 
 Our first approach has been the direct policy one since it represents a natural way to tackle the problem: since our goal is to find the allocation 
@@
 
-We try to investigate the following hyper-parameters: the FFNN architecture, in particular which feature between the depth and the width of the network is more important, the activation function and the learning rate of the optimizer.
+We try to investigate the following hyper-parameters: the FFNN architecture, in particular which feature between the depth and the width of the network is more important, the activation function\change{,} and the learning rate of the optimizer.
 Moreover, we compare the performances of two well-known optimizers in Machine Learning literature: Nadam \cite{Dozat2016} and RMSprop \cite{Hiton2012}.
@@
 
-\subsection{Local \change{V}olatility \change{D}ynamics}
+\subsection{Local volatility dynamics}
 In this section, we study the TVS control problem assuming a local volatility model for the dynamics of the risky assets.
@@
 In all the PPO cases, the trained agents seem to be stuck in the sub-optimal maximum drift strategy policy.
-For the case of the immediate reward function (Figure~\ref{fig:PPO_results} on the right), we have that the learning curves converge to a saturation value that is not compatible with the baseline price; this is due to the fact that with $\lambda,\gamma \neq 1$ we introduce a bias in the problem.
-Because of that, we need to run a test MC simulation with freezed agents and $\lambda=\gamma=1$ to obtain a result to compare with the baseline price.
+For the case of the immediate reward function (Figure~\ref{fig:PPO_results} on the right), we \remove{have that the learning curves converge to a saturation value that is not compatible with the baseline price; this is due to the fact that with $\lambda,\gamma \neq 1$ we introduce a bias in the problem} \change{can not compare the saturation values reached by the learning curves with the baseline price, since the former are discounted due to $\lambda,\gamma \neq 1$}.
+\remove{Because of that, we need to run a test MC simulation with freezed agents and $\lambda=\gamma=1$ to obtain a result to compare with the baseline price} \change{
+Thus, to compare the performance of PPO with the baseline price in the test phase and ensure consistency in unit of measurement, it is necessary to conduct a MC simulation with freezed agents and set $\lambda$ and $\gamma$ values to 1}.
 We report the results in Table~\ref{tab:mc_results_lv}.
 The \remove{Reinforcement Learning} \change{RL} agents we have trained with PPO give all prices compatible with the baseline one.
+
 \begin{table}
@@
 
-\section{Conclusion and \change{F}urther \change{D}evelopments}
+\section{Conclusion and further developments}
 In this paper, we described a non-trivial control problem related to derivative contracts on target volatility strategies.

### 2023-03-05 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
Update section 3.2

@@
 \newcommand{\removeB}[1]{{\color{blue} \soutpars{{#1}}}}%{}
+\newcommand{\changeE}[1]{{\color{purple} {#1}}}%{{#1}}
+\newcommand{\removeE}[1]{{\color{purple} \soutpars{{#1}}}}%{}
 \newcommand\hcancel[2][red]{\setbox0=\hbox{$#2$}%
@@
 \email{\{Roberto.Daluiso, Andrea.Pallavicini\}@intesasanpaolo.com} \and
-Financial Engineering, Marketz S.p.A.,\\
-Corso Europa, 2, 20122 Milan, Italy\\
-\email{Emanuele.Nastasi@marketz.eu} \and \change{Financial Risk Methodologies, Unicredit S.p.A.,\\ Piazza Gae Aulenti, 3, 20154 Milan, Italy \\ \email{Stefano.Polo@unicredit.eu}}
+\changeE{Digital Development Sales \& Trading Milan, Unicredit Bank AG Milano,}\\
+\changeE{Piazza Gae Aulenti, 4, 20154 Milan, Italy}\\
+\email{\changeE{Emanuele.Nastasi@unicredit.eu}} \and \change{Financial Risk Methodologies, Unicredit S.p.A.,\\ Piazza Gae Aulenti, 3, 20154 Milan, Italy \\ \email{Stefano.Polo@unicredit.eu}}
 }
@@
 \label{eq:recursion}\end{equation}
-where $V_{T_k}$ is the option value at time $T_k$ and $X$ is any Markovian state such that the drift and the diffusion coefficient of $I_t$ are a function of $\left(X_t,I_t,\alpha_t\right)$.
-We calculate $I_{T_k}\left(\alpha_{T_{k-1}}\right)$ for any given strategy $\alpha_{T_{k-1}}$ by a suitable discretization of \eqref{eq:TVS_last} starting from $X_{T_{k-1}}$ and $I_{T_{k-1}}$.
-
-Thus the derivative price is given by:
+where $V_{T_k}$ is the option value at time $T_k$ and $X$ is any Markovian state such that the drift and the diffusion coefficient of $I_t$ are a function of $\left(X_t,I_t,\alpha_t\right)$ \changeE{and we indicated with $I_{T_k}\left(\alpha\right)$ the value of the strategy on the point $T_{k}$ for a given choice of the weights $\alpha$ at the previous step on the time grid.} We calculate \changeE{the strategy value for the optimal weights $\alpha_{T_{k-1}}$, namely} $I_{T_k}\left(\alpha_{T_{k-1}}\right)$ \removeE{for any given strategy $\alpha_{T_{k-1}}$} by a suitable discretization of \eqref{eq:TVS_last} starting from $X_{T_{k-1}}$ and $I_{T_{k-1}}$\removeE{.} \changeE{and}
+\removeE{Thus} \changeE{thus} the derivative price is given by:
 \begin{equation}

### 2023-03-06 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
Expand on over-fitting detection

@@
 Firstly we have performed a grid search on the learning rate starting value and we have found $10^{-3}$ a good choice in terms of speed of learning and avoiding over-fitting.
-\changeB{Note that since our environment is simulated, over-fitting is easily detected by cross-validation on a set of freshly-generated scenarios which were not seen during optimization.}
+\changeB{Note that to detect over-fitting one typically compares the performance on one or more training sets with that on corresponding validation sets.
+Since our environment is simulated, there is no need to explicitly set aside training data for such task, nor to perform $K$-fold validation: indeed we can always generate for the purpose a set of new scenarios which were not seen during optimization.}
  \begin{figure}[h!]

### 2023-03-21 tex/m2ef/dnpp_referee_response.tex
Expand on over-fitting detection
minor fixes

@@
 
-We list here a detailed list of responses Referee's major remarks. We address the minor remarks directly in the text.
+We list here a detailed list of responses to Referee's major remarks. We address the minor remarks directly in the text.
 
@@
 
- We have conducted a comprehensive review of the crucial components of the Abstract and Introduction sections to ensure that the reader is presented with a lucid and concise overview of the problem being addressed.
-In the Introduction, our focus was primarily on establishing a stronger connection between TVSs and the derivatives market, which are based on these underlying assets.
-Furthermore, we have improved the description of the problem addressed in this paper, with a particular emphasis on the methods employed and the resultant findings.
-It is hoped that this will enable the reader to contextualize our work more effectively within the pertinent literature.
+ In the revised manuscript, we conduct a comprehensive review of the crucial components of the Abstract and Introduction sections to ensure that the reader is presented with a lucid and concise overview of the problem being addressed.
+In the Introduction, our focus is primarily on establishing a stronger connection between TVSs and the derivatives market, which are based on these underlying assets.
+Furthermore, we improve the description of the problem addressed in this paper, with a particular emphasis on the methods employed and the results.
+We also welcome the referee's suggestion in the minor remarks to improve the title of the manuscript and the related Keywords.
+We hope that this will enable the reader to contextualize our work more effectively within the pertinent literature.
 
@@
 The analytical solution applies only for a specific choice of the model for the dynamics of the underlying  assets and under suitable hypotheses on the payoff.
-We added equation (52) and a sentence in the first paragraph of Section 4 to stress this point.
-Furthermore, we have enhanced the clarity of this point in the Introduction section where we discuss the contribution of this work.
+We add equation (52) and a sentence in the first paragraph of Section 4 to stress this point.
+Furthermore, we enhance the clarity of this point in the Introduction section where we discuss the contribution of this work.
 
@@
  The three \quotations{intuitive strategies} presented in this paper represent plausible expert-based approaches that a practitioner may choose to adopt when addressing the control problem.
-We have provided further elucidation on this matter in Subsection 3.3, where we have also incorporated the financial implications of these strategies.
+We provide further elucidation on this matter in Subsection 3.3, where we also incorporate the financial implications of these strategies.
 
@@
 
- The intention behind the sentence was to apprise the reader that the first algorithm was developed by the authors themselves to tackle the problem at hand.
+ We rephrase the given sentence to clarify that we adopt an existing algorithm well-known in the Reinforcement Learning literature to tackle the problem at hand.
 In formal terms, it can be classified as a direct policy search algorithm that relies on gradient-based optimization of the Monte Carlo outcome obtained from numerical simulations.
-This kind of algorithms is well-known in the Reinforcement Learning literature.
-We have revised this statement to ensure that its meaning is conveyed with greater clarity.
 
@@
 
-We added to the second paragraph of Section 5.1 a sentence which explains how over-fitting is detected.
+We add to the second paragraph of Section 5.1 a sentence which explains that over-fitting is detected on a generated validation set.
 

### 2023-03-21 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
minor fixes

@@
 \email{\{Roberto.Daluiso, Andrea.Pallavicini\}@intesasanpaolo.com} \and
-\changeE{Digital Development Sales \& Trading Milan, Unicredit Bank AG Milano,}\\
-\changeE{Piazza Gae Aulenti, 4, 20154 Milan, Italy}\\
-\email{\changeE{Emanuele.Nastasi@unicredit.eu}} \and \change{Financial Risk Methodologies, Unicredit S.p.A.,\\ Piazza Gae Aulenti, 3, 20154 Milan, Italy \\ \email{Stefano.Polo@unicredit.eu}}
-}
+\change{Capital Market Technologies, swissQuant Group AG,}\\
+\change{Kuttelgasse 7, 8001 Z\"urich, Switzerland}\\
+\email{\change{Nastasi@swissquant.com}} \and Financial Engineering,
+X-Numeris S.r.l.,\\ Via Turro, 7 , 20127 Milan, Italy\\
+\change{\email{Sp.Stefano.Polo@gmail.com}}}
 
@@
 
-We highlight that, to derive the dynamics expressed in Equation~\eqref{eq:TVS_last}, we have not made any assumptions on the risky allocation strategy; thus all this argument is valid for any constraints on the process $\alpha_t$\change{ as, for instance, the case of contracts in which the fund manager is restricted to holding only long positions meaning that $\alpha_t^i>0$ $\forall i = 1,\dots,n$, or situations where the combined total of long and short trades must be capped at a specific contractual amount}.
+We highlight that, to derive the dynamics expressed in Equation~\eqref{eq:TVS_last}, we have not made any assumptions on the risky allocation strategy; thus all this argument is valid for any constraints on the process $\alpha_t$\change{ as, for instance, the case of contracts in which the fund manager is restricted to holding only long positions meaning that $\alpha_t^i\geq0$ $\forall i = 1,\dots,n$, or situations where the combined total of long and short trades must be capped at a specific contractual amount}.
 
@@
 \label{eq:recursion}\end{equation}
-where $V_{T_k}$ is the option value at time $T_k$ and $X$ is any Markovian state such that the drift and the diffusion coefficient of $I_t$ are a function of $\left(X_t,I_t,\alpha_t\right)$ \changeE{and we indicated with $I_{T_k}\left(\alpha\right)$ the value of the strategy on the point $T_{k}$ for a given choice of the weights $\alpha$ at the previous step on the time grid.} We calculate \changeE{the strategy value for the optimal weights $\alpha_{T_{k-1}}$, namely} $I_{T_k}\left(\alpha_{T_{k-1}}\right)$ \removeE{for any given strategy $\alpha_{T_{k-1}}$} by a suitable discretization of \eqref{eq:TVS_last} starting from $X_{T_{k-1}}$ and $I_{T_{k-1}}$\removeE{.} \changeE{and}
-\removeE{Thus} \changeE{thus} the derivative price is given by:
+where $V_{T_k}$ is the option value at time $T_k$ and $X$ is any Markovian state such that the drift and the diffusion coefficient of $I_t$ are a function of $\left(X_t,I_t,\alpha_t\right)$ \change{and we indicated with $I_{T_k}\left(\alpha\right)$ the value of the strategy at time $T_{k}$ for a given choice of the weights $\alpha$ at time $T_{k-1}$.} We calculate \change{the strategy value for the optimal weights $\alpha_{T_{k-1}}$, namely} $I_{T_k}\left(\alpha_{T_{k-1}}\right)$ \remove{for any given strategy $\alpha_{T_{k-1}}$} by a suitable discretization of \eqref{eq:TVS_last} starting from $X_{T_{k-1}}$ and $I_{T_{k-1}}$.
+\change{Thus, once collected the elements $\{X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\}$,} the derivative price is given by:
 \begin{equation}
@@
 \end{proposition}
-The absence of stochastic elements in Equation~\eqref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no \change{numerical} simulation needed \change{to solve the control problem}; in fact\change{,} \remove{one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities} \change{the functions $\mu(t)$ and $\nu(t)$ can be directly derived from the market quotes of the risky assets and of its derivative contracts, then we can apply Equation~\eqref{eq:BS_optimal_strategy} $\forall t \in \mathcal{T}$  to retrieve the deterministic optimal strategy}.
+The absence of stochastic elements in Equation~\eqref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no \change{numerical} simulation needed \change{to solve the control problem}; in fact\change{,} \remove{one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities} \change{the functions $\mu(t)$ and $\nu(t)$ can be directly derived from the market quotes of the risky assets and of its derivative contracts, and then we can apply Equation~\eqref{eq:BS_optimal_strategy} $\forall t \in \mathcal{T}$  to retrieve the deterministic optimal strategy}.
 Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
@@
 \item S$_B$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)$; \change{in this case we consider a fund manager who rebalances the weights $\alpha(t)$ over the life of the portfolio by selecting the asset with the highest expected return at a given point in time;}
-\item S$_C$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)/\|\nu(t)\|$\remove{.}\change{; the strategy is akin to S$_B$, with the exception being that the manager endeavors to maximize the expected return that has been adjusted to account for the target volatility mechanism..}
+\item S$_C$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)/\|\nu(t)\|$\remove{.}\change{; the strategy is akin to S$_B$, with the exception being that the manager endeavors to maximize the expected return that has been adjusted to account for the target volatility mechanism.}
 \end{itemize}
@@
 \subsubsection{Active \change{a}sset (or \change{b}ang\change{-b}ang) \change{s}olution}
-A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry \change{$\mu$} are nonnegative\change{, as they are costs,} and the only constraint on portfolio weights is nonnegativity, which would mean a long-only strategy by the fund manager.
+A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry \change{$\mu$} are nonnegative\change{, so they represent true hedging costs and not benefits,} and the only constraint on portfolio weights is nonnegativity, which would mean a long-only strategy by the fund manager.
 
@@
 Conversely, if we deal with a dynamic model for which the derivative contract $V$ depends on $X$ then the volatility versor term, namely the second one, in the $\max$ operator of Equation~\eqref{eq:HJB_general} is no longer zero and thus one must solve the entire control problem numerically.
-\changeB{Indeed, one can check by direct substitution that the natural generalization of \eqref{eq:HJB_solution}
+\change{Indeed, one can check by direct substitution that the natural generalization of \eqref{eq:HJB_solution}
 \begin{equation}
@@
 
-In \removeB{this work}\changeB{what follows} we tackle the non-trivial case of a local volatility \remove{(LV)} model for the $S_t$-dynamics, such that $\nu_t = \nu(t,S_t)$.
+In \remove{this work}\change{what follows} we tackle the non-trivial case of a local volatility \remove{(LV)} model for the $S_t$-dynamics, such that $\nu_t = \nu(t,S_t)$.
 We have chosen the LV model since it is well known in financial literature and among practitioners \change{\cite{Gatheral2006}}.
@@
 \section{Reinforcement learning}\label{sec:RL}
-As we have discussed in the previous sections, \changeB{the analytical solution described in Proposition \ref{prop:bs} crucially relies on the assumption that the underlyings evolve according to a Black and Scholes model, and on a monotonicity assumption on the payoff.
+As we have discussed in the previous sections, \change{the analytical solution described in Proposition \ref{prop:bs} crucially relies on the assumption that the underlyings evolve according to a Black and Scholes model, and on a monotonicity assumption on the payoff.
 Therefore,}
@@
 This is obtained by updating the agent policy $\pi$ which is a mapping from the environment states to the set of actions.
-Thus RL is independent of pre-collected data as opposed to other Machine Learning techniques\change{, that have to be fed a pre-existing dataset to learn from}.
+Thus RL is independent of pre-collected data as opposed to other Machine Learning techniques\change{, that have to be fed with a pre-existing dataset to learn from}.
 Because of its nature, RL has been successful in quantitative finance for solving control problems; among the most important RL applications in this field, we refer to \cite{Deng2017} as the pioneers in studying self-taught reinforcement trading problems, while to \cite{Halperin2020} and \cite{Kolm2019} for hedging derivatives with RL under market frictions.
@@
 In this algorithm, we parameterize the agent policy with a feed forward neural network (FFNN), such that $\theta_j$ coincides with the hidden weights, $s_k$ with the \remove{inputs} \change{input neurons}, and $a_k$ with the output \change{ones}.
-In this way, we are dealing with a deterministic policy, \change{that maps each state directly to a single action without any addition randomness or probability distribution;} \remove{where its} \change{the map} functional form is given by the neural network.
+In this way, we are dealing with a deterministic policy, \change{that maps each state directly to a single action without any additional randomness or probability distribution;} \remove{where its} \change{the map} functional form is given by the neural network.
 The parameters update is performed as follows: the agent collects a finite batch of experiences interacting with the environment in a set of episodes $\tau$, then the loss-\change{value} function is evaluated as
@@
 In this case, the reward function does not suffer of nil values for $0<T_k<T$: the RL agent always gets a feedback from the environment for its choices.
-\remove{The} \change{Note that the} hyper-parameter $\gamma$ plays the role of a discount factor in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future.
+\remove{The} \change{Notice that the} hyper-parameter $\gamma$ plays the role of a discount factor in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future.
 If we take the cumulative reward per episode and set the PPO parameters\footnote{We refer to \cite{Schulman2016} for a more detailed description for the generalized advantage estimation parameter $\lambda$.} $\gamma =\lambda =1$ we obtain
@@
 In the \textsc{OpenAI Baselines} implementation of the PPO, the agent policy is parameterized again by a neural network; as for the previous method we have chosen an FFNN.
-\removeB{In particular with PPO we deal with a stochastic policy, whose functional form is} \changeB{More precisely, training is based on a stochastic policy, i.e.
+\remove{In particular with PPO we deal with a stochastic policy, whose functional form is} \change{More precisely, training is based on a stochastic policy, i.e.
 the action is drawn from a probability law, and the neural network is used to determine its parameters.
-In settings like ours where the action has a continuous domain, such law is} a multivariate diagonal Gaussian distribution where the mean $\mu^\theta(s_k)$ is the output vector of the FFNN and the log-standard deviation $\log\sigma$ is an \removeB{external} \changeB{additional trainable} parameter:
+In settings like ours where the action has a continuous domain, such law is} a multivariate diagonal Gaussian distribution where the mean $\mu^\theta(s_k)$ is the output vector of the FFNN and the log-standard deviation $\log\sigma$ is an \remove{external} \change{additional trainable} parameter:
 \begin{equation}
@@
 The idea is that the log-standard deviation will be higher at the beginning of the training phase in order to guarantee a good exploration of the action space while it will be lower at the end to avoid too much noise in the proximity of the optimal policy.
-\changeB{Eventually at the end of training, the standard deviation is put to zero, so that the action becomes deterministically equal to the output of the FFNN.}
+\change{Eventually at the end of training, the standard deviation is put to zero, so that the action becomes deterministically equal to the output of the FFNN.}
 
@@
 Firstly we have performed a grid search on the learning rate starting value and we have found $10^{-3}$ a good choice in terms of speed of learning and avoiding over-fitting.
-\changeB{Note that to detect over-fitting one typically compares the performance on one or more training sets with that on corresponding validation sets.
+\change{Note that to detect over-fitting one typically compares the performance on one or more training sets with that on corresponding validation sets.
 Since our environment is simulated, there is no need to explicitly set aside training data for such task, nor to perform $K$-fold validation: indeed we can always generate for the purpose a set of new scenarios which were not seen during optimization.}
@@
 This additional dependency of volatility makes the problem of finding the optimal strategy non-trivial\change{, as a closed form formula cannot be derived under this scenario}; in fact if we consider the whole \remove{securities} \change{volatility} smiles \change{of the securities,} then the second-order term in the HJB equation \eqref{eq:HJB_general} \remove{is not zero and thus a closed formula for $\alpha^*$ is not available anymore} \change{does not elide and consequently the only way to derive $\alpha^*$ is through a numerical method}.
-\remove{Because of that, one must resort to numerical techniques to recover the problem solution.} \remove{Unlike the BS model where the strategy depends only on time, in LV dynamics there is no unnecessary information provided by the state block in Equation~\eqref{eq:state} to take the optimal action} \change{The additional dependence on the state $S_t$ under LV dynamics will cause the agent to consider also the $X_t$ component value in the  state block \eqref{eq:state}, unlike under the BS scenario where this information is redundant as the optimal strategy was function only of time \eqref{eq:TVS_last}}.
+\remove{Because of that, one must resort to numerical techniques to recover the problem solution.} \remove{Unlike the BS model where the strategy depends only on time, in LV dynamics there is no unnecessary information provided by the state block in Equation~\eqref{eq:state} to take the optimal action} \change{The additional dependence on the state $S_t$ under LV dynamics will cause the agent to consider also the $X_t$ component value in the  state block \eqref{eq:state}, unlike under the BS scenario where this information is redundant as the optimal strategy was function only of time (see Proposition \ref{prop:bs})}.
 Here we consider the same market data ($r_t$, $\mu_t$ and $\nu_t$) as in the Black and Scholes environment to study how the optimal solution changes with the dynamics model.
@@
 Thus we train another agent with the direct policy learning by parameterizing its action with the baseline strategy and choosing a smaller learning rate of $10^{-4}$.
-With this parameterization, \change{the RL agent actions are computed by summing} at each observational time $T_k\in\mathcal{T}$ \remove{the risky allocation strategy is obtained by summing} the FFNN output with the Equation~\eqref{eq:analytical_solution_free_strategy}\change{, giving rise to the manager's risky allocation strategy}.
+With this parameterization, \change{the RL agent actions, constituting the risky allocation strategy, are computed by summing} at each observational time $T_k\in\mathcal{T}$ \remove{the risky allocation strategy is obtained by summing} the FFNN output with the Equation~\eqref{eq:analytical_solution_free_strategy}.
 Even in this case, the corresponding learning curve (solid black line in Figure~\ref{fig:result_lv_direct_policy}) is stuck in the maximum drift strategy.

### 2023-07-03 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
start removing corrections

@@
 
-\title{\change{Optimal strategies for options on target volatility funds}\thanks{\change{Our sincere thanks go to Marco Bianchetti and Diego Pierluigi Giovannini of Intesa Sanpaolo Milan, as well as to the Italian computing center Cineca for its support on our numerical simulations.
-The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.}}}
+	\title{Optimal strategies for options on target volatility funds\thanks{Our sincere thanks go to Marco Bianchetti and Diego Pierluigi Giovannini of Intesa Sanpaolo Milan, as well as to the Italian computing center Cineca for its support on our numerical simulations.
+The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.}}
 
-\titlerunning{\change{Optimal strategies for options on target volatility funds}}
+	\titlerunning{Optimal strategies for options on target volatility funds}
 
@@
 	\email{\{Roberto.Daluiso, Andrea.Pallavicini\}@intesasanpaolo.com} \and
-\change{Capital Market Technologies, swissQuant Group AG,}\\
-\change{Kuttelgasse 7, 8001 Z\"urich, Switzerland}\\
-\email{\change{Nastasi@swissquant.com}} \and Financial Engineering,
+	Capital Market Technologies, swissQuant Group AG,\\
+	Kuttelgasse 7, 8001 Z\"urich, Switzerland\\
+	\email{Nastasi@swissquant.com} \and Financial Engineering,
 	X-Numeris S.r.l.,\\ Via Turro, 7 , 20127 Milan, Italy\\
-\change{\email{Sp.Stefano.Polo@gmail.com}}}
+	\email{Sp.Stefano.Polo@gmail.com}}
 
@@
 	\begin{abstract}
-We deal with \remove{the funding costs} \change{a stochastic optimal control problem} rising from hedging the risky securities underlying a target volatility strategy \remove{(TVS)}, a portfolio \remove{of risky assets and a risk-free one dynamically rebalanced in order to keep the} \change{whose asset-allocation is adjusted to maintain the} realized volatility of the portfolio \remove{on} \change{at} a certain level.
-\change{We consider the point of view of a derivative writer selling an option contract as protection to a portfolio manager on the invested capital.} The uncertainty in the \remove{TVS} risky portfolio composition along with the difference in hedging costs \change{of its} \remove{for each} component\change{s} requires to \remove{solve a control problem to evaluate the option} \change{adjust the protection} price\remove{s} \change{to include these costs in the worst-case scenario for the seller}.
-We derive an analytical solution of the problem in \remove{the} \change{a} Black and Scholes \remove{(BS)} scenario.
-Then\change{,} we use Reinforcement Learning \remove{(RL)} techniques to determine the fund composition leading to the \remove{most conservative price} \change{optimal policy} under the local volatility \remove{(LV)} model, for which an {\it a priori} solution is not available.
-We show how the performances of the \remove{RL agents} \change{numerical solution}  are compatible with those obtained by applying path-wise the \remove{BS} analytical \change{solution previously derived.} \remove{strategy to the TVS dynamics, which therefore appears competitive also in the LV scenario}
+	We deal with a stochastic optimal control problem rising from hedging the risky securities underlying a target volatility strategy, a portfolio whose asset-allocation is adjusted to maintain the realized volatility of the portfolio at a certain level.
+We consider the point of view of a derivative writer selling an option contract as protection to a portfolio manager on the invested capital.
+The uncertainty in the risky portfolio composition along with the difference in hedging costs of its components requires to adjust the protection price to include these costs in the worst-case scenario for the seller.
+We derive an analytical solution of the problem in a Black and Scholes scenario.
+Then, we use Reinforcement Learning techniques to determine the fund composition leading to the optimal policy under the local volatility model, for which an {\it a priori} solution is not available.
+We show how the performances of the numerical solution are compatible with those obtained by applying path-wise the analytical solution previously derived.
 
@@
 
-\noindent {\bf Keywords.} \change{Stochastic control problem, reinforcement learning, target volatility, option pricing, hedging costs.}
+	\noindent {\bf Keywords.} Stochastic control problem, reinforcement learning, target volatility, option pricing, hedging costs.
 
@@
 	\section{Introduction}
-In the recent years portfolio managers \remove{are} \change{were} exposed to very low interest rates and quickly changing market volatilities.
-An effective solution to control risks under such an environment is given by target volatility strategies (TVSs)\change{,} \remove{(}also known as constant volatility targeting\remove{)}\change{,} which are able to preserve the portfolio at a predetermined level of volatility.
-A TVS is \change{a strategy for managing} a portfolio of risky assets (typically  equities) and a risk-free \change{one} \remove{asset} dynamically re-balanced with the aim of maintaining the overall portfolio volatility level \remove{closed} \change{close} to some target value.
-\change{The strategy is designed to achieve a stable level of volatility for the underlying portfolio in any market scenario, and to allow a free selection of the relative allocation weights among the risky assets.
-}The constant volatility approach can help investors to obtain desired risk exposures over the short and long term and \change{possibly} it increases the risk-adjusted performance of the portfolio.
+	In the recent years portfolio managers were exposed to very low interest rates and quickly changing market volatilities.
+An effective solution to control risks under such an environment is given by target volatility strategies (TVSs), also known as constant volatility targeting, which are able to preserve the portfolio at a predetermined level of volatility.
+A TVS is a strategy for managing a portfolio of risky assets (typically  equities) and a risk-free one dynamically re-balanced with the aim of maintaining the overall portfolio volatility level close to some target value.
+The strategy is designed to achieve a stable level of volatility for the underlying portfolio in any market scenario, and to allow a free selection of the relative allocation weights among the risky assets.
+The constant volatility approach can help investors to obtain desired risk exposures over the short and long term and possibly it increases the risk-adjusted performance of the portfolio.
 
-\remove{These products} \change{Financial products embedding this strategy} were initially offered in the Asian markets, see for instance \remove{the reports of} \cite{Chew2011} and \cite{Xue2012}, which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \cite{Morrison2013}.
-At the present day\change{,} we can observe some new market indices based on the mechanism of the target volatility strategy such as Dow Jones Volatility Control Index, and S\&P 500 Risk Control Index.
-In the recent literature\change{,} \remove{TVSs are tested to investigate their performances} \change{TVS-based portfolios are investigated with respect to} their performances in term\change{s} of realized returns, see for instance \cite{Hocquard2013} and \cite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim2018}\change{, where the authors propose the use of artificial neural networks for volatility forecasting to enhance the performance of an asset allocation strategy}.
+	Financial products embedding this strategy were initially offered in the Asian markets, see for instance \cite{Chew2011} and \cite{Xue2012}, which highlight the pros and cons for investors, to be adopted in the following years in many other markets in North America and Europe as depicted in \cite{Morrison2013}.
+At the present day, we can observe some new market indices based on the mechanism of the target volatility strategy such as Dow Jones Volatility Control Index, and S\&P 500 Risk Control Index.
+	In the recent literature, TVS-based portfolios are investigated with respect to their performances in terms of realized returns, see for instance \cite{Hocquard2013} and \cite{Perchet2016}, and the soundness of the volatility targeting algorithm, as described in \cite{Kim2018}, where the authors propose the use of artificial neural networks for volatility forecasting to enhance the performance of an asset allocation strategy.
 
-\remove{Moreover, in the present day, in the derivative market new structured financial products that see TVSs as underlying are becoming widely used by practitioners.
-Those derivative instruments are known as target volatility options (TVOs) have gained great interest in the derivative pricing literature, see in particular} \change{The success of TVS-based products has led to the emergence of derivative contracts, known as target volatility options (TVO), where the TVS itself is used as underlying asset of the contract.
-In the derivative pricing literature TVOs are reviewed in } \cite{Albeverio2019}, \cite{DiGraziano2012}, \cite{DiPersio2019}, and \cite{Grasselli2016}.
+	The success of TVS-based products has led to the emergence of derivative contracts, known as target volatility options (TVO), where the TVS itself is used as underlying asset of the contract.
+In the derivative pricing literature TVOs are reviewed in \cite{Albeverio2019}, \cite{DiGraziano2012}, \cite{DiPersio2019}, and \cite{Grasselli2016}.
 
-This paper \change{aims to enrich the pricing framework of the TVOs by studying for the first time,} \remove{is the first attempt} to the best of our knowledge, \remove{to study} the funding costs coming from hedging \change{(or simply hedging costs)} the risky assets underlying the target volatility strategy.
-In particular, we consider the point of view of \remove{a bank} \change{an option writer (for example a bank)} selling a call option to a portfolio manager as protection on the capital invested in a TVS.
-\change{During her activity the} \remove{The} portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS \change{depending on how the market evolves}.
-Since \change{the fund manager's decisions cannot be known {\it a priori} and since} the risky assets have different hedging costs, the \remove{bank} \change{option writer} shall adjust the price of the protection to include \remove{them} \change{these costs} in the worst-case scenario, i.e.
-the most expensive strategy in term of the \remove{financing} \change{hedging} costs.
-Hence, the pricing problem becomes a continuous dynamical control problem over the risky portfolio composition\change{: the problem of finding the manager's optimal strategy that leads to the maximum protection price}.
+	This paper aims to enrich the pricing framework of the TVOs by studying for the first time, to the best of our knowledge, the funding costs coming from hedging (or simply hedging costs) the risky assets underlying the target volatility strategy.
+In particular, we consider the point of view of an option writer (for example a bank) selling a call option to a portfolio manager as protection on the capital invested in a TVS.
+During her activity the portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS depending on how the market evolves.
+Since the fund manager's decisions cannot be known {\it a priori} and since the risky assets have different hedging costs, the option writer shall adjust the price of the protection to include these costs in the worst-case scenario, i.e.
+the most expensive strategy in term of the hedging costs.
+Hence, the pricing problem becomes a continuous dynamical control problem over the risky portfolio composition: the problem of finding the manager's optimal strategy that leads to the maximum protection price.
 
-In our work\change{,} we provide \remove{to} the reader \change{with} a formal description of the stochastic control problem \change{and we show that it can be solved only numerically considering a general state-dependent dynamics for the risky assets.} \change{However, we prove that} \remove{and most importantly we derive} an analytical solution to the problem \change{exists,} assuming that the risky asset\change{,} dynamics underlying the TVS \change{portfolio}  follow a Black and Scholes (BS) model \cite{BlackScholes1973} and that the derivative contract is a European-style option.
-\change{As our first contribution, we derive the BS closed solution in two different ways: either applying the Gy\"ongy Lemma or the Hamilton-Jacobi-Bellman equation.}
+	In our work, we provide the reader with a formal description of the stochastic control problem and we show that it can be solved only numerically considering a general state-dependent dynamics for the risky assets.
+However, we prove that an analytical solution to the problem exists, assuming that the risky asset, dynamics underlying the TVS portfolio follow a Black and Scholes (BS) model \cite{BlackScholes1973} and that the derivative contract is a European-style option.
+As our first contribution, we derive the BS closed solution in two different ways: either applying the Gy\"ongy Lemma or the Hamilton-Jacobi-Bellman equation.
 
-\change{In the second part of our work we tackle the problem in the general case of a state-dependent dynamics for the risky assets for which only the numerical approach is admissible.}
-\remove{Although from a theoretical point of view this solution represents a sub-optimal strategy when dealing with generic dynamics for the risky assets, we show that the BS solution provides a strong baseline as price to practitioners when testing it against numerical methods with different levels of sophistication for solving the control problem}.
-More precisely we numerically study the problem \remove{in the general case of  a local volatility dynamics \cite{Derman1994}, \cite{Dupire1994} by adopting a Reinforcement Learning approach, in particular an \textit{ad hoc} vanilla direct policy algorithm and the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}} \change{assuming a local volatility (LV) dynamics \cite{Derman1994}, \cite{Dupire1994} \change{for the risky assets}.
+	In the second part of our work we tackle the problem in the general case of a state-dependent dynamics for the risky assets for which only the numerical approach is admissible.
+	More precisely we numerically study the problem assuming a local volatility (LV) dynamics \cite{Derman1994}, \cite{Dupire1994} for the risky assets.
 The reason for selecting the LV model is its widespread adoption in the industry to value derivatives, especially those related to asset baskets \cite{Gatheral2006}.
@@
 In particular, we compare the performances of a plain direct policy search gradient-based algorithm \cite{Sutton2018} with those of the more sophisticated proximal policy optimization technique developed in \cite{Schulman2017}.
-At the end of our study, we show that the solution found in the BS scenario provides a good approximation of the LV solution, which could be favored by practitioners if computational time is valued.}
+At the end of our study, we show that the solution found in the BS scenario provides a good approximation of the LV solution, which could be favored by practitioners if computational time is valued.
 	\\
@@
 	The paper is organized as follows.
-In Section \ref{sec:TVS} we introduce more in detail the \remove{manager-bank} contract \change{between the option-writer and the manager} and provide the description of the TVS dynamics in presence of valuation adjustments such as the hedging costs.
+In Section \ref{sec:TVS} we introduce more in detail the contract between the option-writer and the manager and provide the description of the TVS dynamics in presence of valuation adjustments such as the hedging costs.
 Then, in Section \ref{sec:Derivative} we introduce the structured class of derivative contracts linked to TVSs, where we describe the arising dynamical control problem for pricing those options.
-Moreover\change{,} in this section we derive the BS closed solution for European TVOs in two different ways: one applying the Gy\"ongy Lemma and the other \change{through} \remove{by writing} the Hamilton-Jacobi-Bellman equation.
+Moreover, in this section we derive the BS closed solution for European TVOs in two different ways: one applying the Gy\"ongy Lemma and the other through the Hamilton-Jacobi-Bellman equation.
 In Section \ref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a description of the algorithm we have built.
 We conclude the paper with Section \ref{sec:Numerical_investigations} where we present the numerical results obtained in this work for the Black and Scholes model and the local volatility one.
-\\
-
-\remove{A part of this work has been developed during the Master thesis of one of us (SP), where we had the opportunity to collaborate with Marco Bianchetti and Diego Pierluigi Giovannini from Intesa Sanpaolo Milan, that we thank.
-Moreover we wish to thank the Italian computing centre Cineca, which approved our ISCRA C project and provided us with the high-computing resources of Marconi100 for the numerical simulations of this contribution.
-\\
-
-The opinions here expressed are solely those of the authors and do not represent in any way those of their employers.}
 
 	\section{Target volatility strategy}\label{sec:TVS}
-As mentioned before, this work aims to enrich the TVS pricing literature by studying the aspect related to the funding costs coming from hedging \remove{the risky assets underlying a target volatility portfolio} \change{the underlying risky assets.
-In the following, we will refer to these costs simply as hedging costs}.
-\remove{Thus we} \change{We} consider the following scenario: \remove{a bank} \change{an option writer} selling a protection to a portfolio manager who has \remove{his} \change{her} capital invested \change{according to} \remove{in} a TVS.
+	As mentioned before, this work aims to enrich the TVS pricing literature by studying the aspect related to the funding costs coming from hedging the underlying risky assets.
+In the following, we will refer to these costs simply as hedging costs.
+We consider the following scenario: an option writer selling a protection to a portfolio manager who has her capital invested according to a TVS.
 In our case the fund manager has the freedom of changing the relative weights of the risky asset during the life of the TVS; once the allocation strategy is selected then the volatility targeting algorithm rebalances the risky component of the portfolio with the risk-free one in order to keep the overall portfolio volatility close to a target value.
@@
 
-The fund manager usually buys from a \remove{bank} \change{counterpart} an option on the TVS to ensure capital protection.
+	The fund manager usually buys from a counterpart an option on the TVS to ensure capital protection.
 For instance, the capital can be protected by buying a put option.
-In this case, we can write the \change{undiscounted} net asset value (NAV) $A_t$ of the strategy as given by\footnote{Here we neglect discounting factors \change{for sake of clarity}.}
+In this case, we can write the undiscounted net asset value (NAV) $A_t$ of the strategy as given by\footnote{Here we neglect discounting factors for sake of clarity.}
 	\begin{equation}
@@
 
-In this way, the TVS is only defined in the two contracts client-fund and fund-\change{option writer} \remove{fund-bank}.
-The fund manager is not implementing the strategy by trading in the market, and \change{s}he  is not subject to additional costs to access the market.
-On the other way, the \remove{bank} \change{writer} is paying such costs since she is actively hedging the call option sold to the manager.
-The \remove{bank} \change{writer} trading activity implemented to \remove{actively} hedge the option requires funding the collateral procedures of the hedging instruments along with any lending/borrowing fee.
-\change{We refer to \cite{Gabrielli2020} for a discussion of how to explicitly calculate hedging costs in equity trading.}
+	In this way, the TVS is only defined in the two contracts client-fund and fund-option writer.
+The fund manager is not implementing the strategy by trading in the market, and she is not subject to additional costs to access the market.
+On the other way, the writer is paying such costs since she is actively hedging the call option sold to the manager.
+The writer trading activity implemented to hedge the option requires funding the collateral procedures of the hedging instruments along with any lending/borrowing fee.
+We refer to \cite{Gabrielli2020} for a discussion of how to explicitly calculate hedging costs in equity trading.
 
-We remark that\change{, upon entering into the contract,}  the choices of the manager trading activity \remove{are} \change{can be seen as} stochastic processes since they \change{are not known {\it a priori} and }will depend on the market evolution.
-Thus, neither the manager strategy nor the \remove{bank financing costs for hedging the option} \change{writer hedging costs} can be written in the fund-\change{writer} \remove{bank} contract\change{: the only contractual elements are the protection strike and maturity, the market in which the manager can trade and the target volatility of the TVS portfolio}.
-For this reason, the price of a financial product sold by the \remove{bank} \change{writer} is adjusted to include any valuation adjustment due to the trading activity\change{, such as the hedging costs}.
+	We remark that, upon entering into the contract,  the choices of the manager trading activity can be seen as stochastic processes since they are not known {\it a priori} and will depend on the market evolution.
+Thus, neither the manager strategy nor the writer hedging costs can be written in the fund-writer contract: the only contractual elements are the protection strike and maturity, the market in which the manager can trade and the target volatility of the TVS portfolio.
+For this reason, the price of a financial product sold by the writer is adjusted to include any valuation adjustment due to the trading activity, such as the hedging costs.
 
-Our aim is to find the most expensive investing strategy from the point of view of \remove{financing} \change{hedging} costs for the \remove{bank} \change{writer} that the manager could choose in the market.
-In other words, we want to determine the worst-case scenario for the \remove{bank} \change{option-seller}.
+	Our aim is to find the most expensive investing strategy from the point of view of hedging costs for the writer that the manager could choose in the market.
+In other words, we want to determine the worst-case scenario for the option-seller.
 In the current section, we proceed by defining the price process of the TVS so that we can highlight the impact of valuation adjustments.
 
-\subsection{The strategy \change{p}rice \change{p}rocess}
+	\subsection{The strategy price process}
 	We work on a filtered probability space $\left(\Omega, \mathcal{F}, \{\mathcal{F}_t\}_{t\geq0}, \mathbb{P}  \right)$ satisfying the usual assumptions for a market model, where $\mathbb{P}$ is the physical probability measure \remove{representing the actual distribution of supply and demand shocks on equities prices}.
 
-We consider a fund trading a basket of $n$ risky securities with price process \change{$S_t = \{S_t^i, \, i=1,\dots,n\}$} \remove{$S_t = (S_t^i, \, i=1,\dots,n)$} funded with a cash account $B_t$ accruing at $r_t$.
+	We consider a fund trading a basket of $n$ risky securities with price process $S_t = \{S_t^i, \, i=1,\dots,n\}$ funded with a cash account $B_t$ accruing at $r_t$.
 Any dividend paid by the securities is re-invested in the fund.
@@
 	where $\pi^i_t$ represents the cumulative contractual-coupon process paid by the security, and $\psi_t^i$ represents the cumulative valuation adjustments.
-Since fund managers allocating TVS usually rely on equity assets, here we use the results of \cite{Gabrielli2020} that \remove{analyzes} \change{analyze} the valuation adjustments for equity products.
-We can write
+	Since fund managers allocating TVS usually rely on equity assets, here we use the results of \cite{Gabrielli2020} that analyze the valuation adjustments for equity products. We can write 
 	\begin{equation}
@@
 
-Then, we introduce the strategy price process $V_t$, and we define the deflated gain process $\Bar{G}_t^V$ as \remove{given by}
+	Then, we introduce the strategy price process $V_t$, and we define the deflated gain process $\Bar{G}_t^V$ as
 	\begin{equation}
@@
 	\end{equation}
-where $\phi_t$ are the running fees earned by the fund manager for \remove{his} \change{her} activity. We assume that the strategy is self-financing, so that we can write
+	where $\phi_t$ are the running fees earned by the fund manager for her activity. We assume that the strategy is self-financing, so that we can write
 	\begin{equation}
@@
 	Now, in order to prevent arbitrages, we assume the existence of a risk-neutral measure $\mathbb{Q}$ equivalent to $\mathbb{P}$ under which the deflated gain processes of all traded securities are martingales.
-Under this assumption, we are able to derive the drift conditions on the security price processes, and in turn on the strategy price process\change{, that is}
+Under this assumption, we are able to derive the drift conditions on the security price processes, and in turn on the strategy price process, that is
 	\begin{equation}
@@
 	\label{eq:risk_neutral}\end{equation}
-where $M_t^i$ are martingale\change{s} under $\mathbb{Q}$.
-If we substitute this expression for the security dynamics into the definition of the strategy we can check that the price process of the strategy is accruing at \change{the following} \remove{a} cash account rate $r_t$ compensated for the fund manager fees
+	where $M_t^i$ are martingales under $\mathbb{Q}$.
+If we substitute this expression for the security dynamics into the definition of the strategy we can check that the price process of the strategy is accruing at the following cash account rate $r_t$ compensated for the fund manager fees
 	\begin{equation}
@@
 	with $M_t^I$ martingale under $\mathbb{Q}$.
-In this case\change{,} we observe that $I_t$ depends explicitly both on the valuation adjustments and on the allocation strategy.
+In this case, we observe that $I_t$ depends explicitly both on the valuation adjustments and on the allocation strategy.
 Indeed, if we substitute the valuation adjustments with their explicit expression \eqref{eq:XVA_equity}, we get
@@
 
-\subsection{The \change{v}olatility \change{t}argeting \change{c}onstraint}
+	\subsection{The volatility targeting constraint}
 	In a typical TVS, the fund manager selects a risky-asset portfolio with a specific time-dependent allocation strategy expressed by means of the vector of relative weights $\alpha_t$, along with a risk-free asset, which we can identify with the bank account $B_t$.
-In the following, for sake of simplicity in exposition, we consider only total-return securities, namely we set $\pi_t=0$ \change{, which means that the security is not paying dividends}.
+In the following, for sake of simplicity in exposition, we consider only total-return securities, namely we set $\pi_t=0$, which means that the security is not paying dividends.
 Thus we can write Equation~\eqref{eq:TVS_first} as given by
@@
 	where $\Bar{\sigma}$ is the target volatility value.
-In practice, this means that the fund manager will select a risky-portfolio choosing $\alpha_t$ equities from the universe where \change{s}he can trade\change{,} and after that, \change{her} \remove{his} choices will be scaled by the automatic target volatility algorithm $\omega_t$ \change{\eqref{eq:general_tvs_wght}}.\footnote{We recall that the universe of assets where the manager can trade and the value of $\bar{\sigma}$ are written in the contract.}
+	In practice, this means that the fund manager will select a risky-portfolio choosing $\alpha_t$ equities from the universe where she can trade, and after that, her choices will be scaled by the automatic target volatility algorithm $\omega_t$ \eqref{eq:general_tvs_wght}.\footnote{We recall that the universe of assets where the manager can trade and the value of $\bar{\sigma}$ are written in the contract.}
 
-\remove{To derive the expression for $\omega_t$ we need to assume} \change{It is possible to derive the expression for $\omega_t$ assuming} a generic continuous semi-martingales dynamics under the risk-neutral measure for the underlying securities, so that \remove{we can write} Equation~\eqref{eq:risk_neutral} \change{can be written} as
+	It is possible to derive the expression for $\omega_t$ assuming a generic continuous semi-martingales dynamics under the risk-neutral measure for the underlying securities, so that Equation~\eqref{eq:risk_neutral} can be written as
 	\begin{equation}
@@
 	\label{eq:TVS_last}\end{equation}
-where we can see, as expected, that the strategy grows at the risk-free rate but \remove{for adjustments due to valuation adjustments and fees} \change{for the valuation adjustments, given by the hedging costs expressed by the term proportional to $\mu_t$ in \eqref{eq:TVS_last}, and the fees, given by the term $\phi_t$ in \eqref{eq:TVS_last}}.
+	where we can see, as expected, that the strategy grows at the risk-free rate but for the valuation adjustments, given by the hedging costs expressed by the term proportional to $\mu_t$ in \eqref{eq:TVS_last}, and the fees, given by the term $\phi_t$ in \eqref{eq:TVS_last}.
 
-We highlight that, to derive the dynamics expressed in Equation~\eqref{eq:TVS_last}, we have not made any assumptions on the risky allocation strategy; thus all this argument is valid for any constraints on the process $\alpha_t$\change{ as, for instance, the case of contracts in which the fund manager is restricted to holding only long positions meaning that $\alpha_t^i\geq0$ $\forall i = 1,\dots,n$, or situations where the combined total of long and short trades must be capped at a specific contractual amount}.
+	We highlight that, to derive the dynamics expressed in Equation~\eqref{eq:TVS_last}, we have not made any assumptions on the risky allocation strategy; thus all this argument is valid for any constraints on the process $\alpha_t$ as, for instance, the case of contracts in which the fund manager is restricted to holding only long positions meaning that $\alpha_t^i\geq0$ $\forall i = 1,\dots,n$, or situations where the combined total of long and short trades must be capped at a specific contractual amount.
 
-\section{Derivative pricing\change{: target volatility options}}\label{sec:Derivative}
-In this section, we analyze \remove{derivative} \change{TVO contracts, namely} contracts linked to the TVS described previously.
-In particular\change{,} we will focus on European style options, and we will show that under appropriate assumptions it is possible to find a closed form solution for the optimal allocation strategy which maximizes the contract price.
+	\section{Derivative pricing: target volatility options}\label{sec:Derivative}
+	In this section, we analyze TVO contracts, namely contracts linked to the TVS described previously.
+In particular, we will focus on European style options, and we will show that under appropriate assumptions it is possible to find a closed form solution for the optimal allocation strategy which maximizes the contract price.
 
@@
 	where $D(0,T;\zeta_t)$ is the discount factor with rate $\zeta_t$, inclusive of the derivative valuation adjustments, and $\pi_t$ is the cumulative coupon process paid by the derivative, and it depends on the allocation strategy since in turn the TVS depends on it via the valuation adjustments.
-We take the supremum over the strategies since we do not have any information on the future activity of the fund manager \change{and we wish to calculate the worst-case scenario for the option seller, as discussed in the introduction of Section \ref{sec:TVS}}.
+We take the supremum over the strategies since we do not have any information on the future activity of the fund manager and we wish to calculate the worst-case scenario for the option seller, as discussed in the introduction of Section \ref{sec:TVS}.
 
-\subsection{European \change{o}ptions}
-If the derivative contract depends only on the marginal distribution of $I_T$ at maturity (\change{i.e.,} a European payoff), we are able to \remove{calculate an optimal strategy} \change{simplify the pricing problem}.
+	\subsection{European options}
+	If the derivative contract depends only on the marginal distribution of $I_T$ at maturity (i.e., a European payoff), we are able to simplify the pricing problem.
 We consider the following pricing problem
@@
 	\begin{proposition}
-A European \remove{payoff on the TVS} \change{TVO} can be calculated by assuming any allocation in the underlying risky basket if all the underlying
+	A European TVO can be calculated by assuming any allocation in the underlying risky basket if all the underlying
 	securities grow under the risk-neutral measure at the risk-free rate without any valuation adjustment, namely if we can write $\mu_t=0$.
@@
 
-\subsection{Stochastic \change{o}ptimal \change{c}ontrol \change{p}roblem}
+	\subsection{Stochastic optimal control problem}
 	In presence of valuation adjustments, we need to solve the full optimization problem.
@@
 	\label{eq:recursion}\end{equation}
-where $V_{T_k}$ is the option value at time $T_k$ and $X$ is any Markovian state such that the drift and the diffusion coefficient of $I_t$ are a function of $\left(X_t,I_t,\alpha_t\right)$ \change{and we indicated with $I_{T_k}\left(\alpha\right)$ the value of the strategy at time $T_{k}$ for a given choice of the weights $\alpha$ at time $T_{k-1}$.} We calculate \change{the strategy value for the optimal weights $\alpha_{T_{k-1}}$, namely} $I_{T_k}\left(\alpha_{T_{k-1}}\right)$ \remove{for any given strategy $\alpha_{T_{k-1}}$} by a suitable discretization of \eqref{eq:TVS_last} starting from $X_{T_{k-1}}$ and $I_{T_{k-1}}$.
-\change{Thus, once collected the elements $\{X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\}$,} the derivative price is given by:
+	where $V_{T_k}$ is the option value at time $T_k$ and $X$ is any Markovian state such that the drift and the diffusion coefficient of $I_t$ are a function of $\left(X_t,I_t,\alpha_t\right)$ and we indicated with $I_{T_k}\left(\alpha\right)$ the value of the strategy at time $T_{k}$ for a given choice of the weights $\alpha$ at time $T_{k-1}$.
+We calculate the strategy value for the optimal weights $\alpha_{T_{k-1}}$, namely $I_{T_k}\left(\alpha_{T_{k-1}}\right)$ \remove{for any given strategy $\alpha_{T_{k-1}}$} by a suitable discretization of \eqref{eq:TVS_last} starting from $X_{T_{k-1}}$ and $I_{T_{k-1}}$.
+	Thus, once collected the elements $\{X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\}$, the derivative price is given by:
 	\begin{equation}
@@
 
-\subsection{Black and Scholes \change{scenario}}\label{subsec:BS_model}
+	\subsection{Black and Scholes scenario}\label{subsec:BS_model}
 	In the time-dependent Black and Scholes model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$.
@@
 	\end{proposition}
-The absence of stochastic elements in Equation~\eqref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no \change{numerical} simulation needed \change{to solve the control problem}; in fact\change{,} \remove{one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities} \change{the functions $\mu(t)$ and $\nu(t)$ can be directly derived from the market quotes of the risky assets and of its derivative contracts, and then we can apply Equation~\eqref{eq:BS_optimal_strategy} $\forall t \in \mathcal{T}$  to retrieve the deterministic optimal strategy}.
+	The absence of stochastic elements in Equation~\eqref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no numerical simulation needed to solve the control problem; in fact, the functions $\mu(t)$ and $\nu(t)$ can be directly derived from the market quotes of the risky assets and of its derivative contracts, and then we can apply Equation~\eqref{eq:BS_optimal_strategy} $\forall t \in \mathcal{T}$  to retrieve the deterministic optimal strategy.
 Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
@@
 
-In Figure~\ref{fig:strategy_price_comparison} we provide a comparison among the option price obtained with the optimal strategy (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and those with \change{the below expert-based single-asset} \remove{other intuitive} strategies (S$_A$, S$_B$, S$_C$).
+	In Figure~\ref{fig:strategy_price_comparison} we provide a comparison among the option price obtained with the optimal strategy (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and those with the below expert-based single-asset strategies (S$_A$, S$_B$, S$_C$).
 Here we consider the case of an at-the-money call option with spot $I_0 = 1\, \text{EUR}$, maturity $T=5 \, \text{yr}$, target volatility $\bar{\sigma}=5\%$, and a nonnegative constraint on $\alpha$.
@@
 	\begin{itemize}
-\item S$_A$: invest all in the asset with the maximum forward curve at maturity; \change{this strategy consists of selecting a static portfolio that is never rebalanced, such that the TVO moneyness is maximized;}
-\item S$_B$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)$; \change{in this case we consider a fund manager who rebalances the weights $\alpha(t)$ over the life of the portfolio by selecting the asset with the highest expected return at a given point in time;}
-\item S$_C$: \remove{for each market pillar before maturity} \change{$\forall t \in \mathcal{T}$} invest all in the asset with minimum $\mu(t)/\|\nu(t)\|$\remove{.}\change{; the strategy is akin to S$_B$, with the exception being that the manager endeavors to maximize the expected return that has been adjusted to account for the target volatility mechanism.}
+	\item S$_A$: invest all in the asset with the maximum forward curve at maturity; this strategy consists of selecting a static portfolio that is never rebalanced, such that the TVO moneyness is maximized;
+	\item S$_B$: $\forall t \in \mathcal{T}$ invest all in the asset with minimum $\mu(t)$; in this case we consider a fund manager who rebalances the weights $\alpha(t)$ over the life of the portfolio by selecting the asset with the highest expected return at a given point in time;
+	\item S$_C$: $\forall t \in \mathcal{T}$ invest all in the asset with minimum $\mu(t)/\|\nu(t)\|$; the strategy is akin to S$_B$, with the exception being that the manager endeavors to maximize the expected return that has been adjusted to account for the target volatility mechanism.
 	\end{itemize}
-As the reader can observe, our strategy outperforms any other \remove{intuitive} \change{expert-based} one a practitioner might adopt \change{to face the control problem}.
+	As the reader can observe, our strategy outperforms any other expert-based one a practitioner might adopt to face the control problem.
 	\begin{figure}[t]
@@
 		\includegraphics[width=4in]{strategy_price_comparison_grey.eps}
-	\caption{Comparison of plain vanilla prices on the TVS by adopting different allocation strategies: the optimal Black and Scholes solution (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and \change{three expert-based single-asset} strategies (S$_A$, S$_B$, S$_C$)}
+		\caption{Comparison of plain vanilla prices on the TVS by adopting different allocation strategies: the optimal Black and Scholes solution (BS$^*$) of Equation~\eqref{eq:BS_optimal_strategy} and three expert-based single-asset strategies (S$_A$, S$_B$, S$_C$)}
 		\label{fig:strategy_price_comparison}
@@
 
-\change{In the next subsection we derive the analytical solution of the problem \eqref{eq:BS_optimal_strategy} under the assumption of unconstrained allocation strategy $\alpha_t$.}
+	In the next subsection we derive the analytical solution of the problem \eqref{eq:BS_optimal_strategy} under the assumption of unconstrained allocation strategy $\alpha_t$.
 
-\subsubsection{Unconstrained \change{a}llocation \change{s}trategy: \change{c}losed \change{f}orm \change{s}olution}
+	\subsubsection{Unconstrained allocation strategy: closed form solution}
 	In absence of constraints on the allocation strategy, we are able to derive a closed form solution to the BS problem \eqref{eq:BS_optimal_strategy}.
@@
 
-\change{In the following subsection we prove that an analytical solution of \eqref{eq:BS_optimal_strategy} exists also assuming a nonnegative constrained allocation strategy.
+	In the following subsection we prove that an analytical solution of \eqref{eq:BS_optimal_strategy} exists also assuming a nonnegative constrained allocation strategy.
 Under this assumption, we show that the optimal strategy consists in investing in a single asset.
-We name this strategy the bang-bang solution.}
+We name this strategy the bang-bang solution.
 
-\subsubsection{Active \change{a}sset (or \change{b}ang\change{-b}ang) \change{s}olution}
-A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry \change{$\mu$} are nonnegative\change{, so they represent true hedging costs and not benefits,} and the only constraint on portfolio weights is nonnegativity, which would mean a long-only strategy by the fund manager.
+	\subsubsection{Active asset (or bang-bang) solution}
+	A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry $\mu$ are nonnegative, so they represent true hedging costs and not benefits, and the only constraint on portfolio weights is nonnegativity, which would mean a long-only strategy by the fund manager.
 
 	\begin{lemma}
-	Let \remove{$\mu \in \mathbb{R}^n$} \change{$\mu \in \mathbb{R}^n_+$} be a vector with nonnegative components, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix, and $\Sigma = \nu \tr{\nu}$.
-Then
+		Let $\mu \in \mathbb{R}^n_+$ be a vector with nonnegative components, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix, and $\Sigma = \nu \tr{\nu}$. Then
 		\begin{equation}\label{eq:inf}
@@
 
-In full generality, we assume that the time evolution of the Markovian factors governing the problem dynamics is given by  a stochastic multidimensional process in $\mathbb{R}^n$, $X_t$, which is the unique strong solution to the following It$\hat{\text{o}}$ \remove{SDE} \change{stochastic differential equation}
+	In full generality, we assume that the time evolution of the Markovian factors governing the problem dynamics is given by  a stochastic multidimensional process in $\mathbb{R}^n$, $X_t$, which is the unique strong solution to the following It$\hat{\text{o}}$ stochastic differential equation
 	\begin{equation}
@@
 
-In this framework, the TVS price process dynamics is given by the \remove{SDE} \change{stochastic differential equation} 
+	In this framework, the TVS price process dynamics is given by the stochastic differential equation 
 	\begin{equation}
@@
 	\label{eq:HJB_general}\end{equation}
-Equation~\eqref{eq:HJB_general} is the \change{HJB} \remove{Hamilton-Jacobi-Bellman}     equation describing the TVS dynamic problem for a generic dynamics of the risky securities underlying the portfolio.
+	Equation~\eqref{eq:HJB_general} is the HJB equation describing the TVS dynamic problem for a generic dynamics of the risky securities underlying the portfolio.
 
@@
 	\end{equation}
-if the payoff is non-decreasing in $I$\change{,} by homogeneity of the \remove{SDE} \change{stochastic differential equation} we get that $V$ is non-decreasing\change{.}\remove{; t}\change{T}hus the solution is given by
+	if the payoff is non-decreasing in $I$, by homogeneity of the stochastic differential equation we get that $V$ is non-decreasing. Thus the solution is given by
 	\begin{equation}\label{eq:HJB_solution}
@@
 	Conversely, if we deal with a dynamic model for which the derivative contract $V$ depends on $X$ then the volatility versor term, namely the second one, in the $\max$ operator of Equation~\eqref{eq:HJB_general} is no longer zero and thus one must solve the entire control problem numerically.
-\change{Indeed, one can check by direct substitution that the natural generalization of \eqref{eq:HJB_solution}
+Indeed, one can check by direct substitution that the natural generalization of \eqref{eq:HJB_solution}
 	\begin{equation}
@@
 	\end{equation}
-fails to satisfy \eqref{eq:HJB_general}.}
+	fails to satisfy \eqref{eq:HJB_general}.
 
-In \remove{this work}\change{what follows} we tackle the non-trivial case of a local volatility \remove{(LV)} model for the $S_t$-dynamics, such that $\nu_t = \nu(t,S_t)$.
-We have chosen the LV model since it is well known in financial literature and among practitioners \change{\cite{Gatheral2006}}.
+	In what follows we tackle the non-trivial case of a local volatility model for the $S_t$-dynamics, such that $\nu_t = \nu(t,S_t)$.
+We have chosen the LV model since it is well known in financial literature and among practitioners \cite{Gatheral2006}.
 
 	\section{Reinforcement learning}\label{sec:RL}
-As we have discussed in the previous sections, \change{the analytical solution described in Proposition \ref{prop:bs} crucially relies on the assumption that the underlyings evolve according to a Black and Scholes model, and on a monotonicity assumption on the payoff.
-Therefore,}
+	As we have discussed in the previous sections, the analytical solution described in Proposition \ref{prop:bs} crucially relies on the assumption that the underlyings evolve according to a Black and Scholes model, and on a monotonicity assumption on the payoff.
+Therefore,
 	one must resort to numerical approaches to solve the stochastic control problem related to the TVS in the case of general payoffs or risky securities dynamics.
-The standard approach could be to use classical techniques based on \remove{backwards} \change{backward} recursion \eqref{eq:recursion}-\eqref{eq:recursion2} such as American Monte Carlo \cite{Longstaff2001}.
+The standard approach could be to use classical techniques based on backward recursion \eqref{eq:recursion}-\eqref{eq:recursion2} such as American Monte Carlo \cite{Longstaff2001}.
 However, their performances degrade exponentially as the dimension $n$ of the problem increases, making prohibitively costly finding the solution to the problem.
@@
 
-\remove{Reinforcement Learning} \change{RL} is a branch of Machine Learning which allows an artificial agent to interact with an environment through actions and observations in order to maximize some notion of cumulative reward \change{called value function}.
-In RL\change{,} the agent is not told which actions to take but instead, it \remove{must} \change{has to} discover by trial and error which are the behaviors  yielding the highest reward.
+	RL is a branch of Machine Learning which allows an artificial agent to interact with an environment through actions and observations in order to maximize some notion of cumulative reward called value function.
+In RL, the agent is not told which actions to take but instead, it has to discover by trial and error which are the behaviors  yielding the highest reward.
 This is obtained by updating the agent policy $\pi$ which is a mapping from the environment states to the set of actions.
-Thus RL is independent of pre-collected data as opposed to other Machine Learning techniques\change{, that have to be fed with a pre-existing dataset to learn from}.
+Thus RL is independent of pre-collected data as opposed to other Machine Learning techniques, that have to be fed with a pre-existing dataset to learn from.
 Because of its nature, RL has been successful in quantitative finance for solving control problems; among the most important RL applications in this field, we refer to \cite{Deng2017} as the pioneers in studying self-taught reinforcement trading problems, while to \cite{Halperin2020} and \cite{Kolm2019} for hedging derivatives with RL under market frictions.
 
-In our work we adopt two learning strategies to compare their performances in terms of overall reward: \remove{an \textit{ad hoc}} \change{a plain} direct policy \change{search} algorithm \change{\cite{Sutton2018}} and the state of the art proximal policy optimization (PPO) developed in \cite{Schulman2017} and \cite{Schulman2016}.
+	In our work we adopt two learning strategies to compare their performances in terms of overall reward: a plain direct policy search algorithm \cite{Sutton2018} and the state of the art proximal policy optimization (PPO) developed in \cite{Schulman2017} and \cite{Schulman2016}.
 
-The first method is \change{well-known in RL literature: it is based on the direct update of the agent policy via a gradient-based optimization of the value function.}\remove{a specific algorithm developed by us to fit the problem aim to find the optimal option price; this technique can take place as a direct policy optimization method in the wide taxonomy of Reinforcement Learning algorithms}.
+	The first method is well-known in RL literature: it is based on the direct update of the agent policy via a gradient-based optimization of the value function.
 We will go into details in Section~\ref{sec:direct_policy}.
@@
 
-\subsection{Direct policy \change{search} approach}\label{sec:direct_policy}
-\change{In RL, the learning phase takes place within a set of so-called episodes, which refer to a single run or instance of interaction between the RL agent and an environment, where the agent takes a sequence of actions to achieve a specific goal.}
+	\subsection{Direct policy search approach}\label{sec:direct_policy}
+	In RL, the learning phase takes place within a set of so-called episodes, which refer to a single run or instance of interaction between the RL agent and an environment, where the agent takes a sequence of actions to achieve a specific goal.
 
-\change{For our problem, we} \remove{We} consider an episode $\tau$ of length $m+1$ that \remove{takes place on} \change{unfolds over} a discrete time-grid of fixing times expressed in year fractions $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$ maturity of the option.
+	For our problem, we consider an episode $\tau$ of length $m+1$ that unfolds over a discrete time-grid of fixing times expressed in year fractions $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$ maturity of the option.
 At a given episodic time $T_k$ the RL agent interacts with the environment: it receives a representation of the environment called state $s_k$ and on the basis of that it selects an action $a_k$ sampling from the current policy $\pi^{\theta_j}$.
 Here with $\theta_j$ we refer to the set of parameters through which we parameterize the agent policy at the $j$-th algorithm iteration.
-In our case of study\change{,} the \change{RL} agent\change{, representing the fund manager, has the right to select} \remove{can choose} the composition of the risky asset portfolio, so that the policy is the allocation strategy $\alpha$ introduced in Equation~\eqref{eq:TVS_elegant}:
+In our case of study, the RL agent, representing the fund manager, has the right to select the composition of the risky asset portfolio, so that the policy is the allocation strategy $\alpha$ introduced in Equation~\eqref{eq:TVS_elegant}:
 	\begin{equation}
@@
 	\label{eq:state}\end{equation}
-In this way, the state contains all the information needed by the agent to \remove{take an} \change{select the} optimal action, leading to the maximum plain vanilla TVO price.
+	In this way, the state contains all the information needed by the agent to select the optimal action, leading to the maximum plain vanilla TVO price.
 
-\change{Once the agent has selected the action $a_k$ sampled from the current policy, it receives at the next time $T_{k+1}$ a reward $r_{k+1}$ generated by the environment.
-In this algorithm we have defined the reward function as follows}
+	Once the agent has selected the action $a_k$ sampled from the current policy, it receives at the next time $T_{k+1}$ a reward $r_{k+1}$ generated by the environment.
+In this algorithm we have defined the reward function as follows
 	\begin{equation}
@@
 
- \change{Therefore, during the whole episode, the agent receives a nil reward except at maturity when the reward coincides with the option intrinsic value.
+	Therefore, during the whole episode, the agent receives a nil reward except at maturity when the reward coincides with the option intrinsic value.
 This choice may seem too daring because the agent receives a real feedback of its actions only at the end of the episode, which could result in a slower learning.
-However, if the agent has learnt the optimal policy $\pi^* = \pi^{\theta^*}$, the average cumulative reward per episode will coincide with the optimal TVO price.}
+However, if the agent has learnt the optimal policy $\pi^* = \pi^{\theta^*}$, the average cumulative reward per episode will coincide with the optimal TVO price.
 
-In this algorithm, we parameterize the agent policy with a feed forward neural network (FFNN), such that $\theta_j$ coincides with the hidden weights, $s_k$ with the \remove{inputs} \change{input neurons}, and $a_k$ with the output \change{ones}.
+	In this algorithm, we parameterize the agent policy with a feed forward neural network (FFNN), such that $\theta_j$ coincides with the hidden weights, $s_k$ with the input neurons, and $a_k$ with the output \change{ones}.
 In this way, we are dealing with a deterministic policy, \change{that maps each state directly to a single action without any additional randomness or probability distribution;} \remove{where its} \change{the map} functional form is given by the neural network.
@@
 	\begin{equation}
-\change{L(\theta_j) =\hat{\mathbb{E}}\left[ \sum_{k=0}^{m-1} r_{k+1} | \theta_j \right]} \; ,
+	L(\theta_j) =\hat{\mathbb{E}}\left[ \sum_{k=0}^{m-1} r_{k+1} | \theta_j \right] \; ,
 	\label{eq:direct_policy_loss}\end{equation}
@@
 This additional dependency of volatility makes the problem of finding the optimal strategy non-trivial\change{, as a closed form formula cannot be derived under this scenario}; in fact if we consider the whole \remove{securities} \change{volatility} smiles \change{of the securities,} then the second-order term in the HJB equation \eqref{eq:HJB_general} \remove{is not zero and thus a closed formula for $\alpha^*$ is not available anymore} \change{does not elide and consequently the only way to derive $\alpha^*$ is through a numerical method}.
-\remove{Because of that, one must resort to numerical techniques to recover the problem solution.} \remove{Unlike the BS model where the strategy depends only on time, in LV dynamics there is no unnecessary information provided by the state block in Equation~\eqref{eq:state} to take the optimal action} \change{The additional dependence on the state $S_t$ under LV dynamics will cause the agent to consider also the $X_t$ component value in the  state block \eqref{eq:state}, unlike under the BS scenario where this information is redundant as the optimal strategy was function only of time (see Proposition \ref{prop:bs})}.
+\remove{Because of that, one must resort to numerical techniques to recover the problem solution.} \remove{Unlike the BS model where the strategy depends only on time, in LV dynamics there is no unnecessary information provided by the state block in Equation~\eqref{eq:state} to take the optimal action} The additional dependence on the state $S_t$ under LV dynamics will cause the agent to consider also the $X_t$ component value in the  state block \eqref{eq:state}, unlike under the BS scenario where this information is redundant as the optimal strategy was function only of time (see Proposition \ref{prop:bs}).
 Here we consider the same market data ($r_t$, $\mu_t$ and $\nu_t$) as in the Black and Scholes environment to study how the optimal solution changes with the dynamics model.
@@
 		\includegraphics[width=4in]{LV_direct_policy_approach_grey.eps}
-	\caption{Learning curves of the direct policy algorithm applied to the local volatility \change{scenario}.
+		\caption{Learning curves of the direct policy algorithm applied to the local volatility scenario.
 The solid line with triangles is the learning curve of an agent with free actions, while the solid one is the learning curve of an agent whose actions are parameterized with the maximum drift strategy.
@@
 To try to measure the performance of the policy selected by the agent, we can build a n\"aif strategy called \quotations{baseline}.
-By looking at the Markovian projection in \eqref{eq:markovian_projection}-\eqref{eq:local_drift} and the first term in the HJB \change{equation} \eqref{eq:HJB_general}, \change{we see that the control problem is similar to the BS one with an additional second order term.
-Thus} a natural choice for the baseline is to maximize the TVS local drift \remove{.
-In other words this coincides} by applying the path-wise the Black and Scholes solution of Equation~\eqref{eq:BS_optimal_strategy}.
-\change{S}ince here we deal with a free allocation strategy, we can simply use our analytical result \eqref{eq:analytical_solution_free_strategy}.
+By looking at the Markovian projection in \eqref{eq:markovian_projection}-\eqref{eq:local_drift} and the first term in the HJB equation \eqref{eq:HJB_general}, we see that the control problem is similar to the BS one with an additional second order term.
+Thus a natural choice for the baseline is to maximize the TVS local drift by applying the path-wise the Black and Scholes solution of Equation~\eqref{eq:BS_optimal_strategy}.
+Since here we deal with a free allocation strategy, we can simply use our analytical result \eqref{eq:analytical_solution_free_strategy}.
 In the same Figure~\ref{fig:result_lv_direct_policy} we report the 99\% confidence interval of the MC price obtained with the maximum drift baseline as two red-dashed lines.
@@
 Following the theoretical result of the HJB equation, we can assert that the agent has learnt a sub-optimal policy.
-Since we LV model differs from the BS one for a corrective term in the HJB \change{equation} \eqref{eq:HJB_general}, we expect that the optimal solution  in the LV framework will be in a close region of the maximum drift strategy.
+Since we LV model differs from the BS one for a corrective term in the HJB equation \eqref{eq:HJB_general}, we expect that the optimal solution  in the LV framework will be in a close region of the maximum drift strategy.
 Thus we train another agent with the direct policy learning by parameterizing its action with the baseline strategy and choosing a smaller learning rate of $10^{-4}$.
-With this parameterization, \change{the RL agent actions, constituting the risky allocation strategy, are computed by summing} at each observational time $T_k\in\mathcal{T}$ \remove{the risky allocation strategy is obtained by summing} the FFNN output with the Equation~\eqref{eq:analytical_solution_free_strategy}.
+With this parameterization, the RL agent actions, constituting the risky allocation strategy, are computed by summing at each observational time $T_k\in\mathcal{T}$ \remove{the risky allocation strategy is obtained by summing} the FFNN output with the Equation~\eqref{eq:analytical_solution_free_strategy}.
 Even in this case, the corresponding learning curve (solid black line in Figure~\ref{fig:result_lv_direct_policy}) is stuck in the maximum drift strategy.
@@
 Again we experienced that deeper FFNNs outperform shallow ones and tanh is the most preferable.
-We report for completeness the values of the other PPO hyper-parameters, for the description of which we refer to \cite{Schulman2017}: learning rate $3\times 10^{-4}$, $\lambda=0.95$, $\epsilon=0.2$, $c_1=0.7$, $c_2=0$ and mini-batch\footnote{\change{A minibatch is a subset of a larger dataset, which is used to train a machine learning model.
+We report for completeness the values of the other PPO hyper-parameters, for the description of which we refer to \cite{Schulman2017}: learning rate $3\times 10^{-4}$, $\lambda=0.95$, $\epsilon=0.2$, $c_1=0.7$, $c_2=0$ and mini-batch\footnote{A minibatch is a subset of a larger dataset, which is used to train a machine learning model.
 The model updates its parameters based on the error calculated from the predictions it makes on each minibatch.
-This allows for more efficient memory use, faster convergence of the model, and better generalization to new data.}} size of 2048 episodes.
+This allows for more efficient memory use, faster convergence of the model, and better generalization to new data.} size of 2048 episodes.
 
-We have trained a 5 layer FFNN with 8 neurons each with the PPO algorithm in two different environments: one environment generates the reward according to Equation~\eqref{eq:reward1}, while the other exploits the reward function \eqref{eq:reward2} where we set $\gamma=0.98$ as discount factor to make the agent prefer immediate rewards \change{more than in the previous simulations}.
+	We have trained a 5 layer FFNN with 8 neurons each with the PPO algorithm in two different environments: one environment generates the reward according to Equation~\eqref{eq:reward1}, while the other exploits the reward function \eqref{eq:reward2} where we set $\gamma=0.98$ as discount factor to make the agent prefer immediate rewards more than in the previous simulations.
 Again, for both environments, we train two different agents: one whose action is given by Equation~\eqref{eq:action}, while the other implements the action parameterization with the baseline.
@@
 		\includegraphics[width=3.4in]{LV_PPO_reward_b_learning_curve.eps}
-	\caption{Learning curves of the PPO algorithm applied to the local volatility \change{scenario}.
+		\caption{Learning curves of the PPO algorithm applied to the local volatility scenario.
 On the horizontal axis the number of training episodes.
@@
 In all the PPO cases, the trained agents seem to be stuck in the sub-optimal maximum drift strategy policy.
-For the case of the immediate reward function (Figure~\ref{fig:PPO_results} on the right), we \remove{have that the learning curves converge to a saturation value that is not compatible with the baseline price; this is due to the fact that with $\lambda,\gamma \neq 1$ we introduce a bias in the problem} \change{can not compare the saturation values reached by the learning curves with the baseline price, since the former are discounted due to $\lambda,\gamma \neq 1$}.
-\remove{Because of that, we need to run a test MC simulation with freezed agents and $\lambda=\gamma=1$ to obtain a result to compare with the baseline price} \change{
-Thus, to compare the performance of PPO with the baseline price in the test phase and ensure consistency in unit of measurement, it is necessary to conduct a MC simulation with freezed agents and set $\lambda$ and $\gamma$ values to 1}.
+For the case of the immediate reward function (Figure~\ref{fig:PPO_results} on the right), we can not compare the saturation values reached by the learning curves with the baseline price, since the former are discounted due to $\lambda,\gamma \neq 1$.
+	Thus, to compare the performance of PPO with the baseline price in the test phase and ensure consistency in unit of measurement, it is necessary to conduct a MC simulation with freezed agents and set $\lambda$ and $\gamma$ values to 1.
 We report the results in Table~\ref{tab:mc_results_lv}.
-The \remove{Reinforcement Learning} \change{RL} agents we have trained with PPO give all prices compatible with the baseline one.
+The RL agents we have trained with PPO give all prices compatible with the baseline one.
 
@@
 	\begin{center}
-\caption{Comparison of TVO prices under local volatility \change{scenario}.
+	\caption{Comparison of TVO prices under local volatility scenario.
 The baseline price is obtained by applying path-wise the maximum drift strategy \eqref{eq:analytical_solution_free_strategy}.
@@
 	In this paper, we described a non-trivial control problem related to derivative contracts on target volatility strategies.
-In particular, we have considered a\change{n option writer} \remove{bank} selling a call option to a fund manager as protection on the capital invested on the TVS.
-We showed how the presence of different funding costs coming from hedging the risky assets underlying the TVS, obliges the \remove{bank} \change{writer} to solve a stochastic optimal control problem to price the protection.
-This is due to the fact that the \remove{bank} \change{option-seller's} strategy is not self-financing.
+In particular, we have considered an option writer selling a call option to a fund manager as protection on the capital invested on the TVS.
+We showed how the presence of different funding costs coming from hedging the risky assets underlying the TVS, obliges the writer to solve a stochastic optimal control problem to price the protection.
+This is due to the fact that the option-seller's strategy is not self-financing.
 This kind of control problem is hard to solve because here the control process affects both drift and diffusive coefficients of the controlled process.
 Despite its complexity, our first contribution is the derivation of a closed form solution of the control problem in a Black and Scholes framework, which could represent a useful tool for practitioners since it outperforms intuitive trading strategies.
-We have derived this solution in two different ways: first by applying the Gy\"ongy  Lemma and then by writing the \remove{Hamilton-Jacobi-Bellman} \change{HJB} equation.
+We have derived this solution in two different ways: first by applying the Gy\"ongy  Lemma and then by writing the HJB equation.
 We numerically studied the problem in the more general local volatility model where the solution is not available and thus a numerical investigation is needed.
-We tackled the problem by means of the novel \remove{Reinforcement Learning} \change{RL} techniques, by both the direct policy learning and the proximal policy optimization one.
+We tackled the problem by means of the novel RL techniques, by both the direct policy learning and the proximal policy optimization one.
 We used the BS model, where the solution is \textit{a priori} known, as benchmark to perform a series of fine-tuning of the RL algorithm hyper-parameters, such as the artificial neural network architecture.
@@
 	\bibitem{Gatheral2006}
-\change{Gatheral, J.: The volatility surface: a practitioner's guide. Wiley Finance (2006)}
+	Gatheral, J.: The volatility surface: a practitioner's guide. Wiley Finance (2006)
 	%
 	\bibitem{Goodfellow2016}
-\change{Goodfellow, I., Bengio, Y., Courville, A.: Deep learning. MIT Press (2016)}
+	Goodfellow, I., Bengio, Y., Courville, A.: Deep learning. MIT Press (2016)
 	%

### 2023-07-04 tex/m2ef/daluiso_nastasi_pallavicini_polo_rl_for_options_on_tvf.tex
paper final version

@@
 	\subsection{The strategy price process}
-	We work on a filtered probability space $\left(\Omega, \mathcal{F}, \{\mathcal{F}_t\}_{t\geq0}, \mathbb{P}  \right)$ satisfying the usual assumptions for a market model, where $\mathbb{P}$ is the physical probability measure \remove{representing the actual distribution of supply and demand shocks on equities prices}.
+	We work on a filtered probability space $\left(\Omega, \mathcal{F}, \{\mathcal{F}_t\}_{t\geq0}, \mathbb{P}  \right)$ satisfying the usual assumptions for a market model, where $\mathbb{P}$ is the physical probability measure.
 
@@
 	\label{eq:Equity_process}\end{equation}
-	where $\nu_t$ is an adapted matrix process ensuring the existence of a solution for the stochastic differential equation \remove{(SDE)} and $W_t$ is a $n$-dimensional vector of Brownian motions under $\mathbb{Q}$.
+	where $\nu_t$ is an adapted matrix process ensuring the existence of a solution for the stochastic differential equation and $W_t$ is a $n$-dimensional vector of Brownian motions under $\mathbb{Q}$.
 Under these assumptions we can derive an expression for $\omega_t$, and we get\footnote{In all formulae the norm for a vector $a$ is defined as $\|a\|\coloneqq\sqrt{a\cdot a}$.}
@@
 	where $V_{T_k}$ is the option value at time $T_k$ and $X$ is any Markovian state such that the drift and the diffusion coefficient of $I_t$ are a function of $\left(X_t,I_t,\alpha_t\right)$ and we indicated with $I_{T_k}\left(\alpha\right)$ the value of the strategy at time $T_{k}$ for a given choice of the weights $\alpha$ at time $T_{k-1}$.
-We calculate the strategy value for the optimal weights $\alpha_{T_{k-1}}$, namely $I_{T_k}\left(\alpha_{T_{k-1}}\right)$ \remove{for any given strategy $\alpha_{T_{k-1}}$} by a suitable discretization of \eqref{eq:TVS_last} starting from $X_{T_{k-1}}$ and $I_{T_{k-1}}$.
+We calculate the strategy value for the optimal weights $\alpha_{T_{k-1}}$, namely $I_{T_k}\left(\alpha_{T_{k-1}}\right)$ by a suitable discretization of \eqref{eq:TVS_last} starting from $X_{T_{k-1}}$ and $I_{T_{k-1}}$.
 	Thus, once collected the elements $\{X_{T_{k}}, I_{T_{k}}\left(\alpha_{T_{k-1}}\right)\}$, the derivative price is given by:
@@
 	\begin{lemma}
-		Let \remove{be} $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \tr{\nu}$.
+		Let $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \tr{\nu}$.
 Then the closed solution of the optimization problem \eqref{eq:BS_optimal_strategy} is
@@
 	\subsection{Hamilton-Jacobi-Bellman equation for target volatility options}\label{subsec:HJB}
-	In this section, we \remove{want to} provide to the reader a formal description of the dynamic problem associated with options on target volatility strategies by writing the Hamilton-Jacobi-Bellman (HJB) equation for the derivative price.
+	In this section, we provide to the reader a formal description of the dynamic problem associated with options on target volatility strategies by writing the Hamilton-Jacobi-Bellman (HJB) equation for the derivative price.
 We prove that from this equation we can recover the same closed formula \eqref{eq:BS_optimal_strategy} derived above from the Gy\"ongy Lemma for the time-dependent BS model.
@@
 	\begin{equation}
-		\change{r_{k+1}\change{\coloneqq}\left\{\begin{array}{ll}
+		r_{k+1}\coloneqq\left\{\begin{array}{ll}
 			D(0,T;\zeta)(I_T(\alpha(\theta_j)) - K)^+ & \text { if } T_{k+1} = T \\
@@
 		\end{array}\right.
-	\label{eq:reward1} \; .} \end{equation}
+	\label{eq:reward1} \; . \end{equation}
 
@@
 
-	In this algorithm, we parameterize the agent policy with a feed forward neural network (FFNN), such that $\theta_j$ coincides with the hidden weights, $s_k$ with the input neurons, and $a_k$ with the output \change{ones}.
-In this way, we are dealing with a deterministic policy, \change{that maps each state directly to a single action without any additional randomness or probability distribution;} \remove{where its} \change{the map} functional form is given by the neural network.
-The parameters update is performed as follows: the agent collects a finite batch of experiences interacting with the environment in a set of episodes $\tau$, then the loss-\change{value} function is evaluated as
+	In this algorithm, we parameterize the agent policy with a feed forward neural network (FFNN), such that $\theta_j$ coincides with the hidden weights, $s_k$ with the input neurons, and $a_k$ with the output ones.
+In this way, we are dealing with a deterministic policy, that maps each state directly to a single action without any additional randomness or probability distribution; the map functional form is given by the neural network.
+The parameters update is performed as follows: the agent collects a finite batch of experiences interacting with the environment in a set of episodes $\tau$, then the loss-value function is evaluated as
 	\begin{equation}
@@
 	\label{eq:direct_policy_loss}\end{equation}
-	\begin{equation}
-	\hcancel[red]{\change{L(\theta_j) = D(0,T;\zeta)\hat{\mathbb{E}}\left[ \Phi(I_T(\alpha(\theta_j))) \right]}} \; ,
-	\end{equation}  
 	where the expectation $\hat{\mathbb{E}}[\dots]$ indicates the empirical average on the batch.
-Then the parameters are updated by plugging the policy into a stochastic gradient ascent algorithm \change{\cite{Sutton2018} with the aim of maximizing Equation~\eqref{eq:direct_policy_loss}}.
-\remove{The choice of Equation~\eqref{eq:direct_policy_loss} is justified by the fact that, when the algorithm will find the optimal set $\theta^*$, then $L(\theta^*)$ will be a good proxy of the optimal option price.}
+Then the parameters are updated by plugging the policy into a stochastic gradient ascent algorithm \cite{Sutton2018} with the aim of maximizing Equation~\eqref{eq:direct_policy_loss}.
 
@@
 	\subsection{Proximal policy optimization approach}
-	As for the direct policy \change{search} approach, we model the pricing problem considering an episode that takes place on the time-grid $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$.
+	As for the direct policy search approach, we model the pricing problem considering an episode that takes place on the time-grid $\mathcal{T}\coloneqq \{T_0,...,T_k,...,T_m\}$ with $T_0\coloneqq 0$ and $T_m \coloneqq T$.
 Again we choose as observation state the block defined in Equation~\eqref{eq:state} and the agent policy coincides with the risky allocation strategy $\alpha$ \eqref{eq:action}.
-\remove{An important difference from the previous method is that: once the agent has selected the action $a_k$ sampled from the current policy, it receives at the next time $T_{k+1}$ a reward $r_{k+1}$ generated by the environment.}
-
-	\remove{In our work, we have defined two different reward functions with the purpose to analyze which could help the agent to learn more efficiently the optimal policy $\pi^* = \pi^{\theta^*}$. 
-	Our first definition is} 
-	\begin{equation}
-		\hcancel[red]{\change{r_{k+1} =\left\{\begin{array}{ll}
-			D(0,T;\zeta)(I_T(\alpha(\theta_j)) - K)^+ & \text { if } T_{k+1} = T \\
-			0 & \text { otherwise }
-		\end{array}\right.}}
-	\; . \end{equation}
-	\remove{Therefore, during the whole episode, the agent receives a nil reward except at maturity when the reward coincides with the option intrinsic value.
-This choice may seem too daring because the agent receives a real feedback of its actions only at the end of the episode, increasing the probability to obtain a slow learning.
-However, if the agent has learnt $\pi^*$, the average cumulative reward per episode will coincide with the optimal TVO price.}
-
-	\change{As the PPO algorithm is more sophisticated than the previous one, we have decided to test its performance by means of two different reward functions.
-The former is the same as defined by Equation~\eqref{eq:reward1}, while the latter is given by}
-	\remove{The second reward function we have defined is}
+	As the PPO algorithm is more sophisticated than the previous one, we have decided to test its performance by means of two different reward functions.
+The former is the same as defined by Equation~\eqref{eq:reward1}, while the latter is given by
 	\begin{equation}
@@
 In this case, the reward function does not suffer of nil values for $0<T_k<T$: the RL agent always gets a feedback from the environment for its choices.
-\remove{The} \change{Notice that the} hyper-parameter $\gamma$ plays the role of a discount factor in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future.
+Notice that the hyper-parameter $\gamma$ plays the role of a discount factor in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future.
 	If we take the cumulative reward per episode and set the PPO parameters\footnote{We refer to \cite{Schulman2016} for a more detailed description for the generalized advantage estimation parameter $\lambda$.} $\gamma =\lambda =1$ we obtain
@@
 	In the \textsc{OpenAI Baselines} implementation of the PPO, the agent policy is parameterized again by a neural network; as for the previous method we have chosen an FFNN.
-\remove{In particular with PPO we deal with a stochastic policy, whose functional form is} \change{More precisely, training is based on a stochastic policy, i.e.
+More precisely, training is based on a stochastic policy, i.e.
 the action is drawn from a probability law, and the neural network is used to determine its parameters.
-In settings like ours where the action has a continuous domain, such law is} a multivariate diagonal Gaussian distribution where the mean $\mu^\theta(s_k)$ is the output vector of the FFNN and the log-standard deviation $\log\sigma$ is an \remove{external} \change{additional trainable} parameter:
+In settings like ours where the action has a continuous domain, such law is a multivariate diagonal Gaussian distribution where the mean $\mu^\theta(s_k)$ is the output vector of the FFNN and the log-standard deviation $\log\sigma$ is an additional trainable parameter:
 	\begin{equation}
@@
 The idea is that the log-standard deviation will be higher at the beginning of the training phase in order to guarantee a good exploration of the action space while it will be lower at the end to avoid too much noise in the proximity of the optimal policy.
-\change{Eventually at the end of training, the standard deviation is put to zero, so that the action becomes deterministically equal to the output of the FFNN.}
+Eventually at the end of training, the standard deviation is put to zero, so that the action becomes deterministically equal to the output of the FFNN.
 
@@
 	\begin{equation}\nonumber
-		I_0 = K = 1 \, \text{EUR} \; , \quad T=2 \, \text{yr} \; , \quad \bar{\sigma} = 5\% \; \change{,}
+		I_0 = K = 1 \, \text{EUR} \; , \quad T=2 \, \text{yr} \; , \quad \bar{\sigma} = 5\% \; ,
 	\end{equation}
-	\change{where $I_0$ is the starting value of the TVS-based portfolio.}
-	Moreover\change{,} without loss of generality\change{,} we consider the case of a\change{n unconstrained} \remove{completely free} allocation strategy $\alpha$.
-The extension to the constrained case is easy.
+	where $I_0$ is the starting value of the TVS-based portfolio.
+	Moreover, without loss of generality, we consider the case of an unconstrained allocation strategy $\alpha$. The extension to the constrained case is easy.
 
@@
 
-	Although in Section~\ref{subsec:BS_model} we have proved that under the Black and Scholes model the Equation~\eqref{eq:BS_optimal_strategy} solves the control problem, we want to take advantage of this \textit{a priori} solution as a benchmark to gather evidence on the robustness of our \remove{Reinforcement Learning} \change{RL} approach and to check if our analytical result is correct.
+	Although in Section~\ref{subsec:BS_model} we have proved that under the Black and Scholes model the Equation~\eqref{eq:BS_optimal_strategy} solves the control problem, we want to take advantage of this \textit{a priori} solution as a benchmark to gather evidence on the robustness of our RL approach and to check if our analytical result is correct.
 Moreover, we use the BS model as numerical laboratory to perform fine-tuning tests for the RL algorithms hyper-parameters and analyze how they impact the final results and performances.
 
-	We recall that in both the algorithms we parameterize the agent policy with an FFNN; thus this is completely characterized by the following hyper-parameters: number of hidden layers, number of neurons per hidden layer, \change{and} activation function per hidden layer.
-This is due to the fact that in this \remove{Reinforcement Learning} \change{RL} problem the number of the input neurons is equal to the state space dimension \eqref{eq:state}, while the number of the output ones coincides with the action space dimension \eqref{eq:action}.
+	We recall that in both the algorithms we parameterize the agent policy with an FFNN; thus this is completely characterized by the following hyper-parameters: number of hidden layers, number of neurons per hidden layer, and activation function per hidden layer.
+This is due to the fact that in this RL problem the number of the input neurons is equal to the state space dimension \eqref{eq:state}, while the number of the output ones coincides with the action space dimension \eqref{eq:action}.
 It is well known in the literature that neural networks give better performances in the training phase if the input data are well normalized \cite{Puheim2014}, \cite{Sola1997}.
@@
 In Equation~\eqref{eq:input_state} we have chosen as Markovian state $X_{T_k}$ the martingale term of the securities dynamics.
-In this way\change{,} we have that in the input block the variables have the same order of magnitude.
+In this way, we have that in the input block the variables have the same order of magnitude.
 
 	\subsection{Black and Scholes: hyper-parameters fine tuning}
-	We use the BS environment as \change{a} toy model to understand which parameters of the RL algorithm play key roles in the training and testing phase. 
+	We use the BS environment as a toy model to understand which parameters of the RL algorithm play key roles in the training and testing phase. 
 	Our first approach has been the direct policy one since it represents a natural way to tackle the problem: since our goal is to find the allocation 
@@
 
-	We try to investigate the following hyper-parameters: the FFNN architecture, in particular which feature between the depth and the width of the network is more important, the activation function\change{,} and the learning rate of the optimizer.
+	We try to investigate the following hyper-parameters: the FFNN architecture, in particular which feature between the depth and the width of the network is more important, the activation function, and the learning rate of the optimizer.
 Moreover, we compare the performances of two well-known optimizers in Machine Learning literature: Nadam \cite{Dozat2016} and RMSprop \cite{Hiton2012}.
-Since we deal with a free allocation strategy $\alpha$ that can assume negative values, we have chosen among the wide variety of activation functions \change{\cite{Goodfellow2016}} the tanh and the elu.
+Since we deal with a free allocation strategy $\alpha$ that can assume negative values, we have chosen among the wide variety of activation functions\cite{Goodfellow2016} the tanh and the elu.
 In this way, we can analyze the performance of a saturating activation function and a non-saturating one.
 Firstly we have performed a grid search on the learning rate starting value and we have found $10^{-3}$ a good choice in terms of speed of learning and avoiding over-fitting.
-\change{Note that to detect over-fitting one typically compares the performance on one or more training sets with that on corresponding validation sets.
-Since our environment is simulated, there is no need to explicitly set aside training data for such task, nor to perform $K$-fold validation: indeed we can always generate for the purpose a set of new scenarios which were not seen during optimization.}
+Note that to detect over-fitting one typically compares the performance on one or more training sets with that on corresponding validation sets.
+Since our environment is simulated, there is no need to explicitly set aside training data for such task, nor to perform $K$-fold validation: indeed we can always generate for the purpose a set of new scenarios which were not seen during optimization.
 	\begin{figure}[h!]
@@
 		\includegraphics[width=3.4in]{BS_fine_tuning_Nadam_elu_grey.eps}
-		\caption{Learning curves of the direct policy algorithm applied to the Black and Scholes \change{scenario}.
+		\caption{Learning curves of the direct policy algorithm applied to the Black and Scholes scenario.
 The solid lines are the loss functions defined in Equation~\eqref{eq:direct_policy_loss} in function of the number of training epochs.
@@
 		\includegraphics[width=3.4in]{BS_fine_tuning_RMSprop_elu_grey.eps}
-		\caption{Learning curves of the direct policy algorithm applied to the Black and Scholes \change{scenario}.
+		\caption{Learning curves of the direct policy algorithm applied to the Black and Scholes scenario.
 The solid lines are the loss values defined in Equation~\eqref{eq:direct_policy_loss} in function of the number of training epochs.
@@
 Each learning curve is compared with the optimal option price (grey-dashed line) that we have computed with the closed form solution we have derived for the BS model in Section~\ref{subsec:BS_model}.
-We can see that all the learning curves converge to our expected price, providing a numerical \remove{demonstration} \change{finding} of our theoretical result.
+We can see that all the learning curves converge to our expected price, providing a numerical finding of our theoretical result.
 More deeply, from the plots we observe that the RMSprop optimizer outperforms the Nadam.
@@
 	\begin{center}
-	\caption{Comparison of TVO prices under Black and Scholes \change{scenario}: analytical solution price and direct policy reinforcement learning.
+	\caption{Comparison of TVO prices under Black and Scholes scenario: analytical solution price and direct policy reinforcement learning.
 The option parameters are: $I_0=K=1\,[\text{EUR}]$, $T=2\,[\text{yr}]$ and $\bar{\sigma}=5\%$}
@@
 the spot price, $\nu_t=\nu(t,S_t)$.
-This additional dependency of volatility makes the problem of finding the optimal strategy non-trivial\change{, as a closed form formula cannot be derived under this scenario}; in fact if we consider the whole \remove{securities} \change{volatility} smiles \change{of the securities,} then the second-order term in the HJB equation \eqref{eq:HJB_general} \remove{is not zero and thus a closed formula for $\alpha^*$ is not available anymore} \change{does not elide and consequently the only way to derive $\alpha^*$ is through a numerical method}.
-\remove{Because of that, one must resort to numerical techniques to recover the problem solution.} \remove{Unlike the BS model where the strategy depends only on time, in LV dynamics there is no unnecessary information provided by the state block in Equation~\eqref{eq:state} to take the optimal action} The additional dependence on the state $S_t$ under LV dynamics will cause the agent to consider also the $X_t$ component value in the  state block \eqref{eq:state}, unlike under the BS scenario where this information is redundant as the optimal strategy was function only of time (see Proposition \ref{prop:bs}).
+This additional dependency of volatility makes the problem of finding the optimal strategy non-trivial, as a closed form formula cannot be derived under this scenario; in fact if we consider the whole volatility smiles of the securities, then the second-order term in the HJB equation \eqref{eq:HJB_general} does not elide and consequently the only way to derive $\alpha^*$ is through a numerical method.
+The additional dependence on the state $S_t$ under LV dynamics will cause the agent to consider also the $X_t$ component value in the  state block \eqref{eq:state}, unlike under the BS scenario where this information is redundant as the optimal strategy was function only of time (see Proposition \ref{prop:bs}).
 Here we consider the same market data ($r_t$, $\mu_t$ and $\nu_t$) as in the Black and Scholes environment to study how the optimal solution changes with the dynamics model.
@@
 Thus we train another agent with the direct policy learning by parameterizing its action with the baseline strategy and choosing a smaller learning rate of $10^{-4}$.
-With this parameterization, the RL agent actions, constituting the risky allocation strategy, are computed by summing at each observational time $T_k\in\mathcal{T}$ \remove{the risky allocation strategy is obtained by summing} the FFNN output with the Equation~\eqref{eq:analytical_solution_free_strategy}.
+With this parameterization, the RL agent actions, constituting the risky allocation strategy, are computed by summing at each observational time $T_k\in\mathcal{T}$ the FFNN output with the Equation~\eqref{eq:analytical_solution_free_strategy}.
 Even in this case, the corresponding learning curve (solid black line in Figure~\ref{fig:result_lv_direct_policy}) is stuck in the maximum drift strategy.
@@
 
+\cleardoublepage
 	\section{Conclusion and further developments}
